---
title: "compile_datasheets_HR"
author: "Devin Gamble"
date: "2022-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



### Load and Clean Data   

**Fitness Plant Data**  

Recipient IDs to double-check:
- HR_067 (HR_067B)

- Add in Donor information
```{r message = FALSE}
HR_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "HR_fitnessplants_23.csv")) %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient))  #rename - letters were added to mark position

HR_donors <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "HR_donor_info.csv")) 

#Fix Recipient and Donor names
HR_donors2 <- HR_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("HR_", Donor_num))) %>% 
  select(-Recipient_num, -Donor_num)



#Multiple matches, merge only unique pairs for Donor
HR_Fitness_2 <- merge(HR_Fitness, unique(HR_donors2[c("Gen", "Donor", "Recipient")]), by = c("Recipient", "Gen"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Sequence, SegPos)

#Save as V2 Fitness Plant Sheet
write_csv(HR_Fitness_2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "HR_fitnessplants_23_V2.csv"), na = "")


```

<br>  

### Trait Data  

Trait data gathered and processed from the 2022-2023 field season is loaded and organized below.  


#### Flower Check & Phenology  
-- Organize flowering data
-- Add in Sow & Plant dates

#########
Crucial Step!!  
######### 
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R (and lubridate) much easier!  

```{r message = FALSE}
HR_flr_df <- read_csv(file = here::here( "data_sheets", "2023_year2_data", "flowering_checks", "HR_flowercheck_data_2023-02-23 - HR_flowercheck_data_v1.csv"), na = c("", "NA")) 


HR_flr2 <- HR_flr_df %>% 
  mutate(SegPos = paste(Segment, side, sep = "")) %>% 
  select(Sequence, Transect, SegPos, everything(), -c(Segment, side, FD_06, FD_07)) %>% 
  arrange(Sequence, Transect, SegPos) %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient)) #rename- letters were added to mark position
  #Gopher notes "G" inside FD columns removed manually

#Check for mismatched fitness plants
HR_flr_test <- HR_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))

## Fix Dates!

#Check obs first using unique(HR_flr3$FD_01) and make corrections

HR_flr3 <- HR_flr2 %>%
  mutate_at(.vars = vars(FD_01:FD_05), ~as_date(.))

```

Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

HR_flr4 <- HR_flr3 %>% 
  mutate(Sow_Date = as_date("2022-11-10"),
         Plant_Date = as_date("2022-11-13"))

```


Merge in Fitness plant Data  
- Donor

```{r}
#Check to make sure fitness plants line up
HR_flr5 <- merge(HR_flr4, HR_Fitness_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Transect, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>% 
  select(-c(Fit_plant.x, Fit_plant.y)) #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  
  
#####  
#Check to make sure fitness plants line up
  # HR_flr_test <- HR_flr5 %>% 
  #   filter(!is.na(Fit_plant.x)) %>%
  #   mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
  #                           Fit_plant.x != Fit_plant.y ~ FALSE))

```


**Phenology Calculations**    

Survivorship calculated later

```{r}
#Same as last year

HR_flowering <- HR_flr5 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())


#TEMP FILE
write_csv(HR_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "HR_phenology_2023.csv"), na = "")

```

<br>  



#### Fruit and Seed counts    

Fruit & seed counts were recorded on the same spreadsheet, unlike last year. Loaded here and checked to make sure all ID info lines up with the Fitness data sheet.

```{r message = FALSE}
#Load Fruit & Seeds count data

HR_fruit_seeds <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "HR_fruits_seeds_2023 - HR_fruit_seed_counts.csv"), skip = 1) %>% 
  rename(SegPos = Seg)


# #Remove rows where no fruits collected
# HR_fruit_seeds_2 <- HR_fruit_seeds %>% 
#   filter(total_fruits != 0)


##Check Fitness plants
HR_merge_test <- merge(HR_Fitness_2, HR_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient", "Fit_plant"), all.y = TRUE) 


HR_fruit_seeds_3 <- merge(HR_Fitness_2, HR_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient", "Fit_plant"), all.x = TRUE) %>% 
  select(-Sequence.y) %>% 
  rename(Sequence = Sequence.x) %>% 
  rename(Plant_ID = Fit_plant)

```

<br>  

#### Skeleton Weights  

Load and merge-in these data sheets
```{r message = FALSE}

skel_weights_HR <- read_csv(here::here("data_sheets", "2023_year2_data",  "skeleton_data", "Skeleton Data 2023 - HR Skeletons.csv")) %>% 
  rename(position = side, Gen = 4, Fitness = 7, skel_weight_mg = 8, complete_incomplete = 9, selected = 10) %>% 
  arrange(Transect, Segment, position) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Transect, SegPos, Gen,  Recipient, Plant_ID, Fitness, skel_weight_mg) #complete/selected not included

skel_weights_HR_2 <- skel_weights_HR %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient)) #rename- letters were added to mark position

#Check that IDs line up
skelcheck <- merge(HR_Fitness_2, skel_weights_HR_2, by = c("Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>% 
  mutate(FTRUE = case_when(Fit_plant == Plant_ID ~ TRUE,
                           TRUE ~ FALSE))


#Merge with full data set
#HR_merged_df5 <- merge(HR_merged_df4, skel_weights_HR, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)


```



**TEMP MERGED DATA FRAME**  
Merge trait data processed thus far

```{r}

HR_merged_1 <- merge(HR_flowering, skel_weights_HR, by = c("Transect", "SegPos", "Gen", "Recipient", "Plant_ID"), all.x = TRUE) %>% select(-c(Skeleton:Fitness)) %>% arrange(Transect, Sequence, SegPos)

HR_merged_2 <- merge(HR_merged_1, HR_fruit_seeds_3, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient", "Donor", "Plant_ID"), all.x = TRUE) %>% select(-c(FD_01:FD_05, N_CLOSED_fruit_D1:notes)) %>% arrange(Transect, Sequence, SegPos) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, everything())

```



<br>  


#### Seed Weights  

Update: Use revised seed counts in seed weight data sheet. Viable seed counts were corrected in this data sheet, so these values should be used instead of the ones from the seed count data sheet.

Note: 
- Sequence is broken in this data sheet
Day 5 (collection NA/missing) data empty

**Manual Corrections TBD**
- RE-WEIGH SEEDS or Convert 0s to NAs for D1 (T5 s19R; T9 s11L), D2 (T9 s10R; T4 s4L)

```{r}

#Load Data
#Rename columns to standard format

HR_seedmass_2023 <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "HR_Seed_Weights_2023-All_Data.csv"), skip = 1) %>% 
  select(-c(5)) %>% #rm Symbol column
  rename(SegPos = Seg, Plant_ID = Fit_plant, msm_all = 7, 
         closed_fruits_d1 = 8, viable_seeds_d1 = 9, seeds_weighed_d1 = 10, seed_mass_mg_d1 = 11, mean_mass_mg_d1 = 12, 
         closed_fruits_d2 = 13, viable_seeds_d2 = 14, seeds_weighed_d2 = 15, seed_mass_mg_d2 = 16, mean_mass_mg_d2 = 17,
         closed_fruits_d3 = 18, viable_seeds_d3 = 19, seeds_weighed_d3 = 20, seed_mass_mg_d3 = 21, mean_mass_mg_d3 = 22,
         closed_fruits_d4 = 23, viable_seeds_d4 = 24, seeds_weighed_d4 = 25, seed_mass_mg_d4 = 26, mean_mass_mg_d4 = 27) %>% 
  select(-c(1, 28:33)) %>% 
  filter(!is.na(Recipient))


#Calculate number of viable seeds 
#Calculate SD of grand mean seed mass

HR_seedmass_2023_B <- HR_seedmass_2023 %>% 
  mutate(ALL_filled_seeds_N = rowSums(across(c(viable_seeds_d1, viable_seeds_d2, viable_seeds_d3, viable_seeds_d4)), na.rm = TRUE)) %>% #sum viable seed counts across dates (N for NEW)
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(mean_mass_mg_d1, mean_mass_mg_d2, mean_mass_mg_d3, mean_mass_mg_d4)), na.rm = TRUE)) %>%  #calculate total standard deviation of mean seed mass
  select(Transect:Plant_ID, ALL_filled_seeds_N, msm_all, sd_msm_all, everything())
#SD only available when viable seeds from two or more different dates were weighed 
#Each statistical 'unit' is the mean seed weight calculated on a given date
         
  
##Merge with fruit/seed count data frame

HR_merged_2.5 <- merge(HR_merged_2, HR_seedmass_2023_B, by = c("Transect", "SegPos", "Gen", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  select(Sequence, everything()) %>% #put Sequence at beginning
  arrange(Sequence, Transect, SegPos, Plant_ID)
#all.x means that any fitness plant x ID combos that don't match up won't be included.
#Can toggle to all.y = TRUE to confirm merge success (e.g., all rows have matching donor, F_plant status)


#Check merge success
#Only fitness plants should have seed weight data
HR_merge_check <- HR_merged_2.5 %>% 
  filter(!is.na(msm_all))

HR_merge_check2 <- HR_merged_2.5 %>% 
  filter(!is.na(ALL_filled_seeds_N)) #Should match original seed count Nrows

## Correct viable seed counts
HR_merged_3 <- HR_merged_2.5 %>% 
  mutate(seed_recount_diff = ALL_filled_seeds_N - ALL_filled_seeds) #%>% #Check data sheets for high diffs
  select(-c(seed_recount_diff, ALL_filled_seeds)) #comment out this line to check before proceeding

```

<br>  




#### Corolla measurements  
[[UPDATE]]
- Format dates in excel prior to importing
```{r}
corollas_HR <- read_csv(here::here("data_sheets", "2023_year2_data", "corolla_leaf_data", "Corolla Diameters 2023 - HR_corollas.csv")) %>% select(-c(4,10, 11)) %>% 
  rename(corolla_notes = Notes, photo_date = `Photo date`) %>% 
  filter(!grepl("REMOVE", corolla_notes)) ## Filter out notes with "REMOVE"

corollas_HR2 <- corollas_HR %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually 
  arrange(Transect, Recipient, Plant_ID, photo_date) %>% 
  distinct(Transect, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates
#43 duplicates removed

## Merge with main data 


##Create segment-wide corolla diameter column
  
```

<br>  


#### Leaf Area & Weight (SLA)  
[[OLD - UPDATE]]

```{r}
leaves_HR <- read_csv(here::here("data_sheets", "leaf_data", "HR_SLA_data_20230522.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) %>% #exlcude NA leaf areas?
  mutate(leaf_dryweight_mg = as.numeric(leaf_dryweight_mg))


#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_HR_duped <- leaves_HR %>% 
  mutate(duped1 = duplicated(select(leaves_HR, c(Block, Transect, Recipient))),
         duped2 = duplicated(select(leaves_HR, c(Block, Transect, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!

#This seems to work the best!
leaves_HR_unique <- leaves_HR_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) %>% #Don't need plant ID...these were all non-fitness plants
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg,
         LMA = 1 / SLA)

#Merge in to main data 

HR_merged_df7 <- merge(HR_merged_df6, leaves_HR_unique, by = c("Block", "Transect", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo


```

<br>  



#### d13C data (WUE)  
[[OLD - UPDATE]]

```{r}
Nemo_d13C <- read_csv(here::here("data_sheets", "isotope_data", "Nemophila_d13C_230511.csv")) %>% 
  select(1,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error

#filter to HR samples
Nemo_d13C_HR <- Nemo_d13C %>% 
  filter(grepl('HR', sample_ID))

#unravel sample ID
Nemo_d13_HR2 <- Nemo_d13C_HR %>% 
  mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 6)), 
         Transect = as.numeric(str_remove(substr(sample_ID, 8, 10), "^0+")),
         Plant_ID = as.numeric(substr(sample_ID, 12,13))) %>% 
  select(Transect, Recipient, Plant_ID, d13C, sample_weight_mg, -Plant_ID)

#Merge

HR_merged_df_fl3 <- merge(HR_merged_df_fl2, Nemo_d13_HR2, by = c("Recipient", "Transect"), all.x = TRUE)
#Merge by Recipient and Transect only: genotype-wide estimates of d13C since some plants were NF



#check that all data made it (266 obs)
testdf <- HR_merged_df_fl3 %>% 
  filter(!is.na(d13C))
```



**Error checks**  

Checking a variety of IDs and trait data to make sure everything lines up.  

```{r eval = FALSE}
##Check for mismatched (fitness) plants

HR_Echeck1 <- HR_merged_df6 %>% 
  filter(any_FitP == FALSE & seed_ct != 0) #do the same for seed count, using `>` and `!=`

HR_Echeck2 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
HR_Echeck3 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))


HR_Echeck4 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))

## Check for NA seed counts for fitness plants (when fruit production > 0)
HR_Echeck5 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))


#Seed mass
HR_Echeck5 <- HR_merged_df6 %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


### Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd`  

Phenology calculations performed above.


```{r}
#Survival measures
HR_merged_df1 <- HR_merged_2 %>% 
  mutate(surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(F_plant == TRUE & total_fruits > 0 ~ 1, #plants that survived to produce fruit
                                       F_plant == TRUE & total_fruits == 0 ~ 0),
         prod_closed_fruits = case_when(F_plant == TRUE & ALL_closed_fruit > 0 ~ 1,
                                            F_plant == TRUE & ALL_closed_fruit == 0 ~ 0),
         prod_filled_seeds = case_when(F_plant == TRUE & ALL_filled_seeds > 0 ~ 1, #plants produced filled seeds
                                      F_plant == TRUE & ALL_filled_seeds == 0 ~ 0)) #else = 0 (applies to seed_ct = NA)

         
HR_merged_df2 <- HR_merged_df1 %>% 
  rename(mean_seeds_per_fruit = avg_seeds_per_fruit) %>% 
  mutate(across(c(est_fecundity, mean_seeds_per_fruit), ~ as.numeric(na_if(., "#DIV/0!")))) %>% #Converts to NA 
  mutate(est_fecundity = case_when(F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit == 0 ~ 0,
                                   F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit >= 1 ~ NA,
                                               F_plant == TRUE & !is.na(est_fecundity) ~  est_fecundity)) 

# estimated fecundity is 0 when 0 closed or open fruits collected, or when 0 filled seeds collected
# estimated fecundity is NA when 0 closed fruits but 1 or more open fruits collected

```

<br>  



### Master sheets (2023)  

Full and summarized data sheets for all data collected in the 2022-2023 field season. 
These are merged with the year 1 data sheets elsewhere.  

```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...


######
#TEMP VERSION
######

HR_merged_full_1 <- HR_merged_df2 %>% 
  filter(F_plant == TRUE)
##NOTES - there are still fitness plants that never flowered... keep as 0s


write_csv(HR_merged_full_1, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "HR_traits_Y2_incomplete_20230922.csv"), na = "")

######
######
```


**TO BE UPDATED**  
```{r eval = FALSE}
HR_mastersheet_v1 <- HR_merged_df_fl3 %>% 
  mutate(Year = 2022, Gen = "G1") %>% 
  select(Year, Gen, Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

write_csv(HR_mastersheet_v1, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "HR_mastersheet_full_Y2_2023-08-07.csv"))

#Create meta data sheet



#
## Fitness Plants Only
#

HR_mastersheet_F_mains <-  HR_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only single fitness plants

HR_mastersheet_F_all <- HR_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treate these as 0 fitness or as NAs to remove???


write_csv(HR_mastersheet_F_mains, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "HR_mastersheet_Fitness-mains_Y2_2023-08-07.csv"))

write_csv(HR_mastersheet_F_all, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "HR_mastersheet_Fitness-all_Y2_2023-08-07.csv"))

```



