---
title: "compile_datasheets_HR"
author: "Devin Gamble"
date: "2022-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



## Load and Clean Data   

**Load Fitness Plant Data**  

- Add in Donor information
```{r}
HR_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "HR_fitnessplants_23.csv")) %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient))  #rename - letters were added to mark position

HR_donors <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "HR_donor_info.csv")) 

#Fix Recipient and Donor names
HR_donors2 <- HR_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("HR_", Donor_num))) %>% 
  select(-Recipient_num, -Donor_num)



#Multiple matches, merge only unique pairs for Donor
HR_Fitness_2 <- merge(HR_Fitness, unique(HR_donors2[c("Gen", "Donor", "Recipient")]), by = c("Recipient", "Gen"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Sequence, SegPos)

#Save as V2 Fitness Plant Sheet
write_csv(HR_Fitness_2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "HR_fitnessplants_23_V2.csv"), na = "")


```






**Flower Check sheets**  
-- Organize flowering data
-- Add in Sow & Plant dates


##################
**Crucial Step!!**  
##################  
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R much easier!  



```{r}
##HR##

HR_flr_df <- read_csv(file = here::here( "data_sheets", "2023_year2_data", "flowering_checks", "HR_flowercheck_data_2023-02-23 - HR_flowercheck_data_v1.csv"), na = c("", "NA")) 


HR_flr2 <- HR_flr_df %>% 
  mutate(SegPos = paste(Segment, side, sep = "")) %>% 
  select(Sequence, Transect, SegPos, everything(), -c(Segment, side, FD_06, FD_07)) %>% 
  arrange(Sequence, Transect, SegPos) %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient)) #rename- letters were added to mark position
  #Gopher notes "G" inside FD columns removed manually

#Check for mismatched fitness plants
HR_flr_test <- HR_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))

## Fix Dates!

#Check obs first using unique(HR_flr3$FD_01) and make corrections

HR_flr3 <- HR_flr2 %>%
  mutate_at(.vars = vars(FD_01:FD_05), ~as_date(.))

```


Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

HR_flr4 <- HR_flr3 %>% 
  mutate(Sow_Date = as_date("2022-11-10"),
         Plant_Date = as_date("2022-11-13"))

```


Merge in Fitness plant Data  
- Donor

```{r}
#Check to make sure fitness plants line up
HR_flr5 <- merge(HR_flr4, HR_Fitness_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Transect, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>% 
  select(-c(Fit_plant.x, Fit_plant.y)) #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  
  
#####  
#Check to make sure fitness plants line up
  # HR_flr_test <- HR_flr5 %>% 
  #   filter(!is.na(Fit_plant.x)) %>%
  #   mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
  #                           Fit_plant.x != Fit_plant.y ~ FALSE))

```


#### Phenology Calculations  
<<Move these lower later??>>  

**ADD IN SURVIVORSHIP MEASURES**  




```{r}
#Same as last year

HR_flowering <- HR_flr5 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())



#TEMP FILE - delete later
write_csv(HR_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "HR_phenology_2023.csv"), na = "")

```




**Fruit and Seed counts**  

Due to the structure of excel data sheets, it seems easiest to upload fruit and seed count data separately since different spreadsheets were used to record each. Info on plants' Block, Transect, sequence, etc. will be pulled from the fruit count sheets (assumed to be up-to-date). For sheets intended to record fruit production only, counts of viable seeds, mean seed weight, etc. that are copied over are ignored and instead pulled straight from sheets where they were first entered later on. 

Data on fitness plants will be more accurate in the fruit-cound data sheet, while the seed-count sheet will have a handful of rows with non-fitness plants wihtout seed counts.

Date added: **Feb 06, 2023**  


```{r}
#Load Fruit & Seeds count data

HR_fruit_seeds <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "HR_fruits_seeds_2023 - HR_fruit_seed_counts.csv"), skip = 1) %>% 
  rename(SegPos = Seg)


#Remove rows where no fruits collected
HR_fruit_seeds_2 <- HR_fruit_seeds %>% 
  filter(total_fruits != 0)



##Check Fitness plants
HR_merge_test <- merge(HR_Fitness_2, HR_fruit_seeds_2, by = c("Transect", "SegPos", "Gen", "Recipient"), all.y = TRUE) 


#Remove rows 

```




*NOTE*: Update with later versions of data sheet when more seed counts have been entered
```{r message = FALSE}
#
# Fruit count data
# Data on *when* fruits were collected will be omitted here; only the total fruit counts (closed, open, non-fitness) are retained from the csv

HR_fruitprod_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "HR-fruit-production_2023-02-05.csv"), skip = 1) %>% 
  select(1:10) #%>% #omit sow and flower date, seed count, and date-specific fruit counts by column index
  #filter_all(any_vars(!is.na(.))) #omit empty rows at end of df

HR_fruitprod_2 <- HR_fruitprod_1 %>% 
  select(-c(Recipient)) %>% # Sequence in this data set is incorrect
  rename(Recipient = `Recipient for sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, closed_fruits_F = `total_closed_fruits`, open_fruits_F = `total_open_fruits`, damaged_fruits_F = `eaten_damaged_fruits`, total_fruits_F = `total_fruits (open and closed)`, nonfitness_fruits = `nonfitness_fruits`) 

#Note: total fruits should not include eaten/damaged fruit



#
# Seed count data (viable seeds)
# Similarly, only seed count data are retained from this spreadsheet to minimize errors

HR_seedct_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "HR-seed-count_2023-02-05.csv"), skip = 1) %>% 
  select(-c(3, 7:33)) %>% 
  rename(Recipient = `Recipient for sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, seed_ct = `viable_seeds`) %>% 
  mutate(Plant_ID = as.numeric(Plant_ID)) 
  

#merge data
HR_fruitseed_ct1 <- left_join(HR_fruitprod_2, HR_seedct_1, by = c("Transect", "Recipient", "Plant_ID"))
#Sequence data may not be accurate - add in later
#Rows without matches in fruit_prod data sheet are not included


# summary variables, tidying up
HR_fruitseed_ct2 <- HR_fruitseed_ct1 %>% 
  select(Transect, Recipient, Plant_ID, closed_fruits_F, open_fruits_F, damaged_fruits_F, total_fruits_F, nonfitness_fruits, everything()) %>% 
  select(-c(Sequence.x, Sequence.y)) %>% 
  arrange(Transect, Recipient)

## NOTE: total number of non-fitness fruits per segment may be under-represented here since not all counts were matched to fitness plants and therefore merged in.

```



**Seed Weights**  


```{r}
# Read in only # seeds weighed and total weight
# Calculate mean weight per seed

HR_seedmass_1 <- read_csv(here::here("data_sheets", "seed_mass", "HR_Seed_Mass_2023-02-05.csv"), skip = 1) %>% 
  select(-2) %>% slice(1:453)
  

colnames(HR_seedmass_1) <- c("Transect", "Recipient", "Plant_ID", "closed_fruits_d1", "viable_seeds_d1", "seeds_weighed_d1", "seed_mass_mg_d1", "closed_fruits_d2", "viable_seeds_d2", "seeds_weighed_d2", "seed_mass_mg_d2", "closed_fruits_d3", "viable_seeds_d3", "seeds_weighed_d3", "seed_mass_mg_d3", "closed_fruits_d4", "viable_seeds_d4", "seeds_weighed_d4", "seed_mass_mg_d4", "closed_fruits_d5", "viable_seeds_d5", "seeds_weighed_d5", "seed_mass_mg_d5", "closed_fruits_d6", "viable_seeds_d6", "seeds_weighed_d6", "seed_mass_mg_d6", 
                             "closed_fruits_d7", "viable_seeds_d7", "seeds_weighed_d7", "seed_mass_mg_d7",
                             "closed_fruits_d8", "viable_seeds_d8", "seeds_weighed_d8", "seed_mass_mg_d8", 
                             "closed_fruits_d9", "viable_seeds_d9", "seeds_weighed_d9", "seed_mass_mg_d9")


HR_seedmass_2 <- HR_seedmass_1 %>% 
  select(-starts_with(c("closed_fruits_", "viable_seeds_"))) %>% 
  mutate(msm_d1 = seed_mass_mg_d1/seeds_weighed_d1,
         msm_d2 = seed_mass_mg_d2/seeds_weighed_d2,
         msm_d3 = seed_mass_mg_d3/seeds_weighed_d3,
         msm_d4 = seed_mass_mg_d4/seeds_weighed_d4,
         msm_d5 = seed_mass_mg_d5/seeds_weighed_d5,
         msm_d6 = seed_mass_mg_d6/seeds_weighed_d6,
         msm_d7 = seed_mass_mg_d7/seeds_weighed_d7,
         msm_d8 = seed_mass_mg_d8/seeds_weighed_d8,
         msm_d9 = seed_mass_mg_d9/seeds_weighed_d9)
         

HR_seedmass_3 <- HR_seedmass_2 %>%
  mutate(msm_all = rowMeans(select(HR_seedmass_2, starts_with("msm_")), na.rm = TRUE)) %>% 
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(msm_d1:msm_d9)), na.rm = TRUE))
         
  
##Merge with fruit/seed count data frame

HR_fruitseed_ct3 <- merge(HR_fruitseed_ct2, HR_seedmass_3, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```


Merge everything together:  
```{r}
HR_merged_df1 <- merge(HR_GermFlow_3, HR_fruitseed_ct3, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, everything()) %>%
  arrange(Block, Transect, Sequence, Recipient, Plant_ID) %>% 
  mutate(closed_fruits_F = as.numeric(closed_fruits_F),
         total_fruits_F = as.numeric(total_fruits_F)) #correct counts to numeric values


#Correct FruitNAs to 0s in cases where no fruit data was entered (fitness and non-fitness plants):

#
HR_merged_df2 <- HR_merged_df1 %>% 
  mutate(closed_fruits_F = replace_na(closed_fruits_F, 0),
         total_fruits_F = replace_na(total_fruits_F, 0),
         seed_ct = replace_na(seed_ct, 0))

#Calculate mean seeds per fruit, total_est_seed prod
HR_merged_df3 <- HR_merged_df2 %>% 
  mutate(mean_seeds_per_fruit = seed_ct/closed_fruits_F,
         total_est_seed_production = mean_seeds_per_fruit*total_fruits_F) %>% 
  mutate(mean_seeds_per_fruit = case_when(any_FitP == TRUE & closed_fruits_F == 0 ~ 0, 
                                          TRUE ~ mean_seeds_per_fruit)) %>%  #else, original value
  mutate(total_est_seed_production = case_when(any_FitP == TRUE & mean_seeds_per_fruit == 0 ~ 0,
                                               TRUE ~ total_est_seed_production))


#mean seeds per fruit and total est fruit production have NAs/NaNs for when there were no closed fruits/seeds (fitness plants only)



#Correct formatting issues

HR_merged_df4 <- HR_merged_df3 %>% 
  mutate(Fl_3.16 = as_date(Fl_3.16),
         Fl_3.24 = as_date(Fl_3.24),
         Fl_3.31 = as_date(Fl_3.31),
         Fl_4.7 = as_date(Fl_4.7),
         Fl_4.14 = as_date(Fl_4.14),
         Fl_4.20 = as_date(Fl_4.20),
         Fl_4.26 = as_date(Fl_4.26),
         Fl_5.20 = as_date(Fl_5.20))


```


**Skeleton Weights & Corolla area and diameter**  

Load and merge-in these data sheets
```{r}

skel_weights_HR <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeleton_weights_HR_2023_03_01.csv")) %>% select(-Units, -7, -8) %>% 
  filter(!grepl("EXCLUDE.", Notes)) %>% ## Filter out notes with "Exclude"
  rename(AG_biomass_mg = `Skeleton Weights (mg)`)

#Watered points
skel_weights_HR_W <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeletons_watered_HR_2023_02_08.csv")) %>% 
  select(-`Units`) %>% 
  filter(!grepl("EXCLUDE.", Notes)) %>% ## Filter out notes with "Exclude"
  rename(Plant_ID = `Plant ID`, AG_biomass_mg = `Skeleton Weights`)

#Combine
skel_weights_all <- rbind(skel_weights_HR, skel_weights_HR_W) %>% 
  select(-Notes) %>% 
  filter(!is.na(AG_biomass_mg)) %>% 
  mutate(AG_biomass_mg = as.numeric(AG_biomass_mg))


#Merge with full data set
HR_merged_df5 <- merge(HR_merged_df4, skel_weights_all, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)


```


Corolla measurements
```{r}
corollas_HR <- read_csv(here::here("data_sheets", "corolla_measures", "corollameasures_HR_2023-02-28.csv")) %>% select(1, 3, 4, 6, 13, 16, 18) %>% 
  rename(photo_date = 4, corolla_diam_mm = 5, corolla_area_mm2 = 6, notes_corolla = 7) %>% 
  filter(!grepl("REMOVE", notes_corolla)) ## Filter out notes with "REMOVE"

corollas_HR2 <- corollas_HR %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  arrange(Transect, Recipient, Plant_ID, photo_date) %>% 
  distinct(Transect, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates


## Merge with main data 
HR_merged_df6 <- merge(HR_merged_df5, corollas_HR2, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)
  
```


**Leaf Area & Weight**  
```{r}
leaves_HR <- read_csv(here::here("data_sheets", "leaf_data", "HR_SLA_data_20230522.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) %>% #exlcude NA leaf areas?
  mutate(leaf_dryweight_mg = as.numeric(leaf_dryweight_mg))


#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_HR_duped <- leaves_HR %>% 
  mutate(duped1 = duplicated(select(leaves_HR, c(Block, Transect, Recipient))),
         duped2 = duplicated(select(leaves_HR, c(Block, Transect, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!

#This seems to work the best!
leaves_HR_unique <- leaves_HR_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) %>% #Don't need plant ID...these were all non-fitness plants
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg,
         LMA = 1 / SLA)

#Merge in to main data 

HR_merged_df7 <- merge(HR_merged_df6, leaves_HR_unique, by = c("Block", "Transect", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo


```

Error checks:

```{r eval = FALSE}
##Check for mismatched (fitness) plants

HR_Echeck1 <- HR_merged_df6 %>% 
  filter(any_FitP == FALSE & seed_ct != 0) #do the same for seed count, using `>` and `!=`

HR_Echeck2 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
HR_Echeck3 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))


HR_Echeck4 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))

## Check for NA seed counts for fitness plants (when fruit production > 0)
HR_Echeck5 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))


#Seed mass
HR_Echeck5 <- HR_merged_df6 %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


## Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd`  


#### Phenology  

```{r}

HR_merged_df_fl <- HR_merged_df7 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('Fl_')), na.rm = TRUE)), #First flowering date
         days_germ2flower = as.numeric(FFD - Germ_Date), #days from germination to flowering
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from sow to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('Fl_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) #Flowering duration


#Survival measures
HR_merged_df_fl2 <- HR_merged_df_fl %>% 
  mutate(surv_to_germ = case_when(!is.na(Germ_Date) ~ 1, #plants that germinated
                                  is.na(Germ_Date) ~ 0),
         surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(total_fruits_F > 0 ~ 1, #plants that survived to produce fruit
                                       total_fruits_F == 0 ~ 0),
         surv_to_seedprod = case_when(seed_ct > 0 ~ 1, #plants that survived to produce fruit with viable seeds
                                       TRUE ~ 0)) #else = 0 (applies to seed_ct = NA)

         


## Note: should flowering duration remain unchanged when a plant went several weeks without flowering (e.g., had two temporally distinct blooms during the growing season)?
```

<br>  


### d13C data  

```{r}
Nemo_d13C <- read_csv(here::here("data_sheets", "isotope_data", "Nemophila_d13C_230511.csv")) %>% 
  select(1,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error

#filter to HR samples
Nemo_d13C_HR <- Nemo_d13C %>% 
  filter(grepl('HR', sample_ID))

#unravel sample ID
Nemo_d13_HR2 <- Nemo_d13C_HR %>% 
  mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 6)), 
         Transect = as.numeric(str_remove(substr(sample_ID, 8, 10), "^0+")),
         Plant_ID = as.numeric(substr(sample_ID, 12,13))) %>% 
  select(Transect, Recipient, Plant_ID, d13C, sample_weight_mg, -Plant_ID)

#Merge

HR_merged_df_fl3 <- merge(HR_merged_df_fl2, Nemo_d13_HR2, by = c("Recipient", "Transect"), all.x = TRUE)
#Merge by Recipient and Transect only: genotype-wide estimates of d13C since some plants were NF



#check that all data made it (266 obs)
testdf <- HR_merged_df_fl3 %>% 
  filter(!is.na(d13C))
```


## Master sheets (v1)  


```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...


  HR_mastersheet_v1 <- HR_merged_df_fl3 %>% 
  mutate(Year = 2022, Gen = "G1") %>% 
  select(Year, Gen, Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

write_csv(HR_mastersheet_v1, here::here("data_sheets", "compiled_sheets", "HR_mastersheet_full_2023-08-07.csv"))

#Create meta data sheet



#
## Fitness Plants Only
#

HR_mastersheet_F_mains <-  HR_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only single fitness plants

HR_mastersheet_F_all <- HR_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treate these as 0 fitness or as NAs to remove???


write_csv(HR_mastersheet_F_mains, here::here("data_sheets", "compiled_sheets", "HR_mastersheet_Fitness-mains_2023-08-07.csv"))

write_csv(HR_mastersheet_F_all, here::here("data_sheets", "compiled_sheets", "HR_mastersheet_Fitness-all_2023-08-07.csv"))

```



