---
title: "compile_datasheets_BO"
author: "Devin Gamble"
date: "2022-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



### Load and Clean Data   


**Load Fitness Plant Data**  

- Add in Donor information
```{r message = FALSE}
BO_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "BO_fitnessplants_23.csv")) %>% 
  slice(-620)

#Donor info loaded from 2021-22 data
BO_donors <- read_csv(here::here("data_sheets", "tray_positions_2021_year1","BO_tray_rand.csv"), skip = 1) %>% 
  select(Donor_ID:Recipient_ID) %>% 
  rename(Donor = Donor_ID, Recipient = Recipient_ID)

#Fix Recipient and Donor names
BO_donors2 <- BO_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("BO_", Donor_num))) %>% 
  select(-Recipient_num, -Donor_num)



#Multiple matches, merge only unique pairs for Donor
BO_Fitness_2 <- merge(BO_Fitness, unique(BO_donors2[c("Donor", "Recipient")]), by = c("Recipient"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Sequence, SegPos)

#Save as V2 Fitness Plant Sheet
write_csv(BO_Fitness_2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "BO_fitnessplants_23_V2.csv"), na = "")


```

<br>  


### Trait Data  

Trait data gathered and processed from the 2022-2023 field season is loaded and organized below.  


#### Flower Check & Phenology  
-- Organize flowering data
-- Add in Sow & Plant dates

#########
Crucial Step!!  
######### 
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R (and lubridate) much easier!  

```{r message = FALSE}

BO_flr_df <- read_csv(file = here::here( "data_sheets", "2023_year2_data", "flowering_checks", "BO_flowercheck_data_2023-02-24 - BO_flowercheck_data_v1.csv"), na = c("", "NA")) %>% 
  rename(F_Num_02 = 19)


BO_flr2 <- BO_flr_df %>% 
  mutate(SegPos = paste(Segment, side, sep = "")) %>% 
  select(Sequence, Transect, SegPos, everything(), -c(Segment, side, FD_06, FD_07)) %>% 
  slice(-c(698:700, 779:780, 3096:3102)) %>% #Remove individuals not planted and NA rows
  arrange(Sequence, Transect, SegPos)
  

#Check for mismatched fitness plants
BO_flr_test <- BO_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))


## Fix Dates!

#Check obs first using unique(BO_flr3$FD_01) and make corrections

BO_flr3 <- BO_flr2 %>%
  mutate_at(.vars = vars(FD_01:FLR_P), ~ as_date(.))

```


Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

BO_flr4 <- BO_flr3 %>% 
  mutate(Sow_Date = as_date("2022-10-31"),
         Plant_Date = as_date("2022-11-03"))

```


Merge in Fitness plant Data  
- Donor
```{r}
#Check to make sure fitness plants line up
BO_flr5 <- merge(BO_flr4, BO_Fitness_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Transect, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>% 
  select(-c(Fit_plant.x, Fit_plant.y)) #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  
  
#####  
#Check to make sure fitness plants line up
# BO_flr_test <- BO_flr5 %>% 
#   filter(!is.na(Fit_plant.x)) %>%
#   mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
#                            Fit_plant.x != Fit_plant.y ~ FALSE))
# #
```



**Phenology Calculations**  
 

```{r}
#Same as last year

BO_flowering <- BO_flr5 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())



#TEMP FILE - delete later
write_csv(BO_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BO_phenology_2023.csv"), na = "")

```


<br>  


#### Fruit and Seed counts  

Fruit & seed counts were recorded on the same spreadsheet, unlike last year. Loaded here and checked to make sure all ID info lines up with the Fitness data sheet.


```{r message = FALSE}
#Load Fruit & Seeds count data

BO_fruit_seeds <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "BO_fruits_seeds_2023 - BO_fruit_seed_counts.csv"), skip = 1) %>% 
  rename(SegPos = Seg)


# #Remove rows where no fruits collected
# BO_fruit_seeds_2 <- BO_fruit_seeds %>% 
#   filter(total_fruits != 0)



##Check Fitness plants
BO_merge_test <- merge(BO_Fitness_2, BO_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient"), all.y = TRUE) %>% 
  mutate(FITTRUE = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
                             Fit_plant.x != Fit_plant.y ~ FALSE))


BO_fruit_seeds_3 <- merge(BO_Fitness_2, BO_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient", "Fit_plant"), all.x = TRUE) %>% 
  select(-Sequence.y) %>% 
  rename(Sequence = Sequence.x) %>% 
  rename(Plant_ID = Fit_plant)
```

<br>  

#### Skeleton Weights  

```{r message = FALSE}

skel_weights_BO <- read_csv(here::here("data_sheets", "2023_year2_data",  "skeleton_data", "Skeleton Data 2023 - BO Skeletons.csv")) %>% 
  rename(position = side, Gen = 4, Fitness = 7, skel_weight_mg = 8, complete_incomplete = 9, selected = 10) %>% 
  arrange(Transect, Segment, position) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Transect, SegPos, Gen,  Recipient, Plant_ID, Fitness, skel_weight_mg) #complete/selected not included


#Check that IDs line up
skelcheck <- merge(BO_Fitness_2, skel_weights_BO, by = c("Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE)


#Merge with full data set
#BO_merged_df5 <- merge(BO_merged_df4, skel_weights_BO, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)


```


**TEMP MERGED DATA FRAME**  
Merge trait data processed thus far

```{r}

BO_merged_1 <- merge(BO_flowering, skel_weights_BO, by = c("Transect", "SegPos", "Gen", "Recipient", "Plant_ID"), all.x = TRUE) %>% select(-c(Skeleton:Fitness)) %>% arrange(Transect, Sequence, SegPos)

BO_merged_2 <- merge(BO_merged_1, BO_fruit_seeds_3, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient", "Donor", "Plant_ID"), all.x = TRUE) %>% select(-c(FD_01:FD_05, N_CLOSED_fruit_D1:notes)) %>% arrange(Transect, Sequence, SegPos) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, everything())

```



<br>  



#### Seed Weights  
[[OLD - UPDATE]]  

```{r}
# Read in only # seeds weighed and total weight
# Calculate mean weight per seed

BO_seedmass_1 <- read_csv(here::here("data_sheets", "seed_mass", "BO_Seed_Mass_2023-02-05.csv"), skip = 1) %>% 
  select(-Donor) %>% slice(1:339)
  

colnames(BO_seedmass_1) <- c("Transect", "Sequence", "Block", "Recipient", "Plant_ID", "closed_fruits_d1", "viable_seeds_d1", "seeds_weighed_d1", "seed_mass_mg_d1", "closed_fruits_d2", "viable_seeds_d2", "seeds_weighed_d2", "seed_mass_mg_d2", "closed_fruits_d3", "viable_seeds_d3", "seeds_weighed_d3", "seed_mass_mg_d3", "closed_fruits_d4", "viable_seeds_d4", "seeds_weighed_d4", "seed_mass_mg_d4", "closed_fruits_d5", "viable_seeds_d5", "seeds_weighed_d5", "seed_mass_mg_d5", "closed_fruits_d6", "viable_seeds_d6", "seeds_weighed_d6", "seed_mass_mg_d6", 
                             "closed_fruits_d7", "viable_seeds_d7", "seeds_weighed_d7", "seed_mass_mg_d7",
                             "closed_fruits_d8", "viable_seeds_d8", "seeds_weighed_d8", "seed_mass_mg_d8", 
                             "closed_fruits_d9", "viable_seeds_d9", "seeds_weighed_d9", "seed_mass_mg_d9")


BO_seedmass_2 <- BO_seedmass_1 %>% 
  select(-c(42:48)) %>% 
  select(-starts_with(c("closed_fruits_", "viable_seeds_"))) %>% 
  mutate(msm_d1 = seed_mass_mg_d1/seeds_weighed_d1,
         msm_d2 = seed_mass_mg_d2/seeds_weighed_d2,
         msm_d3 = seed_mass_mg_d3/seeds_weighed_d3,
         msm_d4 = seed_mass_mg_d4/seeds_weighed_d4,
         msm_d5 = seed_mass_mg_d5/seeds_weighed_d5,
         msm_d6 = seed_mass_mg_d6/seeds_weighed_d6,
         msm_d7 = seed_mass_mg_d7/seeds_weighed_d7,
         msm_d8 = seed_mass_mg_d8/seeds_weighed_d8,
         msm_d9 = seed_mass_mg_d9/seeds_weighed_d9)
         

BO_seedmass_3 <- BO_seedmass_2 %>%
  mutate(msm_all = rowMeans(select(BO_seedmass_2, starts_with("msm_")), na.rm = TRUE)) %>% 
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(msm_d1:msm_d9)), na.rm = TRUE)) %>% 
  select(-c(Block, Sequence))
         
  
##Merge with fruit/seed count data frame

BO_fruitseed_ct3 <- merge(BO_fruitseed_ct2, BO_seedmass_3, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```

<br>  



#### Corolla measurements  
[[OLD - UPDATE]]

```{r}

corollas_BO <- read_csv(here::here("data_sheets", "corolla_measures", "corollameasures_BO_2023-02-28.csv")) %>% select(1, 3, 4, 6, 8, 9, 10) %>% 
  rename(Plant_ID = 3, photo_date = 4, corolla_diam_mm = 6, corolla_area_mm2 = 5, notes_corolla = 7) %>% 
  filter(!grepl("REMOVE", notes_corolla)) ## Filter out notes with "REMOVE"

corollas_BO2 <- corollas_BO %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  arrange(Transect, Recipient, Plant_ID, photo_date) %>% 
  distinct(Transect, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates


## Merge with main data 
BO_merged_df6 <- merge(BO_merged_df5, corollas_BO2, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)

```

<br>  


#### Leaf Area & Weight  
[[OLD - UPDATE]]

```{r}
leaves_BO <- read_csv(here::here("data_sheets", "leaf_data", "BO_SLA_data_20230522.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) #exlcude NA leaf areas?

#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_BO_duped <- leaves_BO %>% 
  mutate(duped1 = duplicated(select(leaves_BO, c(Block, Transect, Recipient))),
         duped2 = duplicated(select(leaves_BO, c(Block, Transect, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!

#This seems to work the best!
leaves_BO_unique <- leaves_BO_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) %>% #Don't need plant ID...these were all non-fitness plants
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg,
         LMA = 1 / SLA)


#Merge in to main data 

BO_merged_df7 <- merge(BO_merged_df6, leaves_BO_unique, by = c("Block", "Transect", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo


```

<br>  


#### d13C data  
[[OLD - UPDATE]]

```{r}
Nemo_d13C <- read_csv(here::here("data_sheets", "isotope_data", "Nemophila_d13C_230511.csv")) %>% 
  select(1,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error

#filter to BO samples
Nemo_d13C_BO <- Nemo_d13C %>% 
  slice(-54) %>% 
  filter(grepl('BO', sample_ID))

#unravel sample ID
Nemo_d13_BO2 <- Nemo_d13C_BO %>% 
  mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 6)), 
         Transect = as.numeric(str_remove(substr(sample_ID, 8, 10), "^0+")),
         Plant_ID = as.numeric(substr(sample_ID, 12,13))) %>% 
  select(Transect, Recipient, Plant_ID, d13C, sample_weight_mg, -Plant_ID)

#Merge

BO_merged_df_fl3 <- merge(BO_merged_df_fl2, Nemo_d13_BO2, by = c("Recipient", "Transect"), all.x = TRUE)

#check that all data made it (196 obs)
testdf <- BO_merged_df_fl3 %>% 
  filter(!is.na(d13C))

#One sample missing from main data - Tr 5, BO_118 plant 8 -- no such recipient in Tr 5, leave out of data

```

<br>  


**Error Checks**  

```{r eval = FALSE}
##Check for mismatched (fitness) plants

BO_Echeck1 <- BO_merged_df6 %>% 
  filter(any_FitP == FALSE & total_fruits_F != 0) #do the same for seed count, using `>` and `!=`

BO_Echeck2 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
BO_Echeck3 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))


BO_Echeck4 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))

## Check for NA seed counts for fitness plants (when fruit production > 0)
BO_Echeck5 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))


#Seed mass
BO_Echeck5 <- BO_merged_df6 %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


### Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd`  


```{r}

#Survival measures
BO_merged_df1 <- BO_merged_2 %>% 
  mutate(surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(F_plant == TRUE & total_fruits > 0 ~ 1, #plants that survived to produce fruit
                                       F_plant == TRUE & total_fruits == 0 ~ 0),
         prod_closed_fruits = case_when(F_plant == TRUE & ALL_closed_fruit > 0 ~ 1,
                                            F_plant == TRUE & ALL_closed_fruit == 0 ~ 0),
         prod_filled_seeds = case_when(F_plant == TRUE & ALL_filled_seeds > 0 ~ 1, #plants produced filled seeds
                                      F_plant == TRUE & ALL_filled_seeds == 0 ~ 0)) #else = 0 (applies to seed_ct = NA)

         
BO_merged_df2 <- BO_merged_df1 %>% 
  rename(mean_seeds_per_fruit = avg_seeds_per_fruit) %>% 
  mutate(across(c(est_fecundity, mean_seeds_per_fruit), ~ as.numeric(na_if(., "#DIV/0!")))) %>% #Converts to NA 
  mutate(est_fecundity = case_when(F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit == 0 ~ 0,
                                   F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit >= 1 ~ NA,
                                               F_plant == TRUE & !is.na(est_fecundity) ~  est_fecundity)) 

# estimated fecundity is 0 when 0 closed or open fruits collected, or when 0 filled seeds collected
# estimated fecundity is NA when 0 closed fruits but 1 or more open fruits collected


```

<br>  



### Master sheets (2023)  

Full and summarized data sheets for all data collected in the 2022-2023 field season. 
These are merged with the year 1 data sheets elsewhere.  

```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...


######
#TEMP VERSION
######

BO_merged_full_1 <- BO_merged_df2 %>% 
  filter(F_plant == TRUE)
##NOTES - there are still fitness plants that never flowered... keep as 0s


write_csv(BO_merged_full_1, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BO_traits_Y2_incomplete_20230922.csv"), na = "")

######
######



BO_mastersheet_v1 <- BO_merged_df_fl3 %>% 
  mutate(Year = 2022, Gen = "G1") %>% 
  select(Year, Gen, Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

write_csv(BO_mastersheet_v1, here::here("data_sheets", "compiled_sheets", "BO_mastersheet_full_2023-08-07.csv"))

#Create meta data sheet



#
## Fitness Plants Only
#

BO_mastersheet_F_mains <-  BO_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only single fitness plants

BO_mastersheet_F_all <- BO_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treate these as 0 fitness or as NAs to remove???


write_csv(BO_mastersheet_F_mains, here::here("data_sheets", "compiled_sheets", "BO_mastersheet_Fitness-mains_2023-08-07.csv"))

write_csv(BO_mastersheet_F_all, here::here("data_sheets", "compiled_sheets", "BO_mastersheet_Fitness-all_2023-08-07.csv"))

```



