---
title: "compile_datasheets_HR"
author: "Devin Gamble"
date: "2022-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



### Load and Clean Data   

**Fitness Plant Data**  

Recipient IDs to double-check:
- HR_067 (HR_067B)

- Add in Donor information
```{r message = FALSE}
HR_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "HR_fitnessplants_23.csv")) %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient))  #rename - letters were added to mark position

HR_donors <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "HR_donor_info.csv")) 

#Fix Recipient and Donor names
HR_donors2 <- HR_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("HR_", Donor_num))) %>% 
  select(-Recipient_num, -Donor_num)



#Multiple matches, merge only unique pairs for Donor
HR_Fitness_2 <- merge(HR_Fitness, unique(HR_donors2[c("Gen", "Donor", "Recipient")]), by = c("Recipient", "Gen"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Sequence, SegPos)

#Save as V2 Fitness Plant Sheet
write_csv(HR_Fitness_2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "HR_fitnessplants_23_V2.csv"), na = "")


```

<br>  

### Trait Data  

Trait data gathered and processed from the 2022-2023 field season is loaded and organized below.  


#### Flower Check & Phenology  
-- Organize flowering data
-- Add in Sow & Plant dates

#########
Crucial Step!!  
######### 
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R (and lubridate) much easier!  

```{r message = FALSE}
HR_flr_df <- read_csv(file = here::here( "data_sheets", "2023_year2_data", "flowering_checks", "HR_flowercheck_data_2023-02-23 - HR_flowercheck_data_v1.csv"), na = c("", "NA")) 


HR_flr2 <- HR_flr_df %>% 
  mutate(SegPos = paste(Segment, side, sep = "")) %>% 
  select(Sequence, Transect, SegPos, everything(), -c(Segment, side, FD_06, FD_07)) %>% 
  arrange(Sequence, Transect, SegPos) %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient)) #rename- letters were added to mark position
  #Gopher notes "G" inside FD columns removed manually

#Check for mismatched fitness plants
HR_flr_test <- HR_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))

## Fix Dates!

#Check obs first using unique(HR_flr3$FD_01) and make corrections

HR_flr3 <- HR_flr2 %>%
  mutate_at(.vars = vars(FD_01:FD_05), ~as_date(.))

```

Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

HR_flr4 <- HR_flr3 %>% 
  mutate(Sow_Date = as_date("2022-11-10"),
         Plant_Date = as_date("2022-11-13"))

```


Merge in Fitness plant Data  
- Donor

```{r}
#Check to make sure fitness plants line up
HR_flr5 <- merge(HR_flr4, HR_Fitness_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Transect, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>% 
  select(-c(Fit_plant.x, Fit_plant.y)) #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  
  
#####  
#Check to make sure fitness plants line up
  # HR_flr_test <- HR_flr5 %>% 
  #   filter(!is.na(Fit_plant.x)) %>%
  #   mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
  #                           Fit_plant.x != Fit_plant.y ~ FALSE))

```


**Phenology Calculations**    

Survivorship calculated later

```{r}
#Same as last year

HR_flowering <- HR_flr5 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())


#TEMP FILE
write_csv(HR_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "HR_phenology_2023.csv"), na = "")

```

<br>  



#### Fruit and Seed counts    

Fruit & seed counts were recorded on the same spreadsheet, unlike last year. Loaded here and checked to make sure all ID info lines up with the Fitness data sheet.

```{r message = FALSE}
#Load Fruit & Seeds count data

HR_fruit_seeds <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "HR_fruits_seeds_2023 - HR_fruit_seed_counts.csv"), skip = 1) %>% 
  rename(SegPos = Seg)


# #Remove rows where no fruits collected
# HR_fruit_seeds_2 <- HR_fruit_seeds %>% 
#   filter(total_fruits != 0)


##Check Fitness plants
HR_merge_test <- merge(HR_Fitness_2, HR_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient", "Fit_plant"), all.y = TRUE) 


HR_fruit_seeds_3 <- merge(HR_Fitness_2, HR_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient", "Fit_plant"), all.x = TRUE) %>% 
  select(-Sequence.y) %>% 
  rename(Sequence = Sequence.x) %>% 
  rename(Plant_ID = Fit_plant)

```

<br>  

#### Skeleton Weights  

Load and merge-in these data sheets
```{r message = FALSE}

skel_weights_HR <- read_csv(here::here("data_sheets", "2023_year2_data",  "skeleton_data", "Skeleton Data 2023 - HR Skeletons.csv")) %>% 
  rename(position = side, Gen = 4, Fitness = 7, skel_weight_mg = 8, complete_incomplete = 9, selected = 10) %>% 
  arrange(Transect, Segment, position) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Transect, SegPos, Gen,  Recipient, Plant_ID, Fitness, skel_weight_mg) #complete/selected not included

skel_weights_HR_2 <- skel_weights_HR %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient)) #rename- letters were added to mark position

#Check that IDs line up
skelcheck <- merge(HR_Fitness_2, skel_weights_HR_2, by = c("Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>% 
  mutate(FTRUE = case_when(Fit_plant == Plant_ID ~ TRUE,
                           TRUE ~ FALSE))


#Merge with full data set
#HR_merged_df5 <- merge(HR_merged_df4, skel_weights_HR, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)


```



**TEMP MERGED DATA FRAME**  
Merge trait data processed thus far

```{r}

HR_merged_1 <- merge(HR_flowering, skel_weights_HR, by = c("Transect", "SegPos", "Gen", "Recipient", "Plant_ID"), all.x = TRUE) %>% select(-c(Skeleton:Fitness)) %>% arrange(Transect, Sequence, SegPos)

HR_merged_2 <- merge(HR_merged_1, HR_fruit_seeds_3, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient", "Donor", "Plant_ID"), all.x = TRUE) %>% select(-c(FD_01:FD_05, N_CLOSED_fruit_D1:notes)) %>% arrange(Transect, Sequence, SegPos) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, everything())

```



<br>  


#### Seed Weights  
[[OLD - UPDATE]]

```{r}
# Read in only # seeds weighed and total weight
# Calculate mean weight per seed

HR_seedmass_1 <- read_csv(here::here("data_sheets", "seed_mass", "HR_Seed_Mass_2023-02-05.csv"), skip = 1) %>% 
  select(-2) %>% slice(1:453)
  

colnames(HR_seedmass_1) <- c("Transect", "Recipient", "Plant_ID", "closed_fruits_d1", "viable_seeds_d1", "seeds_weighed_d1", "seed_mass_mg_d1", "closed_fruits_d2", "viable_seeds_d2", "seeds_weighed_d2", "seed_mass_mg_d2", "closed_fruits_d3", "viable_seeds_d3", "seeds_weighed_d3", "seed_mass_mg_d3", "closed_fruits_d4", "viable_seeds_d4", "seeds_weighed_d4", "seed_mass_mg_d4", "closed_fruits_d5", "viable_seeds_d5", "seeds_weighed_d5", "seed_mass_mg_d5", "closed_fruits_d6", "viable_seeds_d6", "seeds_weighed_d6", "seed_mass_mg_d6", 
                             "closed_fruits_d7", "viable_seeds_d7", "seeds_weighed_d7", "seed_mass_mg_d7",
                             "closed_fruits_d8", "viable_seeds_d8", "seeds_weighed_d8", "seed_mass_mg_d8", 
                             "closed_fruits_d9", "viable_seeds_d9", "seeds_weighed_d9", "seed_mass_mg_d9")


HR_seedmass_2 <- HR_seedmass_1 %>% 
  select(-starts_with(c("closed_fruits_", "viable_seeds_"))) %>% 
  mutate(msm_d1 = seed_mass_mg_d1/seeds_weighed_d1,
         msm_d2 = seed_mass_mg_d2/seeds_weighed_d2,
         msm_d3 = seed_mass_mg_d3/seeds_weighed_d3,
         msm_d4 = seed_mass_mg_d4/seeds_weighed_d4,
         msm_d5 = seed_mass_mg_d5/seeds_weighed_d5,
         msm_d6 = seed_mass_mg_d6/seeds_weighed_d6,
         msm_d7 = seed_mass_mg_d7/seeds_weighed_d7,
         msm_d8 = seed_mass_mg_d8/seeds_weighed_d8,
         msm_d9 = seed_mass_mg_d9/seeds_weighed_d9)
         

HR_seedmass_3 <- HR_seedmass_2 %>%
  mutate(msm_all = rowMeans(select(HR_seedmass_2, starts_with("msm_")), na.rm = TRUE)) %>% 
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(msm_d1:msm_d9)), na.rm = TRUE))
         
  
##Merge with fruit/seed count data frame

HR_fruitseed_ct3 <- merge(HR_fruitseed_ct2, HR_seedmass_3, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```

<br>  




#### Corolla measurements  
[[OLD - UPDATE]]

```{r}
corollas_HR <- read_csv(here::here("data_sheets", "corolla_measures", "corollameasures_HR_2023-02-28.csv")) %>% select(1, 3, 4, 6, 13, 16, 18) %>% 
  rename(photo_date = 4, corolla_diam_mm = 5, corolla_area_mm2 = 6, notes_corolla = 7) %>% 
  filter(!grepl("REMOVE", notes_corolla)) ## Filter out notes with "REMOVE"

corollas_HR2 <- corollas_HR %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  arrange(Transect, Recipient, Plant_ID, photo_date) %>% 
  distinct(Transect, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates


## Merge with main data 
HR_merged_df6 <- merge(HR_merged_df5, corollas_HR2, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)
  
```

<br>  


#### Leaf Area & Weight (SLA)  
[[OLD - UPDATE]]

```{r}
leaves_HR <- read_csv(here::here("data_sheets", "leaf_data", "HR_SLA_data_20230522.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) %>% #exlcude NA leaf areas?
  mutate(leaf_dryweight_mg = as.numeric(leaf_dryweight_mg))


#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_HR_duped <- leaves_HR %>% 
  mutate(duped1 = duplicated(select(leaves_HR, c(Block, Transect, Recipient))),
         duped2 = duplicated(select(leaves_HR, c(Block, Transect, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!

#This seems to work the best!
leaves_HR_unique <- leaves_HR_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) %>% #Don't need plant ID...these were all non-fitness plants
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg,
         LMA = 1 / SLA)

#Merge in to main data 

HR_merged_df7 <- merge(HR_merged_df6, leaves_HR_unique, by = c("Block", "Transect", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo


```

<br>  



#### d13C data (WUE)  
[[OLD - UPDATE]]

```{r}
Nemo_d13C <- read_csv(here::here("data_sheets", "isotope_data", "Nemophila_d13C_230511.csv")) %>% 
  select(1,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error

#filter to HR samples
Nemo_d13C_HR <- Nemo_d13C %>% 
  filter(grepl('HR', sample_ID))

#unravel sample ID
Nemo_d13_HR2 <- Nemo_d13C_HR %>% 
  mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 6)), 
         Transect = as.numeric(str_remove(substr(sample_ID, 8, 10), "^0+")),
         Plant_ID = as.numeric(substr(sample_ID, 12,13))) %>% 
  select(Transect, Recipient, Plant_ID, d13C, sample_weight_mg, -Plant_ID)

#Merge

HR_merged_df_fl3 <- merge(HR_merged_df_fl2, Nemo_d13_HR2, by = c("Recipient", "Transect"), all.x = TRUE)
#Merge by Recipient and Transect only: genotype-wide estimates of d13C since some plants were NF



#check that all data made it (266 obs)
testdf <- HR_merged_df_fl3 %>% 
  filter(!is.na(d13C))
```



**Error checks**  

Checking a variety of IDs and trait data to make sure everything lines up.  

```{r eval = FALSE}
##Check for mismatched (fitness) plants

HR_Echeck1 <- HR_merged_df6 %>% 
  filter(any_FitP == FALSE & seed_ct != 0) #do the same for seed count, using `>` and `!=`

HR_Echeck2 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
HR_Echeck3 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))


HR_Echeck4 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))

## Check for NA seed counts for fitness plants (when fruit production > 0)
HR_Echeck5 <- HR_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))


#Seed mass
HR_Echeck5 <- HR_merged_df6 %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


### Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd`  

Phenology calculations performed above.


```{r}

#Survival measures
HR_merged_df1 <- HR_merged_2 %>% 
  mutate(surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(F_plant == TRUE & total_fruits > 0 ~ 1, #plants that survived to produce fruit
                                       F_plant == TRUE & total_fruits == 0 ~ 0),
         surv_to_seedprod = case_when(F_plant == TRUE & ALL_filled_seeds > 0 ~ 1, #plants produced filled seeds
                                      F_plant == TRUE & ALL_filled_seeds == 0 ~ 0)) #else = 0 (applies to seed_ct = NA)

         
HR_merged_df2 <- HR_merged_df1 %>% 
  rename(total_est_seed_production = est_fecundity, mean_seeds_per_fruit = avg_seeds_per_fruit) %>% 
  mutate(across(c(total_est_seed_production, mean_seeds_per_fruit), ~ as.numeric(na_if(., "#DIV/0!")))) %>% #Converts to NA when 0 closed fruit were collected
  mutate(total_est_seed_production = case_when(F_plant == TRUE & is.na(total_est_seed_production) ~ 0,
                                               F_plant == TRUE & !is.na(total_est_seed_production) ~  total_est_seed_production)) %>% 
    mutate(mean_seeds_per_fruit = case_when(F_plant == TRUE & is.na(mean_seeds_per_fruit) ~ 0,
                                            F_plant == TRUE & !is.na(mean_seeds_per_fruit) ~  mean_seeds_per_fruit))#Set seeds per fruit & est_fecundity to 0 when NA, regardless of # closed fruits collected

```

<br>  



### Master sheets (2023)  

Full and summarized data sheets for all data collected in the 2022-2023 field season. 
These are merged with the year 1 data sheets elsewhere.  

```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...


######
#TEMP VERSION
######

HR_merged_full_1 <- HR_merged_df2 %>% 
  filter(F_plant == TRUE)
##NOTES - there are still fitness plants that never flowered... keep as 0s


write_csv(HR_merged_full_1, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "HR_traits_Y2_incomplete_20230919.csv"), na = "")

######
######



  HR_mastersheet_v1 <- HR_merged_df_fl3 %>% 
  mutate(Year = 2022, Gen = "G1") %>% 
  select(Year, Gen, Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

write_csv(HR_mastersheet_v1, here::here("data_sheets", "compiled_sheets", "HR_mastersheet_full_2023-08-07.csv"))

#Create meta data sheet



#
## Fitness Plants Only
#

HR_mastersheet_F_mains <-  HR_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only single fitness plants

HR_mastersheet_F_all <- HR_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treate these as 0 fitness or as NAs to remove???


write_csv(HR_mastersheet_F_mains, here::here("data_sheets", "compiled_sheets", "HR_mastersheet_Fitness-mains_2023-08-07.csv"))

write_csv(HR_mastersheet_F_all, here::here("data_sheets", "compiled_sheets", "HR_mastersheet_Fitness-all_2023-08-07.csv"))

```



