---
title: "Compile FLower, Fruit, and Seed count data sheets - AC Fitness plants"
output: html_notebook
author: Devin Gamble
---


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```

###ISSUE####
Genotype Segments that had 0 survivors and NO fitness plant selected are EXCLUDED as of now... These should probably be included but with an assignment of 0 fitness


### Load and Clean Data   

Recipient IDs to double-check:
- AC_159 (AC_159A/B)
- IDs not planted

**Load Fitness Plant Data**  

```{r message = FALSE}
AC_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "AC_fitness_23.csv")) %>% 
  slice(-639) %>% 
  mutate(Recipient = case_when(Recipient %in% c("AC_159", "AC_159A", "AC_159B") ~ "AC_159",
                               TRUE ~ Recipient)) %>%  #rename 159 - letters were added to mark position
  slice(-423) #AC_213 in Tr 6 seg 33L was never planted - remove entirely

AC_donors <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "AC_donor_info.csv")) %>% 
  select(-position) 

#Fix Recipient and Donor names
AC_donors2 <- AC_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("AC_", Donor_num))) %>% 
  select(-Recipient_num, -Donor_num)



#Multiple matches, merge only unique pairs for Donor
AC_Fitness_2 <- merge(AC_Fitness, unique(AC_donors2[c("Gen", "Donor", "Recipient")]), by = c("Recipient", "Gen"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Sequence, SegPos)

#Save as V2 Fitness Plant Sheet
write_csv(AC_Fitness_2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "AC_fitnessplants_23_V2.csv"), na = "")


```



### Trait Data  

Trait data gathered and processed from the 2022-2023 field season is loaded and organized below.  


#### Flower Check & Phenology  
-- Organize flowering data
-- Add in Sow & Plant dates

#########
Crucial Step!!  
######### 
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R (and lubridate) much easier!  

```{r message = FALSE}

AC_flr_df <- read_csv(file = here::here( "data_sheets", "2023_year2_data", "flowering_checks", "AC_flowercheck_data_2023-03-15 - AC_flowercheck_data_v1.csv"), na = c("", "NA")) 


AC_flr2 <- AC_flr_df %>% 
  mutate(SegPos = paste(Segment, side, sep = "")) %>% 
  select(Sequence, Transect, SegPos, everything(), -c(Segment, side, FD_06, FD_07)) %>% 
  slice(-c(1355, 1600, 2111:2115, 3191:3195)) %>% #Remove rows not planted
  arrange(Sequence, Transect, SegPos) %>% 
  mutate(Recipient = case_when(Recipient %in% c("AC_159", "AC_159A", "AC_159B") ~ "AC_159",
                               TRUE ~ Recipient)) #rename 159 - letters were added to mark position
  

#Check for mismatched fitness plants
AC_flr_test <- AC_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))

## Fix Dates!

#Check obs first using unique(AC_flr3$FD_01) and make corrections

AC_flr3 <- AC_flr2 %>%
  mutate_at(.vars = vars(FD_01:FD_05), ~as_date(.)) %>% 
  mutate(FD_05 = as_date(case_when(FD_05 == "2023-06-07" ~ "2023-06-08",
                           FD_05 == "2023-06-08" ~ "2023-06-08",
                           FD_05 == "2023-06-09" ~ "2023-06-09",
                           TRUE ~ as.character(NA)))) #Some date typos - one day early

```

Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

AC_flr4 <- AC_flr3 %>% 
  mutate(Sow_Date = as_date("2022-10-17"),
         Plant_Date = as_date("2022-10-21"))

```


Merge in Fitness plant Data  
- Donor
```{r message = FALSE}
#Check to make sure fitness plants line up
AC_flr5 <- merge(AC_flr4, AC_Fitness_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Transect, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>% 
  select(-c(Fit_plant.x, Fit_plant.y)) #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  
  
#####  
#Check to make sure fitness plants line up
  #filter(!is.na(Fit_plant.x)) %>%
  #mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
  #                          Fit_plant.x != Fit_plant.y ~ FALSE))
#
```


**Phenology Calculations**  

```{r}
#Same as last year

AC_flowering <- AC_flr5 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())



#TEMP FILE - delete later
write_csv(AC_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "AC_phenology_2023.csv"), na = "")

```

<br>  


#### Fruit and Seed counts  

Fruit & seed counts were recorded on the same spreadsheet, unlike last year. Loaded here and checked to make sure all ID info lines up with the Fitness data sheet.


```{r message = FALSE}
#Load Fruit & Seeds count data

AC_fruit_seeds <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "AC_fruits_seeds_2023 - AC_fruit_seed_counts.csv"), skip = 1) %>% 
  rename(SegPos = Seg)


# #Remove rows where no fruits collected (optional)
# AC_fruit_seeds_2 <- AC_fruit_seeds %>% 
#   filter(total_fruits != 0)


##Check Fitness plants
AC_merge_test <- merge(AC_Fitness_2, AC_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient"), all.y = TRUE) %>% 
  mutate(FITTRUE = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
                             Fit_plant.x != Fit_plant.y ~ FALSE))


AC_fruit_seeds_3 <- merge(AC_Fitness_2, AC_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient", "Fit_plant"), all.x = TRUE) %>% 
  select(-Sequence.y) %>% 
  rename(Sequence = Sequence.x,
         Plant_ID = Fit_plant) #Rename Plant ID for merging below AFTER double-checking fitness plants
```

Note: Some Segments not represented by any Fitness plants (no plants survived) -- Fecundity values are currently NA but will be adjusted to 0 later. 

<br>  



#### Skeleton Weights  

```{r message = FALSE}

skel_weights_AC <- read_csv(here::here("data_sheets", "2023_year2_data",  "skeleton_data", "Skeleton Data 2023 - AC Skeletons.csv")) %>% 
  rename(position = side, Gen = 4, Fitness = 7, skel_weight_mg = 8, complete_incomplete = 9, selected = 10) %>% 
  arrange(Transect, Segment, position) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Transect, SegPos, Gen,  Recipient, Plant_ID, Fitness, skel_weight_mg) #complete/selected not included

skel_weights_AC_2 <- skel_weights_AC %>% 
    mutate(Recipient = case_when(Recipient %in% c("AC_159", "AC_159A", "AC_159B") ~ "AC_159",
                               TRUE ~ Recipient)) #rename 159 - letters were added to mark position

#Check that IDs line up
skelcheck <- merge(AC_Fitness_2, skel_weights_AC, by = c("Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE)


#Merge with full data set
#AC_merged_df5 <- merge(AC_merged_df4, skel_weights_AC, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)


```

<br>  

**TEMP MERGED DATA FRAME**  

Merge trait data processed thus far

**NOTE** SHOULD THOSE COLUMNS BE PRESENT/ABSENT IN THE FINAL DATA SHEET??

```{r}

AC_merged_1 <- merge(AC_flowering, skel_weights_AC_2, by = c("Transect", "SegPos", "Gen", "Recipient", "Plant_ID"), all.x = TRUE) %>% select(-c(Skeleton:Fitness)) %>% arrange(Transect, Sequence, SegPos)

AC_merged_2 <- merge(AC_merged_1, AC_fruit_seeds_3, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient", "Donor", "Plant_ID"), all.x = TRUE) %>% select(-c(FD_01:FD_05, N_CLOSED_fruit_D1:notes)) %>% arrange(Transect, Sequence, SegPos) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, everything())

```



#### Seed Weights  
[[OLD - UPDATE]]  

```{r eval = FALSE}
# Read in only # seeds weighed and total weight
# Calculate mean weight per seed

AC_seedmass_1 <- read_csv(here::here("data_sheets", "seed_mass", "AC_Seed_Mass_2023-02-05.csv"), skip = 2)

colnames(AC_seedmass_1) <- c("Sequence", "Transect", "Recipient_old", "Recipient", "Plant_ID", "closed_fruits_d1", "viable_seeds_d1", "seeds_weighed_d1", "seed_mass_mg_d1", "closed_fruits_d2", "viable_seeds_d2", "seeds_weighed_d2", "seed_mass_mg_d2", "closed_fruits_d3", "viable_seeds_d3", "seeds_weighed_d3", "seed_mass_mg_d3", "closed_fruits_d4", "viable_seeds_d4", "seeds_weighed_d4", "seed_mass_mg_d4", "closed_fruits_d5", "viable_seeds_d5", "seeds_weighed_d5", "seed_mass_mg_d5")


#Unsure how to account for number of seeds weighed - each statistical 'unit' is the mean seed weight calculated on a given date

AC_seedmass_2 <- AC_seedmass_1 %>% 
  select(-starts_with(c("closed_fruits_", "viable_seeds_")), -Recipient_old) %>% 
  mutate(msm_d1 = seed_mass_mg_d1/seeds_weighed_d1,
         msm_d2 = seed_mass_mg_d2/seeds_weighed_d2,
         msm_d3 = seed_mass_mg_d3/seeds_weighed_d3,
         msm_d4 = seed_mass_mg_d4/seeds_weighed_d4,
         msm_d5 = seed_mass_mg_d5/seeds_weighed_d5)
         

AC_seedmass_3 <- AC_seedmass_2 %>%
  mutate(msm_all = rowMeans(select(AC_seedmass_2, starts_with("msm_")), na.rm = TRUE)) %>% 
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(msm_d1:msm_d5)), na.rm = TRUE))
         
  
##Merge with fruit/seed count data frame

AC_fruitseed_ct3 <- merge(AC_fruitseed_ct2, AC_seedmass_3, by = c("Transect", "Sequence", "Recipient", "Plant_ID"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```





#### Corolla measurements  
[[UPDATE]]

- Format dates in excel prior to importing
```{r}
corollas_AC <- read_csv(here::here("data_sheets", "2023_year2_data", "corolla_leaf_data", "Corolla Diameters 2023 - AC_corollas.csv")) %>% select(-c(4,10, 11)) %>% 
  rename(corolla_notes = Notes, photo_date = `Photo date`) 


corollas_AC2 <- corollas_AC %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  arrange(Transect, Recipient, Plant_ID, photo_date) %>% 
  distinct(Transect, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates
#13 duplicate observations removed


## Merge with main data 





##Create segment-wide corolla diameter column


#Merge to identify MAIN fitness plants
# corollas_seg1 <- AC_merged_df5 %>% 
#   filter(!is.na(corolla_diam_mm)) %>% #diameter prioritized - main trait to measure (& fewer NAs)
#   select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, corolla_diam_mm, corolla_area_mm2) 
# 
# 
# #Highlight duplicates
# corollas_duped <- corollas_seg1 %>% 
#   mutate(duped1 = duplicated(subset(corollas_seg1, select = c("Block", "Transect", "Sequence", "Donor", "Recipient"))),
#          duped2 = duplicated(subset(corollas_seg1, select = c("Block", "Transect", "Sequence", "Donor", "Recipient")), fromLast = TRUE),
#          duped = case_when(duped1|duped2 == TRUE ~ TRUE,
#                            TRUE ~ FALSE))
# 
# 
# #Filter out Replicated Plants when a Main plant is available (non-watered transects only)
# #Note: Duplicates (2+) of replicated plants are removed; merged back in after random selection below
# corollas_unduped1 <- corollas_duped %>% 
#   filter(case_when(duped == TRUE & Replicated == "R" & Block != "W" ~ FitP == TRUE,
#                    TRUE ~ any_FitP == TRUE)) %>% #keep all main FitPs when duped+replicated, else keep all
#   filter(Block != "W") #241 obs
# 
# 
# #Randomly selected F plants - one measure per available segment
# corollas_seg_RAND <- corollas_duped %>% 
#   group_by(Block, Transect, Recipient) %>% 
#   slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
#   select(-c(duped, Plant_ID)) #%>% 
#   #filter(Block != "W")
# ##309 total obs; 267 obs excluding W block - number of segment measurements there should be
# 
# 
# #Anti-join to get missing obs
# corollas_unduped_comp <- anti_join(corollas_seg_RAND, corollas_unduped1, by = c("Block", "Transect", "Sequence", "Donor", "Recipient")) %>% 
#   select(-c(Replicated:any_FitP, duped1, duped2))
# 
# 
# #Main plant measures - remove unnecessary columns
# corollas_unduped1b <- corollas_unduped1 %>% 
#   select(-c(Replicated:any_FitP, duped, duped1, duped2)) #Make sure columns match up
# 
# #Combine with missing complement - measurements from replicated/NF plants where no main plant available
# corollas_SEG <- rbind(corollas_unduped1b, corollas_unduped_comp) %>% 
#   rename(corolla_diam_mm_SEG = corolla_diam_mm,
#          corolla_area_mm2_SEG = corolla_area_mm2) #rename vars
# #double-check that obs total as expected
  
```

<br>  


#### Leaf Area & Weight  
[[OLD - UPDATE]]

```{r eval = FALSE}
leaves_AC <- read_csv(here::here("data_sheets", "leaf_data", "AC_SLA_data_20230522.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) %>% #exlcude NA leaf areas?
  mutate(leaf_area_mm2 = as.numeric(leaf_area_mm2))



#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_AC_duped <- leaves_AC %>% 
  mutate(duped1 = duplicated(select(leaves_AC, c(Block, Transect, Recipient))),
         duped2 = duplicated(select(leaves_AC, c(Block, Transect, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!


#This seems to work the best!
leaves_AC_unique <- leaves_AC_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) %>%  #Don't need plant ID...these were all non-fitness plants
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg,
         LMA = 1 / SLA)


#Merge in to main data 

AC_merged_df6 <- merge(AC_merged_df5, leaves_AC_unique, by = c("Block", "Transect", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo



```

<br>  


#### d13C data  
[[OLD-UPDATE]]

```{r eval = FALSE}
Nemo_d13C <- read_csv(here::here("data_sheets", "isotope_data", "Nemophila_d13C_230511.csv")) %>% 
  select(1,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error

#filter to AC samples
Nemo_d13C_AC <- Nemo_d13C %>% 
  filter(grepl('AC', sample_ID))

#unravel sample ID
Nemo_d13_AC2 <- Nemo_d13C_AC %>% 
  mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 6)), 
         Transect = as.numeric(str_remove(substr(sample_ID, 8, 10), "^0+")),
         Plant_ID = as.numeric(substr(sample_ID, 12,13))) %>% 
  select(Transect, Recipient, Plant_ID, d13C, sample_weight_mg, -Plant_ID)

#Merge

AC_merged_df_fl3 <- merge(AC_merged_df_fl2, Nemo_d13_AC2, by = c("Recipient", "Transect"), all.x = TRUE)
#Merge by Recipient and Transect only: genotype-wide estimates of d13C since some plants were NF



#check that all data made it (266 obs)
#testdf <- AC_merged_df_fl3 %>% 
#  filter(!is.na(d13C))
```

<br>  


**Error checks**

```{r eval = FALSE}
##Check for mismatched (fitness) plants

AC_Echeck1 <- AC_merged_df5 %>% 
  filter(any_FitP == FALSE & total_fruits_F != 0) #do the same for seed count, using `>` and `!=`

AC_Echeck2 <- AC_merged_df5 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
AC_Echeck3 <- AC_merged_df5 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))

AC_Echeck4 <- AC_merged_df5 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))


## Check for NA seed counts for fitness plants (when fruit production > 0)
#Will be empty when seed_ct NAs are replaced with 0
AC_Echeck5 <- AC_merged_df5 %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))
##NAs present here before setting seed_ct NAs to 0 were from fitness plants that had 0 closed fruit


#Seed mass
AC_Echeck5 <- AC_merged_df5 %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


### Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd.  

Fruits were only collected from fitness plants, so only survival to flowering is calcualted for all plants (fitness and non-fitness).

UPDATE with latest merged df...

```{r}

#Survival measures
#Added var for closed fruit produced
#renamed "survival" to seed production
#
AC_merged_df1 <- AC_merged_2 %>% 
  mutate(surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(F_plant == TRUE & total_fruits > 0 ~ 1, #plants that survived to produce fruit
                                       F_plant == TRUE & total_fruits == 0 ~ 0),
         prod_closed_fruits = case_when(F_plant == TRUE & ALL_closed_fruit > 0 ~ 1,
                                            F_plant == TRUE & ALL_closed_fruit == 0 ~ 0),
         prod_filled_seeds = case_when(F_plant == TRUE & ALL_filled_seeds > 0 ~ 1, #plants produced filled seeds
                                      F_plant == TRUE & ALL_filled_seeds == 0 ~ 0)) #else = 0

#NOTE
#Adjusting fecundity/fitness:
#The issue of how to handle 0s vs missing data in order to accurately calculate fitness has led us to apply the following corrections:
# Estimated fecundity (seeds per fruit x total fruits) should be set to *0* when closed fruits >= 1 & filled seeds == 0; when closed fruits == open fruits == 0
#Set fecundity to NA when closed = 0 & open >= 1

#Mean seeds per fruit left as-is (NA when no closed fruits collected)
         
AC_merged_df2 <- AC_merged_df1 %>% 
  rename(mean_seeds_per_fruit = avg_seeds_per_fruit) %>% 
  mutate(across(c(est_fecundity, mean_seeds_per_fruit), ~ as.numeric(na_if(., "#DIV/0!")))) %>% #Converts to NA
  mutate(est_fecundity = case_when(F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit == 0 ~ 0,
                                   F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit >= 1 ~ NA,
                                               F_plant == TRUE & !is.na(est_fecundity) ~  est_fecundity))

```

<br>  



### Master sheets (2023)  

Full and summarized data sheets for all data collected in the 2022-2023 field season. 
These are merged with the year 1 data sheets elsewhere.  

```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...

######
#TEMP VERSION
######

AC_merged_full_1 <- AC_merged_df2 %>% 
  filter(F_plant == TRUE)
##NOTES - there are still fitness plants that never flowered... keep as 0s


write_csv(AC_merged_full_1, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "AC_traits_Y2_incomplete_20230922.csv"), na = "")


######
######


AC_mastersheet_v1 <- AC_merged_df_fl3 %>% 
  mutate(Year = 2022, Gen = "G1") %>% 
  select(Year, Gen, Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

write_csv(AC_mastersheet_v1, here::here("data_sheets", "compiled_sheets", "AC_mastersheet_full_2023-08-07.csv"))
#Renamed to full - has records of all plants (including non-fitness)



#
## Fitness Plants Only
#

AC_mastersheet_F_mains <-  AC_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only main fitness plants

AC_mastersheet_F_all <- AC_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treate these as 0 fitness or as NAs to remove???


write_csv(AC_mastersheet_F_mains, here::here("data_sheets", "compiled_sheets", "AC_mastersheet_Fitness-mains_2023-08-07.csv"))

write_csv(AC_mastersheet_F_all, here::here("data_sheets", "compiled_sheets", "AC_mastersheet_Fitness-all_2023-08-07.csv"))
```



