---
title: "compile_datasheets_BO"
author: "Devin Gamble"
date: "2022-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



## Load and Clean Data   


**Load Fitness Plant Data**  

- Add in Donor information
```{r}
BO_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "BO_fitnessplants_23.csv")) %>% 
  slice(-620)

#Donor info loaded from 2021-22 data
BO_donors <- read_csv(here::here("data_sheets", "tray_positions_2021_year1","BO_tray_rand.csv"), skip = 1) %>% 
  select(Donor_ID:Recipient_ID) %>% 
  rename(Donor = Donor_ID, Recipient = Recipient_ID)

#Fix Recipient and Donor names
BO_donors2 <- BO_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("BO_", Donor_num))) %>% 
  select(-Recipient_num, -Donor_num)



#Multiple matches, merge only unique pairs for Donor
BO_Fitness_2 <- merge(BO_Fitness, unique(BO_donors2[c("Donor", "Recipient")]), by = c("Recipient"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Sequence, SegPos)

#Save as V2 Fitness Plant Sheet
write_csv(BO_Fitness_2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "BO_fitnessplants_23_V2.csv"), na = "")


```






**Flower Check sheets**  
-- Organize flowering data
-- Add in Sow & Plant dates


##################
**Crucial Step!!**  
##################  
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R much easier!  



```{r}
##BO##

BO_flr_df <- read_csv(file = here::here( "data_sheets", "2023_year2_data", "flowering_checks", "BO_flowercheck_data_2023-02-24 - BO_flowercheck_data_v1.csv"), na = c("", "NA")) %>% 
  rename(F_Num_02 = 19)


BO_flr2 <- BO_flr_df %>% 
  mutate(SegPos = paste(Segment, side, sep = "")) %>% 
  select(Sequence, Transect, SegPos, everything(), -c(Segment, side, FD_06, FD_07)) %>% 
  slice(-c(698:700, 779:780, 3096:3102)) %>% #Remove individuals not planted and NA rows
  arrange(Sequence, Transect, SegPos)
  
  

#Check for mismatched fitness plants
BO_flr_test <- BO_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))


## Fix Dates!

#Check obs first using unique(BO_flr3$FD_01) and make corrections

BO_flr3 <- BO_flr2 %>%
  mutate_at(.vars = vars(FD_01:FLR_P), ~ as_date(.))

```


Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

BO_flr4 <- BO_flr3 %>% 
  mutate(Sow_Date = as_date("2022-10-31"),
         Plant_Date = as_date("2022-11-03"))

```


Merge in Fitness plant Data  
- Donor

```{r}
#Check to make sure fitness plants line up
BO_flr5 <- merge(BO_flr4, BO_Fitness_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Transect, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>% 
  select(-c(Fit_plant.x, Fit_plant.y)) #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  
  
#####  
#Check to make sure fitness plants line up
# BO_flr_test <- BO_flr5 %>% 
#   filter(!is.na(Fit_plant.x)) %>%
#   mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
#                            Fit_plant.x != Fit_plant.y ~ FALSE))
# #
```


#### Phenology Calculations  
<<Move these lower later??>>  

```{r}
#Same as last year

BO_flowering <- BO_flr5 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())



#TEMP FILE - delete later
write_csv(BO_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BO_phenology_2023.csv"), na = "")

```






**Fruit and Seed counts**  

Due to the structure of excel data sheets, it seems easiest to upload fruit and seed count data separately since different spreadsheets were used to record each. Info on plants' Block, Transect, sequence, etc. will be pulled from the fruit count sheets (assumed to be up-to-date). For sheets intended to record fruit production only, counts of viable seeds, mean seed weight, etc. that are copied over are ignored and instead pulled straight from sheets where they were first entered later on. 

Data on fitness plants will be more accurate in the fruit-cound data sheet, while the seed-count sheet will have a handful of rows with non-fitness plants wihtout seed counts.

Date added: **Feb 06, 2023**  

*NOTE*: Update with later versions of data sheet when more seed counts have been entered
```{r message = FALSE}
#
# Fruit count data
# Data on *when* fruits were collected will be omitted here; only the total fruit counts (closed, open, non-fitness) are retained from the csv

BO_fruitprod_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "BO-fruit-production_2023-02-15.csv"), skip = 1) %>% 
  select(1:10) #%>% #omit sow and flower date, seed count, and date-specific fruit counts by column index
  #filter_all(any_vars(!is.na(.))) #omit empty rows at end of df

BO_fruitprod_2 <- BO_fruitprod_1 %>% 
  select(-c(Recipient, Sequence)) %>% # Sequence in this data set is incorrect
  rename(Recipient = `Recipient for sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, closed_fruits_F = `total_closed_fruits`, open_fruits_F = `total_open_fruits`, damaged_fruits_F = `eaten_damaged_fruits`, total_fruits_F = `total_fruits (open and closed)`, nonfitness_fruits = `nonfitness_fruits`) 

#Note: total fruits should not include eaten/damaged fruit



#
# Seed count data (viable seeds)
# Similarly, only seed count data are retained from this spreadsheet to minimize errors

BO_seedct_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "BO-seed-count_2023-02-15.csv"), skip = 1) %>% 
  select(-c(1, 3, 7:33)) %>% 
  rename(Recipient = `Recipient for sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, seed_ct = `viable_seeds`) %>% 
  mutate(Plant_ID = as.numeric(Plant_ID)) 
  

#merge data
BO_fruitseed_ct1 <- left_join(BO_fruitprod_2, BO_seedct_1, by = c("Transect", "Recipient", "Plant_ID"))
#Sequence data in fruit count sheets is unreliable for these data sheets
#Rows without matches in fruit_prod data sheet are not included


# summary variables, tidying up
BO_fruitseed_ct2 <- BO_fruitseed_ct1 %>% 
  select(Transect, Recipient, Plant_ID, closed_fruits_F, open_fruits_F, damaged_fruits_F, total_fruits_F, nonfitness_fruits, everything()) %>% 
  arrange(Transect, Recipient)

## NOTE: total number of non-fitness fruits per segment may be under-represented here since not all counts were matched to fitness plants and therefore merged in.

```



**Seed Weights**  



```{r}
# Read in only # seeds weighed and total weight
# Calculate mean weight per seed

BO_seedmass_1 <- read_csv(here::here("data_sheets", "seed_mass", "BO_Seed_Mass_2023-02-05.csv"), skip = 1) %>% 
  select(-Donor) %>% slice(1:339)
  

colnames(BO_seedmass_1) <- c("Transect", "Sequence", "Block", "Recipient", "Plant_ID", "closed_fruits_d1", "viable_seeds_d1", "seeds_weighed_d1", "seed_mass_mg_d1", "closed_fruits_d2", "viable_seeds_d2", "seeds_weighed_d2", "seed_mass_mg_d2", "closed_fruits_d3", "viable_seeds_d3", "seeds_weighed_d3", "seed_mass_mg_d3", "closed_fruits_d4", "viable_seeds_d4", "seeds_weighed_d4", "seed_mass_mg_d4", "closed_fruits_d5", "viable_seeds_d5", "seeds_weighed_d5", "seed_mass_mg_d5", "closed_fruits_d6", "viable_seeds_d6", "seeds_weighed_d6", "seed_mass_mg_d6", 
                             "closed_fruits_d7", "viable_seeds_d7", "seeds_weighed_d7", "seed_mass_mg_d7",
                             "closed_fruits_d8", "viable_seeds_d8", "seeds_weighed_d8", "seed_mass_mg_d8", 
                             "closed_fruits_d9", "viable_seeds_d9", "seeds_weighed_d9", "seed_mass_mg_d9")


BO_seedmass_2 <- BO_seedmass_1 %>% 
  select(-c(42:48)) %>% 
  select(-starts_with(c("closed_fruits_", "viable_seeds_"))) %>% 
  mutate(msm_d1 = seed_mass_mg_d1/seeds_weighed_d1,
         msm_d2 = seed_mass_mg_d2/seeds_weighed_d2,
         msm_d3 = seed_mass_mg_d3/seeds_weighed_d3,
         msm_d4 = seed_mass_mg_d4/seeds_weighed_d4,
         msm_d5 = seed_mass_mg_d5/seeds_weighed_d5,
         msm_d6 = seed_mass_mg_d6/seeds_weighed_d6,
         msm_d7 = seed_mass_mg_d7/seeds_weighed_d7,
         msm_d8 = seed_mass_mg_d8/seeds_weighed_d8,
         msm_d9 = seed_mass_mg_d9/seeds_weighed_d9)
         

BO_seedmass_3 <- BO_seedmass_2 %>%
  mutate(msm_all = rowMeans(select(BO_seedmass_2, starts_with("msm_")), na.rm = TRUE)) %>% 
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(msm_d1:msm_d9)), na.rm = TRUE)) %>% 
  select(-c(Block, Sequence))
         
  
##Merge with fruit/seed count data frame

BO_fruitseed_ct3 <- merge(BO_fruitseed_ct2, BO_seedmass_3, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```



Merge everything together:  
```{r}
BO_merged_df1 <- merge(BO_GermFlow_3, BO_fruitseed_ct3, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, everything()) %>%
  arrange(Block, Transect, Sequence, Recipient, Plant_ID) %>% 
  mutate(closed_fruits_F = as.numeric(closed_fruits_F),
         total_fruits_F = as.numeric(total_fruits_F)) #correct counts to numeric values


#Correct FruitNAs to 0s in cases where no fruit data was entered (fitness and non-fitness plants):

#
BO_merged_df2 <- BO_merged_df1 %>% 
  mutate(closed_fruits_F = replace_na(closed_fruits_F, 0),
         total_fruits_F = replace_na(total_fruits_F, 0),
         seed_ct = replace_na(seed_ct, 0))

#Calculate mean seeds per fruit, total_est_seed prod
BO_merged_df3 <- BO_merged_df2 %>% 
  mutate(mean_seeds_per_fruit = seed_ct/closed_fruits_F,
         total_est_seed_production = mean_seeds_per_fruit*total_fruits_F) %>% 
  mutate(mean_seeds_per_fruit = case_when(any_FitP == TRUE & closed_fruits_F == 0 ~ 0, 
                                          TRUE ~ mean_seeds_per_fruit)) %>%  #else, original value
  mutate(total_est_seed_production = case_when(any_FitP == TRUE & mean_seeds_per_fruit == 0 ~ 0,
                                               TRUE ~ total_est_seed_production))


#mean seeds per fruit and total est fruit production have NAs/NaNs for when there were no closed fruits/seeds (fitness plants only)



#Correct formatting issues

BO_merged_df4 <- BO_merged_df3 %>% 
  mutate(Fl_4.13 = as_date(Fl_4.13),
         Fl_4.21 = as_date(Fl_4.21),
         Fl_5.25 = as_date(Fl_5.25))


```


**Skeleton Weights & Corolla area and diameter**  

Load and merge-in these data sheets
Skeletons:
```{r}
##
skel_weights_BO <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeleton_weights_BO_2023_03_01.csv")) %>% select(-Units, -7) %>% 
  filter(!grepl("EXCLUDE.", Notes)) %>% ## Filter out notes with "Exclude"
  rename(AG_biomass_mg = `Skeleton Weights`)

#Watered points
skel_weights_BO_W <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeletons_watered_BO_2023_02_08.csv")) %>% 
  select(-`Units`) %>% 
  rename(Plant_ID = `Plant ID`, AG_biomass_mg = `Skeleton Weights`)
#3 observations :(

#Combine
skel_weights_all <- rbind(skel_weights_BO, skel_weights_BO_W) %>% 
  select(-Notes) %>% 
  filter(!is.na(AG_biomass_mg))


#Merge with full data set
BO_merged_df5 <- merge(BO_merged_df4, skel_weights_all, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)

```


Corolla Area & Diameter:
```{r}

corollas_BO <- read_csv(here::here("data_sheets", "corolla_measures", "corollameasures_BO_2023-02-28.csv")) %>% select(1, 3, 4, 6, 8, 9, 10) %>% 
  rename(Plant_ID = 3, photo_date = 4, corolla_diam_mm = 6, corolla_area_mm2 = 5, notes_corolla = 7) %>% 
  filter(!grepl("REMOVE", notes_corolla)) ## Filter out notes with "REMOVE"

corollas_BO2 <- corollas_BO %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  arrange(Transect, Recipient, Plant_ID, photo_date) %>% 
  distinct(Transect, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates


## Merge with main data 
BO_merged_df6 <- merge(BO_merged_df5, corollas_BO2, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)

```
**Leaf Area & Weight**  
```{r}
leaves_BO <- read_csv(here::here("data_sheets", "leaf_data", "BO_SLA_data_20230522.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) #exlcude NA leaf areas?

#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_BO_duped <- leaves_BO %>% 
  mutate(duped1 = duplicated(select(leaves_BO, c(Block, Transect, Recipient))),
         duped2 = duplicated(select(leaves_BO, c(Block, Transect, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!

#This seems to work the best!
leaves_BO_unique <- leaves_BO_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) %>% #Don't need plant ID...these were all non-fitness plants
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg,
         LMA = 1 / SLA)


#Merge in to main data 

BO_merged_df7 <- merge(BO_merged_df6, leaves_BO_unique, by = c("Block", "Transect", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo


```

Error checks:

```{r eval = FALSE}
##Check for mismatched (fitness) plants

BO_Echeck1 <- BO_merged_df6 %>% 
  filter(any_FitP == FALSE & total_fruits_F != 0) #do the same for seed count, using `>` and `!=`

BO_Echeck2 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
BO_Echeck3 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))


BO_Echeck4 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))

## Check for NA seed counts for fitness plants (when fruit production > 0)
BO_Echeck5 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))


#Seed mass
BO_Echeck5 <- BO_merged_df6 %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


## Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd`  


#### Phenology  

```{r}

BO_merged_df_fl <- BO_merged_df7 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('Fl_')), na.rm = TRUE)), #First flowering date
         days_germ2flower = as.numeric(FFD - Germ_Date), #days from germination to flowering
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('Fl_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) #Flowering duration


#Survival measures
BO_merged_df_fl2 <- BO_merged_df_fl %>% 
  mutate(surv_to_germ = case_when(!is.na(Germ_Date) ~ 1, #plants that germinated
                                  is.na(Germ_Date) ~ 0),
         surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(total_fruits_F > 0 ~ 1, #plants that survived to produce fruit
                                       total_fruits_F == 0 ~ 0),
         surv_to_seedprod = case_when(seed_ct > 0 ~ 1, #plants that survived to produce fruit with viable seeds
                                       TRUE ~ 0)) #else = 0 (applies to seed_ct = NA)

BO_merged_df_fl2 <- BO_merged_df_fl2 %>% 
  filter(!duplicated(BO_merged_df_fl2)) #DUPLICATES REMOVED - Discovered 2023-08-07 when adding d13C data  


## Note: should flowering duration remain unchanged when a plant went several weeks without flowering (e.g., had two temporally distinct blooms during the growing season)?
```

<br>  

### d13C data  

```{r}
Nemo_d13C <- read_csv(here::here("data_sheets", "isotope_data", "Nemophila_d13C_230511.csv")) %>% 
  select(1,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error

#filter to BO samples
Nemo_d13C_BO <- Nemo_d13C %>% 
  slice(-54) %>% 
  filter(grepl('BO', sample_ID))

#unravel sample ID
Nemo_d13_BO2 <- Nemo_d13C_BO %>% 
  mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 6)), 
         Transect = as.numeric(str_remove(substr(sample_ID, 8, 10), "^0+")),
         Plant_ID = as.numeric(substr(sample_ID, 12,13))) %>% 
  select(Transect, Recipient, Plant_ID, d13C, sample_weight_mg, -Plant_ID)

#Merge

BO_merged_df_fl3 <- merge(BO_merged_df_fl2, Nemo_d13_BO2, by = c("Recipient", "Transect"), all.x = TRUE)

#check that all data made it (196 obs)
testdf <- BO_merged_df_fl3 %>% 
  filter(!is.na(d13C))

#One sample missing from main data - Tr 5, BO_118 plant 8 -- no such recipient in Tr 5, leave out of data

```


## Master sheets (v1)  


```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...


BO_mastersheet_v1 <- BO_merged_df_fl3 %>% 
  mutate(Year = 2022, Gen = "G1") %>% 
  select(Year, Gen, Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

write_csv(BO_mastersheet_v1, here::here("data_sheets", "compiled_sheets", "BO_mastersheet_full_2023-08-07.csv"))

#Create meta data sheet



#
## Fitness Plants Only
#

BO_mastersheet_F_mains <-  BO_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only single fitness plants

BO_mastersheet_F_all <- BO_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treate these as 0 fitness or as NAs to remove???


write_csv(BO_mastersheet_F_mains, here::here("data_sheets", "compiled_sheets", "BO_mastersheet_Fitness-mains_2023-08-07.csv"))

write_csv(BO_mastersheet_F_all, here::here("data_sheets", "compiled_sheets", "BO_mastersheet_Fitness-all_2023-08-07.csv"))

```



