---
title: "compile_datasheets_HR"
author: "Devin Gamble"
date: "2022-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



## Load and Clean Data   

NOTE: All date data should be standardized prior to importing into R. In Excel, set format to YYYY-MM-DD for consistency. 

NOTE 2: All paths are relative using the 'here' package.  


**Germination Data**  
- Add 'Date Sown' column and data in as well.
- Germination data sheet formatted prior to loading: set all date values in excel as YYYY-MM-DD
```{r message = FALSE}

#Skip first nonsense row (the second row is the header)
HR_germ_df1 <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "HR_germ_data.csv"), skip = 1) %>%
  rename(Block = 'Block (1-3)', Transect = 'Transect (1-12)', Sequence = 'Sequence (1-396)', Donor = "Pollen Donor ID", Recipient = "Maternal Sibship", Stage = 'mm/dd') %>% 
  filter(!Stage == "Flow Date") %>%   #Remove empty flowering date rows
  select(-c(`9`:`Oak leaves removed when ontop of seedlings`)) %>% 
  mutate(`1` = as_date(`1`),
         `2` = as_date(`2`),
         `4` = as_date(`4`),
         `5` = as_date(`5`),
         `6` = as_date(`6`),
         `7` = as_date(`7`),
         `8` = as_date(`8`)) #set as date to allow for pivot_longer 

#
#Reshape data sheet to 'long' format
HR_germ_df2 <- HR_germ_df1 %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'Germ_Date') 
#
##Fix typos in date values
## Excel auto-populated year as 2022 when in fact the year should be 2021!
unique(HR_germ_df2$Germ_Date)

HR_germ_df3 <- HR_germ_df2 %>% 
  mutate(
    Germ_Date = case_when(
    Germ_Date == "2022-12-06" ~ "2021-12-06",
    Germ_Date == "2022-12-17" ~ "2021-12-17",
    Germ_Date == "2022-11-19" ~ "2021-11-19"))

#check again
unique(HR_germ_df3$Germ_Date)
#Use sum(is.na(HR_germ_df2$Germ_Date)) to compare change in number of NA values

#
#Adjust Donor and Recipient names for easy sorting - three digits (leading zeros)
#Specify Germ_Date as 'date' class variable
HR_germ_df4 <- HR_germ_df3 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("HR_", Donor_num))) %>%
  mutate(Germ_Date = as_date(Germ_Date)) %>% #set variable to class = date
  select(Block, Transect, Sequence, Donor, Donor_num, Recipient, Recipient_num, Plant_ID, everything()) %>% 
  select(-c(Stage, Donor_num, Recipient_num))

#
##Add in Sow Date  
HR_germ_df5 <- HR_germ_df4 %>% 
  mutate(Sow_Date = as_date("2021-11-10")) %>% #lab sowing date
  mutate(Plant_Date = as_date("2021-11-19")) %>% #field planting date
  arrange(Block, Transect, Sequence, Donor, Recipient, Plant_ID) #Arrange by Transect, Recipient, Plant_ID
#Plugs planted in ground from November 19-20 ,2021


```


Add in watered transect germination data
```{r message = FALSE}

## Add in watered transects
HR_germ_W1 <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "watered_HR_germ.csv"), skip = 1) %>%
  rename(Sequence = 'Seq (1-42)', Donor = "Donor_ID", Recipient = "Recipient_ID", Stage = 'Date (mm/dd)') %>% 
  filter(!Stage == "Flow") %>%  #Remove empty flowering date rows
  select(-c(Co., `Notes 12/6`:`Notes 12/17`))

  #
#Reshape data sheet to 'long' format
HR_germ_W2 <- HR_germ_W1 %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'Germ_Date') %>% 
  mutate(Sequence = (Sequence + 411)) #Sequence continues following main transects

#
##Fix typos in date values
unique(HR_germ_W2$Germ_Date)

HR_germ_W3 <- HR_germ_W2 %>% 
  mutate(Germ_Date = case_when(
    Germ_Date == "2022-12-06" ~ "2021-12-06",
    Germ_Date == "2022-12-17" ~ "2021-12-17",
    Germ_Date == "2022-11-19" ~ "2021-11-19",
    ))

#
#Adjust Donor and Recipient names for easy sorting - three digits (leading zeros)
#Specify Germ_Date as 'date' class variable
HR_germ_W4 <- HR_germ_W3 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("HR_", Donor_num))) %>%
  mutate(Germ_Date = as_date(Germ_Date)) %>% #set variable to class = date
  mutate(Plant_ID = as.numeric(Plant_ID)) %>% #set Plant ID to numeric for sorting purposes
  select(-c(Stage, Donor_num, Recipient_num))

#
##Add in Sow Date & add block ('W')
HR_germ_W5 <- HR_germ_W4 %>% 
  mutate(Sow_Date = as_date("2021-11-10")) %>% #lab sowing date
  mutate(Plant_Date = as_date("2021-11-19")) %>% #field planting date
  mutate(Block = "W") %>% 
  select(Block, Transect, Donor, Recipient, Plant_ID, everything()) %>% 
  arrange(Transect, Recipient, Plant_ID) #Arrange by Transect, Recipient, Plant_ID


#
## Combine watered and main experiment germination data
HR_germ_full <- rbind(HR_germ_df5, HR_germ_W5) %>% 
    mutate(Plant_ID = as.numeric(Plant_ID))

#write_csv(HR_germ_full, here::here("data_sheets", "compiled_sheets", "HR_germination_full_2021-2022.csv"))

```



**Flower Check sheets**  
-- Organize flowering data


*Important Step!!*  
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R much easier!  

NOTE: Opening CSVs after doing this step appears to reset the standardization back to M/D/YYYY !

```{r}

HRflr_df <- read_csv(file = here::here( "data_sheets", "season_end_flowerchecks_2022",  "HR_flowercheck_2022_03_14.csv"), col_types = cols(`10-Mar...5` = col_date(), `10-Mar...23` = col_date(), `16-Mar...6` = col_date()), na = c("", "NA", "G", "D")) #specify date columns
##Ignore 'F' plants - more up-to-date fitness plant data is matched in below 
#parse warning - one typo coerced to NA -OK

#split groups
f1_HRlong <- HRflr_df[c(1:18)] # Group 1: plants 1-4
f2_HRlong <- HRflr_df[c(1, 20:36)] # Group 2: plants 5-8

#Revise names
colnames(f1_HRlong) <- c("Transect", "Recipient", "Plant_ID", "F.", "Fl_3.10", "Fl_3.16", "Fl_3.24", "Fl_3.31", "Fl_4.7", "Fl_4.14", "Fl_4.20", "Fl_4.26", "Fl_5.4", "Fl_5.11", "Fl_5.20", "Fl_5.24", "LEAF", "flr_P")
colnames(f2_HRlong) <- c("Transect", "Recipient", "Plant_ID", "F.",  "Fl_3.10", "Fl_3.16", "Fl_3.24", "Fl_3.31", "Fl_4.7", "Fl_4.14", "Fl_4.20", "Fl_4.26", "Fl_5.4", "Fl_5.11", "Fl_5.20", "Fl_5.24", "LEAF", "flr_P")

#rbind, tidy up and remove unnecessary columns
f_HRlong <- rbind(f1_HRlong, f2_HRlong) %>% arrange(Transect, Recipient, Plant_ID) %>% select(-c(F.)) 


#Format recipient IDs for easier sorting - three digits (includes leading zeros)
f_HRlong_2 <- f_HRlong %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  select(Transect, Recipient, Recipient_num, Plant_ID, everything()) %>% 
  select(-Recipient_num)


#### Fix Dates
#Individually check each column (e.g., Fl_3.3) for the correct date format (unique(df$col)). Include LEAF and flr_p. Multiple dates are okay.

f_HRlong_3 <- f_HRlong_2 %>% 
  mutate(Fl_3.16 = case_when(Fl_3.16 == 19067 ~ "2022-03-16",
                             TRUE ~ as.character(NA))) %>% 
  mutate(Fl_3.31 = case_when(Fl_3.24 == "2022-03-31" ~ "2022-03-31",
                             Fl_3.31 == "2022-03-31" ~ "2022-03-31",
                             TRUE ~ as.character(NA))) %>% 
  mutate(Fl_3.24 = case_when(Fl_3.24 == "2022-03-24" ~ "2022-03-24",
                             TRUE ~ as.character(NA))) %>% 
  mutate(Fl_4.7 = case_when(Fl_4.7 == "2022-04-07" ~ "2022-04-07",
                             Fl_4.14 == "2022-04-07" ~ "2022-04-07",
                             TRUE ~ as.character(NA))) %>% 
  mutate(Fl_4.14 = case_when(Fl_4.14 == "2022-04-14" ~ "2022-04-14",
                             TRUE ~ as.character(NA))) %>% 
  mutate(Fl_4.26 = case_when(Fl_4.20 == "2022-04-27" ~ "2022-04-27",
                             Fl_4.26 == "2022-04-27" ~ "2022-04-27",
                             Fl_4.26 == "2022-04-26" ~ "2022-04-26",
                             TRUE ~ as.character(NA))) %>% 
  mutate(Fl_4.20 = case_when(Fl_4.20 == "2022-04-20" ~ "2022-04-20",
                             TRUE ~ as.character(NA))) %>% 
  mutate(Fl_5.20 = case_when(Fl_5.20 == "2022-05-20" ~ "2022-05-20",
                             TRUE ~ as.character(NA)))

```


Combine Germination and Flowering data frames
```{r}
#Merge by Recipient, Transect, and Plant ID
HR_GermFlow_1 <- left_join(HR_germ_full, f_HRlong_3, by = c("Transect", "Recipient", "Plant_ID")) #


##Correct date columns!
```


<br>  


#### Fitness Plant data  
- Indicate which plants were fitness plants, including those replicated in the watering experiment

```{r message = FALSE}
HR_Fplants <- read_csv(here::here("data_sheets", "fitness_plants_7-21-2022", "HR_F_plants.csv"), col_types = cols(`ALL Fitness Plants` = col_character())) %>% #retain commas for column with multiple fitness plants
  rename(F_plant = `SINGLE Fitness Plant`, F_multi = `ALL Fitness Plants`, F_plant_notes = NOTES) %>% 
  select(Transect:F_plant_notes)

#Adjust Recipient names to match three-digit format
HR_Fplants2 <- HR_Fplants %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  rename(Replicated = `Replicated?`) %>% 
  mutate(Replicated = case_when(Replicated == "R" ~ "R",
                                is.na(Replicated) ~ "n")) #change NAs to 'n' (no) for non-rep'd Recipient IDs

```


Index fitness plants & Replicated (y/n) in germination and flowering data
```{r message = FALSE}
# Merge the two dataframes 
# Create new row (Plant ID) specific variables for main and replicated fitness plants
HR_GermFlow_2 <- merge(HR_GermFlow_1, HR_Fplants2) %>% 
  mutate(FitP = case_when(F_plant == Plant_ID ~ TRUE,
                           F_plant != Plant_ID ~ FALSE)) %>% 
  mutate(Rep_FitP = case_when(Replicated == "R" & Plant_ID == as.numeric(str_match(F_multi, as.character(Plant_ID))) ~ TRUE,   
                              TRUE ~ FALSE)) %>%  #if else, set value to FALSE
  mutate(any_FitP = case_when(FitP == TRUE | Rep_FitP == TRUE ~ TRUE, #plant is any form of fitness plant
                              TRUE ~ FALSE))


#Tidy up
HR_GermFlow_3 <- HR_GermFlow_2 %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, everything()) %>%
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

## Fit_P has NA values for when a fitness plant was never chosen

```



**Fruit and Seed counts**  

Due to the structure of excel data sheets, it seems easiest to upload fruit and seed count data separately since different spreadsheets were used to record each. Info on plants' Block, Transect, sequence, etc. will be pulled from the fruit count sheets. As typos were discovered in fruit and seed counts, corrections were made in batches.

Data on fitness plants will be more accurate in the fruit-count data sheet.

Date added: **Feb 06, 2023**  

*NOTE*:
```{r message = FALSE}
# Fruit count data
# Data on *when* fruits were collected will be omitted here

HR_fruitprod_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "HR-fruit-production_2023-02-05.csv"), skip = 1) %>% 
  select(1:10) #%>% #omit sow and flower date, seed count, and date-specific fruit counts by column index
  #filter_all(any_vars(!is.na(.))) #omit empty rows at end of df

HR_fruitprod_2 <- HR_fruitprod_1 %>% 
  select(-c(Recipient)) %>% # Sequence in this data set is incorrect
  rename(Recipient = `Recipient for sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, closed_fruits_F = `total_closed_fruits`, open_fruits_F = `total_open_fruits`, damaged_fruits_F = `eaten_damaged_fruits`, total_fruits_F = `total_fruits (open and closed)`, nonfitness_fruits = `nonfitness_fruits`) 

#Note: total fruits do not include eaten/damaged fruit


#
# Seed count data (viable seeds)
# Similarly, only seed count data are retained from this spreadsheet to minimize errors

HR_seedct_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "HR-seed-count_2023-02-05.csv"), skip = 1) %>% 
  select(-c(3, 7:33)) %>% 
  rename(Recipient = `Recipient for sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, seed_ct = `viable_seeds`) %>% 
  mutate(Plant_ID = as.numeric(Plant_ID)) 
  

#merge data
HR_fruitseed_ct1 <- left_join(HR_fruitprod_2, HR_seedct_1, by = c("Transect", "Recipient", "Plant_ID"))
#Sequence data may not be accurate - add in later
#Rows without matches in fruit_prod data sheet are not included


# tidying up
HR_fruitseed_ct2 <- HR_fruitseed_ct1 %>% 
  select(Transect, Recipient, Plant_ID, closed_fruits_F, open_fruits_F, damaged_fruits_F, total_fruits_F, nonfitness_fruits, everything()) %>% 
  select(-c(Sequence.x, Sequence.y)) %>% 
  arrange(Transect, Recipient)

## NOTE: total number of non-fitness fruits per segment may be under-represented here since not all counts were matched to fitness plants and therefore merged in.

```


**Seed Weights**  
```{r}
# Read in only # seeds weighed and total weight
# Calculate mean weight per seed

HR_seedmass_1 <- read_csv(here::here("data_sheets", "seed_mass", "HR_Seed_Mass_2023-02-05.csv"), skip = 1) %>%
  select(-2)
  

colnames(HR_seedmass_1) <- c("Transect", "Recipient", "Plant_ID", "closed_fruits_d1", "viable_seeds_d1", "seeds_weighed_d1", "seed_mass_mg_d1", "closed_fruits_d2", "viable_seeds_d2", "seeds_weighed_d2", "seed_mass_mg_d2", "closed_fruits_d3", "viable_seeds_d3", "seeds_weighed_d3", "seed_mass_mg_d3", "closed_fruits_d4", "viable_seeds_d4", "seeds_weighed_d4", "seed_mass_mg_d4", "closed_fruits_d5", "viable_seeds_d5", "seeds_weighed_d5", "seed_mass_mg_d5", "closed_fruits_d6", "viable_seeds_d6", "seeds_weighed_d6", "seed_mass_mg_d6", 
                             "closed_fruits_d7", "viable_seeds_d7", "seeds_weighed_d7", "seed_mass_mg_d7",
                             "closed_fruits_d8", "viable_seeds_d8", "seeds_weighed_d8", "seed_mass_mg_d8", 
                             "closed_fruits_d9", "viable_seeds_d9", "seeds_weighed_d9", "seed_mass_mg_d9")


HR_seedmass_2 <- HR_seedmass_1 %>% 
  select(-starts_with(c("closed_fruits_", "viable_seeds_"))) %>% 
  mutate(msm_d1 = seed_mass_mg_d1/seeds_weighed_d1,
         msm_d2 = seed_mass_mg_d2/seeds_weighed_d2,
         msm_d3 = seed_mass_mg_d3/seeds_weighed_d3,
         msm_d4 = seed_mass_mg_d4/seeds_weighed_d4,
         msm_d5 = seed_mass_mg_d5/seeds_weighed_d5,
         msm_d6 = seed_mass_mg_d6/seeds_weighed_d6,
         msm_d7 = seed_mass_mg_d7/seeds_weighed_d7,
         msm_d8 = seed_mass_mg_d8/seeds_weighed_d8,
         msm_d9 = seed_mass_mg_d9/seeds_weighed_d9)
         

HR_seedmass_3 <- HR_seedmass_2 %>%
  mutate(msm_all = rowMeans(select(HR_seedmass_2, starts_with("msm_")), na.rm = TRUE)) %>% 
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(msm_d1:msm_d9)), na.rm = TRUE))
         
  
##Merge with fruit/seed count data frame
HR_fruitseed_ct3 <- merge(HR_fruitseed_ct2, HR_seedmass_3, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```



Merge everything together:  
```{r}
HR_merged_df1 <- merge(HR_GermFlow_3, HR_fruitseed_ct3, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, everything()) %>%
  arrange(Block, Transect, Sequence, Recipient, Plant_ID) %>% 
  mutate(closed_fruits_F = as.numeric(closed_fruits_F),
         total_fruits_F = as.numeric(total_fruits_F)) #correct counts to numeric values


#Correct FruitNAs to 0s in cases where no fruit data was entered (fitness and non-fitness plants):
HR_merged_df2 <- HR_merged_df1 %>% 
  mutate(closed_fruits_F = replace_na(closed_fruits_F, 0),
         total_fruits_F = replace_na(total_fruits_F, 0),
         seed_ct = replace_na(seed_ct, 0))

#Calculate mean seeds per fruit, total_est_seed prod
#variables renamed later
HR_merged_df3 <- HR_merged_df2 %>% 
  mutate(mean_seeds_per_fruit = seed_ct/closed_fruits_F,
         total_est_seed_production = mean_seeds_per_fruit*total_fruits_F) #%>% 
 
#mean seeds per fruit and total est fruit production have NAs/NaNs for when there were no closed fruits/seeds (fitness plants only)


#Correct formatting issues
HR_merged_df4 <- HR_merged_df3 %>% 
  mutate(Fl_3.16 = as_date(Fl_3.16),
         Fl_3.24 = as_date(Fl_3.24),
         Fl_3.31 = as_date(Fl_3.31),
         Fl_4.7 = as_date(Fl_4.7),
         Fl_4.14 = as_date(Fl_4.14),
         Fl_4.20 = as_date(Fl_4.20),
         Fl_4.26 = as_date(Fl_4.26),
         Fl_5.20 = as_date(Fl_5.20))


```


**Skeleton Weights & Corolla area and diameter**  

Load and merge-in these data sheets
```{r}

skel_weights_HR <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeleton_weights_HR_2023_03_01.csv")) %>% select(-Units, -7, -8) %>% 
  filter(!grepl("EXCLUDE.", Notes)) %>% ## Filter out notes with "Exclude"
  rename(AG_biomass_mg = `Skeleton Weights (mg)`)

#Watered points
skel_weights_HR_W <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeletons_watered_HR_2023_02_08.csv")) %>% 
  select(-`Units`) %>% 
  filter(!grepl("EXCLUDE.", Notes)) %>% ## Filter out notes with "Exclude"
  rename(Plant_ID = `Plant ID`, AG_biomass_mg = `Skeleton Weights`)

#Combine
skel_weights_all <- rbind(skel_weights_HR, skel_weights_HR_W) %>% 
  select(-Notes) %>% 
  filter(!is.na(AG_biomass_mg)) %>% 
  mutate(AG_biomass_mg = as.numeric(AG_biomass_mg))


#Merge with full data set
HR_merged_df5 <- merge(HR_merged_df4, skel_weights_all, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)

```


Corolla measurements
- to avoid parsing issues, format date as yyyy-mm-dd in excel beforehand
```{r}
corollas_HR <- read_csv(here::here("data_sheets", "corolla_measures", "corollameasures_HR_2023-02-28.csv"), col_types = cols(`Date_Flower_Photo_Taken_yr-mo-day` = col_date())) %>% select(1, 3, 4, 6, 13, 16, 18) %>% 
  rename(photo_date = 4, corolla_diam_mm = 5, corolla_area_mm2 = 6, notes_corolla = 7) %>% 
  filter(!grepl("REMOVE", notes_corolla)) ## Filter out notes with "REMOVE"
#parse warning - notes coerced to NA in columns that are dropped -OK

corollas_HR2 <- corollas_HR %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  #mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  arrange(Transect, Recipient, Plant_ID, photo_date) %>% 
  distinct(Transect, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates


## Merge with main data 
HR_merged_df6 <- merge(HR_merged_df5, corollas_HR2, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)
  
```


**Leaf Area & Weight**  
```{r}
leaves_HR <- read_csv(here::here("data_sheets", "leaf_data", "HR_SLA_data_20230522.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) %>% #exlcude NA leaf areas?
  mutate(leaf_dryweight_mg = as.numeric(leaf_dryweight_mg))


#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_HR_duped <- leaves_HR %>% 
  mutate(duped1 = duplicated(select(leaves_HR, c(Block, Transect, Recipient))),
         duped2 = duplicated(select(leaves_HR, c(Block, Transect, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!

#This seems to work the best!
leaves_HR_unique <- leaves_HR_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) %>% 
  mutate(SLA_SEG = leaf_area_mm2 / leaf_dryweight_mg,
         LMA_SEG = 1 / SLA_SEG) %>% 
  rename(leaf_area_mm2_SEG = leaf_area_mm2,
         leaf_dryweight_mg_SEG = leaf_dryweight_mg)


#Merge in to main data 
HR_merged_df7 <- merge(HR_merged_df6, leaves_HR_unique, by = c("Block", "Transect", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo


#
##Include plant-level leaf measurements as well:
leaves_HR_ind <- leaves_HR %>% 
  select(Block, Transect, Recipient, Plant_ID, leaf_area_mm2, leaf_dryweight_mg) %>% #ignore notes
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg,
         LMA = 1 / SLA) 

#Merge - sheet 7b
HR_merged_df7b <- merge(HR_merged_df7, leaves_HR_ind, by = c("Block", "Transect", "Recipient", "Plant_ID"), all.x = TRUE) #measures of individual plants. Total # obs unchanged

# make sure all measurements merged in
HR_NFleaves <- HR_merged_df7b %>% 
  filter(!is.na(leaf_dryweight_mg)) #fewer than 501 obs as 4 entries were mislabeled - not included in merged df

```

Error checks:

```{r eval = FALSE}
##Check for mismatched (fitness) plants

HR_Echeck1 <- HR_merged_df7b %>% 
  filter(any_FitP == FALSE & seed_ct != 0) #do the same for seed count, using `>` and `!=`

HR_Echeck2 <- HR_merged_df7b %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
HR_Echeck3 <- HR_merged_df7b %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))


HR_Echeck4 <- HR_merged_df7b %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))

## Check for NA seed counts for fitness plants (when fruit production > 0)
HR_Echeck5 <- HR_merged_df7b %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))


#Seed mass
HR_Echeck5 <- HR_merged_df7b %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


## Trait calculations 

#### Phenology  

```{r}
HR_merged_df_fl <- HR_merged_df7b %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('Fl_')), na.rm = TRUE)), #First flowering date
         days_germ2flower = as.numeric(FFD - Germ_Date), #days from germination to flowering
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from sow to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('Fl_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) #Flowering duration
```


**Revise Calculations of Fitness Data to match Year 2**

```{r}
HR_merged_df_flB <- HR_merged_df_fl %>% 
  mutate(closed_fruits_F = case_when(any_FitP == TRUE & is.na(closed_fruits_F) ~ 0,
                                     any_FitP == TRUE & !is.na(closed_fruits_F) ~ closed_fruits_F,
                                     TRUE ~ closed_fruits_F),
         open_fruits_F = case_when(any_FitP == TRUE & is.na(open_fruits_F) ~ 0,
                                   any_FitP == TRUE & !is.na(open_fruits_F) ~ open_fruits_F,
                                   TRUE ~ open_fruits_F),
         total_fruits_F = case_when(any_FitP == TRUE & is.na(total_fruits_F) ~ 0,
                                   any_FitP == TRUE & !is.na(total_fruits_F) ~ total_fruits_F,
                                   TRUE ~ total_fruits_F))

#Survival measures
HR_merged_df_flC <- HR_merged_df_flB %>% 
  mutate(surv_to_germ = case_when(!is.na(Germ_Date) ~ 1, #plants that germinated
                                  is.na(Germ_Date) ~ 0),
         surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(any_FitP == TRUE & total_fruits_F > 0 ~ 1, #plants that survived to produce fruit
                                       any_FitP == TRUE & total_fruits_F == 0 ~ 0),
         prod_closed_fruits = case_when(any_FitP == TRUE & closed_fruits_F > 0 ~ 1,
                                        any_FitP == TRUE & closed_fruits_F == 0 ~ 0),
         prod_filled_seeds = case_when(any_FitP == TRUE & seed_ct > 0 ~ 1, #plants that survived to produce fruit with viable seeds
                                       any_FitP == TRUE & seed_ct == 0 ~ 0))  #else = 0 



#Calculate Fecundity 
HR_merged_df_fl2 <- HR_merged_df_flC %>% 
  rename(est_fecundity = total_est_seed_production) %>% #rename to match Y2
  mutate(across(c(est_fecundity, mean_seeds_per_fruit), ~ as.numeric(na_if(., NaN)))) %>% #Converts to NA
  mutate(mean_seeds_per_fruit = case_when(any_FitP == TRUE & closed_fruits_F == 0 ~ NA, #NA when 0 closed fruits
                                          TRUE ~ mean_seeds_per_fruit)) %>%  #else, original value
  mutate(est_fecundity = case_when(any_FitP == TRUE & closed_fruits_F == 0 & open_fruits_F == 0 ~ 0,
                                               any_FitP == TRUE & closed_fruits_F == 0 & open_fruits_F >= 1 ~ NA,
                                               any_FitP == TRUE & !is.na(est_fecundity) ~  est_fecundity))

```

<br>  


### d13C data  

```{r}
Nemo_d13C <- read_csv(here::here("data_sheets", "isotope_data", "Nemophila_d13C_230511.csv")) %>% 
  select(1,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error

#filter to HR samples
Nemo_d13C_HR <- Nemo_d13C %>% 
  filter(grepl('HR', sample_ID))

#unravel sample ID
Nemo_d13_HR2 <- Nemo_d13C_HR %>% 
  mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 6)), 
         Transect = as.numeric(str_remove(substr(sample_ID, 8, 10), "^0+")),
         Plant_ID = as.numeric(substr(sample_ID, 12,13))) %>% 
  select(Transect, Recipient, Plant_ID, d13C, sample_weight_mg)

#Merge
HR_merged_df_fl3 <- merge(HR_merged_df_fl2, Nemo_d13_HR2, by = c("Recipient", "Transect", "Plant_ID"), all.x = TRUE)
#Merge by Recipient and Transect only: genotype-wide estimates of d13C since some plants were NF


#check that all data made it (266 obs)
testdf <- HR_merged_df_fl3 %>% 
  filter(!is.na(d13C))


#
#Create segment-wide estimate of d13C:
Nemo_d13_HR3 <- Nemo_d13_HR2 %>% 
  rename(d13C_SEG = d13C) %>% #rename for seg-wide measure
  select(Transect, Recipient, d13C_SEG) #un-select Plant_ID, dryweight of sample


HR_merged_df_fl3b <- merge(HR_merged_df_fl3, Nemo_d13_HR3, by = c("Recipient", "Transect"), all.x = TRUE) 
#Merge by Recipient and Transect only: segment-wide estimates of d13C since some plants were NF


#Check merge
d13C_check_HR <- HR_merged_df_fl3b %>% 
  filter(FitP == TRUE) %>% 
  filter(!is.na(d13C))

```



##### Segment-wide estimates for skeleton weight, corolla area & diameter  

**Skeleton Weight**  
```{r}
#Merge to identify MAIN fitness plants
skel_weights_seg1 <- merge(HR_GermFlow_3, skel_weights_all, by = c("Recipient", "Transect", "Plant_ID"), all.x = TRUE) %>% 
  filter(!is.na(AG_biomass_mg)) %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, AG_biomass_mg)


#Highlight ALL DUPES (multiple measures within a segment)
skel_weights_duped <- skel_weights_seg1 %>% 
  mutate(duped1 = duplicated(subset(skel_weights_seg1, select = c("Block", "Transect", "Sequence", "Donor", "Recipient"))),
         duped2 = duplicated(subset(skel_weights_seg1, select = c("Block", "Transect", "Sequence", "Donor", "Recipient")), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE))

#Filter out Replicated Plants when a Main plant is available (non-watered transects only)
#Note: Non-main FitPs (replicated) retained when only when a single measure taken from a segment, but duplicates (2+) of replicated plants are removed
skel_weights_unduped1 <- skel_weights_duped %>% 
  filter(case_when(duped == TRUE & Replicated == "R" & Block != "W" ~ FitP == TRUE,
                   TRUE ~ any_FitP == TRUE)) %>% #keep all main FitPs when duped+replicated,else keep all
  filter(Block != "W") #407 total obs; 317 non-watered


#Randomly selected F plants - one measure per available segment
skel_weights_seg_RAND <- skel_weights_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID))  
#Confirms # obs = # segments (360)


#Anti-join to see missing obs
skel_weights_unduped_comp <- anti_join(skel_weights_seg_RAND, skel_weights_unduped1, by = c("Block", "Transect", "Sequence", "Donor", "Recipient")) %>% 
  select(-c(Replicated:any_FitP, duped1, duped2))

#Main plant measures - remove unnecessary columns
skel_weights_unduped1b <- skel_weights_unduped1 %>% 
  select(-c(Replicated:any_FitP, duped, duped1, duped2)) #Make sure columns match up

#Combine with missing complement - measurements from replicated/NF plants where no main plant available
skel_weights_SEG <- rbind(skel_weights_unduped1b, skel_weights_unduped_comp) %>% 
  rename(AG_biomass_mg_SEG = AG_biomass_mg)
#End result: One measure per segment (from which a skeleton was collected), with main fitness plants retained where available and NF/Replicated FitPs selected "at random" (using slice_sample(n=1))


#Merge with main data sheet
HR_merged_df_fl4 <- merge(HR_merged_df_fl3b, skel_weights_SEG, by = c("Block", "Transect", "Sequence", "Donor", "Recipient"), all.x = TRUE)

```



**Corolla Diameter & Area**  
Use same approach as above

```{r}
#Merge to identify MAIN fitness plants
corollas_seg1 <- HR_merged_df6 %>% 
  filter(!is.na(corolla_diam_mm)) %>% #diameter prioritized - main trait to measure (& fewer NAs)
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, corolla_diam_mm, corolla_area_mm2) 


#Highlight duplicates
corollas_duped <- corollas_seg1 %>% 
  mutate(duped1 = duplicated(subset(corollas_seg1, select = c("Block", "Transect", "Sequence", "Donor", "Recipient"))),
         duped2 = duplicated(subset(corollas_seg1, select = c("Block", "Transect", "Sequence", "Donor", "Recipient")), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE))


#Filter out Replicated Plants when a Main plant is available (non-watered transects only)
#Note: Duplicates (2+) of replicated plants are removed; merged back in after random selection below
corollas_unduped1 <- corollas_duped %>% 
  filter(case_when(duped == TRUE & Replicated == "R" & Block != "W" ~ FitP == TRUE,
                   TRUE ~ any_FitP == TRUE)) %>% #keep all main FitPs when duped+replicated, else keep all
  filter(Block != "W") #353 obs


#Randomly selected F plants - one measure per available segment
corollas_seg_RAND <- corollas_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) 
##420 total obs


#Anti-join to get missing obs
corollas_unduped_comp <- anti_join(corollas_seg_RAND, corollas_unduped1, by = c("Block", "Transect", "Sequence", "Donor", "Recipient")) %>% 
  select(-c(Replicated:any_FitP, duped1, duped2))


#Main plant measures - remove unnecessary columns
corollas_unduped1b <- corollas_unduped1 %>% 
  select(-c(Replicated:any_FitP, duped, duped1, duped2)) #Make sure columns match up

#Combine with missing complement - measurements from replicated/NF plants where no main plant available
corollas_SEG <- rbind(corollas_unduped1b, corollas_unduped_comp) %>% 
  rename(corolla_diam_mm_SEG = corolla_diam_mm,
         corolla_area_mm2_SEG = corolla_area_mm2) #rename vars
#double-check that obs total as expected


#Merge
HR_merged_df_fl5 <- merge(HR_merged_df_fl4, corollas_SEG, by = c("Block", "Transect", "Sequence", "Donor", "Recipient"), all.x = TRUE)

```



**Duplicate & Final Error Check**  

```{r}
#All rows
HR_merged_df_fl5 %>% 
  select(Recipient:Plant_ID) %>% 
  get_dupes()


#Fitness Plants
HR_merged_df_fl5 %>% 
  filter(any_FitP == TRUE) %>% 
  get_dupes()



#Main fitness plants
HR_merged_df_fl5 %>% 
  filter(FitP == TRUE) %>% 
  get_dupes()

```



## Master sheets (v1)  


```{r message = FALSE}
## Create a Full (long) version with all data - one row per plant 
HR_mastersheet_v1 <- HR_merged_df_fl5 %>% 
  mutate(Year = 2022, Gen = "G1") %>% 
  select(Year, Gen, Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

#Rename and organize variables to match Year 2 data:
HR_mastersheet_v2 <- HR_mastersheet_v1 %>% 
  rename(F_plant_ID = F_plant, #number of F plant
         F_plant = FitP, #T/F is F plant
         filled_seeds = seed_ct,
         open_fruits = open_fruits_F,
         closed_fruits = closed_fruits_F,
         total_fruits = total_fruits_F,
         skel_dryweight_mg = AG_biomass_mg, 
         skel_dryweight_mg_SEG = AG_biomass_mg_SEG) %>% 
  mutate(est_fecundity_int = round(est_fecundity)) #round to nearest whole number


#Add in Greenhouse Cohort IDs
HR_tray21 <- read_csv(here::here("data_sheets", "tray_positions_2021_year1", "HR_tray_rand.csv"), skip = 1) %>% 
  select(Cohort, Donor_ID, Recipient_ID) %>% #Block and Transect not necessary to save here
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient_ID)))) %>% #Fix Recipient/Donor names
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor_ID)))) %>% 
  mutate(Donor = as.character(str_c("HR_", Donor_num))) %>%
  select(-c(Donor_num, Recipient_num, Donor_ID, Recipient_ID)) %>% 
  distinct()


#merge
HR_mastersheet_v3 <- left_join(HR_mastersheet_v2, HR_tray21, by = c("Donor", "Recipient")) %>% 
  select(-c(F_plant_notes, Fl_3.10:flr_P, damaged_fruits_F, seeds_weighed_d1:msm_d9, photo_date, notes_corolla, leafarea_notes, leafweight_notes)) %>% #Remove unnecessary columns 
  select(Year, Sequence, Cohort, Block, Transect, Donor, Recipient, Gen, Replicated, Plant_ID, F_plant:Plant_Date, FFD, LFD, fl_duration, days_germ2flower:days_plant2flower, corolla_diam_mm, corolla_diam_mm_SEG, corolla_area_mm2, corolla_area_mm2_SEG, leaf_area_mm2:LMA, leaf_area_mm2_SEG:LMA_SEG, skel_dryweight_mg, skel_dryweight_mg_SEG, d13C, d13C_SEG, sample_weight_mg, closed_fruits:est_fecundity, est_fecundity_int, everything()) #Reorder columns

```


**Writing CSVs**  

```{r}
#Save full (long) version
write_csv(HR_mastersheet_v3, here::here("data_sheets", "compiled_sheets", "HR_mastersheet_full_2023-11-06.csv"))


## Fitness Plants Only
#
HR_mastersheet_F_mains <-  HR_mastersheet_v3 %>% 
  filter(F_plant == TRUE) #Only main fitness plants

HR_mastersheet_F_all <- HR_mastersheet_v3 %>% 
  filter(any_FitP == TRUE) #main and multiple fitness plants (IDs replicated in watering experiment)


write_csv(HR_mastersheet_F_mains, here::here("data_sheets", "compiled_sheets", "HR_mastersheet_Fitness-mains_2023-11-06.csv"))

write_csv(HR_mastersheet_F_all, here::here("data_sheets", "compiled_sheets", "HR_mastersheet_Fitness-all_2023-11-06.csv"))

```



