---
title: "compile_datasheets_BO"
author: "Devin Gamble"
date: "2022-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



### Load and Clean Data   


**Load Fitness Plant Data**  

- Add in Donor information
```{r message = FALSE}
BO_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "BO_fitnessplants_23.csv")) %>% 
  slice(-620)

#Donor info loaded from 2021-22 data
BO_donors <- read_csv(here::here("data_sheets", "tray_positions_2021_year1","BO_tray_rand.csv"), skip = 1) %>% 
  select(Donor_ID:Recipient_ID) %>% 
  rename(Donor = Donor_ID, Recipient = Recipient_ID)

#Fix Recipient and Donor names
BO_donors2 <- BO_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("BO_", Donor_num))) %>% 
  select(-Recipient_num, -Donor_num)



#Multiple matches, merge only unique pairs for Donor
BO_Fitness_2 <- merge(BO_Fitness, unique(BO_donors2[c("Donor", "Recipient")]), by = c("Recipient"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Sequence, SegPos)

#Save as V2 Fitness Plant Sheet
write_csv(BO_Fitness_2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "BO_fitnessplants_23_V2.csv"), na = "")


```

<br>  


### Trait Data  

Trait data gathered and processed from the 2022-2023 field season is loaded and organized below.  


#### Flower Check & Phenology  
-- Organize flowering data
-- Add in Sow & Plant dates

#########
Crucial Step!!  
######### 
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R (and lubridate) much easier!  

```{r message = FALSE}

BO_flr_df <- read_csv(file = here::here( "data_sheets", "2023_year2_data", "flowering_checks", "BO_flowercheck_data_2023-02-24 - BO_flowercheck_data_v1.csv"), na = c("", "NA")) %>% 
  rename(F_Num_02 = 19)


BO_flr2 <- BO_flr_df %>% 
  mutate(SegPos = paste(Segment, side, sep = "")) %>% 
  select(Sequence, Transect, SegPos, everything(), -c(Segment, side, FD_06, FD_07)) %>% 
  slice(-c(698:700, 779:780, 3096:3102)) %>% #Remove individuals not planted and NA rows
  arrange(Sequence, Transect, SegPos)
  

#Check for mismatched fitness plants
BO_flr_test <- BO_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))


## Fix Dates!

#Check obs first using unique(BO_flr3$FD_01) and make corrections

BO_flr3 <- BO_flr2 %>%
  mutate_at(.vars = vars(FD_01:FLR_P), ~ as_date(.))

```


Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

BO_flr4 <- BO_flr3 %>% 
  mutate(Sow_Date = as_date("2022-10-31"),
         Plant_Date = as_date("2022-11-03"))

```


Merge in Fitness plant Data  
- Donor
```{r}
#Check to make sure fitness plants line up
BO_flr5 <- merge(BO_flr4, BO_Fitness_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Transect, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>% 
  select(-c(Fit_plant.x, Fit_plant.y)) #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  
  
#####  
#Check to make sure fitness plants line up
# BO_flr_test <- BO_flr5 %>% 
#   filter(!is.na(Fit_plant.x)) %>%
#   mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
#                            Fit_plant.x != Fit_plant.y ~ FALSE))
# #
```



**Phenology Calculations**  
 

```{r}
#Same as last year

BO_flowering <- BO_flr5 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())



#TEMP FILE - delete later
write_csv(BO_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BO_phenology_2023.csv"), na = "")

```


<br>  


#### Fruit and Seed counts  

Fruit & seed counts were recorded on the same spreadsheet, unlike last year. Loaded here and checked to make sure all ID info lines up with the Fitness data sheet.


```{r message = FALSE}
#Load Fruit & Seeds count data

BO_fruit_seeds <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "BO_fruits_seeds_2023 - BO_fruit_seed_counts.csv"), skip = 1) %>% 
  rename(SegPos = Seg)


# #Remove rows where no fruits collected
# BO_fruit_seeds_2 <- BO_fruit_seeds %>% 
#   filter(total_fruits != 0)



##Check Fitness plants
BO_merge_test <- merge(BO_Fitness_2, BO_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient"), all.y = TRUE) %>% 
  mutate(FITTRUE = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
                             Fit_plant.x != Fit_plant.y ~ FALSE))


BO_fruit_seeds_3 <- merge(BO_Fitness_2, BO_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient", "Fit_plant"), all.x = TRUE) %>% 
  select(-Sequence.y) %>% 
  rename(Sequence = Sequence.x) %>% 
  rename(Plant_ID = Fit_plant)
```

<br>  

#### Skeleton Weights  

```{r message = FALSE}

skel_weights_BO <- read_csv(here::here("data_sheets", "2023_year2_data",  "skeleton_data", "Skeleton Data 2023 - BO Skeletons.csv")) %>% 
  rename(position = side, Gen = 4, Fitness = 7, skel_weight_mg = 8, complete_incomplete = 9, selected = 10) %>% 
  arrange(Transect, Segment, position) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Transect, SegPos, Gen,  Recipient, Plant_ID, Fitness, skel_weight_mg) #complete/selected not included


#Check that IDs line up
skelcheck <- merge(BO_Fitness_2, skel_weights_BO, by = c("Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE)


#Merge with full data set
#BO_merged_df5 <- merge(BO_merged_df4, skel_weights_BO, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)


```


**TEMP MERGED DATA FRAME**  
Merge trait data processed thus far

```{r}

BO_merged_1 <- merge(BO_flowering, skel_weights_BO, by = c("Transect", "SegPos", "Gen", "Recipient", "Plant_ID"), all.x = TRUE) %>% select(-c(Skeleton:Fitness)) %>% arrange(Transect, Sequence, SegPos)

BO_merged_2 <- merge(BO_merged_1, BO_fruit_seeds_3, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient", "Donor", "Plant_ID"), all.x = TRUE) %>% select(-c(FD_01:FD_05, N_CLOSED_fruit_D1:notes)) %>% arrange(Transect, Sequence, SegPos) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, everything())

```



<br>  



#### Seed Weights  

Update: Use revised seed counts in seed weight data sheet. Viable seed counts were corrected in this data sheet, so these values should be used instead of the ones from the seed count data sheet.

**Manual Corrections** - RE-WEIGH SEEDS or Convert 0 to NA for mean-mass_d6 Tr9 Seg 18R BO_125 AND Tr 10, Seg 30R BO_018
Day 6 weights (Date NA/missing) only has one entry with missing seed weight

```{r}

#Load Data
#Rename columns to standard format

BO_seedmass_2023 <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "BO_Seed_Weights_2023-All_Data.csv"), skip = 1) %>% 
  select(-c(5)) %>% #rm Symbol column
  rename(SegPos = Seg, Plant_ID = Fit_plant, msm_all = 7, 
         closed_fruits_d1 = 8, viable_seeds_d1 = 9, seeds_weighed_d1 = 10, seed_mass_mg_d1 = 11, mean_mass_mg_d1 = 12, 
         closed_fruits_d2 = 13, viable_seeds_d2 = 14, seeds_weighed_d2 = 15, seed_mass_mg_d2 = 16, mean_mass_mg_d2 = 17,
         closed_fruits_d3 = 18, viable_seeds_d3 = 19, seeds_weighed_d3 = 20, seed_mass_mg_d3 = 21, mean_mass_mg_d3 = 22,
         closed_fruits_d4 = 23, viable_seeds_d4 = 24, seeds_weighed_d4 = 25, seed_mass_mg_d4 = 26, mean_mass_mg_d4 = 27,
         closed_fruits_d5 = 28, viable_seeds_d5 = 29, seeds_weighed_d5 = 30, seed_mass_mg_d5 = 31, mean_mass_mg_d5 = 32,
         closed_fruits_d6 = 33, viable_seeds_d6 = 34, seeds_weighed_d6 = 35, seed_mass_mg_d6 = 36, mean_mass_mg_d6 = 37) %>% 
  filter(!is.na(Recipient))


#Calculate number of viable seeds 
#Calculate SD of grand mean seed mass

BO_seedmass_2023_B <- BO_seedmass_2023 %>% 
  mutate(ALL_filled_seeds_N = rowSums(across(c(viable_seeds_d1, viable_seeds_d2, viable_seeds_d3, viable_seeds_d4, viable_seeds_d5, viable_seeds_d6)), na.rm = TRUE)) %>% #sum viable seed counts across dates (N for NEW)
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(mean_mass_mg_d1, mean_mass_mg_d2, mean_mass_mg_d3, mean_mass_mg_d4, mean_mass_mg_d5, mean_mass_mg_d6)), na.rm = TRUE)) %>%  #calculate total standard deviation of mean seed mass
  select(Sequence:Plant_ID, ALL_filled_seeds_N, msm_all, sd_msm_all, everything())
#SD only available when viable seeds from two or more different dates were weighed 
#Each statistical 'unit' is the mean seed weight calculated on a given date
         
  
##Merge with fruit/seed count data frame

BO_merged_2.5 <- merge(BO_merged_2, BO_seedmass_2023_B, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient", "Plant_ID"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.
#Can toggle to all.y = TRUE to confirm merge success (e.g., all rows have matching donor, F_plant status)


#Check merge success
#Only fitness plants should have seed weight data
BO_merge_check <- BO_merged_2.5 %>% 
  filter(!is.na(msm_all))

BO_merge_check2 <- BO_merged_2.5 %>% 
  filter(!is.na(ALL_filled_seeds_N)) #Should match original seed count Nrows

## Correct viable seed counts
BO_merged_3 <- BO_merged_2.5 %>% 
  mutate(seed_recount_diff = ALL_filled_seeds_N - ALL_filled_seeds) %>% #Check data sheets for high diffs
  select(-c(seed_recount_diff, ALL_filled_seeds)) #comment out this line to check before proceeding
```

<br>  



#### Corolla measurements  
[[UPDATE]]
- Format dates in excel prior to importing
```{r}

corollas_BO <- read_csv(here::here("data_sheets", "2023_year2_data", "corolla_leaf_data", "Corolla Diameters 2023 - BO_corollas.csv")) %>% select(-c(4,10, 11)) %>% 
  rename(corolla_notes = Notes, photo_date = `Photo date`) %>% 
  filter(!grepl("REMOVE", corolla_notes)) ## Filter out notes with "REMOVE"

corollas_BO2 <- corollas_BO %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually 
  arrange(Transect, Recipient, Plant_ID, photo_date) %>% 
  distinct(Transect, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates
#0 duplicates found

## Merge with main data 


##Create segment-wide corolla diameter column

```

<br>  


#### Leaf Area & Weight  
[[OLD - UPDATE]]

```{r}
leaves_BO <- read_csv(here::here("data_sheets", "leaf_data", "BO_SLA_data_20230522.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) #exlcude NA leaf areas?

#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_BO_duped <- leaves_BO %>% 
  mutate(duped1 = duplicated(select(leaves_BO, c(Block, Transect, Recipient))),
         duped2 = duplicated(select(leaves_BO, c(Block, Transect, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!

#This seems to work the best!
leaves_BO_unique <- leaves_BO_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) %>% #Don't need plant ID...these were all non-fitness plants
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg,
         LMA = 1 / SLA)


#Merge in to main data 

BO_merged_df7 <- merge(BO_merged_df6, leaves_BO_unique, by = c("Block", "Transect", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo


```

<br>  


#### d13C data  
[[OLD - UPDATE]]

```{r}
Nemo_d13C <- read_csv(here::here("data_sheets", "isotope_data", "Nemophila_d13C_230511.csv")) %>% 
  select(1,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error

#filter to BO samples
Nemo_d13C_BO <- Nemo_d13C %>% 
  slice(-54) %>% 
  filter(grepl('BO', sample_ID))

#unravel sample ID
Nemo_d13_BO2 <- Nemo_d13C_BO %>% 
  mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 6)), 
         Transect = as.numeric(str_remove(substr(sample_ID, 8, 10), "^0+")),
         Plant_ID = as.numeric(substr(sample_ID, 12,13))) %>% 
  select(Transect, Recipient, Plant_ID, d13C, sample_weight_mg, -Plant_ID)

#Merge

BO_merged_df_fl3 <- merge(BO_merged_df_fl2, Nemo_d13_BO2, by = c("Recipient", "Transect"), all.x = TRUE)

#check that all data made it (196 obs)
testdf <- BO_merged_df_fl3 %>% 
  filter(!is.na(d13C))

#One sample missing from main data - Tr 5, BO_118 plant 8 -- no such recipient in Tr 5, leave out of data

```

<br>  


**Error Checks**  

```{r eval = FALSE}
##Check for mismatched (fitness) plants

BO_Echeck1 <- BO_merged_df6 %>% 
  filter(any_FitP == FALSE & total_fruits_F != 0) #do the same for seed count, using `>` and `!=`

BO_Echeck2 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
BO_Echeck3 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))


BO_Echeck4 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))

## Check for NA seed counts for fitness plants (when fruit production > 0)
BO_Echeck5 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))


#Seed mass
BO_Echeck5 <- BO_merged_df6 %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


### Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd`  


```{r}

#Survival measures
BO_merged_df1 <- BO_merged_2 %>% 
  mutate(surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(F_plant == TRUE & total_fruits > 0 ~ 1, #plants that survived to produce fruit
                                       F_plant == TRUE & total_fruits == 0 ~ 0),
         prod_closed_fruits = case_when(F_plant == TRUE & ALL_closed_fruit > 0 ~ 1,
                                            F_plant == TRUE & ALL_closed_fruit == 0 ~ 0),
         prod_filled_seeds = case_when(F_plant == TRUE & ALL_filled_seeds > 0 ~ 1, #plants produced filled seeds
                                      F_plant == TRUE & ALL_filled_seeds == 0 ~ 0)) #else = 0 (applies to seed_ct = NA)

         
BO_merged_df2 <- BO_merged_df1 %>% 
  rename(mean_seeds_per_fruit = avg_seeds_per_fruit) %>% 
  mutate(across(c(est_fecundity, mean_seeds_per_fruit), ~ as.numeric(na_if(., "#DIV/0!")))) %>% #Converts to NA 
  mutate(est_fecundity = case_when(F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit == 0 ~ 0,
                                   F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit >= 1 ~ NA,
                                               F_plant == TRUE & !is.na(est_fecundity) ~  est_fecundity)) 

# estimated fecundity is 0 when 0 closed or open fruits collected, or when 0 filled seeds collected
# estimated fecundity is NA when 0 closed fruits but 1 or more open fruits collected


```

<br>  



### Master sheets (2023)  

Full and summarized data sheets for all data collected in the 2022-2023 field season. 
These are merged with the year 1 data sheets elsewhere.  

```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...


######
#TEMP VERSION
######

BO_merged_full_1 <- BO_merged_df2 %>% 
  filter(F_plant == TRUE)
##NOTES - there are still fitness plants that never flowered... keep as 0s


write_csv(BO_merged_full_1, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BO_traits_Y2_incomplete_20230922.csv"), na = "")

######
```

**TO BE UPDATED** 
```{r eval = FALSE}
######



BO_mastersheet_v1 <- BO_merged_df_fl3 %>% 
  mutate(Year = 2022, Gen = "G1") %>% 
  select(Year, Gen, Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

write_csv(BO_mastersheet_v1, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BO_mastersheet_full_Y2_2023-08-07.csv"))

#Create meta data sheet



#
## Fitness Plants Only
#

BO_mastersheet_F_mains <-  BO_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only single fitness plants

BO_mastersheet_F_all <- BO_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treate these as 0 fitness or as NAs to remove???


write_csv(BO_mastersheet_F_mains, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BO_mastersheet_Fitness-mains_Y2_2023-08-07.csv"))

write_csv(BO_mastersheet_F_all, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BO_mastersheet_Fitness-all_Y2_2023-08-07.csv"))

```



