---
title: "Compile FLower, Fruit, and Seed count data sheets - AC Fitness plants"
output: html_notebook
author: Devin Gamble
---


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



## Load and Clean Data   



**Load Fitness Plant Data**  

- Add in Donor information
```{r}
AC_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "AC_fitness_23.csv")) %>% 
  slice(-639) %>% 
  mutate(Recipient = case_when(Recipient %in% c("AC_159", "AC_159A", "AC_159B") ~ "AC_159",
                               TRUE ~ Recipient)) %>%  #rename 159 - letters were added to mark position
  slice(-423) #AC_213 in Tr 6 seg 33L was never planted - remove entirely

AC_donors <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "AC_donor_info.csv")) %>% 
  select(-position) 

#Fix Recipient and Donor names
AC_donors2 <- AC_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("AC_", Donor_num))) %>% 
  select(-Recipient_num, -Donor_num)



#Multiple matches, merge only unique pairs for Donor
AC_Fitness_2 <- merge(AC_Fitness, unique(AC_donors2[c("Gen", "Donor", "Recipient")]), by = c("Recipient", "Gen"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Sequence, SegPos)

#Save as V2 Fitness Plant Sheet
write_csv(AC_Fitness_2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "AC_fitnessplants_23_V2.csv"), na = "")


```






**Flower Check sheets**  
-- Organize flowering data
-- Add in Sow & Plant dates


##################
**Crucial Step!!**  
##################  
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R much easier!  



```{r}
##AC##

AC_flr_df <- read_csv(file = here::here( "data_sheets", "2023_year2_data", "flowering_checks", "AC_flowercheck_data_2023-03-15 - AC_flowercheck_data_v1.csv"), na = c("", "NA")) 


AC_flr2 <- AC_flr_df %>% 
  mutate(SegPos = paste(Segment, side, sep = "")) %>% 
  select(Sequence, Transect, SegPos, everything(), -c(Segment, side, FD_06, FD_07)) %>% 
  slice(-c(1355, 1600, 2111:2115, 3191:3195)) %>% #Remove rows not planted
  arrange(Sequence, Transect, SegPos) %>% 
  mutate(Recipient = case_when(Recipient %in% c("AC_159", "AC_159A", "AC_159B") ~ "AC_159",
                               TRUE ~ Recipient)) #rename 159 - letters were added to mark position
  

#Check for mismatched fitness plants
AC_flr_test <- AC_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))


## Fix Dates!

#Check obs first using unique(AC_flr3$FD_01) and make corrections

AC_flr3 <- AC_flr2 %>%
  mutate_at(.vars = vars(FD_01:FD_05), ~as_date(.)) %>% 
  mutate(FD_05 = as_date(case_when(FD_05 == "2023-06-07" ~ "2023-06-08",
                           FD_05 == "2023-06-08" ~ "2023-06-08",
                           FD_05 == "2023-06-09" ~ "2023-06-09",
                           TRUE ~ as.character(NA)))) #Some date typos - one day early

```


Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

AC_flr4 <- AC_flr3 %>% 
  mutate(Sow_Date = as_date("2022-10-17"),
         Plant_Date = as_date("2022-10-21"))

```


Merge in Fitness plant Data  
- Donor

```{r}
#Check to make sure fitness plants line up
AC_flr5 <- merge(AC_flr4, AC_Fitness_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Transect, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>% 
  select(-c(Fit_plant.x, Fit_plant.y)) #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  
  
#####  
#Check to make sure fitness plants line up
  #filter(!is.na(Fit_plant.x)) %>%
  #mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
  #                          Fit_plant.x != Fit_plant.y ~ FALSE))
#
```


#### Phenology Calculations  
<<Move these lower later??>>  

```{r}
#Same as last year

AC_flowering <- AC_flr5 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())



#TEMP FILE - delete later
write_csv(AC_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "AC_phenology_2023.csv"), na = "")

```




<br>  












**Fruit and Seed counts**  

Due to the structure of excel data sheets, it seems easiest to upload fruit and seed count data separately since different spreadsheets were used to record each. Info on plants' Block, Transect, sequence, etc. will be pulled from the fruit count sheets (should be up-to-date). For sheets intended to record fruit production only, counts of viable seeds, mean seed weight, etc. that are copied over are ignored and instead pulled straight from sheets where they were first entered later double-checked.

Date added: **Feb 05, 2023**  

*NOTE*: All typos believed to have been removed. Individual rows of fruit/seed data cross-checked against listed fitness plant and look good. 
```{r message = FALSE}
#
# Fruit count data
# Data on *when* fruits were collected will be omitted here; only the total fruit counts (closed, open, non-fitness) are retained from the csv

AC_fruitprod_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "AC-fruit-production_2023-02-05.csv"), skip = 2) %>% 
  select(1:11) #omit sow and flower date, seed count, and date-specific fruit counts by column index

AC_fruitprod_2 <- AC_fruitprod_1 %>% 
  select(-c(Recipient, 9)) %>% 
  rename(Recipient = `Recipient for Sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, closed_fruits_F = `Number of closed fruits used to estimate mean seeds per fruit (include multiple dates)`, open_fruits_F = `Open Fruits`, damaged_fruits_F = `Eaten or damaged fruits`, total_fruits_F = `Total number of all fruits produced by the plant (open + closed)`, nonfitness_fruits = `Total number of non-fitness fruits collected from segment`) 

#Note: total fruits should not include eaten/damaged fruit


#
# Seed count data (viable seeds)
# Similarly, only seed count data are retained from this spreadsheet to minimize errors

AC_seedct_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "AC-seed-count_2023-02-05.csv"), skip = 2) %>% 
  select(-c(3, 6, 7, 9:29)) %>% 
  rename(Recipient = `Recipient for Sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, seed_ct = `Total number of viable seeds counted across multiple dates`)



#merge data
AC_fruitseed_ct1 <- left_join(AC_fruitprod_2, AC_seedct_1, by = c("Sequence", "Transect", "Recipient", "Plant_ID"))
#Sequence data in fruit count sheets may be unreliable for some sites (e.g., BO)


# summary variables, tidying up
AC_fruitseed_ct2 <- AC_fruitseed_ct1 %>% 
  select(Transect, Sequence, Recipient, Plant_ID, closed_fruits_F, open_fruits_F, damaged_fruits_F, total_fruits_F, nonfitness_fruits, everything()) %>% 
  arrange(Transect, Sequence, Recipient)

## NOTE: total number of non-fitness fruits per segment may be under-represented here since not all counts were matched to fitness plants and therefore merged in.

```



**Seed Weights**  

*Note*: Prior to csv reading:
- header row for seed weights was moved down to line up with other headers


```{r}
# Read in only # seeds weighed and total weight
# Calculate mean weight per seed

AC_seedmass_1 <- read_csv(here::here("data_sheets", "seed_mass", "AC_Seed_Mass_2023-02-05.csv"), skip = 2)

colnames(AC_seedmass_1) <- c("Sequence", "Transect", "Recipient_old", "Recipient", "Plant_ID", "closed_fruits_d1", "viable_seeds_d1", "seeds_weighed_d1", "seed_mass_mg_d1", "closed_fruits_d2", "viable_seeds_d2", "seeds_weighed_d2", "seed_mass_mg_d2", "closed_fruits_d3", "viable_seeds_d3", "seeds_weighed_d3", "seed_mass_mg_d3", "closed_fruits_d4", "viable_seeds_d4", "seeds_weighed_d4", "seed_mass_mg_d4", "closed_fruits_d5", "viable_seeds_d5", "seeds_weighed_d5", "seed_mass_mg_d5")


#Unsure how to account for number of seeds weighed - each statistical 'unit' is the mean seed weight calculated on a given date

AC_seedmass_2 <- AC_seedmass_1 %>% 
  select(-starts_with(c("closed_fruits_", "viable_seeds_")), -Recipient_old) %>% 
  mutate(msm_d1 = seed_mass_mg_d1/seeds_weighed_d1,
         msm_d2 = seed_mass_mg_d2/seeds_weighed_d2,
         msm_d3 = seed_mass_mg_d3/seeds_weighed_d3,
         msm_d4 = seed_mass_mg_d4/seeds_weighed_d4,
         msm_d5 = seed_mass_mg_d5/seeds_weighed_d5)
         

AC_seedmass_3 <- AC_seedmass_2 %>%
  mutate(msm_all = rowMeans(select(AC_seedmass_2, starts_with("msm_")), na.rm = TRUE)) %>% 
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(msm_d1:msm_d5)), na.rm = TRUE))
         
  
##Merge with fruit/seed count data frame

AC_fruitseed_ct3 <- merge(AC_fruitseed_ct2, AC_seedmass_3, by = c("Transect", "Sequence", "Recipient", "Plant_ID"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```



Merge everything together:  
```{r}
AC_merged_df1 <- merge(AC_GermFlow_3, AC_fruitseed_ct3, by = c("Transect", "Sequence", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID) %>% 
  mutate(closed_fruits_F = as.numeric(closed_fruits_F),
         total_fruits_F = as.numeric(total_fruits_F)) #correct counts to numeric values


#Correct Fruit NAs (but not seed count NAs) to 0s in cases where no fruit data was entered (fitness and non-fitness plants):

AC_merged_df2 <- AC_merged_df1 %>% 
  mutate(closed_fruits_F = replace_na(closed_fruits_F, 0),
         total_fruits_F = replace_na(total_fruits_F, 0),
         seed_ct = replace_na(seed_ct, 0)) #REPLACE ALL seed_ct NAs with 0 


#Calculate mean seeds per fruit, total_est_seed prod
AC_merged_df3 <- AC_merged_df2 %>% 
  mutate(mean_seeds_per_fruit = seed_ct/closed_fruits_F,
         total_est_seed_production = mean_seeds_per_fruit*total_fruits_F) %>% 
  mutate(mean_seeds_per_fruit = case_when(any_FitP == TRUE & closed_fruits_F == 0 ~ 0, 
                                          TRUE ~ mean_seeds_per_fruit)) %>%  #else, original value
  mutate(total_est_seed_production = case_when(any_FitP == TRUE & mean_seeds_per_fruit == 0 ~ 0,
                                               TRUE ~ total_est_seed_production))
#These values will be 0 when seed_ct was recorded as NA (no closed fruits or non-fitness plant) or 0


##commented out code was first used to replace NAs with 0s when fitness plants had closed fruits for mean seeds & est seed production


#mean seeds per fruit and total est fruit production have NAs/NaNs for when there were no closed fruits/seeds (fitness plants only)

```



**Skeleton Weights & Corolla area and diameter**  

Load and merge-in these data sheets
```{r}
##
skel_weights_AC <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeleton_weights_AC_2023_03_01.csv")) %>% select(-Units, -7, -8) %>% 
  filter(!grepl("EXCLUDE", Notes)) %>% ## Filter out notes with "Exclude"
  rename(Plant_ID = `Fitness Plant_ID`, AG_biomass_mg = `Skeleton Weights (mg)`)

#Watered points
skel_weights_AC_W <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeletons_watered_AC_2023_02_08.csv")) %>% 
  select(-`Units (mg)`) %>% 
  rename(Plant_ID = `Plant ID`, AG_biomass_mg = `Skeleton Weights`)

#Combine
skel_weights_all <- rbind(skel_weights_AC, skel_weights_AC_W) %>% 
  select(-Notes) %>% 
  filter(!is.na(AG_biomass_mg))


#Merge with full data set
AC_merged_df4 <- merge(AC_merged_df3, skel_weights_all, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)


```


Corolla measurements
```{r}
corollas_AC <- read_csv(here::here("data_sheets", "corolla_measures", "corollameasures_AC_2023-02-28.csv")) %>% select(1, 3, 4, 6, 7, 9, 11) %>% 
  rename(photo_date = 4, corolla_diam_mm = 5, corolla_area_mm2 = 6, notes_corolla = 7) %>% 
  filter(!grepl("REMOVE", notes_corolla)) ## Filter out notes with "REMOVE"

corollas_AC2 <- corollas_AC %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  arrange(Transect, Recipient, Plant_ID, photo_date) %>% 
  distinct(Transect, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates
#Only about 6 duplicate observations removed


## Merge with main data 
AC_merged_df5 <- merge(AC_merged_df4, corollas_AC2, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)
  
```




**Leaf Area & Weight**  
```{r}
leaves_AC <- read_csv(here::here("data_sheets", "leaf_data", "AC_SLA_data_20230522.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) %>% #exlcude NA leaf areas?
  mutate(leaf_area_mm2 = as.numeric(leaf_area_mm2))



#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_AC_duped <- leaves_AC %>% 
  mutate(duped1 = duplicated(select(leaves_AC, c(Block, Transect, Recipient))),
         duped2 = duplicated(select(leaves_AC, c(Block, Transect, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!


#This seems to work the best!
leaves_AC_unique <- leaves_AC_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) %>%  #Don't need plant ID...these were all non-fitness plants
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg,
         LMA = 1 / SLA)


#Merge in to main data 

AC_merged_df6 <- merge(AC_merged_df5, leaves_AC_unique, by = c("Block", "Transect", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo



```



Error checks:

```{r eval = FALSE}
##Check for mismatched (fitness) plants

AC_Echeck1 <- AC_merged_df5 %>% 
  filter(any_FitP == FALSE & total_fruits_F != 0) #do the same for seed count, using `>` and `!=`

AC_Echeck2 <- AC_merged_df5 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
AC_Echeck3 <- AC_merged_df5 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))

AC_Echeck4 <- AC_merged_df5 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))


## Check for NA seed counts for fitness plants (when fruit production > 0)
#Will be empty when seed_ct NAs are replaced with 0
AC_Echeck5 <- AC_merged_df5 %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))
##NAs present here before setting seed_ct NAs to 0 were from fitness plants that had 0 closed fruit


#Seed mass
AC_Echeck5 <- AC_merged_df5 %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


## Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd`  


#### Phenology  

```{r}

AC_merged_df_fl <- AC_merged_df6 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('Fl_')), na.rm = TRUE)), #First flowering date
         days_germ2flower = as.numeric(FFD - Germ_Date), #days from germination to flowering
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('Fl_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) #Flowering duration


#Survival measures
AC_merged_df_fl2 <- AC_merged_df_fl %>% 
  mutate(surv_to_germ = case_when(!is.na(Germ_Date) ~ 1, #plants that germinated
                                  is.na(Germ_Date) ~ 0),
         surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(total_fruits_F > 0 ~ 1, #plants that survived to produce fruit
                                       total_fruits_F == 0 ~ 0),
         surv_to_seedprod = case_when(seed_ct > 0 ~ 1, #plants that survived to produce fruit with viable seeds
                                       TRUE ~ 0)) #else = 0 (applies to seed_ct = NA)

         


## Note: should flowering duration remain unchanged when a plant went several weeks without flowering (e.g., had two temporally distinct blooms during the growing season)?
```

<br>  


### d13C data  

```{r}
Nemo_d13C <- read_csv(here::here("data_sheets", "isotope_data", "Nemophila_d13C_230511.csv")) %>% 
  select(1,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error

#filter to AC samples
Nemo_d13C_AC <- Nemo_d13C %>% 
  filter(grepl('AC', sample_ID))

#unravel sample ID
Nemo_d13_AC2 <- Nemo_d13C_AC %>% 
  mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 6)), 
         Transect = as.numeric(str_remove(substr(sample_ID, 8, 10), "^0+")),
         Plant_ID = as.numeric(substr(sample_ID, 12,13))) %>% 
  select(Transect, Recipient, Plant_ID, d13C, sample_weight_mg, -Plant_ID)

#Merge

AC_merged_df_fl3 <- merge(AC_merged_df_fl2, Nemo_d13_AC2, by = c("Recipient", "Transect"), all.x = TRUE)
#Merge by Recipient and Transect only: genotype-wide estimates of d13C since some plants were NF



#check that all data made it (266 obs)
#testdf <- AC_merged_df_fl3 %>% 
#  filter(!is.na(d13C))
```





## Master sheets (v1)  


```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...


AC_mastersheet_v1 <- AC_merged_df_fl3 %>% 
  mutate(Year = 2022, Gen = "G1") %>% 
  select(Year, Gen, Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

write_csv(AC_mastersheet_v1, here::here("data_sheets", "compiled_sheets", "AC_mastersheet_full_2023-08-07.csv"))
#Renamed to full - has records of all plants (including non-fitness)



#
## Fitness Plants Only
#

AC_mastersheet_F_mains <-  AC_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only main fitness plants

AC_mastersheet_F_all <- AC_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treate these as 0 fitness or as NAs to remove???


write_csv(AC_mastersheet_F_mains, here::here("data_sheets", "compiled_sheets", "AC_mastersheet_Fitness-mains_2023-08-07.csv"))

write_csv(AC_mastersheet_F_all, here::here("data_sheets", "compiled_sheets", "AC_mastersheet_Fitness-all_2023-08-07.csv"))
```



