---
title: "compile_datasheets_BB"
author: "Devin Gamble"
date: "2022-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



### Load and Clean Data  


Recipient IDs to double-check:
- BB_205 (see note)
- BB_147 (BB_147a/b)
- BB_070 (BB_070A/B)
- IDs not planted (SegPos 208L)(see below)


NOTE: Two iterations of a maternal family ID (names fixed before saving csv):
**BB_205_A** occurs at locations xx, xx, and xxx and has Donor BB_217 (greenhouse cohort 5)
**BB_205_B** occurs at locations xx, xx, and xx and has Donor BB_188 (cohort 6)


**Fitness Plant Data**  
Update Fitness plant data
- Add in Donor information

```{r message = FALSE}
BB_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "BB_fitnessplants_23.csv")) %>% 
  mutate(Recipient = case_when(Recipient %in% c("BB_205", "BB_205A", "BB_205B") ~ "BB_205",
                               TRUE ~ Recipient)) %>%  #reset 205 - A/B specified below based on donor
  mutate(Recipient = case_when(Recipient %in% c("BB_147", "BB_147a", "BB_147b") ~ "BB_147",
                               TRUE ~ Recipient)) %>% # 147 renamed based on positions. Reset here.
  mutate(Recipient = case_when(Recipient %in% c("BB_070", "BB_070A", "BB_070B") ~ "BB_070",
                               TRUE ~ Recipient)) %>% 
  slice(-c(415, 644)) #BB_157 in Pt 70 seg 208L was never planted - remove entirely. Last row NULL
#BB_050 at 283L plant 5 removed (never planted) below in flowering sheet

BB_donors <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "BB_donor_info.csv")) 

#Fix Recipient and Donor names
BB_donors2 <- BB_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BB_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("BB_", Donor_num))) %>% 
  select(-Recipient_num, -Donor_num) 



#Multiple matches, merge only unique pairs for Donor
BB_Fitness_2 <- merge(BB_Fitness, unique(BB_donors2[c("Sequence", "Gen", "Donor", "Recipient")]), by = c("Sequence", "Recipient", "Gen"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Sequence, position, sep = "")) %>% 
  select(Sequence, Location, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Location, SegPos) 


#Mismatched rows?
#which(duplicated(BB_Fitness_2[,1:4]))


#Add in Location coordinates
BB_planting <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "BB_planting_locations_fall22.csv"), skip = 1) %>% rename(Location = Point, latitude = 8, longitude = 9) %>% 
  filter(!is.na(Location)) %>% slice(-c(109:111)) %>% 
  select(Location, latitude, longitude)

BB_Fitness_2a <- merge(BB_Fitness_2, BB_planting, by = "Location", all.x = TRUE)


#Use above Fitness df version for merging with flowering data. 
#Use following version (including 205 As and Bs) to update Fitness csv
BB_Fitness_2.5 <- BB_Fitness_2a %>% 
  mutate(Recipient = case_when(Donor == "BB_217" & Recipient == "BB_205" ~ "BB_205A",
                               Donor == "BB_188" & Recipient == "BB_205" ~ "BB_205B",
                               TRUE ~ Recipient)) #Fix 205 A/B labeling based on donor

#Save as V2 Fitness Plant Sheet
write_csv(BB_Fitness_2.5, here::here("data_sheets", "2023_year2_data", "fitness_plants", "BB_fitnessplants_23_V2.csv"), na = "")

```

### Trait Data  

Trait data gathered and processed from the 2022-2023 field season is loaded and organized below.  


#### Flower Check & Phenology  
-- Organize flowering data
-- Add in Sow & Plant dates


#########
Crucial Step!!  
######### 
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R (and lubridate) much easier!  

```{r message = FALSE}

BB_flr_df <- read_csv(file = here::here("data_sheets", "2023_year2_data", "flowering_checks", "BB_flowercheck_data_2023-02-15 - BB_flowercheck_data_v1.csv"), na = c("", "NA")) 


BB_flr2 <- BB_flr_df %>% 
  mutate(SegPos = paste(Sequence, side, sep = "")) %>% 
  select(Sequence, Location, SegPos, everything()) %>% 
  slice(-c(2214:2218, 2882, 3216:3220)) %>% #Remove rows not planted
  arrange(Sequence, Location, SegPos) %>% 
  mutate(Recipient = case_when(Recipient %in% c("BB_205", "BB_205A", "BB_205B") ~ "BB_205",
                               TRUE ~ Recipient)) %>%  #reset 205 - A/B specifiedbased on donor
  mutate(Recipient = case_when(Recipient %in% c("BB_147", "BB_147a", "BB_147b") ~ "BB_147",
                               TRUE ~ Recipient)) %>%  # 147 renamed based on positions. Reset here.
  mutate(Recipient = case_when(Recipient %in% c("BB_070", "BB_070A", "BB_070B") ~ "BB_070",
                               TRUE ~ Recipient)) %>% 
  mutate_at(.vars = vars(FD_01:FD_09), ~na_if(., "G")) #Remove Gopher notes



#Check for mismatched fitness plants
BB_flr_test <- BB_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))


## Fix Dates!

#Check obs first using unique(BB_flr2$FD_01) and make corrections
#Remember to format dates first in the csv file!

BB_flr3 <- BB_flr2 %>%
  mutate_at(.vars = vars(FD_01:FD_09), ~ as_date(.))

```


Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

BB_flr4 <- BB_flr3 %>% 
  mutate(Sow_Date = as_date("2022-10-03"),
         Plant_Date = as_date("2022-10-06"))

```


Merge in Fitness plant Data  
- Donor

```{r}
#Check to make sure fitness plants line up
BB_flr5 <- merge(BB_flr4, BB_Fitness_2a, by = c("Sequence", "Location", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Location, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>%
  mutate(Recipient = case_when(Donor == "BB_217" & Recipient == "BB_205" ~ "BB_205A",
                               Donor == "BB_188" & Recipient == "BB_205" ~ "BB_205B",
                               TRUE ~ Recipient)) %>% #Fix 205 A/B labeling based on donor
  select(-c(Fit_plant.x, Fit_plant.y, side)) %>% #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  slice(-c(2071:2075))# Remove Segment that was never actually planted
  
  
#####  
#Check to make sure fitness plants line up
# BB_flrt <- BB_flr5 %>% 
#   filter(!is.na(Fit_plant.x)) %>%
#   mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
#                             Fit_plant.x != Fit_plant.y ~ FALSE))

```


**Phenology Calculations**  

Survivorship calculated later

```{r}
#Same as last year

BB_flowering <- BB_flr5 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  select(Sequence, Location, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())


#TEMP FILE - delete later
write_csv(BB_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BB_phenology_2023.csv"), na = "")

```

<br>  


#### Fruit and Seed counts  

Fruit & seed counts were recorded on the same spreadsheet, unlike last year. Loaded here and checked to make sure all ID info lines up with the Fitness data sheet.


*NOTE*: Update with later versions of data sheet when more seed counts have been entered  

```{r message = FALSE}
#Load Fruit & Seeds count data

BB_fruit_seeds <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "BB_fruits_seeds_2023 - BB_fruit_seed_counts.csv"), skip = 1) %>% 
  rename(SegPos = Seg)

#Remove rows where no fruits collected (optional)
# BB_fruit_seeds_2 <- BB_fruit_seeds %>% 
#   filter(total_fruits != 0)


##Check Fitness plants
BB_merge_test <- merge(BB_Fitness_2.5, BB_fruit_seeds, by = c("Location", "SegPos", "Gen", "Recipient", "Fit_plant"), all.y = TRUE) 
#Fit datasheets with updated labels used here



BB_fruit_seeds_3 <- merge(BB_Fitness_2.5, BB_fruit_seeds, by = c("Location", "SegPos", "Gen", "Recipient", "Fit_plant"), all.x = TRUE) %>% 
  rename(Plant_ID = Fit_plant) #Rename Plant ID for merging below AFTER double-checking fitness plants)

```

<br>  


#### Skeleton Weights  

```{r message = FALSE}
skel_weights_BB <- read_csv(here::here("data_sheets", "2023_year2_data",  "skeleton_data", "Skeleton Data 2023 - BB Skeletons.csv")) %>% 
  rename(position = side, Sequence = 2, Gen = 4, Fitness = 7, skel_weight_mg = 8, complete_incomplete = 9, selected = 10) %>% 
  arrange(Location, Sequence, position) %>% 
  mutate(SegPos = paste(Sequence, position, sep = "")) %>% 
  select(Location, SegPos, Gen,  Recipient, Plant_ID, Fitness, skel_weight_mg) #complete/selected not included

skel_weights_BB_2 <- skel_weights_BB %>% 
  mutate(Recipient = case_when(Recipient %in% c("BB_147", "BB_147a", "BB_147b") ~ "BB_147",
                               TRUE ~ Recipient)) %>% # 147 renamed based on positions. Reset here.
  mutate(Recipient = case_when(Recipient %in% c("BB_070", "BB_070A", "BB_070B") ~ "BB_070",
                               TRUE ~ Recipient)) #rename- letters were added to mark position

#Check that IDs line up
skelcheck <- merge(BB_Fitness_2, skel_weights_BB_2, by = c("Location", "SegPos", "Gen", "Recipient"), all.x = TRUE)
#Use Fitness_2.5 once 205A/B s added

#Merge with full data set
#BB_merged_df5 <- merge(BB_merged_df4, skel_weights_BB, by = c("Location", "Recipient", "Plant_ID"), all.x = TRUE)
```


**TEMP MERGED DATA FRAME**  
Merge trait data processed thus far

```{r}

BB_merged_1 <- merge(BB_flowering, skel_weights_BB_2, by = c("Location", "SegPos", "Gen", "Recipient", "Plant_ID"), all.x = TRUE) %>% select(-c(Skeleton:Fitness)) %>% arrange(Location, Sequence, SegPos)

BB_merged_2 <- merge(BB_merged_1, BB_fruit_seeds_3, by = c("Sequence", "Location", "SegPos", "Gen", "Recipient", "Donor", "Plant_ID"), all.x = TRUE) %>% select(-c(FD_01:FD_05, N_CLOSED_fruit_D1:N_total_seeds_D8)) %>% arrange(Location, Sequence, SegPos) %>% 
  select(Sequence, Location, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, everything(), -36)

```

<br>  


#### Seed Weights  
[[OLD - UPDATE]]  

```{r eval = FALSE}
# Read in only # seeds weighed and total weight
# Calculate mean weight per seed

BB_seedmass_1 <- read_csv(here::here("data_sheets", "seed_mass", "BB_Seed_Mass_2023-02-05.csv"), skip = 1) 
  

colnames(BB_seedmass_1) <- c("Block", "Location", "Recipient", "Plant_ID", "closed_fruits_d1", "viable_seeds_d1", "seeds_weighed_d1", "seed_mass_mg_d1", "closed_fruits_d2", "viable_seeds_d2", "seeds_weighed_d2", "seed_mass_mg_d2", "closed_fruits_d3", "viable_seeds_d3", "seeds_weighed_d3", "seed_mass_mg_d3", "closed_fruits_d4", "viable_seeds_d4", "seeds_weighed_d4", "seed_mass_mg_d4", "closed_fruits_d5", "viable_seeds_d5", "seeds_weighed_d5", "seed_mass_mg_d5", "closed_fruits_d6", "viable_seeds_d6", "seeds_weighed_d6", "seed_mass_mg_d6", 
                             "closed_fruits_d7", "viable_seeds_d7", "seeds_weighed_d7", "seed_mass_mg_d7",
                             "closed_fruits_d8", "viable_seeds_d8", "seeds_weighed_d8", "seed_mass_mg_d8", 
                             "closed_fruits_d9", "viable_seeds_d9", "seeds_weighed_d9", "seed_mass_mg_d9", "ffd")

#Fix Recipient IDs
BB_seedmass_1.5 <- BB_seedmass_1 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BB_", Recipient_num))) %>% 
  mutate(Plant_ID = as.numeric(Plant_ID)) %>% #set Plant ID to numeric for sorting purposes
  select(-c(Recipient_num, ffd))


BB_seedmass_2 <- BB_seedmass_1.5 %>% 
  select(-starts_with(c("closed_fruits_", "viable_seeds_"))) %>% 
  mutate(msm_d1 = seed_mass_mg_d1/seeds_weighed_d1,
         msm_d2 = seed_mass_mg_d2/seeds_weighed_d2,
         msm_d3 = seed_mass_mg_d3/seeds_weighed_d3,
         msm_d4 = seed_mass_mg_d4/seeds_weighed_d4,
         msm_d5 = seed_mass_mg_d5/seeds_weighed_d5,
         msm_d6 = seed_mass_mg_d6/seeds_weighed_d6,
         msm_d7 = seed_mass_mg_d7/seeds_weighed_d7,
         msm_d8 = seed_mass_mg_d8/seeds_weighed_d8,
         msm_d9 = seed_mass_mg_d9/seeds_weighed_d9)
         

BB_seedmass_3 <- BB_seedmass_2 %>%
  mutate(msm_all = rowMeans(select(BB_seedmass_2, starts_with("msm_")), na.rm = TRUE)) %>% 
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(msm_d1:msm_d9)), na.rm = TRUE))
         
  
##Merge with fruit/seed count data frame

BB_fruitseed_ct3 <- merge(BB_fruitseed_ct2, BB_seedmass_3, by = c("Block", "Location", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  mutate(Block = case_when(Block %in% c("W1","W2") ~ "W",
                           TRUE ~ Block)) #Correct Black names to "W" to avoid losing data below
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```

<br>  

#### Corolla measurements  
[[OLD - UPDATE]]

```{r eval = FALSE}
corollas_BB <- read_csv(here::here("data_sheets", "corolla_measures", "corollameasures_BB_2023-02-28.csv")) %>% select(1, 2, 3, 4, 6, 8, 9, 10) %>% 
  rename(Location = 2, Recipient = 3, Plant_ID = 4, photo_date = 5, corolla_diam_mm = 7, corolla_area_mm2 = 6, notes_corolla = 8) %>% 
  filter(!grepl("REMOVE", notes_corolla)) ## Filter out notes with "REMOVE"

corollas_BB2 <- corollas_BB %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("BB_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  mutate(Block = case_when(Block %in% c("W1","W2") ~ "W", #Fix blocks
                           TRUE ~ Block)) %>% 
  arrange(Block, Location, Recipient, Plant_ID, photo_date) %>% 
  distinct(Location, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates
#Only about 6 duplicate observations removed


## Merge with main data 
BB_merged_df6 <- merge(BB_merged_df5, corollas_BB2, by = c("Block", "Location", "Recipient", "Plant_ID"), all.x = TRUE)
  
```

<br>  


#### Leaf Area & Weight  
[[OLD - UPDATE]]

```{r eval = FALSE}
leaves_BB <- read_csv(here::here("data_sheets", "leaf_data", "BB_SLA_data_20230522.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) #exlcude NA leaf areas?



#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_BB_duped <- leaves_BB %>% 
  mutate(duped1 = duplicated(select(leaves_BB, c(Block, Location, Recipient))),
         duped2 = duplicated(select(leaves_BB, c(Block, Location, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!


#This seems to work the best!
leaves_BB_unique <- leaves_BB_duped %>% 
  group_by(Block, Location, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) %>% #Don't need plant ID...these were all non-fitness plants
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg,
         LMA = 1/SLA)


#Merge in to main data 

BB_merged_df7 <- merge(BB_merged_df6, leaves_BB_unique, by = c("Block", "Location", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo



```

<br>  


#### d13C data (WUE)
[[OLD - UPDATE]]

```{r eval = FALSE}
Nemo_d13C <- read_csv(here::here("data_sheets", "isotope_data", "Nemophila_d13C_230511.csv")) %>% 
  select(1,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error

#filter to BB samples
Nemo_d13C_BB <- Nemo_d13C %>% 
  filter(grepl('BB', sample_ID))

#Note! Two 205A samples from points 99 and 154 (rows 132, 199)

#unravel sample ID
Nemo_d13_BB2 <- Nemo_d13C_BB %>% 
  slice(-c(132, 199)) %>% 
  mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 6)), 
         Location = as.numeric(str_remove(substr(sample_ID, 8, 10), "^0+")),
         Plant_ID = as.numeric(substr(sample_ID, 12,13))) %>% 
  select(Location, Recipient, Plant_ID, d13C, sample_weight_mg, -Plant_ID)

#special cases - 205As
Nemo_205As <- Nemo_d13C_BB %>% 
  slice(132, 199) %>% 
  mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 7)), 
         Location = as.numeric(str_remove(substr(sample_ID, 9, 11), "^0+")),
         Plant_ID = as.numeric(substr(sample_ID, 13,14))) %>% 
  select(Location, Recipient, Plant_ID, d13C, sample_weight_mg, -Plant_ID)

Nemo_d13_BB2.5 <- rbind(Nemo_d13_BB2, Nemo_205As) %>% 
  mutate(Recipient = gsub('205A', '205', Recipient)) #Remove 'A' - is specified later...


#Merge
BB_merged_df_fl3 <- merge(BB_merged_df_fl2, Nemo_d13_BB2.5, by = c("Recipient", "Location"), all.x = TRUE)

#check that all data made it (203 obs)
testdf <- BB_merged_df_fl3 %>% 
  filter(!is.na(d13C))
```

<br>  


**Error checks**:

```{r eval = FALSE}
##Check for mismatched (fitness) plants

BB_Echeck1 <- BB_merged_df6 %>% 
  filter(any_FitP == FALSE & total_fruits_F != 0) #do the same for seed count, using `>` and `!=`

BB_Echeck2 <- BB_merged_df6 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
BB_Echeck3 <- BB_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))
######Loc. 16 ID BB_166 plant 5 shows up as NA for seed count??

BB_Echeck4 <- BB_merged_df6 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))

## Check for NA seed counts for fitness plants (when fruit production > 0)
BB_Echeck5 <- BB_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))


#Seed mass
BB_Echeck5 <- BB_merged_df6 %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


### Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd`  


```{r}
#Survival measures
BB_merged_df1 <- BB_merged_2 %>% 
  mutate(surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(F_plant == TRUE & total_fruits > 0 ~ 1, #plants that survived to produce fruit
                                       F_plant == TRUE & total_fruits == 0 ~ 0),
         surv_to_seedprod = case_when(F_plant == TRUE & ALL_filled_seeds > 0 ~ 1, #plants produced filled seeds
                                      F_plant == TRUE & ALL_filled_seeds == 0 ~ 0)) #else = 0 (applies to seed_ct = NA)

         
BB_merged_df2 <- BB_merged_df1 %>% 
  rename(total_est_seed_production = est_fecundity, mean_seeds_per_fruit = avg_seeds_per_fruit) %>% 
  mutate(across(c(total_est_seed_production, mean_seeds_per_fruit), ~ as.numeric(na_if(., "#DIV/0!")))) %>% #Converts to NA when 0 closed fruit were collected
  mutate(total_est_seed_production = case_when(F_plant == TRUE & is.na(total_est_seed_production) ~ 0,
                                               F_plant == TRUE & !is.na(total_est_seed_production) ~  total_est_seed_production)) %>% 
    mutate(mean_seeds_per_fruit = case_when(F_plant == TRUE & is.na(mean_seeds_per_fruit) ~ 0,
                                            F_plant == TRUE & !is.na(mean_seeds_per_fruit) ~  mean_seeds_per_fruit))#Set seeds per fruit & est_fecundity to 0 when NA, regardless of # closed fruits collected


```

<br>  




### Master sheets (2023)  

Full and summarized data sheets for all data collected in the 2022-2023 field season. 
These are merged with the year 1 data sheets elsewhere.  

```{r}

######
#TEMP VERSION
######

BB_merged_full_1 <- BB_merged_df2 %>% 
  filter(F_plant == TRUE)

write_csv(BB_merged_full_1, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BB_traits_Y2_incomplete_20230914.csv"), na = "")


######
######


## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...


BB_mastersheet_v1 <- BB_merged_df_fl3 %>% 
  mutate(Year = 2022, Gen = "G1") %>% 
  select(Year, Gen, Block, Location, Sequence, Co., Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Location, Sequence, Co., Recipient, Plant_ID) %>% #Correct Recipient names - two mat fams with different donors
  mutate(Recipient = case_when(Donor == "BB_217" & Location %in% c("33", "99", "154") ~ "BB_205_A",
                               Donor == "BB_188" & Location %in% c("48", "96", "126") ~ "BB_205_B",
                               TRUE ~ Recipient))

write_csv(BB_mastersheet_v1, here::here("data_sheets", "compiled_sheets", "BB_mastersheet_full_2023-08-07.csv"))

#Create meta data sheet



#
## Fitness Plants Only
#

BB_mastersheet_F_mains <-  BB_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only single fitness plants

BB_mastersheet_F_all <- BB_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treat these as 0 fitness or as NAs to remove???


write_csv(BB_mastersheet_F_mains, here::here("data_sheets", "compiled_sheets", "BB_mastersheet_Fitness-mains_2023-08-07.csv"))

write_csv(BB_mastersheet_F_all, here::here("data_sheets", "compiled_sheets", "BB_mastersheet_Fitness-all_2023-08-07.csv"))

```



