---
title: "compile_datasheets_BO"
author: "Devin Gamble"
date: "2022-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



## Load and Clean Data   

NOTE: All date data should be standardized prior to importing into R. In Excel, set format to YYYY-MM-DD for consistency. Unfortunately, this process in Excel will auto-populate the year (when not already present) based on system settings. For germination data, this means that the year should be corrected to 2021!!


NOTE 2: All paths are relative thanks to the 'here' package.  


**Germination Data**  
- Add 'Date Sown' column and data in as well.


```{r message = FALSE}
#Load csv

#Skip first nonsense row (the second row is the header)
BO_germ_df1 <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "BO_germ_data.csv"), skip = 1) %>%
  rename(Block = 'Block (1-3)', Transect = 'Transect (1-12)', Sequence = 'Sequence (1-396)', Donor = "Pollen Donor ID", Recipient = "Maternal Sibship", Stage = 'mm/dd') %>% 
  filter(!Stage == "Flow Date") %>%   #Remove empty flowering date rows
  select(-c(`9`:`Notes 12/28-29`)) %>% 
  mutate(`4` = as_date(`4`),
         `5` = as_date(`5`)) #set as date to allow for pivot_longer 

#
#Reshape data sheet to 'long' format
BO_germ_df2 <- BO_germ_df1 %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'Germ_Date') 


#
##Fix typos in date values
##MAJOR TYPO: Excel auto-populated year as 2022 when in fact the year should be 2021!
unique(BO_germ_df2$Germ_Date)

BO_germ_df3 <- BO_germ_df2 %>% 
  mutate(
    Germ_Date = case_when(
    Germ_Date == "2022-12-06" ~ "2021-12-06",
    Germ_Date == "2022-12-15" ~ "2021-12-15",
    Germ_Date == "2022-12-21" ~ "2021-12-21",
    Germ_Date == "2022-12-28" ~ "2021-12-28",
    Germ_Date == "2022-12-29" ~ "2021-12-29"))

#check again
unique(BO_germ_df3$Germ_Date)
#Use sum(is.na(BO_germ_df2$Germ_Date)) to compare change in number of NA values


#
#Adjust Donor and Recipient names for easy sorting - three digits (leading zeros)
#Specify Germ_Date as 'date' class variable
BO_germ_df4 <- BO_germ_df3 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("BO_", Donor_num))) %>%
  mutate(Germ_Date = as_date(Germ_Date)) %>% #set variable to class = date
  select(Block, Transect, Sequence, Donor, Donor_num, Recipient, Recipient_num, Plant_ID, everything()) %>% 
  select(-c(Stage, Donor_num, Recipient_num))


#
##Add in Sow Date  
BO_germ_df5 <- BO_germ_df4 %>% 
  mutate(Sow_Date = as_date("2021-11-11")) %>% #lab sowing date 11/11 - 11/12
  mutate(Plant_Date = as_date("2021-11-17")) %>% #field planting date
  arrange(Block, Transect, Sequence, Donor, Recipient, Plant_ID) #Arrange by Transect, Recipient, Plant_ID
#Plugs planted in ground from November 17-18 ,2021


```


Add in watered transect germination data
```{r message = FALSE}

## Add in watered transects

BO_germ_W1 <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "watered_BO_germ.csv"), skip = 1) %>%
  rename(Sequence = 'Seq (1-42)', Donor = "Donor_ID", Recipient = "Recipient_ID", Stage = 'Date (mm/dd)') %>% 
  filter(!Stage == "Flow") %>%  #Remove empty flowering date rows
  select(-c(Co., `Notes - 12/6`:`Notes 12/28-29`))

  
#
#Reshape data sheet to 'long' format
BO_germ_W2 <- BO_germ_W1 %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'Germ_Date') %>% 
  mutate(Sequence = (Sequence + 396)) #Sequence continues following main transects

#
##Fix typos in date values
##MAJOR TYPO: Excel auto-populated year as 2022 when in fact the year should be 2021!
unique(BO_germ_W2$Germ_Date)

BO_germ_W3 <- BO_germ_W2 %>% 
  mutate(Germ_Date = case_when(
    Germ_Date == "2022-12-06" ~ "2021-12-06",
    Germ_Date == "2022-12-15" ~ "2021-12-15",
    Germ_Date == "2022-12-21" ~ "2021-12-21",
    Germ_Date == "2022-12-30" ~ "2021-12-30",
    ))

#
#Adjust Donor and Recipient names for easy sorting - three digits (leading zeros)
#Specify Germ_Date as 'date' class variable
BO_germ_W4 <- BO_germ_W3 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("BO_", Donor_num))) %>%
  mutate(Germ_Date = as_date(Germ_Date)) %>% #set variable to class = date
  mutate(Plant_ID = as.numeric(Plant_ID)) %>% #set Plant ID to numeric for sorting purposes
  select(-c(Stage, Donor_num, Recipient_num))


#
##Add in Sow Date & add block ('W')
BO_germ_W5 <- BO_germ_W4 %>% 
  mutate(Sow_Date = as_date("2021-11-11")) %>% #lab sowing date 11/11 - 11/12
  mutate(Plant_Date = as_date("2021-11-17")) %>% #field planting date
  mutate(Block = "W") %>% 
  select(Block, Transect, Donor, Recipient, Plant_ID, everything()) %>% 
  arrange(Transect, Recipient, Plant_ID) #Arrange by Transect, Recipient, Plant_ID


#
## Combine watered and main experiment germination data

BO_germ_full <- rbind(BO_germ_df5, BO_germ_W5) %>% 
    mutate(Plant_ID = as.numeric(Plant_ID))

write_csv(BO_germ_full, here::here("data_sheets", "compiled_sheets", "BO_germination_full_2021-2022.csv"))

```



**Flower Check sheets**  
-- Organize flowering data
-- Do not trust the fitness plant data in the flowering check data sheets (add in later).  

*Important Step!!*  
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R much easier!  



```{r}

BOflr_df <- read_csv(file = here::here( "data_sheets", "season_end_flowerchecks_2022",  "BO_flowercheck_2022_05_15.csv"), na = c("", "NA", "G", "G?")) %>% 
  mutate(across(c(5:19, 23:37), as_date)) #specify date columns
##Ignore 'F' plants - more up-to-date fitness plant data is matched in below 

#split groups
f1_BOlong <- BOflr_df[c(1:19)] # Group 1: plants 1-4
f2_BOlong <- BOflr_df[c(1, 20:37)] # Group 2: plants 5-8

#Revise names
colnames(f1_BOlong) <- c("Transect", "Recipient", "Plant_ID", "F.", "Fl_3.11", "Fl_3.17", "Fl_3.25", "Fl_4.1", "Fl_4.8", "Fl_4.13", "Fl_4.21", "Fl_4.28", "Fl_5.5", "Fl_5.12", "Fl_5.20", "Fl_5.25", "RMV", "LEAF", "flr_P")
colnames(f2_BOlong) <- c("Transect", "Recipient", "Plant_ID", "F.",  "Fl_3.11", "Fl_3.17", "Fl_3.25", "Fl_4.1", "Fl_4.8", "Fl_4.13", "Fl_4.21", "Fl_4.28", "Fl_5.5", "Fl_5.12", "Fl_5.20", "Fl_5.25", "RMV", "LEAF", "flr_P")

#rbind, tidy up and remove unnecessary columns
f_BOlong <- rbind(f1_BOlong, f2_BOlong) %>% arrange(Transect, Recipient, Plant_ID) %>% select(-c(F., RMV)) 


#Format recipient IDs for easier sorting - three digits (includes leading zeros)
f_BOlong_2 <- f_BOlong %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  select(Transect, Recipient, Recipient_num, Plant_ID, everything()) %>% 
  select(-Recipient_num)


#### Fix Dates

#Individually check each column (e.g., Fl_3.3) for the correct date format (unique(df$col)). Include LEAF and flr_p. Multiple dates are okay.

f_BOlong_3 <- f_BOlong_2 %>% 
  mutate(Fl_4.21 = case_when(Fl_4.13 == "2022-04-21" ~ "2022-04-21",
                             Fl_4.21 == "2022-04-21" ~ "2022-04-21",
                             Fl_4.21 == "2022-04-20" ~ "2022-04-20",
                             TRUE ~ as.character(NA))) %>% 
  mutate(Fl_4.13 = case_when(Fl_4.13 == "2022-04-13" ~ "2022-04-13",
                             TRUE ~ as.character(NA))) %>% 
  mutate(Fl_5.25 = case_when(Fl_5.25 == "2022-05-25" ~ "2022-05-25",
                             Fl_5.25 == "2022-05-26" ~ "2022-05-26",
                             TRUE ~ as.character(NA)))

```


Combine Germination and Flowering data frames
```{r}
#Merge by Recipient, Transect, and Plant ID
BO_GermFlow_1 <- left_join(BO_germ_full, f_BOlong_3, by = c("Transect", "Recipient", "Plant_ID")) #

```


<br>  


#### Fitness Plant data  
- Indicate which plants were fitness plants
- Replicated? Indicate which maternal families were present in the main and watering experiment  

```{r message = FALSE}
BO_Fplants <- read_csv(here::here("data_sheets", "fitness_plants_7-21-2022", "BO_F_plants.csv"), col_types = cols(`ALL Fitness Plants` = col_character())) %>% #retain commas for column with multiple fitness plants
  rename(F_plant = `SINGLE Fitness Plant`, F_multi = `ALL Fitness Plants`, F_plant_notes = NOTES) %>% 
  select(Transect:F_plant_notes)

#Adjust Recipient names to match three-digit format
BO_Fplants2 <- BO_Fplants %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  rename(Replicated = `Replicated?`) %>% 
  mutate(Replicated = case_when(Replicated == "R" ~ "R",
                                is.na(Replicated) ~ "n")) #change NAs to 'n' (no) for non-rep'd Recipient IDs

```


Index fitness plants & Replicated (y/n) in germination and flowering data
```{r message = FALSE}
# Merge the two dataframes 
# Create new row (Plant ID) specific variables for main and replicated fitness plants
BO_GermFlow_2 <- merge(BO_GermFlow_1, BO_Fplants2) %>% 
  mutate(FitP = case_when(F_plant == Plant_ID ~ TRUE,
                           F_plant != Plant_ID ~ FALSE)) %>% 
  mutate(Rep_FitP = case_when(Replicated == "R" & Plant_ID == as.numeric(str_match(F_multi, as.character(Plant_ID))) ~ TRUE,   
                              TRUE ~ FALSE)) %>%  #if else, set value to FALSE
  mutate(any_FitP = case_when(FitP == TRUE | Rep_FitP == TRUE ~ TRUE, #plant is any form of fitness plant
                              TRUE ~ FALSE))


#Tidy up
BO_GermFlow_3 <- BO_GermFlow_2 %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, everything()) %>%
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

## Fit_P has NA values for when a fitness plant was never chosen

```



**Fruit and Seed counts**  

Due to the structure of excel data sheets, it seems easiest to upload fruit and seed count data separately since different spreadsheets were used to record each. Info on plants' Block, Transect, sequence, etc. will be pulled from the fruit count sheets (assumed to be up-to-date). For sheets intended to record fruit production only, counts of viable seeds, mean seed weight, etc. that are copied over are ignored and instead pulled straight from sheets where they were first entered later on. 

Data on fitness plants will be more accurate in the fruit-cound data sheet, while the seed-count sheet will have a handful of rows with non-fitness plants wihtout seed counts.

Date added: **Feb 06, 2023**  

*NOTE*: Update with later versions of data sheet when more seed counts have been entered
```{r message = FALSE}
#
# Fruit count data
# Data on *when* fruits were collected will be omitted here; only the total fruit counts (closed, open, non-fitness) are retained from the csv

BO_fruitprod_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "BO-fruit-production_2023-02-15.csv"), skip = 1) %>% 
  select(1:10) #%>% #omit sow and flower date, seed count, and date-specific fruit counts by column index
  #filter_all(any_vars(!is.na(.))) #omit empty rows at end of df

BO_fruitprod_2 <- BO_fruitprod_1 %>% 
  select(-c(Recipient, Sequence)) %>% # Sequence in this data set is incorrect
  rename(Recipient = `Recipient for sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, closed_fruits_F = `total_closed_fruits`, open_fruits_F = `total_open_fruits`, damaged_fruits_F = `eaten_damaged_fruits`, total_fruits_F = `total_fruits (open and closed)`, nonfitness_fruits = `nonfitness_fruits`) 

#Note: total fruits should not include eaten/damaged fruit



#
# Seed count data (viable seeds)
# Similarly, only seed count data are retained from this spreadsheet to minimize errors

BO_seedct_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "BO-seed-count_2023-02-15.csv"), skip = 1) %>% 
  select(-c(1, 3, 7:33)) %>% 
  rename(Recipient = `Recipient for sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, seed_ct = `viable_seeds`) %>% 
  mutate(Plant_ID = as.numeric(Plant_ID)) 
  

#merge data
BO_fruitseed_ct1 <- left_join(BO_fruitprod_2, BO_seedct_1, by = c("Transect", "Recipient", "Plant_ID"))
#Sequence data in fruit count sheets is unreliable for these data sheets
#Rows without matches in fruit_prod data sheet are not included


# summary variables, tidying up
BO_fruitseed_ct2 <- BO_fruitseed_ct1 %>% 
  select(Transect, Recipient, Plant_ID, closed_fruits_F, open_fruits_F, damaged_fruits_F, total_fruits_F, nonfitness_fruits, everything()) %>% 
  arrange(Transect, Recipient)

## NOTE: total number of non-fitness fruits per segment may be under-represented here since not all counts were matched to fitness plants and therefore merged in.

```



**Seed Weights**  



```{r}
# Read in only # seeds weighed and total weight
# Calculate mean weight per seed

BO_seedmass_1 <- read_csv(here::here("data_sheets", "seed_mass", "BO_Seed_Mass_2023-02-05.csv"), skip = 1) %>% 
  select(-Donor) %>% slice(1:339)
  

colnames(BO_seedmass_1) <- c("Transect", "Sequence", "Block", "Recipient", "Plant_ID", "closed_fruits_d1", "viable_seeds_d1", "seeds_weighed_d1", "seed_mass_mg_d1", "closed_fruits_d2", "viable_seeds_d2", "seeds_weighed_d2", "seed_mass_mg_d2", "closed_fruits_d3", "viable_seeds_d3", "seeds_weighed_d3", "seed_mass_mg_d3", "closed_fruits_d4", "viable_seeds_d4", "seeds_weighed_d4", "seed_mass_mg_d4", "closed_fruits_d5", "viable_seeds_d5", "seeds_weighed_d5", "seed_mass_mg_d5", "closed_fruits_d6", "viable_seeds_d6", "seeds_weighed_d6", "seed_mass_mg_d6", 
                             "closed_fruits_d7", "viable_seeds_d7", "seeds_weighed_d7", "seed_mass_mg_d7",
                             "closed_fruits_d8", "viable_seeds_d8", "seeds_weighed_d8", "seed_mass_mg_d8", 
                             "closed_fruits_d9", "viable_seeds_d9", "seeds_weighed_d9", "seed_mass_mg_d9")


BO_seedmass_2 <- BO_seedmass_1 %>% 
  select(-c(42:48)) %>% 
  select(-starts_with(c("closed_fruits_", "viable_seeds_"))) %>% 
  mutate(msm_d1 = seed_mass_mg_d1/seeds_weighed_d1,
         msm_d2 = seed_mass_mg_d2/seeds_weighed_d2,
         msm_d3 = seed_mass_mg_d3/seeds_weighed_d3,
         msm_d4 = seed_mass_mg_d4/seeds_weighed_d4,
         msm_d5 = seed_mass_mg_d5/seeds_weighed_d5,
         msm_d6 = seed_mass_mg_d6/seeds_weighed_d6,
         msm_d7 = seed_mass_mg_d7/seeds_weighed_d7,
         msm_d8 = seed_mass_mg_d8/seeds_weighed_d8,
         msm_d9 = seed_mass_mg_d9/seeds_weighed_d9)
         

BO_seedmass_3 <- BO_seedmass_2 %>%
  mutate(msm_all = rowMeans(select(BO_seedmass_2, starts_with("msm_")), na.rm = TRUE)) %>% 
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(msm_d1:msm_d9)), na.rm = TRUE)) %>% 
  select(-c(Block, Sequence))
         
  
##Merge with fruit/seed count data frame

BO_fruitseed_ct3 <- merge(BO_fruitseed_ct2, BO_seedmass_3, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```



Merge everything together:  
```{r}
BO_merged_df1 <- merge(BO_GermFlow_3, BO_fruitseed_ct3, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, everything()) %>%
  arrange(Block, Transect, Sequence, Recipient, Plant_ID) %>% 
  mutate(closed_fruits_F = as.numeric(closed_fruits_F),
         total_fruits_F = as.numeric(total_fruits_F)) #correct counts to numeric values


#Correct FruitNAs to 0s in cases where no fruit data was entered (fitness and non-fitness plants):

#
BO_merged_df2 <- BO_merged_df1 %>% 
  mutate(closed_fruits_F = replace_na(closed_fruits_F, 0),
         total_fruits_F = replace_na(total_fruits_F, 0),
         seed_ct = replace_na(seed_ct, 0))

#Calculate mean seeds per fruit, total_est_seed prod
BO_merged_df3 <- BO_merged_df2 %>% 
  mutate(mean_seeds_per_fruit = seed_ct/closed_fruits_F,
         total_est_seed_production = mean_seeds_per_fruit*total_fruits_F) %>% 
  mutate(mean_seeds_per_fruit = case_when(any_FitP == TRUE & closed_fruits_F == 0 ~ 0, 
                                          TRUE ~ mean_seeds_per_fruit)) %>%  #else, original value
  mutate(total_est_seed_production = case_when(any_FitP == TRUE & mean_seeds_per_fruit == 0 ~ 0,
                                               TRUE ~ total_est_seed_production))


#mean seeds per fruit and total est fruit production have NAs/NaNs for when there were no closed fruits/seeds (fitness plants only)



#Correct formatting issues

BO_merged_df4 <- BO_merged_df3 %>% 
  mutate(Fl_4.13 = as_date(Fl_4.13),
         Fl_4.21 = as_date(Fl_4.21),
         Fl_5.25 = as_date(Fl_5.25))


```


**Skeleton Weights & Corolla area and diameter**  

Load and merge-in these data sheets
Skeletons:
```{r}
##
skel_weights_BO <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeleton_weights_BO_2023_03_01.csv")) %>% select(-Units, -7) %>% 
  filter(!grepl("EXCLUDE.", Notes)) %>% ## Filter out notes with "Exclude"
  rename(AG_biomass_mg = `Skeleton Weights`)

#Watered points
skel_weights_BO_W <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeletons_watered_BO_2023_02_08.csv")) %>% 
  select(-`Units`) %>% 
  rename(Plant_ID = `Plant ID`, AG_biomass_mg = `Skeleton Weights`)
#3 observations :(

#Combine
skel_weights_all <- rbind(skel_weights_BO, skel_weights_BO_W) %>% 
  select(-Notes) %>% 
  filter(!is.na(AG_biomass_mg))


#Merge with full data set
BO_merged_df5 <- merge(BO_merged_df4, skel_weights_all, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)

```


Corolla Area & Diameter:
```{r}

corollas_BO <- read_csv(here::here("data_sheets", "corolla_measures", "corollameasures_BO_2023-02-28.csv")) %>% select(1, 3, 4, 6, 8, 9, 10) %>% 
  rename(Plant_ID = 3, photo_date = 4, corolla_diam_mm = 6, corolla_area_mm2 = 5, notes_corolla = 7) %>% 
  filter(!grepl("REMOVE", notes_corolla)) ## Filter out notes with "REMOVE"

corollas_BO2 <- corollas_BO %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  arrange(Transect, Recipient, Plant_ID, photo_date) %>% 
  distinct(Transect, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates


## Merge with main data 
BO_merged_df6 <- merge(BO_merged_df5, corollas_BO2, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)

```
**Leaf Area & Weight**  
```{r}
leaves_BO <- read_csv(here::here("data_sheets", "leaf_data", "BO_SLA_data_20230508.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) #exlcude NA leaf areas?

#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_BO_duped <- leaves_BO %>% 
  mutate(duped1 = duplicated(select(leaves_BO, c(Block, Transect, Recipient))),
         duped2 = duplicated(select(leaves_BO, c(Block, Transect, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!

#This seems to work the best!
leaves_BO_unique <- leaves_BO_duped %>% 
  group_by(Block, Transect, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) #Don't need plant ID...these were all non-fitness plants


#Merge in to main data 

BO_merged_df7 <- merge(BO_merged_df6, leaves_BO_unique, by = c("Block", "Transect", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo


```

Error checks:

```{r eval = FALSE}
##Check for mismatched (fitness) plants

BO_Echeck1 <- BO_merged_df6 %>% 
  filter(any_FitP == FALSE & total_fruits_F != 0) #do the same for seed count, using `>` and `!=`

BO_Echeck2 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
BO_Echeck3 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))


BO_Echeck4 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))

## Check for NA seed counts for fitness plants (when fruit production > 0)
BO_Echeck5 <- BO_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))


#Seed mass
BO_Echeck5 <- BO_merged_df6 %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


## Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd`  


#### Phenology  

```{r}

BO_merged_df_fl <- BO_merged_df7 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('Fl_')), na.rm = TRUE)), #First flowering date
         days_germ2flower = as.numeric(FFD - Germ_Date), #days from germination to flowering
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('Fl_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) #Flowering duration


#Survival measures
BO_merged_df_fl2 <- BO_merged_df_fl %>% 
  mutate(surv_to_germ = case_when(!is.na(Germ_Date) ~ 1, #plants that germinated
                                  is.na(Germ_Date) ~ 0),
         surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(total_fruits_F > 0 ~ 1, #plants that survived to produce fruit
                                       total_fruits_F == 0 ~ 0),
         surv_to_seedprod = case_when(seed_ct > 0 ~ 1, #plants that survived to produce fruit with viable seeds
                                       TRUE ~ 0)) #else = 0 (applies to seed_ct = NA)

         


## Note: should flowering duration remain unchanged when a plant went several weeks without flowering (e.g., had two temporally distinct blooms during the growing season)?
```

<br>  

## Master sheets (v1)  


```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...


BO_mastersheet_v1 <- BO_merged_df_fl2 %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

write_csv(BO_mastersheet_v1, here::here("data_sheets", "compiled_sheets", "BO_mastersheet_full_2023-03-01.csv"))

#Create meta data sheet



#
## Fitness Plants Only
#

BO_mastersheet_F_mains <-  BO_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only single fitness plants

BO_mastersheet_F_all <- BO_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treate these as 0 fitness or as NAs to remove???


write_csv(BO_mastersheet_F_mains, here::here("data_sheets", "compiled_sheets", "BO_mastersheet_Fitness-mains_2023-05-08.csv"))

write_csv(BO_mastersheet_F_all, here::here("data_sheets", "compiled_sheets", "BO_mastersheet_Fitness-all_2023-05-08.csv"))

```



