---
title: "compile_datasheets_BB"
author: "Devin Gamble"
date: "2022-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



## Load and Clean Data   

NOTE: All date data should be standardized prior to importing into R. In Excel, set format to YYYY-MM-DD for consistency. Unfortunately, this process in Excel will auto-populate the year (when not already present) based on system settings. For germination data, this means that the year should be corrected to 2021!!


NOTE 2: All paths are relative thanks to the 'here' package.  


NOTE 3: Two iterations of a maternal family ID (names fixed before saving csv):
**BB_205_A** occurs at locations 33, 99, and 154 and has Donor BB_217 (greenhouse cohort 5)
**BB_205_B** occurs at locations 48, 96, and 126 and has Donor BB_188 (cohort 6)


**Germination Data**  
- Add 'Date Sown' column and data in as well.


```{r message = FALSE}
#Load csv

#Skip first nonsense row (the second row is the header)
BB_germ_df1 <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "BB_germ_data.csv"), skip = 1) %>%
  rename(Block = 'Block (1-3)', Location = 'Location (1-156)', Sequence = 'Seq (1-441)', Donor = "Donor_ID", Recipient = "Recipient_ID", Stage = 'mm/dd') %>% 
  filter(!Stage == "Flow Date")   #Remove empty flowering date rows

#
#Reshape data sheet to 'long' format
BB_germ_df2 <- BB_germ_df1 %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'Germ_Date') 


#
##Fix typos in date values
##MAJOR TYPO: Excel auto-populated year as 2022 when in fact the year should be 2021!
unique(BB_germ_df2$Germ_Date)

BB_germ_df3 <- BB_germ_df2 %>% 
  mutate(Germ_Date = replace(Germ_Date, Germ_Date %in% c("??"), NA)) %>% #specify values to assign 'NA'
  mutate(
    Germ_Date = case_when(
    Germ_Date %in% c("12/12/2022", "12/12/2021") ~ "2021-12-12",
    Germ_Date %in% c("12/13/2022", "12/13/2021 ?") ~ "2021-12-13",
    Germ_Date == "12/19/2022" ~ "2021-12-19",
    Germ_Date == "12/20/2022" ~ "2021-12-20",
    Germ_Date == "12/30/2022" ~ "2021-12-30"))

#check again
unique(BB_germ_df3$Germ_Date)
#Use sum(is.na(BB_germ_df2$Germ_Date)) to compare change in number of NA values


#
#Adjust Donor and Recipient names for easy sorting - three digits (leading zeros)
#Specify Germ_Date as 'date' class variable
BB_germ_df4 <- BB_germ_df3 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BB_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("BB_", Donor_num))) %>%
  mutate(Germ_Date = as_date(Germ_Date)) %>% #set variable to class = date
  select(Block, Location, Sequence, Donor, Donor_num, Recipient, Recipient_num, Plant_ID, everything()) %>% 
  select(-c(Stage, Donor_num, Recipient_num, ...16:`Notes 12/30`))


#
##Add in Sow Date  
BB_germ_df5 <- BB_germ_df4 %>% 
  mutate(Sow_Date = as_date("2021-11-29")) %>% #actual lab-sow date
  mutate(Plant_Date = as_date("2021-12-02")) %>% #actual field planting date
  arrange(Block, Location, Sequence, Donor, Recipient, Plant_ID) #Arrange by Transect, Recipient, Plant_ID
#Plugs planted in ground from December 2-4,2021, but all seeds were sown on the same day (or two) in the lab


```


Add in watered transect germination data
```{r message = FALSE}

## Add in watered transects

BB_germ_W1 <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "watered_BB_germ.csv"), skip = 1) %>%
  rename(Sequence = 'Seq (1-42)', Donor = "Donor_ID", Recipient = "Recipient_ID", Stage = 'Date (mm/dd)') %>% 
  filter(!Stage == "Flow") #Remove empty flowering date rows

  
#
#Reshape data sheet to 'long' format
BB_germ_W2 <- BB_germ_W1 %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'Germ_Date') 

#
##Fix typos in date values
##MAJOR TYPO: Excel auto-populated year as 2022 when in fact the year should be 2021!
unique(BB_germ_W2$Germ_Date) #no typos here

BB_germ_W3 <- BB_germ_W2 %>% 
  mutate(Germ_Date = case_when(
    Germ_Date == "2022-12-12" ~ "2021-12-12",
    Germ_Date == "2022-12-13" ~ "2021-12-13",
    Germ_Date == "2022-12-20" ~ "2021-12-20",
    Germ_Date == "2022-12-30" ~ "2021-12-30",
    ))


#
#Adjust Donor and Recipient names for easy sorting - three digits (leading zeros)
#Specify Germ_Date as 'date' class variable
BB_germ_W4 <- BB_germ_W3 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BB_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("BB_", Donor_num))) %>%
  mutate(Germ_Date = as_date(Germ_Date)) %>% #set variable to class = date
  mutate(Plant_ID = as.numeric(Plant_ID)) %>% #set Plant ID to numeric for sorting purposes
  select(-c(Stage, Donor_num, Recipient_num, `Notes 12/12`:`Notes 12/30`))


#
##Add in Sow Date & Update Sequence number & add block ('W')
BB_germ_W5 <- BB_germ_W4 %>% 
  mutate(Sow_Date = as_date("2021-11-29")) %>% #actual lab-sow date
  mutate(Plant_Date = as_date("2021-12-02")) %>% #actual field planting date
  mutate(Sequence = NA) %>% # Sequence in this data will not align with Sequence in the main points
  mutate(Block = "W") %>% 
  select(Block, Location, Co., Donor, Recipient, Plant_ID, everything()) %>% 
  arrange(Location, Recipient, Plant_ID) #Arrange by Transect, Recipient, Plant_ID


#
## Combine watered and main experiment germination data

BB_germ_full <- rbind(BB_germ_df5, BB_germ_W5) %>% 
    mutate(Plant_ID = as.numeric(Plant_ID))

#write_csv(BB_germ_full, here::here("data_sheets", "compiled_sheets", "BB_germination_full_2021-2022.csv"))

```



**Flower Check sheets**  
-- Organize flowering data
-- Do not trust the fitness plant data in the flowering check data sheets (add in later).  

*Important Step!!*  
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R much easier!  



```{r}

BBflr_df <- read_csv(file = here::here( "data_sheets", "season_end_flowerchecks_2022",  "BB_flowercheck_2022_05_30.csv"), na = c("", "NA", "missing", "buried"))
##Ignore 'F' plants - more up-to-date fitness plant data is matched in below 

#split groups
f1_BBlong <- BBflr_df[c(1:20)] # Group 1: plants 1-4
f2_BBlong <- BBflr_df[c(1:2, 21:38)] # Group 2: plants 5-8

#Revise names
colnames(f1_BBlong) <- c("Block", "Location", "Recipient", "Plant_ID", "F.", "Fl_2.25", "Fl_3.3", "Fl_3.11", "Fl_3.17", "Fl_3.25", "Fl_4.1", "Fl_4.9", "Fl_4.22", "Fl_4.28", "Fl_5.4", "Fl_5.12", "Fl_5.19", "Fl_5.25", "LEAF", "flr_P")
colnames(f2_BBlong) <- c("Block", "Location", "Recipient", "Plant_ID", "F.", "Fl_2.25", "Fl_3.3", "Fl_3.11", "Fl_3.17", "Fl_3.25", "Fl_4.1", "Fl_4.9", "Fl_4.22", "Fl_4.28", "Fl_5.4", "Fl_5.12", "Fl_5.19", "Fl_5.25", "LEAF", "flr_P")

#rbind, tidy up and remove unnecessary columns
f_BBlong <- rbind(f1_BBlong, f2_BBlong) %>% arrange(Block, Location, Recipient, Plant_ID) %>% select(-F.)


#Format recipient IDs for easier sorting - three digits (includes leading zeros)
f_BBlong_2 <- f_BBlong %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BB_", Recipient_num))) %>% 
  select(Block, Location, Recipient, Recipient_num, Plant_ID, everything()) %>% 
  select(-Recipient_num)


#### Fix Dates

#Individually check each column (e.g., Fl_3.3) for the correct date format (unique(df$col)). Multiple dates are okay.

f_BBlong_3 <- f_BBlong_2 %>% 
  mutate(Fl_2.25 = case_when(!is.na(Fl_2.25) ~ "2022-02-25")) %>% 
  mutate(Fl_3.25 = case_when(Fl_3.25 == "2022-03-25" ~ "2022-03-25",
                             Fl_3.25 == "2022-03-26" ~ "2022-03-26",
                             TRUE ~ as.character(NA))) %>% 
  mutate(Fl_4.1 = case_when(Fl_4.1 %in% c("2022-04-01", "19083") ~ "2022-04-01",
                            Fl_4.1 %in% c("2022-04-02", "19084") ~ "2022-04-02",
                            TRUE ~ as.character(NA))) %>% 
  mutate(Fl_4.22 = case_when(Fl_4.22 %in% c("2022-04-22", "19104") ~ "2022-04-22",
                            Fl_4.22 %in% c("2022-04-23", "19105") ~ "2022-04-23",
                            TRUE ~ as.character(NA))) %>% 
  mutate(Fl_5.4 = case_when(Fl_5.4 == 19116 ~ "2022-05-04",
                            TRUE ~ as.character(NA))) %>%  #only one plant checked on this date
  mutate(Fl_5.12 = case_when(Fl_5.12 %in% c("2022-05-12", "19124") ~ "2022-05-12",
                             TRUE ~ as.character(NA))) %>% 
  mutate(Block = case_when(Location %in% c("30", "40", "50", "60", "70", "80.5", "90", "100", "110", "120", "130", "140", "150", "160") ~ "W",
                            TRUE ~ Block)) #Rename blocks so they match across .csvs


```


Combine Germination and Flowering data frames
```{r}
#Merge by Recipient, Transect, and Plant ID
BB_GermFlow_1 <- left_join(BB_germ_full, f_BBlong_3, by = c("Block", "Location", "Recipient", "Plant_ID")) #

```



<br>  


#### Fitness Plant data  
- Indicate which plants were fitness plants
- Replicated? Indicate which maternal families were present in the main and watering experiment  

```{r message = FALSE}
BB_Fplants <- read_csv(here::here("data_sheets", "fitness_plants_7-21-2022", "BB_F_plants.csv"), col_types = cols(`ALL Fitness Plants` = col_character())) %>% #retain commas for column with multiple fitness plants
  rename(F_plant = `SINGLE Fitness Plant`, F_multi = `ALL Fitness Plants`, F_plant_notes = NOTES) %>% 
  select(Block:F_plant_notes)

#Adjust Recipient names to match three-digit format
BB_Fplants2 <- BB_Fplants %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BB_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  rename(Replicated = `Replicated?`) %>% 
  mutate(Replicated = case_when(Replicated == "R" ~ "R",
                                is.na(Replicated) ~ "n")) %>% #change NAs to 'n' (no) for non-rep'd Recipient IDs
  mutate(Block = case_when(Location %in% c("30", "40", "50", "60", "70", "80.5", "90", "100", "110", "120", "130", "140", "150", "160") ~ "W",
                            TRUE ~ Block)) #Rename blocks so they match across .csvs
```


Index fitness plants & Replicated (y/n) in germination and flowering data
```{r message = FALSE}
# Merge the two dataframes 
# Create new row (Plant ID) specific variables for main and replicated fitness plants
BB_GermFlow_2 <- merge(BB_GermFlow_1, BB_Fplants2) %>% 
  mutate(FitP = case_when(F_plant == Plant_ID ~ TRUE,
                           F_plant != Plant_ID ~ FALSE)) %>% 
  mutate(Rep_FitP = case_when(Replicated == "R" & Plant_ID == as.numeric(str_match(F_multi, as.character(Plant_ID))) ~ TRUE,   
                              TRUE ~ FALSE)) %>%  #if else, set value to FALSE
  mutate(any_FitP = case_when(FitP == TRUE | Rep_FitP == TRUE ~ TRUE, #plant is any form of fitness plant
                              TRUE ~ FALSE))

#Note - added as.numeric() around the str_match calls above since an R update caused the previous version of this code to not work.

#Tidy up
BB_GermFlow_3 <- BB_GermFlow_2 %>% 
  select(Block, Location, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, everything()) %>%
  arrange(Location, Sequence, Recipient, Plant_ID)

## Fit_P has NA values for when a fitness plant was never chosen

```



**Fruit and Seed counts**  

Due to the structure of excel data sheets, it seems easiest to upload fruit and seed count data separately since different spreadsheets were used to record each. Info on plants' Block, Transect, sequence, etc. will be pulled from the fruit count sheets (assumed to be up-to-date). For sheets intended to record fruit production only, counts of viable seeds, mean seed weight, etc. that are copied over are ignored and instead pulled straight from sheets where they were first entered later on. 

Data on fitness plants will be more accurate in the fruit-cound data sheet, while the seed-count sheet will have a handful of rows with non-fitness plants wihtout seed counts.

Date added: **Feb 06, 2023** 

*NOTE*: Update with later versions of data sheet when more seed counts have been entered
```{r message = FALSE}
#
# Fruit count data
# Data on *when* fruits were collected will be omitted here; only the total fruit counts (closed, open, non-fitness) are retained from the csv

BB_fruitprod_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "BB-fruit-production_2023-02-05.csv"), skip = 1) %>% 
  select(1:9, 11, 12) %>% #omit sow and flower date, seed count, and date-specific fruit counts by column index
  filter_all(any_vars(!is.na(.))) #omit empty rows at end of df

BB_fruitprod_2 <- BB_fruitprod_1 %>% 
  select(-c(Recipient, Sequence)) %>% # Sequence in this data set is incorrect
  rename(Recipient = `Recipient for sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, closed_fruits_F = `Number of closed fruits used to estimate mean seeds per fruit (include multiple dates)`, open_fruits_F = `Total number of all Open fruits`, damaged_fruits_F = `Eaten or damaged fruits`, total_fruits_F = `Total number of all fruits produced by the plant (open + closed)`, nonfitness_fruits = `Total number of non-fitness fruits collected from segment`) 

#Note: total fruits should not include eaten/damaged fruit


#
# Seed count data (viable seeds)
# Similarly, only seed count data are retained from this spreadsheet to minimize errors

BB_seedct_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "BB-seed-count_2023-02-05.csv"), skip = 1) %>% 
  select(-c(1, 4, 8:47)) %>% 
  rename(Recipient = `Recipient for sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, seed_ct = `Total number of viable seeds counted across multiple dates...7`) %>% 
  mutate(Plant_ID = as.numeric(Plant_ID)) %>% 
  filter_all(any_vars(!is.na(.)))

##NOTE##: In this sheet, plant numbers were only double checked for those that had any closed fruit and therefore a count (inc. 0s) of seed number


#merge data
BB_fruitseed_ct1 <- left_join(BB_fruitprod_2, BB_seedct_1, by = c("Block", "Location", "Recipient", "Plant_ID"))
#Sequence data in fruit count sheets is unreliable for these data sheets
#Rows without matches in fruit_prod data sheet are not included


# summary variables, tidying up
BB_fruitseed_ct2 <- BB_fruitseed_ct1 %>% 
  select(Block, Location, Recipient, Plant_ID, closed_fruits_F, open_fruits_F, damaged_fruits_F, total_fruits_F, nonfitness_fruits, everything()) %>% 
  arrange(Block, Location, Recipient)

#CHECK for seed_ct NAs. Should only occur when closed fruits = 0.

## NOTE: total number of non-fitness fruits per segment may be under-represented here since not all counts were matched to fitness plants and therefore merged in.

```



**Seed Weights**  

```{r}
# Read in only # seeds weighed and total weight
# Calculate mean weight per seed

BB_seedmass_1 <- read_csv(here::here("data_sheets", "seed_mass", "BB_Seed_Mass_2023-02-05.csv"), skip = 1) 
  

colnames(BB_seedmass_1) <- c("Block", "Location", "Recipient", "Plant_ID", "closed_fruits_d1", "viable_seeds_d1", "seeds_weighed_d1", "seed_mass_mg_d1", "closed_fruits_d2", "viable_seeds_d2", "seeds_weighed_d2", "seed_mass_mg_d2", "closed_fruits_d3", "viable_seeds_d3", "seeds_weighed_d3", "seed_mass_mg_d3", "closed_fruits_d4", "viable_seeds_d4", "seeds_weighed_d4", "seed_mass_mg_d4", "closed_fruits_d5", "viable_seeds_d5", "seeds_weighed_d5", "seed_mass_mg_d5", "closed_fruits_d6", "viable_seeds_d6", "seeds_weighed_d6", "seed_mass_mg_d6", 
                             "closed_fruits_d7", "viable_seeds_d7", "seeds_weighed_d7", "seed_mass_mg_d7",
                             "closed_fruits_d8", "viable_seeds_d8", "seeds_weighed_d8", "seed_mass_mg_d8", 
                             "closed_fruits_d9", "viable_seeds_d9", "seeds_weighed_d9", "seed_mass_mg_d9", "ffd")

#Fix Recipient IDs
BB_seedmass_1.5 <- BB_seedmass_1 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BB_", Recipient_num))) %>% 
  mutate(Plant_ID = as.numeric(Plant_ID)) %>% #set Plant ID to numeric for sorting purposes
  select(-c(Recipient_num, ffd))


BB_seedmass_2 <- BB_seedmass_1.5 %>% 
  select(-starts_with(c("closed_fruits_", "viable_seeds_"))) %>% 
  mutate(msm_d1 = seed_mass_mg_d1/seeds_weighed_d1,
         msm_d2 = seed_mass_mg_d2/seeds_weighed_d2,
         msm_d3 = seed_mass_mg_d3/seeds_weighed_d3,
         msm_d4 = seed_mass_mg_d4/seeds_weighed_d4,
         msm_d5 = seed_mass_mg_d5/seeds_weighed_d5,
         msm_d6 = seed_mass_mg_d6/seeds_weighed_d6,
         msm_d7 = seed_mass_mg_d7/seeds_weighed_d7,
         msm_d8 = seed_mass_mg_d8/seeds_weighed_d8,
         msm_d9 = seed_mass_mg_d9/seeds_weighed_d9)
         

BB_seedmass_3 <- BB_seedmass_2 %>%
  mutate(msm_all = rowMeans(select(BB_seedmass_2, starts_with("msm_")), na.rm = TRUE)) %>% 
  rowwise() %>% 
  mutate(sd_msm_all = sd(c_across(c(msm_d1:msm_d9)), na.rm = TRUE))
         
  
##Merge with fruit/seed count data frame

BB_fruitseed_ct3 <- merge(BB_fruitseed_ct2, BB_seedmass_3, by = c("Block", "Location", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  mutate(Block = case_when(Block %in% c("W1","W2") ~ "W",
                           TRUE ~ Block)) #Correct Black names to "W" to avoid losing data below
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```



Merge everything together:  
```{r}
BB_merged_df1 <- merge(BB_GermFlow_3, BB_fruitseed_ct3, by = c("Block", "Location", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  select(Block, Location, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, everything()) %>%
  arrange(Block, Location, Sequence, Recipient, Plant_ID) %>% 
  mutate(closed_fruits_F = as.numeric(closed_fruits_F),
         total_fruits_F = as.numeric(total_fruits_F)) #correct counts to numeric values


#Correct FruitNAs to 0s in cases where no fruit data was entered (fitness and non-fitness plants):

#
BB_merged_df2 <- BB_merged_df1 %>% 
  mutate(closed_fruits_F = replace_na(closed_fruits_F, 0),
         total_fruits_F = replace_na(total_fruits_F, 0),
         seed_ct = replace_na(seed_ct, 0)) #set seed_ct NAs to 0s

#Calculate mean seeds per fruit, total_est_seed prod
BB_merged_df3 <- BB_merged_df2 %>% 
  mutate(mean_seeds_per_fruit = seed_ct/closed_fruits_F,
         total_est_seed_production = mean_seeds_per_fruit*total_fruits_F) %>% 
  mutate(mean_seeds_per_fruit = case_when(any_FitP == TRUE & closed_fruits_F == 0 ~ 0, 
                                          TRUE ~ mean_seeds_per_fruit)) %>%  #else, original value
  mutate(total_est_seed_production = case_when(any_FitP == TRUE & mean_seeds_per_fruit == 0 ~ 0,
                                               TRUE ~ total_est_seed_production))


#mean seeds per fruit and total est fruit production have NAs/NaNs for when there were no closed fruits/seeds (fitness plants only)



#Correct remaining formatting issues

BB_merged_df4 <- BB_merged_df3 %>% 
  mutate(Fl_2.25 = as_date(Fl_2.25),
         Fl_3.25 = as_date(Fl_3.25),
         Fl_4.1 = as_date(Fl_4.1),
         Fl_4.22 = as_date(Fl_4.22),
         Fl_5.4 = as_date(Fl_5.4),
         Fl_5.12 = as_date(Fl_5.12))

```



**Skeleton Weights & Corolla area and diameter**  

Load and merge-in these data sheets
```{r}
##
skel_weights_BB <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeleton_weights_BB_2023_03_01.csv")) %>% select(-Units, -Block, -8, -9) %>% 
  filter(!grepl("EXCLUDE", Notes)) %>% ## Filter out notes with "Exclude" #None in BB data here
  rename(AG_biomass_mg = `Skeleton Weights (mg)`)

#Watered points
skel_weights_BB_W <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeletons_watered_BB_2023_02_08.csv")) %>% 
  select(-`Units`) %>% 
  rename(Plant_ID = `Plant ID`, AG_biomass_mg = `Skeleton Weights`)

#Combine
skel_weights_all_BB <- rbind(skel_weights_BB, skel_weights_BB_W) %>% 
  select(-Notes) %>% 
  filter(!is.na(AG_biomass_mg))


#Merge with full data set
BB_merged_df5 <- merge(BB_merged_df4, skel_weights_all_BB, by = c("Location", "Recipient", "Plant_ID"), all.x = TRUE)

```


Corolla measurements
```{r}
corollas_BB <- read_csv(here::here("data_sheets", "corolla_measures", "corollameasures_BB_2023-02-28.csv")) %>% select(1, 2, 3, 4, 6, 8, 9, 10) %>% 
  rename(Location = 2, Recipient = 3, Plant_ID = 4, photo_date = 5, corolla_diam_mm = 7, corolla_area_mm2 = 6, notes_corolla = 8) %>% 
  filter(!grepl("REMOVE", notes_corolla)) ## Filter out notes with "REMOVE"

corollas_BB2 <- corollas_BB %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("BB_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  mutate(Block = case_when(Block %in% c("W1","W2") ~ "W", #Fix blocks
                           TRUE ~ Block)) %>% 
  arrange(Block, Location, Recipient, Plant_ID, photo_date) %>% 
  distinct(Location, Recipient, Plant_ID, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates
#Only about 6 duplicate observations removed


## Merge with main data 
BB_merged_df6 <- merge(BB_merged_df5, corollas_BB2, by = c("Block", "Location", "Recipient", "Plant_ID"), all.x = TRUE)
  
```




**Leaf Area & Weight**  
```{r}
leaves_BB <- read_csv(here::here("data_sheets", "leaf_data", "BB_SLA_data_20230522.csv")) %>% 
  rename(leaf_area_mm2 = 5, leafarea_notes = 6, leaf_dryweight_mg = 7, leafweight_notes = 8) %>% 
  filter(!grepl("EXCLUDE", leafarea_notes)) %>% ## Filter out notes with "EXCLUDE"
  filter(!is.na(leaf_area_mm2)) #exlcude NA leaf areas?



#Eliminate Duplicates
#Randomly select one leaf to be analyzed per transect x maternal ID combo
leaves_BB_duped <- leaves_BB %>% 
  mutate(duped1 = duplicated(select(leaves_BB, c(Block, Location, Recipient))),
         duped2 = duplicated(select(leaves_BB, c(Block, Location, Recipient)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# Select dupe rows that have both area and weight!


#This seems to work the best!
leaves_BB_unique <- leaves_BB_duped %>% 
  group_by(Block, Location, Recipient) %>% 
  slice_sample(n = 1) %>%  #select a random row for each  combo of block, transect, recipient
  select(-c(duped, Plant_ID)) %>% #Don't need plant ID...these were all non-fitness plants
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg,
         LMA = 1/SLA)


#Merge in to main data 

BB_merged_df7 <- merge(BB_merged_df6, leaves_BB_unique, by = c("Block", "Location", "Recipient"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo



```




Error checks:

```{r eval = FALSE}
##Check for mismatched (fitness) plants

BB_Echeck1 <- BB_merged_df6 %>% 
  filter(any_FitP == FALSE & total_fruits_F != 0) #do the same for seed count, using `>` and `!=`

BB_Echeck2 <- BB_merged_df6 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
BB_Echeck3 <- BB_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))
######Loc. 16 ID BB_166 plant 5 shows up as NA for seed count??

BB_Echeck4 <- BB_merged_df6 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))

## Check for NA seed counts for fitness plants (when fruit production > 0)
BB_Echeck5 <- BB_merged_df6 %>% 
  filter(any_FitP == TRUE & is.na(seed_ct))


#Seed mass
BB_Echeck5 <- BB_merged_df6 %>% 
  filter(any_FitP == FALSE & !is.na(msm_all))


```

<br>  


## Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd`  


#### Phenology  

```{r}

BB_merged_df_fl <- BB_merged_df7 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('Fl_')), na.rm = TRUE)), #First flowering date
         days_germ2flower = as.numeric(FFD - Germ_Date), #days from germination to flowering
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab-sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('Fl_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) #Flowering duration


#Survival measures
BB_merged_df_fl2 <- BB_merged_df_fl %>% 
  mutate(surv_to_germ = case_when(!is.na(Germ_Date) ~ 1, #plants that germinated
                                  is.na(Germ_Date) ~ 0),
         surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(total_fruits_F > 0 ~ 1, #plants that survived to produce fruit
                                       total_fruits_F == 0 ~ 0),
         surv_to_seedprod = case_when(seed_ct > 0 ~ 1, #plants that survived to produce fruit with viable seeds
                                       TRUE ~ 0)) #else = 0 (applies to seed_ct = NA)

         


## Note: should flowering duration remain unchanged when a plant went several weeks without flowering (e.g., had two temporally distinct blooms during the growing season)?
```

<br>  

## Master sheets (v1)  


```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...


BB_mastersheet_v1 <- BB_merged_df_fl2 %>% 
  select(Block, Location, Sequence, Co., Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Location, Sequence, Co., Recipient, Plant_ID) %>% #Correct Recipient names - two mat fams with different donors
  mutate(Recipient = case_when(Donor == "BB_217" & Location %in% c("33", "99", "154") ~ "BB_205_A",
                               Donor == "BB_188" & Location %in% c("48", "96", "126") ~ "BB_205_B",
                               TRUE ~ Recipient))

write_csv(BB_mastersheet_v1, here::here("data_sheets", "compiled_sheets", "BB_mastersheet_full_2023-07-19.csv"))

#Create meta data sheet



#
## Fitness Plants Only
#

BB_mastersheet_F_mains <-  BB_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only single fitness plants

BB_mastersheet_F_all <- BB_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treat these as 0 fitness or as NAs to remove???


write_csv(BB_mastersheet_F_mains, here::here("data_sheets", "compiled_sheets", "BB_mastersheet_Fitness-mains_2023-07-19.csv"))

write_csv(BB_mastersheet_F_all, here::here("data_sheets", "compiled_sheets", "BB_mastersheet_Fitness-all_2023-07-19.csv"))

```



