---
title: "Compile FLower, Fruit, and Seed count data sheets - Fitness plants"
output: html_notebook
author: Devin Gamble
---


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



## Load and Clean Data   

NOTE: All date data should be standardized prior to importing into R. In Excel, set format to YYYY-MM-DD for consistency. Unfortunately, this process in Excel will auto-populate the year (when not already present) based on system settings. For germination data, this means that the year should be corrected to 2021!!


NOTE 2: All paths are relative thanks to the 'here' package.  


**Germination Data**  
- Add 'Date Sown' column and data in as well.

<<AC>>  
-NOTE: for AC, plant number must be FLIPPED (i.e., plants 1,2,3,4,5,6,7,8,9,10 become plants 10,9,8,7,6,5,4,3,2,1, respectively)  
```{r message = FALSE}
#Load csv

#Skip first nonsense row (the second row is the header)
AC_germ_df1 <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "AC_germ_data.csv"), skip = 1) %>%
  rename(Block = 'Block (1-3)', Transect = 'Transect (1-9)', Sequence = 'Sequence (1-321)', Donor = "Pollen Donor ID", Recipient = "Maternal Sibship", Stage = 'mm/dd') %>% 
  filter(!Stage == "Flow Date")   #Remove empty flowering date rows


#
#Reshape data sheet to 'long' format
AC_germ_df2 <- AC_germ_df1 %>% 
  pivot_longer(cols = c('1':'10'),
               names_to = 'Plant_ID',
               values_to = 'Germ_Date') 


#
##Fix typos in date values
##MAJOR TYPO: Excel auto-populated year as 2022 when in fact the year should be 2021!
unique(AC_germ_df2$Germ_Date)

AC_germ_df3 <- AC_germ_df2 %>% 
  mutate(Germ_Date = replace(Germ_Date, Germ_Date %in% c("Dead 12/5, plug missing", "disturbed by animal (deer)", "? disturbed by animal (deer)"), NA)) %>% #specify values to assign 'NA' first to avoid mismatching text ("disturbed") present in valid dates
mutate(
  Germ_Date = case_when(
    Germ_Date %in% c("2022-12-05", "12/5, ? plug buried", "12/5/2021, ?", "12/5/2021, disturbed by animal (deer)") ~ "2021-12-05",
    Germ_Date %in% c("2022-11-03", "11/3/2021, disturbed by animal (deer)", "11/3 (n=2, n=1 on 12/5)") ~ "2021-11-03")) 

#check again
unique(AC_germ_df3$Germ_Date)
#Use sum(is.na(AC_germ_df2$Date)) to compare change in number of NA values


#
#Adjust Donor and Recipient names for easy sorting - three digits (leading zeros)
#Specify Germ_Date as 'date' class variable
AC_germ_df4 <- AC_germ_df3 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("AC_", Donor_num))) %>%
  mutate(Germ_Date = as_date(Germ_Date)) %>% #set variable to class = date
  select(Block, Transect, Sequence, Donor, Donor_num, Recipient, Recipient_num, Plant_ID, everything()) %>% 
  select(-c(Stage, Donor_num, Recipient_num))


#
##Add in Sow Date  
AC_germ_df5 <- AC_germ_df4 %>% 
  mutate(Sow_Date = as_date("2021-11-04"))
#Some plugs may have been sown on November 5, 2022 (information on which transects not recorded)


#
## Flip Plant numbers
## (AC germination data only) - the orientation in which plants were identified and checked for flowering was reversed
AC_germ_df6 <- AC_germ_df5 %>% 
  mutate(Plant_ID = as.numeric(case_when(
    Plant_ID == "1" ~ "10",
    Plant_ID == "2" ~ "9",
    Plant_ID == "3" ~ "8",
    Plant_ID == "4" ~ "7",
    Plant_ID == "5" ~ "6",
    Plant_ID == "6" ~ "5",
    Plant_ID == "7" ~ "4",
    Plant_ID == "8" ~ "3",
    Plant_ID == "9" ~ "2",
    Plant_ID == "10" ~ "1"))) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID) #Arrange by Transect, Recipient, Plant_ID

```


Add in watered transect germination data
```{r message = FALSE}
#AC
## Add in watered transects
## *Note*: Only 8 seeds sown in these segments instead of 10!
AC_germ_W1 <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "watered_AC_germ.csv"), skip = 1) %>%
  rename(Sequence = 'Seq (1-42)', Donor = "Donor_ID", Recipient = "Recipient_ID", Stage = 'Date (mm/dd)') %>% 
  filter(!Stage == "Flow") %>%    #Remove empty flowering date rows
  mutate(`2` = as_date(`2`), `5` = as_date(`5`), `8` = as_date(`8`)) #these columns not recognized as dates during read-in. These must all be the same class since they'll be reshaped below.
  
#
#Reshape data sheet to 'long' format
AC_germ_W2 <- AC_germ_W1 %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'Germ_Date') 

#
##Fix typos in date values
##MAJOR TYPO: Excel auto-populated year as 2022 when in fact the year should be 2021!
unique(AC_germ_W2$Germ_Date) #no typos here

AC_germ_W3 <- AC_germ_W2 %>% 
  mutate(Germ_Date = case_when(
    Germ_Date == "2022-12-05" ~ "2021-12-05",
    Germ_Date == "2022-11-03" ~ "2021-11-03"))


#
#Adjust Donor and Recipient names for easy sorting - three digits (leading zeros)
#Specify Germ_Date as 'date' class variable
AC_germ_W4 <- AC_germ_W3 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("AC_", Donor_num))) %>%
  mutate(Germ_Date = as_date(Germ_Date)) %>% #set variable to class = date
  mutate(Plant_ID = as.numeric(Plant_ID)) %>% #set Plant ID to numeric for sorting purposes
  select(-c(Stage, Co., Donor_num, Recipient_num))


#
##Add in Sow Date & Update Sequence number & add block ('W')
AC_germ_W5 <- AC_germ_W4 %>% 
  mutate(Sow_Date = as_date("2021-11-04")) %>% 
  mutate(Sequence = (Sequence + 321)) %>% 
  mutate(Block = "W") %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Plant_ID, everything()) %>% 
  arrange(Transect, Sequence, Recipient, Plant_ID) #Arrange by Transect, Recipient, Plant_ID


#
## Flipping Plant numbers not required for watering transects - plugs planted top-to-bottom for ID 1 through 8


#
## Combine watered and main experiment germination data

AC_germ_full <- rbind(AC_germ_df6, AC_germ_W5)

write_csv(AC_germ_full, here::here("data_sheets", "compiled_sheets", "AC_germination_full_2021-2022.csv"))

```




**Flower Check sheets**  
-- Organize flowering data
-- Do not trust the fitness plant data in the flowering check data sheets (add in later).  

*Important Step!!*  
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R much easier!  



```{r}
##AC##

ACflr_df <- read_csv(file = here::here( "data_sheets", "season_end_flowerchecks_2022",  "AC_flowercheck_2022_06_28.csv"), na = c("", "NA"))
##Ignore 'F' plants - more up-to-date fitness plant data is matched in below 

#split groups
f1_AClong <- ACflr_df[c(1:15)] # Group 1: plants 1-4
f2_AClong <- ACflr_df[c(1, 16:29)] # Group 2: plants 5-8

#Revise names
colnames(f1_AClong) <- c("Transect", "Recipient", "Plant_ID", "F.", "Fl_4.9", "Fl_4.21", "Fl_5.3", "Fl_5.13", "Fl_5.25", "Fl_6.8", "Fl_6.22", "D", "D1", "LEAF", "flr_P")
colnames(f2_AClong) <- c("Transect", "Recipient", "Plant_ID", "F.", "Fl_4.9", "Fl_4.21", "Fl_5.3", "Fl_5.13", "Fl_5.25", "Fl_6.8", "Fl_6.22", "D", "D1", "LEAF", "flr_P")

#rbind, tidy up and remove unnecessary columns
f_AClong <- rbind(f1_AClong, f2_AClong) %>% arrange(Transect, Recipient, Plant_ID) %>% select(-c(F., D, D1)) %>% remove_empty("rows") #sort by transect, then Recipient, then Plant ID


#Format recipient IDs for easier sorting - three digits (includes leading zeros)
f_AClong_2 <- f_AClong %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  select(Transect, Recipient, Recipient_num, Plant_ID, everything()) %>% 
  select(-Recipient_num)


#Old - LK
####f_AClongTEST <- f_AClong[complete.cases(f_AClong[ , 5]), ] #This removes plants that didn't flower on 4/9, but also plants that may have flowered later

##f_AClong$ff_date <- with(f_AClong, coalesce(X4.9, X04.21, X5.3, X5.13, X5.25, X6.8, x6.22))
##write.csv(f_AClong, "cAC_07142022.csv" )

```


Combine Germination and Flowering data frames
```{r}
#Merge by Recipient, Transect, and Plant ID
AC_GermFlow_1 <- left_join(AC_germ_full, f_AClong_2, by = c("Recipient", "Transect", "Plant_ID")) %>% 
  mutate(Fl_6.22 = as_date(Fl_6.22)) #correct this column to be class 'date'

```


<br>  


#### Fitness Plant data  
- Indicate which plants were fitness plants
- Replicated? Indicate which maternal families were present in the main and watering experiment  

```{r message = FALSE}
AC_Fplants <- read_csv(here::here("data_sheets", "fitness_plants_7-21-2022", "AC_F_plants.csv"), col_types = cols(`ALL Fitness Plants` = col_character())) %>% #retain commas for column with multiple fitness plants
  rename(F_plant = `SINGLE Fitness Plant`, F_multi = `ALL Fitness Plants`, F_plant_notes = NOTES) %>% 
  select(Transect:F_plant_notes)

#Adjust Recipient names to match three-digit format
AC_Fplants2 <- AC_Fplants %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  mutate(Replicated = case_when(Replicated == "R" ~ "R",
                                is.na(Replicated) ~ "n")) #change NAs to 'n' (no) for non-rep'd Recipient IDs

```


Index fitness plants & Replicated (y/n) in germination and flowering data
```{r message = FALSE}
# Merge the two dataframes - F plant data will be repeated x10 
# Create new row (Plant ID) specific variables for main and replicated fitness plants
AC_GermFlow_2 <- merge(AC_GermFlow_1, AC_Fplants2) %>% 
  mutate(FitP = case_when(F_plant == Plant_ID ~ TRUE,
                           F_plant != Plant_ID ~ FALSE)) %>% 
  mutate(Rep_FitP = case_when(Replicated == "R" & Plant_ID != 1 & Plant_ID == str_match(F_multi, as.character(Plant_ID)) ~ TRUE, #prevent str_match from matching '1' to the 1 in '10'
                              Replicated == "R" & Plant_ID == 1 & Plant_ID == str_match(F_multi, "1(?!0)") ~ TRUE, #cases where Rep_FitP should == 1
                              TRUE ~ FALSE)) %>%  #if else, set value to FALSE
  mutate(any_FitP = case_when(FitP == TRUE | Rep_FitP == TRUE ~ TRUE, #plant is any form of fitness plant
                              TRUE ~ FALSE))


#Tidy up
AC_GermFlow_3 <- AC_GermFlow_2 %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, everything()) %>% 
  arrange(Transect, Sequence, Recipient, Plant_ID)


```



**Fruit and Seed counts**  

Due to the structure of excel data sheets, it seems easiest to upload fruit and seed count data separately since different spreadsheets were used to record each. Info on plants' Block, Transect, sequence, etc. will be pulled from the fruit count sheets (assumed to be up-to-date). For sheets intended to record fruit production only, counts of viable seeds, mean seed weight, etc. that are copied over are ignored and instead pulled straight from sheets where they were first entered later on.

Date added: **7/28/2022**  

*NOTE*: Update with later versions of data sheet when more seed counts have been entered
```{r message = FALSE}
#
# Fruit count data
# Data on *when* fruits were collected will be omitted here; only the total fruit counts (closed, open, non-fitness) are retained from the csv

AC_fruitprod_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "AC-Fruit-Production.csv"), skip = 2) %>% 
  select(1:11) #omit sow and flower date, seed count, and date-specific fruit counts by column index

AC_fruitprod_2 <- AC_fruitprod_1 %>% 
  select(-c(Recipient, 9)) %>% 
  rename(Recipient = `Recipient for Sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, closed_fruits_F = `Number of closed fruits used to estimate mean seeds per fruit (include multiple dates)`, open_fruits_F = `Open Fruits`, damaged_fruits_F = `Eaten or damaged fruits`, total_fruits_F = `Total number of all fruits produced by the plant (open + closed)`, nonfitness_fruits = `Total number of non-fitness fruits collected from segment`) 

#Note: total fruits should not include eaten/damaged fruit


#
# Seed count data (viable seeds)
# Similarly, only seed count data are retained from this spreadsheet to minimize errors

AC_seedct_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "AC-seed-count.csv"), skip = 2) %>% 
  select(-c(3, 6, 7, 9:29)) %>% 
  rename(Recipient = `Recipient for Sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, seed_ct = `Total number of viable seeds counted across multiple dates`)



#merge data
AC_fruitseed_ct1 <- left_join(AC_fruitprod_2, AC_seedct_1, by = c("Sequence", "Transect", "Recipient", "Plant_ID"))
#Sequence data in fruit count sheets may be unreliable for some sites (e.g., BO)


# summary variables, tidying up
AC_fruitseed_ct2 <- AC_fruitseed_ct1 %>% 
  select(Transect, Sequence, Recipient, Plant_ID, closed_fruits_F, open_fruits_F, damaged_fruits_F, total_fruits_F, nonfitness_fruits, everything()) %>% 
  arrange(Transect, Sequence, Recipient)

## NOTE: total number of non-fitness fruits per segment may be unrepresented here since not all counts were matched to fitness plants and therefore merged in.

```



**Seed Weights**  
```{r}
# Add in once weighing completed


```



Merge everything together:  
```{r}
AC_merged_df1 <- merge(AC_GermFlow_3, AC_fruitseed_ct2, by = c("Transect", "Sequence", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID) %>% 
  mutate(closed_fruits_F = as.numeric(closed_fruits_F),
         total_fruits_F = as.numeric(total_fruits_F)) #correct counts to numeric values


#Correct Fruit/seed NAs to 0s in cases where no fruit data was entered (fitness and non-fitness plants):

AC_merged_df2 <- AC_merged_df1 %>% 
  mutate(closed_fruits_F = replace_na(closed_fruits_F, 0),
         total_fruits_F = replace_na(total_fruits_F, 0),
         seed_ct = replace_na(seed_ct, 0)) %>% 
  mutate(mean_seeds_per_fruit = seed_ct/closed_fruits_F,
         total_est_seed_production = mean_seeds_per_fruit*total_fruits_F) 


#Calculate mean seeds per fruit, total_est_seed prod
AC_merged_df3 <- AC_merged_df2 %>% 
  mutate(mean_seeds_per_fruit = seed_ct/closed_fruits_F,
         total_est_seed_production = mean_seeds_per_fruit*total_fruits_F) %>% 
  mutate(mean_seeds_per_fruit = case_when(any_FitP == TRUE & closed_fruits_F == 0 ~ 0, 
                                          TRUE ~ mean_seeds_per_fruit)) %>%  #else, original value
  mutate(total_est_seed_production = case_when(any_FitP == TRUE & mean_seeds_per_fruit == 0 ~ 0,
                                               TRUE ~ total_est_seed_production))

##commented out code was first used to replace NAs with 0s when fitness plants had closed fruits for mean seeds & est seed production


#mean seeds per fruit and total est fruit production have NAs/NaNs for when there were no closed fruits/seeds (fitness plants only)

```


Error checks:

```{r eval = FALSE}
##Check for mismatched (fitness) plants

AC_Echeck1 <- AC_merged_df3 %>% 
  filter(any_FitP == FALSE & total_fruits_F != 0) #do the same for seed count, using `>` and `!=`

AC_Echeck2 <- AC_merged_df3 %>% 
  filter(any_FitP == TRUE) %>% 
  filter(closed_fruits_F == 0 & seed_ct > 0)

##NA/NaN/Info values for mean seeds/est seed prod
AC_Echeck3 <- AC_merged_df3 %>% 
  filter(any_FitP == TRUE & is.na(as.numeric(mean_seeds_per_fruit)))

AC_Echeck4 <- AC_merged_df3 %>% 
  filter(any_FitP == TRUE & !is.finite(mean_seeds_per_fruit))


```

<br>  


## Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd`  


#### Phenology  

```{r}

AC_merged_df_fl <- AC_merged_df3 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('Fl_')), na.rm = TRUE)), #First flowering date
         days_germ2flower = as.numeric(FFD - Germ_Date), #days from germination to flowering
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from germination to flowering
         LFD = do.call(pmax, c(select(., starts_with('Fl_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) #Flowering duration


#Survival measures
AC_merged_df_fl2 <- AC_merged_df_fl %>% 
  mutate(surv_to_germ = case_when(!is.na(Germ_Date) ~ 1, #plants that germinated
                                  is.na(Germ_Date) ~ 0),
         surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(total_fruits_F > 0 ~ 1, #plants that survived to produce fruit
                                       total_fruits_F == 0 ~ 0),
         surv_to_seedprod = case_when(seed_ct > 0 ~ 1, #plants that survived to produce fruit with viable seeds
                                       seed_ct == 0 ~ 0))
  
#Note: it appears that only 1 plant was recorded as having 0 fruits. Is this because 0s were generally not input? 
# Regardless, survival to seed production would be a better metric here
         


## Note 2: should flowering duration remain unchanged when a plant went several weeks without flowering (e.g., had two temporally distinct blooms during the growing season)?
```

<br>  

## Master sheets (v1)  


```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...


AC_mastersheet_v1 <- AC_merged_df_fl2 %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, any_FitP, F_plant, F_multi, F_plant_notes, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

write_csv(AC_mastersheet_v1, here::here("data_sheets", "compiled_sheets", "AC_mastersheet_v1_2022-07-26.csv"))

#Create meta data sheet



#
## Fitness Plants Only
#

AC_mastersheet_Fonly_v1 <-  AC_mastersheet_v1 %>% 
  filter(FitP == TRUE) #Only single fitness plants

AC_mastersheet_Fonly_v2 <- AC_mastersheet_v1 %>% 
  filter(any_FitP == TRUE) #single and multiple fitness plants (replicated IDs)


##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treate these as 0 fitness or as NAs to remove???


write_csv(AC_mastersheet_Fonly_v1, here::here("data_sheets", "compiled_sheets", "AC_mastersheet_v1_Fitness-only_2022-07-26.csv"))
```


```{r eval = FALSE}


# Calculate Maternal- and paterna- genotype fitness by FRIDAY


## UPDATE EVERYTHING
# (0) After all rounds of typo-correcting complete
# (1) After seed weighing complete
# (2) after watered plant data complete (seed count and weight?)

```



