---
title: "Compile FLower, Fruit, and Seed count data sheets - Fitness plants"
output: html_notebook
author: Devin Gamble
---


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



## Load and Clean Data   

NOTE: All date data should be standardized prior to importing into R. In Excel, set format to YYYY-MM-DD for consistency. Unfortunately, this process in Excel will auto-populate the year (when not already present) based on system settings. For germination data, this means that the year should be corrected to 2021!!


NOTE 2: All paths are relative thanks to the 'here' package.  


**Germination Data**  
- Add 'Date Sown' column and data in as well.

<<AC>>  
-NOTE: for AC, plant number must be FLIPPED (i.e., plants 1,2,3,4,5,6,7,8,9,10 become plants 10,9,8,7,6,5,4,3,2,1, respectively)  
```{r message = FALSE}
#Load csv

#Skip first nonsense row (the second row is the header)
AC_germ_df1 <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "AC_germ_data.csv"), skip = 1) %>%
  rename(Block = 'Block (1-3)', Transect = 'Transect (1-9)', Sequence = 'Sequence (1-321)', Donor = "Pollen Donor ID", Recipient = "Maternal Sibship", Stage = 'mm/dd') %>% 
  filter(!Stage == "Flow Date")   #Remove empty flowering date rows


#
#Reshape data sheet to 'long' format
AC_germ_df2 <- AC_germ_df1 %>% 
  pivot_longer(cols = c('1':'10'),
               names_to = 'Plant_ID',
               values_to = 'Germ_Date') 


#
##Fix typos in date values
##MAJOR TYPO: Excel auto-populated year as 2022 when in fact the year should be 2021!
unique(AC_germ_df2$Germ_Date)

AC_germ_df3 <- AC_germ_df2 %>% 
  mutate(Germ_Date = replace(Germ_Date, Germ_Date %in% c("Dead 12/5, plug missing", "disturbed by animal (deer)", "? disturbed by animal (deer)"), NA)) %>% #specify values to assign 'NA' first to avoid mismatching text ("disturbed") present in valid dates
mutate(
  Germ_Date = case_when(
    Germ_Date %in% c("2022-12-05", "12/5, ? plug buried", "12/5/2021, ?", "12/5/2021, disturbed by animal (deer)") ~ "2021-12-05",
    Germ_Date %in% c("2022-11-03", "11/3/2021, disturbed by animal (deer)", "11/3 (n=2, n=1 on 12/5)") ~ "2021-11-03")) 

#check again
unique(AC_germ_df3$Germ_Date)
#Use sum(is.na(AC_germ_df2$Date)) to compare change in number of NA values


#
#Adjust Donor and Recipient names for easy sorting - three digits (leading zeros)
#Specify Germ_Date as 'date' class variable
AC_germ_df4 <- AC_germ_df3 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("AC_", Donor_num))) %>%
  mutate(Germ_Date = as_date(Germ_Date)) %>% #set variable to class = date
  select(Block, Transect, Sequence, Donor, Donor_num, Recipient, Recipient_num, Plant_ID, everything()) %>% 
  select(-c(Stage, Donor_num, Recipient_num))


#
##Add in Sow Date  
AC_germ_df5 <- AC_germ_df4 %>% 
  mutate(Sow_Date = as_date("2021-11-04"))
#Some plugs may have been sown on November 5, 2022 (information on which transects not recorded)


#
## Flip Plant numbers
## (AC germination data only) - the orientation in which plants were identified and checked for flowering was reversed
AC_germ_df6 <- AC_germ_df5 %>% 
  mutate(Plant_ID = as.numeric(case_when(
    Plant_ID == "1" ~ "10",
    Plant_ID == "2" ~ "9",
    Plant_ID == "3" ~ "8",
    Plant_ID == "4" ~ "7",
    Plant_ID == "5" ~ "6",
    Plant_ID == "6" ~ "5",
    Plant_ID == "7" ~ "4",
    Plant_ID == "8" ~ "3",
    Plant_ID == "9" ~ "2",
    Plant_ID == "10" ~ "1"))) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID) #Arrange by Transect, Recipient, Plant_ID

```


Add in watered transect germination data
```{r message = FALSE}
#AC
## Add in watered transects
## *Note*: Only 8 seeds sown in these segments instead of 10!
AC_germ_W1 <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "watered_AC_germ.csv"), skip = 1) %>%
  rename(Sequence = 'Seq (1-42)', Donor = "Donor_ID", Recipient = "Recipient_ID", Stage = 'Date (mm/dd)') %>% 
  filter(!Stage == "Flow") %>%    #Remove empty flowering date rows
  mutate(`2` = as_date(`2`), `5` = as_date(`5`), `8` = as_date(`8`)) #these columns not recognized as dates during read-in. These must all be the same class since they'll be reshaped below.
  
#
#Reshape data sheet to 'long' format
AC_germ_W2 <- AC_germ_W1 %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'Germ_Date') 

#
##Fix typos in date values
##MAJOR TYPO: Excel auto-populated year as 2022 when in fact the year should be 2021!
unique(AC_germ_W2$Germ_Date) #no typos here

AC_germ_W3 <- AC_germ_W2 %>% 
  mutate(Germ_Date = case_when(
    Germ_Date == "2022-12-05" ~ "2021-12-05",
    Germ_Date == "2022-11-03" ~ "2021-11-03"))


#
#Adjust Donor and Recipient names for easy sorting - three digits (leading zeros)
#Specify Germ_Date as 'date' class variable
AC_germ_W4 <- AC_germ_W3 %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("AC_", Donor_num))) %>%
  mutate(Germ_Date = as_date(Germ_Date)) %>% #set variable to class = date
  mutate(Plant_ID = as.numeric(Plant_ID)) %>% #set Plant ID to numeric for sorting purposes
  select(-c(Stage, Co., Donor_num, Recipient_num))


#
##Add in Sow Date & Update Sequence number & add block ('W')
AC_germ_W5 <- AC_germ_W4 %>% 
  mutate(Sow_Date = as_date("2021-11-04")) %>% 
  mutate(Sequence = (Sequence + 321)) %>% 
  mutate(Block = "W") %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Plant_ID, everything()) %>% 
  arrange(Transect, Sequence, Recipient, Plant_ID) #Arrange by Transect, Recipient, Plant_ID


#
## Flipping Plant numbers not required for watering transects - plugs planted top-to-bottom for ID 1 through 8


#
#
## Combine watered and main experiment germination data

AC_germ_full <- rbind(AC_germ_df6, AC_germ_W5)

write_csv(AC_germ_full, here::here("data_sheets", "compiled_sheets", "AC_germination_full_2021-2022.csv"))

```




**Flower Check sheets**  
-- Organize flowering data
-- Do not trust the fitness plant data in the flowering check data sheets (add in later).  

*Important Step!!*  
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R much easier!  



```{r}
##AC##

ACflr_df <- read_csv(file = here::here( "data_sheets", "season_end_flowerchecks_2022",  "AC_flowercheck_2022_06_28.csv"), na = c("", "NA"))
##Ignore 'F' plants - more up-to-date fitness plant data is matched in below 

#split groups
f1_AClong <- ACflr_df[c(1:15)] # Group 1: plants 1-4
f2_AClong <- ACflr_df[c(1, 16:29)] # Group 2: plants 5-8

#Revise names
colnames(f1_AClong) <- c("Transect", "Recipient", "Plant_ID", "F.", "Fl_4.9", "Fl_4.21", "Fl_5.3", "Fl_5.13", "Fl_5.25", "Fl_6.8", "Fl_6.22", "D", "D1", "LEAF", "flr_P")
colnames(f2_AClong) <- c("Transect", "Recipient", "Plant_ID", "F.", "Fl_4.9", "Fl_4.21", "Fl_5.3", "Fl_5.13", "Fl_5.25", "Fl_6.8", "Fl_6.22", "D", "D1", "LEAF", "flr_P")

#rbind, tidy up and remove unnecessary columns
f_AClong <- rbind(f1_AClong, f2_AClong) %>% arrange(Transect, Recipient, Plant_ID) %>% select(-c(F., D, D1)) %>% remove_empty("rows") #sort by transect, then Recipient, then Plant ID


#Format recipient IDs for easier sorting - three digits (includes leading zeros)
f_AClong_2 <- f_AClong %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  select(Transect, Recipient, Recipient_num, Plant_ID, everything()) %>% 
  select(-Recipient_num)


#Old - LK
####f_AClongTEST <- f_AClong[complete.cases(f_AClong[ , 5]), ] #This removes plants that didn't flower on 4/9, but also plants that may have flowered later

##f_AClong$ff_date <- with(f_AClong, coalesce(X4.9, X04.21, X5.3, X5.13, X5.25, X6.8, x6.22))
##write.csv(f_AClong, "cAC_07142022.csv" )

```


Combine Germination and Flowering data frames
```{r}
#Merge by Recipient, Transect, and Plant ID
AC_GermFlow_1 <- left_join(AC_germ_full, f_AClong_2, by = c("Recipient", "Transect", "Plant_ID"))

```


<br>  


#### Fitness Plant data  
- Indicate which plants were fitness plants
- Replicated? Indicate which maternal families were present in the main and watering experiment  

```{r message = FALSE}
AC_Fplants <- read_csv(here::here("data_sheets", "fitness_plants_7-21-2022", "AC_F_plants.csv"), col_types = cols(`ALL Fitness Plants` = col_character())) %>% #retain commas for column with multiple fitness plants
  rename(F_plant = `SINGLE Fitness Plant`, F_multi = `ALL Fitness Plants`, F_plant_notes = NOTES) %>% 
  select(Transect:F_plant_notes)

#Adjust Recipient names to match three-digit format
AC_Fplants2 <- AC_Fplants %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  mutate(Replicated = case_when(Replicated == "R" ~ "R",
                                is.na(Replicated) ~ "n")) #change NAs to 'n' (no) for non-rep'd Recipient IDs

```


Index fitness plants & Replicated (y/n) in germination and flowering data
```{r message = FALSE}
# Merge the two dataframes - F plant data will be repeated x10 
# Create new row (Plant ID) specific variables for main and replicated fitness plants
AC_GermFlow_2 <- merge(AC_GermFlow_1, AC_Fplants2) %>% 
  mutate(FitP = case_when(F_plant == Plant_ID ~ TRUE,
                           F_plant != Plant_ID ~ FALSE)) %>% 
  mutate(Rep_FitP = case_when(Replicated == "R" & Plant_ID == str_match(F_multi, as.character(Plant_ID)) ~ TRUE,
                              TRUE ~ FALSE)) #if else, set value to FALSE

#Tidy up
AC_GermFlow_3 <- AC_GermFlow_2 %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, everything()) %>% 
  arrange(Transect, Sequence, Recipient, Plant_ID)


```



**Fruit and Seed counts**  

Due to the structure of excel data sheets, it seems easiest to upload fruit and seed count data separately since different spreadsheets were used to record each. The same applies to estimates of seed weight. For sheets intended to record fruit production only, counts of viable seeds, mean seed weight, etc. that are copied over are ignored and instead pulled straight from sheets where they were first entered later on.

Date added: 7/25/2022

*NOTE*: Update with later versions of data sheet when more seed counts have been entered
```{r message = FALSE}
#
# Fruit count data
# Data on *when* fruits were collected will be omitted here; only the total fruit counts (closed, open, non-fitness) are retained from the csv

AC_fruitprod_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "AC-Fruit-Production.csv"), skip = 2) %>% 
  select(1:11) #omit sow and flower date, seed count, and date-specific fruit counts by column index

AC_fruitprod_2 <- AC_fruitprod_1 %>% 
  select(-c(Recipient, 9)) %>% 
  rename(Recipient = `Recipient for Sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, closed_fruits_F = `Number of closed fruits used to estimate mean seeds per fruit (include multiple dates)`, open_fruits_F = `Open Fruits`, damaged_fruits_F = `Eaten or damaged fruits`, Total_fruits_F = `Total number of all fruits produced by the plant (open + closed)`, nonfitness_fruits = `Total number of non-fitness fruits collected from segment`) 


#Fix - total fruits does not include eaten/damaged fruits (of which there were very few here)
AC_fruitprod_3 <- AC_fruitprod_2 %>% 
  mutate(Total_fruits_F_NEW = Total_fruits_F + damaged_fruits_F) %>% 
  select(-Total_fruits_F) %>% 
  rename(all_fruits_F = Total_fruits_F_NEW)


#
# Seed count data (viable seeds)
# Similarly, only seed count data are retained from this spreadsheet to minimize errors

AC_seedct_1 <- read_csv(here::here("data_sheets", "fruit_seed_counts_final", "AC-seed-count.csv"), skip = 2) %>% 
  select(-c(3, 6, 7, 9:29)) %>% 
  rename(Recipient = `Recipient for Sorting`, Plant_ID = `Plant number of fitness plant (when chosen)`, seed_ct = `Total number of viable seeds counted across multiple dates`)



#merge data
AC_fruitseed_ct1 <- left_join(AC_fruitprod_3, AC_seedct_1, by = c("Sequence", "Transect", "Recipient", "Plant_ID"))


# descriptive stats, tidying up
AC_fruitseed_ct2 <- AC_fruitseed_ct1 %>% 
  mutate(mean_seeds_per_fruit = seed_ct/closed_fruits_F,
         total_est_seed_production = mean_seeds_per_fruit*all_fruits_F) %>% 
  select(Transect, Sequence, Recipient, Plant_ID, closed_fruits_F, open_fruits_F, damaged_fruits_F, all_fruits_F, nonfitness_fruits, everything()) %>% 
  arrange(Transect, Sequence, Recipient)


```


**Seed Weights**  
```{r}



```



Merge everything together:  
```{r}
AC_merged_df1 <- merge(AC_GermFlow_3, AC_fruitseed_ct2, by = c("Transect", "Sequence", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  select(Block, Transect, Sequence, Donor, Recipient, Replicated, Plant_ID, FitP, Rep_FitP, everything()) %>% 
  arrange(Block, Transect, Sequence, Recipient, Plant_ID)

```


<br>  



## Trait calculations


#### Phenology  

```{r}


```





## Filtering 

Save new csv for each filter constraint  


#### Fitness Plants Only
*** This is a priority  



**Single Fitness Plants**  
For the main experiment  


#### Only Plants that Flowered...?????


**Multiple Fitness Plants**  
For the watered and replicated segments - my experiment
- Match only replicated families in the watered and unwatered treatments





#### Fitness Estimates





```{r}
##Final merged sheets:

## 'Long' version with all data - one row per plant ID


## 'Short' version with one row = one segment



```










