---
title: "datasheet org year 2"
author: "Devin Gamble"
date: "2023-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Description**:  

Organizational Rmd for the creation of data sheets and the fixing of typos. Year 2 (2022-2023) field season data.

Code chunks are set to eval = FALSE once the necessary sheets or changes have been made.  

Load Packages:
```{r}
library(tidyverse)
library(janitor)
library(here)
library(reshape2)
```



### Creating Flowering Check Data sheets  

**AC**  


```{r}
##Load Planting Locations Data

##AC
AC_plantings <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "AC_planting_locations_fall22.csv"), skip = 1) %>% 
  select(-c(4,5,10,11))

colnames(AC_plantings) <- c("Transect", "Segment", "Sequence", "Gen.L", "Recipient.L", "Gen.R", "Recipient.R")

#Note: No lat/long recorded here (somewhere else)?

#Reshape to consolidate Gen and Recipient Variables, noting position
AC_plantings_long <- AC_plantings %>% 
  gather(key, value, -Transect, -Segment, -Sequence) %>% 
  separate(key, c("key", "position"), "\\.") %>% 
  spread(key, value)


#Add multiple rows for individual plants
#Add leading 0s to Recipients

AC_flowersheet1 <- AC_plantings_long %>% 
  mutate("1" = NA, "2" = NA, "3" = NA, "4" = NA, "5" = NA)

AC_flowersheet2 <- pivot_longer(AC_flowersheet1, cols = c("1":"5"), names_to = "Plant_ID", values_to = "FD_01") %>% 
  mutate(FD_02 = NA, FD_03 = NA, FD_04 = NA, FD_05 = NA, FD_06 = NA, FD_07 = NA, FLR_P = NA, Fit_plant = NA, LeafD = NA, Notes = NA) %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  select(-Recipient_num)



##Option - Keep the left and right positions in the same row?
AC_flowersheet2_wide <- AC_plantings %>% 
  mutate("1" = NA, "2" = NA, "3" = NA, "4" = NA, "5" = NA) %>% 
  pivot_longer(cols = c("1":"5"), names_to = "Plant_ID", values_to = "FD_01") %>% 
  rename(Plant_ID.L = Plant_ID) %>% 
  mutate(Plant_ID.R = Plant_ID.L)
#Seems too confusing...



##CHECK comments in excel/google drive version to note which Recipients weren't planted or had fewer than 5 seeds

AC_flowersheet3 <- AC_flowersheet2 %>% 
  mutate(Notes = case_when(Transect == 5 & Segment == 17 & Recipient == "AC_213" ~ "Only 4 seeds planted",
                           Transect == 4 & Segment == 29 & Recipient == "AC_203" ~ "Only 4 seeds planted",
                           Transect == 6 & Segment == 33 & Recipient == "AC_213" ~ "NOT PLANTED")) %>% 
  select(Sequence, Transect, Segment, position, Gen, Recipient, Plant_ID, Fit_plant, FD_01:FD_07, LeafD, FLR_P, Notes)


#Save CSV
write_csv(AC_flowersheet3, here::here("data_sheets", "2023_year2_data", "flowering_checks", "AC_flowercheck_data_v1.csv"), na = "")



```


<br>  


**BB**  


```{r}
##Load Planting Locations Data

##BB
BB_plantings <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "BB_planting_locations_fall22.csv"), skip = 1) %>% 
  select(-c(1,3))

colnames(BB_plantings) <- c("Location", "Gen.L", "Recipient.L", "Gen.R", "Recipient.R", "Point_lat", "Point_lon")

##Auto-fill Point numbers and point lat/lon
BB_plantings_2 <- BB_plantings %>% 
  fill(c(Location, Point_lat, Point_lon)) %>% 
  replace(is.na(.), "NULL") %>% #replace NAs for following reshaping
  mutate(Sequence = seq_along(1:n())) %>%  #Need Sequence number 
  select(-c(Point_lat, Point_lon)) #de-select for reshaping
  
#Reshape to consolidate Gen and Recipient Variables, noting position
BB_plantings_long <- BB_plantings_2 %>% 
  gather(key, value, -Location, -Sequence) %>% 
  separate(key, c("key", "position"), "\\.") %>% 
  spread(key, value)


#Add multiple rows for individual plants
#Add leading 0s to Recipients

BB_flowersheet1 <- BB_plantings_long %>% 
  mutate("1" = NA, "2" = NA, "3" = NA, "4" = NA, "5" = NA)

BB_flowersheet2 <- pivot_longer(BB_flowersheet1, cols = c("1":"5"), names_to = "Plant_ID", values_to = "FD_01") %>% 
  mutate(FD_02 = NA, FD_03 = NA, FD_04 = NA, FD_05 = NA, FD_06 = NA, FD_07 = NA, FLR_P = NA, Fit_plant = NA, LeafD = NA, Notes = NA) %>%
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BB_", Recipient_num))) %>% 
  select(-Recipient_num)



##CHECK comments in excel/google drive version to note which Recipients weren't planted or had fewer than 5 seeds

BB_flowersheet3 <- BB_flowersheet2 %>% 
  mutate(Notes = case_when(Location == 95 & Gen == "G1" & Recipient == "BB_050" ~ "Only 4 seeds planted",
                           Location == 70 & Gen == "G2" & Recipient == "BB_157" ~ "NOT PLANTED")) %>% 
  select(Sequence, Location, position, Gen, Recipient, Plant_ID, Fit_plant, FD_01:FD_07, LeafD, FLR_P, Notes)


#Save CSV
write_csv(BB_flowersheet3, here::here("data_sheets", "2023_year2_data", "flowering_checks", "BB_flowercheck_data_v1.csv"), na = "")



```

<br>  

**BO**  


```{r}
##Load Planting Locations Data

##BO
BO_plantings <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "BO_planting_locations_fall22.csv"), skip = 1) %>% 
  select(-c(4,5))

colnames(BO_plantings) <- c("Transect", "Segment", "Sequence", "Gen.L", "Recipient.L", "Gen.R", "Recipient.R", "Seg_lat", "Seg_lon")

## Remove lat and lon
BO_plantings_2 <- BO_plantings %>% 
  select(-c(Seg_lat, Seg_lon)) #de-select for reshaping

#Reshape to consolidate Gen and Recipient Variables, noting position
BO_plantings_long <- BO_plantings_2 %>% 
  gather(key, value, -Transect, -Segment, -Sequence) %>% 
  separate(key, c("key", "position"), "\\.") %>% 
  spread(key, value)


#Add multiple rows for individual plants
#Add leading 0s to Recipients

BO_flowersheet1 <- BO_plantings_long %>% 
  mutate("1" = NA, "2" = NA, "3" = NA, "4" = NA, "5" = NA)

BO_flowersheet2 <- pivot_longer(BO_flowersheet1, cols = c("1":"5"), names_to = "Plant_ID", values_to = "FD_01") %>% 
  mutate(FD_02 = NA, FD_03 = NA, FD_04 = NA, FD_05 = NA, FD_06 = NA, FD_07 = NA, FLR_P = NA, Fit_plant = NA, LeafD = NA, Notes = NA) %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  select(-Recipient_num)


##CHECK comments in excel/google drive version to note which Recipients weren't planted or had fewer than 5 seeds

BO_flowersheet3 <- BO_flowersheet2 %>% 
  mutate(Notes = case_when(Transect == 2 & Segment == 30 & Recipient == "BO_055" ~ "Only 2 seeds planted",
                           Transect == 3 & Segment == 7 & Recipient == "BO_200" ~ "Only 3 seeds planted")) %>% 
  select(Sequence, Transect, Segment, position, Gen, Recipient, Plant_ID, Fit_plant, FD_01:FD_07, LeafD, FLR_P, Notes)


#Save CSV
write_csv(BO_flowersheet3, here::here("data_sheets", "2023_year2_data", "flowering_checks", "BO_flowercheck_data_v1.csv"), na = "")



```



<br>  

**HR**  



```{r}
##Load Planting Locations Data

##AC
HR_plantings <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "HR_planting_locations_fall22.csv"), skip = 1) %>% 
  select(-c(4,5,10,11))

colnames(HR_plantings) <- c("Transect", "Segment", "Sequence", "Gen.L", "Recipient.L", "Gen.R", "Recipient.R")


#Reshape to consolidate Gen and Recipient Variables, noting position
HR_plantings_long <- HR_plantings %>% 
  gather(key, value, -Transect, -Segment, -Sequence) %>% 
  separate(key, c("key", "position"), "\\.") %>% 
  spread(key, value)


#Add multiple rows for individual plants
#Add leading 0s to Recipients
HR_flowersheet1 <- HR_plantings_long %>% 
  mutate("1" = NA, "2" = NA, "3" = NA, "4" = NA, "5" = NA)

HR_flowersheet2 <- pivot_longer(HR_flowersheet1, cols = c("1":"5"), names_to = "Plant_ID", values_to = "FD_01") %>% 
  mutate(FD_02 = NA, FD_03 = NA, FD_04 = NA, FD_05 = NA, FD_06 = NA, FD_07 = NA, FLR_P = NA, Fit_plant = NA, LeafD = NA, Notes = NA) %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  select(-Recipient_num)


##CHECK comments in excel/google drive version to note which Recipients weren't planted or had fewer than 5 seeds

HR_flowersheet3 <- HR_flowersheet2 %>% 
  select(Sequence, Transect, Segment, position, Gen, Recipient, Plant_ID, Fit_plant, FD_01:FD_07, LeafD, FLR_P, Notes)
#Typo present in segment numbering of Transect 7; comment added in flower check google sheet
#No other notes to add here


#Save CSV
write_csv(HR_flowersheet3, here::here("data_sheets", "2023_year2_data", "flowering_checks", "HR_flowercheck_data_v1.csv"), na = "")



```



#### Create Fruit & Seed Count data sheets  

Also create fitness plant reference sheets

**AC**  

Reminder: Segment number may not line up with sequence number due to mix-ups at AC, BO, HR

```{r}
#Copied from code chunk above for creating flowering sheets
AC_plantings <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "AC_planting_locations_fall22.csv"), skip = 1) %>% 
  select(-c(4,5,10,11))

colnames(AC_plantings) <- c("Transect", "Segment", "Sequence", "Gen.L", "Recipient.L", "Gen.R", "Recipient.R")

#Note: No lat/long recorded here (somewhere else)?

#Reshape to consolidate Gen and Recipient Variables, noting position
AC_plantings_long <- AC_plantings %>% 
  gather(key, value, -Transect, -Segment, -Sequence) %>% 
  separate(key, c("key", "position"), "\\.") %>% 
  spread(key, value) 

AC_plantings_long2 <- AC_plantings_long %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Add leading 0s
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  select(-Recipient_num)


##UPDATE WITH LATEST FLOWERING DATA SHEET
## Add fitness plant data from flowering sheet
AC_flowering <- read_csv(here::here("data_sheets", "2023_year2_data", "flowering_checks", "AC_flowercheck_data_2023-03-15 - AC_flowercheck_data_v1.csv"))

AC_Fitps <- AC_flowering %>% 
  filter(!is.na(Fit_plant)) %>% 
  select(1:6, 8) %>% 
  rename(position = side)

#Fitness plant sheet
AC_Fitness1 <- AC_flowering %>% 
  distinct(pick(c(Sequence:Recipient))) %>% 
  #filter(!is.na(Fit_plant)) %>% 
  rename(position = side)

AC_Fitness2 <- merge(AC_Fitness1, AC_Fitps, all.x = TRUE, by = c("Sequence", "Transect", "Segment", "position", "Gen", "Recipient"))

#Find dupes (Multiple fitness plants in a segment)
#AC_Fitness3 <- AC_Fitness2 %>% 
#  select(Sequence:Recipient) %>% 
#  filter(duplicated(.))


#save csv Fitness plants
write_csv(AC_Fitness2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "fitness_plants", "AC_fitness_23.csv"), na = "")

#Have a student check this fitness plant sheet against the flowering check sheet


AC_template <- AC_plantings_long2 %>% 
  mutate(GenSym = case_when(Gen == "G1" ~ "-",
                            Gen == "G2" ~ "X")) %>% 
  mutate(SegNoPos = paste(Segment, position, sep = "")) %>% 
  mutate(N_CLOSED_fruit_D1 =NA, N_OPEN_fruit_D1 = NA, N_FILLED_seeds_D1 = NA, N_total_seeds_D1 = NA,
         N_CLOSED_fruit_D2 =NA, N_OPEN_fruit_D2 = NA, N_FILLED_seeds_D2 = NA, N_total_seeds_D2 = NA,
         N_CLOSED_fruit_D3 =NA, N_OPEN_fruit_D3 = NA, N_FILLED_seeds_D3 = NA, N_total_seeds_D3 = NA,
         N_CLOSED_fruit_D4 =NA, N_OPEN_fruit_D4 = NA, N_FILLED_seeds_D4 = NA, N_total_seeds_D4 = NA,
         N_CLOSED_fruit_D5 =NA, N_OPEN_fruit_D5 = NA, N_FILLED_seeds_D5 = NA, N_total_seeds_D5 = NA)


#Merge in Fitness plant data
AC_template2 <- left_join(AC_template, AC_Fitness2, by = c("Sequence", "Transect", "Segment", "position", "Gen", "Recipient")) %>% 
  mutate(GT_closed_fruit = NA, GT_open_fruit = NA, GT_filled_seeds = NA, notes = NA) %>% #columns for grand totals
  select(Sequence, Transect, SegNoPos, Gen, GenSym, Recipient, Fit_plant, GT_closed_fruit, GT_open_fruit, GT_filled_seeds, everything(), -c(Segment, position)) %>% 
  filter(!is.na(Fit_plant)) %>%  #filter out where fitness plants never chosen
  rename(Seg = SegNoPos)


#Save csv - Fruit & seed counts
write_csv(AC_template2, here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "AC_fruit_seed_counts_blank_20230712.csv"), na = "")


```


**BB**  
```{r}
## No "segment" in BB data - sequence = segment.
#Copied from code chunk above for creating flowering sheets
BB_plantings <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "BB_planting_locations_fall22.csv"), skip = 1) %>% 
  select(-c(1,3))

colnames(BB_plantings) <- c("Location", "Gen.L", "Recipient.L", "Gen.R", "Recipient.R", "Point_lat", "Point_lon")


##Auto-fill Point numbers and point lat/lon
BB_plantings_2 <- BB_plantings %>% 
  fill(c(Location, Point_lat, Point_lon)) %>% 
  replace(is.na(.), "NULL") %>% #replace NAs for following reshaping
  mutate(Sequence = seq_along(1:n())) %>%  #Need Sequence number 
  select(-c(Point_lat, Point_lon)) #de-select for reshaping
  
#Reshape to consolidate Gen and Recipient Variables, noting position
BB_plantings_long <- BB_plantings_2 %>% 
  gather(key, value, -Location, -Sequence) %>% 
  separate(key, c("key", "position"), "\\.") %>% 
  spread(key, value)

BB_plantings_long2 <- BB_plantings_long %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Add leading 0s
  mutate(Recipient = as.character(str_c("BB_", Recipient_num))) %>% 
  select(-Recipient_num)


##UPDATE WITH LATEST FLOWERING DATA SHEET
## Add fitness plant data from flowering sheet
BB_flowering <- read_csv(here::here("data_sheets", "2023_year2_data", "flowering_checks", "BB_flowercheck_data_2023-02-15 - BB_flowercheck_data_v1.csv"))

BB_Fitps <- BB_flowering %>% 
  filter(!is.na(Fit_plant)) %>% 
  select(1:5, 7) %>% 
  rename(position = side)

#Fitness plant sheet
BB_Fitness1 <- BB_flowering %>% 
  distinct(pick(c(Sequence:Recipient))) %>% 
  #filter(!is.na(Fit_plant)) %>% 
  rename(position = side)

BB_Fitness2 <- merge(BB_Fitness1, BB_Fitps, all.x = TRUE, by = c("Sequence", "Location", "position", "Gen", "Recipient"))


#save csv Fitness plants
write_csv(BB_Fitness2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "BB_fitnessplants_23.csv"), na = "")


##Fruit & Seed Count Data  


BB_template <- BB_plantings_long2 %>% 
  mutate(GenSym = case_when(Gen == "G1" ~ "-",
                            Gen == "G2" ~ "X")) %>% 
  mutate(N_CLOSED_fruit_D1 =NA, N_OPEN_fruit_D1 = NA, N_FILLED_seeds_D1 = NA, N_total_seeds_D1 = NA,
         N_CLOSED_fruit_D2 =NA, N_OPEN_fruit_D2 = NA, N_FILLED_seeds_D2 = NA, N_total_seeds_D2 = NA,
         N_CLOSED_fruit_D3 =NA, N_OPEN_fruit_D3 = NA, N_FILLED_seeds_D3 = NA, N_total_seeds_D3 = NA,
         N_CLOSED_fruit_D4 =NA, N_OPEN_fruit_D4 = NA, N_FILLED_seeds_D4 = NA, N_total_seeds_D4 = NA,
         N_CLOSED_fruit_D5 =NA, N_OPEN_fruit_D5 = NA, N_FILLED_seeds_D5 = NA, N_total_seeds_D5 = NA)


#Merge in Fitness plant data
BB_template2 <- left_join(BB_template, BB_Fitness2, by = c("Location", "Sequence", "position", "Gen", "Recipient")) %>% 
  mutate(ALL_closed_fruit = NA, ALL_open_fruit = NA, ALL_filled_seeds = NA, notes = NA) %>% #columns for grand totals
  mutate(SegNoPos = paste(Sequence, position, sep = "")) %>% 
  select(Location, SegNoPos, Gen, GenSym, Recipient, Fit_plant, ALL_closed_fruit, ALL_open_fruit, ALL_filled_seeds, everything(), -c(Sequence, position)) %>% 
  filter(!is.na(Fit_plant)) %>%  #filter out where fitness plants never chosen
  rename(Seg = SegNoPos)

#Save csv - Fruit & seed counts
write_csv(BB_template2, here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "BB_fruit_seed_counts_blank_20230711.csv"), na = "")


```



**BO**  
```{r}
#BO
BO_plantings <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "BO_planting_locations_fall22.csv"), skip = 1) %>% 
  select(-c(4,5))

colnames(BO_plantings) <- c("Transect", "Segment", "Sequence", "Gen.L", "Recipient.L", "Gen.R", "Recipient.R", "Seg_lat", "Seg_lon")

## Remove lat and lon
BO_plantings_2 <- BO_plantings %>% 
  select(-c(Seg_lat, Seg_lon)) #de-select for reshaping

#Reshape to consolidate Gen and Recipient Variables, noting position
BO_plantings_long <- BO_plantings_2 %>% 
  gather(key, value, -Transect, -Segment, -Sequence) %>% 
  separate(key, c("key", "position"), "\\.") %>% 
  spread(key, value)

BO_plantings_long2 <- BO_plantings_long %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Add leading 0s
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  select(-Recipient_num)


##UPDATE WITH LATEST FLOWERING DATA SHEET
## Add fitness plant data from flowering sheet
BO_flowering <- read_csv(here::here("data_sheets", "2023_year2_data", "flowering_checks", "BO_flowercheck_data_2023-02-24 - BO_flowercheck_data_v1.csv")) %>% 
  slice(-c(3101, 3102))

BO_Fitps <- BO_flowering %>% 
  filter(!is.na(Fit_plant)) %>% 
  select(1:6, 8) %>% 
  rename(position = side)

#Fitness plant sheet
BO_Fitness1 <- BO_flowering %>% 
  distinct(pick(c(Sequence:Recipient))) %>% 
  #filter(!is.na(Fit_plant)) %>% 
  rename(position = side)

BO_Fitness2 <- merge(BO_Fitness1, BO_Fitps, all.x = TRUE, by = c("Sequence", "Transect", "Segment", "position", "Gen", "Recipient"))

#save csv Fitness plants
write_csv(BO_Fitness2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "BO_fitnessplants_23.csv"), na = "")



#Fruit & Seed Count Sheets


BO_template <- BO_plantings_long2 %>% 
  mutate(GenSym = case_when(Gen == "G1" ~ "-",
                            Gen == "G2" ~ "X")) %>% 
  mutate(SegNoPos = paste(Segment, position, sep = "")) %>% 
  mutate(N_CLOSED_fruit_D1 =NA, N_OPEN_fruit_D1 = NA, N_FILLED_seeds_D1 = NA, N_total_seeds_D1 = NA,
         N_CLOSED_fruit_D2 =NA, N_OPEN_fruit_D2 = NA, N_FILLED_seeds_D2 = NA, N_total_seeds_D2 = NA,
         N_CLOSED_fruit_D3 =NA, N_OPEN_fruit_D3 = NA, N_FILLED_seeds_D3 = NA, N_total_seeds_D3 = NA,
         N_CLOSED_fruit_D4 =NA, N_OPEN_fruit_D4 = NA, N_FILLED_seeds_D4 = NA, N_total_seeds_D4 = NA,
         N_CLOSED_fruit_D5 =NA, N_OPEN_fruit_D5 = NA, N_FILLED_seeds_D5 = NA, N_total_seeds_D5 = NA)


#Merge in Fitness plant data
BO_template2 <- left_join(BO_template, BO_Fitness2, by = c("Sequence", "Transect", "Segment", "position", "Gen", "Recipient")) %>% 
  mutate(GT_closed_fruit = NA, GT_open_fruit = NA, GT_filled_seeds = NA, notes = NA) %>% #columns for grand totals
  select(Sequence, Transect, SegNoPos, Gen, GenSym, Recipient, Fit_plant, GT_closed_fruit, GT_open_fruit, GT_filled_seeds, everything(), -c(Segment, position)) %>% 
  filter(!is.na(Fit_plant)) %>%  #filter out where fitness plants never chosen
  rename(Seg = SegNoPos)


#Save csv - Fruit & seed counts
write_csv(BO_template2, here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "BO_fruit_seed_counts_blank_20230712.csv"), na = "")

```



**HR**  
```{r}
#HR
HR_plantings <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "HR_planting_locations_fall22.csv"), skip = 1) %>% 
  select(-c(4,5))

colnames(HR_plantings) <- c("Transect", "Segment", "Sequence", "Gen.L", "Recipient.L", "Gen.R", "Recipient.R", "Seg_lat", "Seg_lon")

## Remove lat and lon
HR_plantings_2 <- HR_plantings %>% 
  select(-c(Seg_lat, Seg_lon)) #de-select for reshaping

#Reshape to consolidate Gen and Recipient Variables, noting position
HR_plantings_long <- HR_plantings_2 %>% 
  gather(key, value, -Transect, -Segment, -Sequence) %>% 
  separate(key, c("key", "position"), "\\.") %>% 
  spread(key, value)

HR_plantings_long2 <- HR_plantings_long %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Add leading 0s
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  select(-Recipient_num)


##UPDATE WITH LATEST FLOWERING DATA SHEET
## Add fitness plant data from flowering sheet
HR_flowering <- read_csv(here::here("data_sheets", "2023_year2_data", "flowering_checks", "HR_flowercheck_data_2023-02-23 - HR_flowercheck_data_v1.csv")) 

HR_Fitps <- HR_flowering %>% 
  filter(!is.na(Fit_plant)) %>% 
  select(1:6, 8) %>% 
  rename(position = side)

#Fitness plant sheet
HR_Fitness1 <- HR_flowering %>% 
  distinct(pick(c(Sequence:Recipient))) %>% 
  #filter(!is.na(Fit_plant)) %>% 
  rename(position = side)

HR_Fitness2 <- merge(HR_Fitness1, HR_Fitps, all.x = TRUE, by = c("Sequence", "Transect", "Segment", "position", "Gen", "Recipient"))

#save csv Fitness plants
write_csv(HR_Fitness2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "HR_fitnessplants_23.csv"), na = "")


#Fruit & Seed Count Sheets

HR_template <- HR_plantings_long2 %>% 
  mutate(GenSym = case_when(Gen == "G1" ~ "-",
                            Gen == "G2" ~ "X")) %>% 
  mutate(N_closed_fruit_D1 =NA, N_open_fruit_D1 = NA, N_other_fruit_D1 = NA, N_total_seeds_D1 = NA, N_filled_seeds_D1 = NA,
         N_closed_fruit_D2 = NA, N_open_fruit_D2 = NA, N_other_fruit_D2 = NA, N_total_seeds_D2 = NA, N_filled_seeds_D2 = NA,
         N_closed_fruit_D3 =NA, N_open_fruit_D3 = NA, N_other_fruit_D2 = NA, N_total_seeds_D3 = NA, N_filled_seeds_D3 = NA,
         N_closed_fruit_D4 =NA, N_open_fruit_D4 = NA, N_other_fruit_D3 = NA, N_total_seeds_D4 = NA, N_filled_seeds_D4 = NA,
         N_closed_fruit_D5 =NA, N_open_fruit_D5 = NA, N_other_fruit_D4 = NA, N_total_seeds_D5 = NA, N_filled_seeds_D5 = NA)


#Merge in Fitness plant data
HR_template2 <- left_join(HR_template, HR_Fitness2, by = c("Sequence", "Transect", "Segment", "position", "Gen", "Recipient")) %>% 
  mutate(ALL_closed_fruit = NA, ALL_open_fruit = NA, ALL_filled_seeds = NA, notes = NA) %>% #columns for grand totals
  mutate(SegNoPos = paste(Segment, position, sep = "")) %>% 
  select(Sequence, Transect, SegNoPos, Gen, GenSym, Recipient, Fit_plant, ALL_closed_fruit, ALL_open_fruit, ALL_filled_seeds, everything(), -c(Segment, position)) %>% 
  filter(!is.na(Fit_plant)) #filter out where fitness plants never chosen


#Save csv - Fruit & seed counts
write_csv(HR_template2, here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "HR_fruit_seed_counts_blank_20230710.csv"), na = "")





```


##### Skeleton Weight sheets  

Should include Transect/Location, Segment + position, Generation, Recipient, Fitness Plant ID, Skeleton weight (mg), Notes (complete/incomplete), selected for pulverization. Non-fitness skeletons may be used instead of fitness plant skeletons?  

(Already Made)

#### Pulverization Data sheets

**AC**  

```{r}

#Load fitness plant data
AC_fitness_23 <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "AC_fitness_23.csv")) %>% 
  arrange(Transect, Segment, position) %>% filter(!is.na(Fit_plant))

#skel weight data
AC_skel_23 <- read_csv(here::here("data_sheets", "2023_year2_data", "skeleton_data", "Skeleton Data 2023 - AC Skeletons.csv")) %>% rename(position = side, Gen = 4, Fitness = 7, skel_weight_mg = 8, complete_incomplete = 9, selected = 10) %>% 
  arrange(Transect, Segment, position)


AC_pulvsheet1 <- merge(AC_skel_23, AC_fitness_23, all.x = TRUE, by = c("Transect", "Segment", "position", "Gen", "Recipient")) %>% filter(Fitness != "missing") #exclude missing skeletons

#Inspect df and correct any erroneously categorized NF plants


#Filter to chosen skeletons
AC_pulvsheet2 <- AC_pulvsheet1 %>% 
  filter(selected == "y") %>% 
  select(-c(7:12)) %>% 
  arrange(Transect, Segment, position) %>% 
  mutate(rowN = seq(1:nrow(.))) #Create rowN var


#Create Sample IDs
#Population_MatFam_Gen_Transect_SegNo+side_plantNo
#EG.
#AC_156_2_001_21R_5
AC_pulvsheet3 <- AC_pulvsheet2 %>% 
  mutate(Transect_00 = sprintf("%03d", Transect), SegNoPos = paste(Segment, position, sep = ""), Plant_ID_0 = sprintf("%01d", Plant_ID)) %>% 
  unite(`Sample_ID`, c(Recipient, Gen, Transect_00, SegNoPos, Plant_ID_0), sep = "-", remove = FALSE) %>% 
  select(-c(Transect_00, Plant_ID_0)) 

#Add Columns
AC_pulvsheet4 <- AC_pulvsheet3 %>% 
  mutate(`Tube ID` = seq_along(1:n()),
         `Total_weight(capsule+sample)(mg)` = NA, `sample weight (mg)` = NA, capsule_size = "8x5mm", capsule_weight_mg = "31.48", `Tray Name` = c(rep(c("Nemo_2023_AC_1", "Nemo_2023_AC_2", "Nemo_2023_AC_3", "Nemo_2023_AC_4"), each=96), rep("Nemo_2023_AC_5", 92)), Well_ID = NA, `Type of Material` = "plant stem tissue", Analysis = "13C, 15N", Enriched = "No") %>% 
         select(rowN, Transect, SegNoPos, Recipient, Gen, Plant_ID, Sample_ID, `Tube ID`, `Total_weight(capsule+sample)(mg)`, `Tray Name`, Well_ID, capsule_size, capsule_weight_mg, `sample weight (mg)`, everything(), -c(Segment, position)) #organize columns
         

#check for dupes
AC_pulvsheet4a <- AC_pulvsheet4 %>% 
  select(Transect:Gen) %>% 
  group_by_all() %>% 
  filter(n()>1)


##Add in for tray number reps
#476 rows

write_csv(AC_pulvsheet4, here::here("data_sheets", "2023_year2_data", "skeleton_data", "AC_pulv_blank_2023.csv"), na = "")

```

**BB**  

```{r message = FALSE, warning = FALSE}
#fitness data
BB_fitness_23 <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "BB_fitnessplants_23.csv")) %>% 
  rename(Segment = Sequence)

#skel weight data
BB_skel_23 <- read_csv(here::here("data_sheets", "2023_year2_data", "skeleton_data", "Skeleton Data 2023 - BB Skeletons.csv")) %>% rename(Segment = 2, position = side, Gen = 4, Fitness = 7, skel_weight_mg = 8, complete_incomplete = 9, selected = 10) 


BB_pulvsheet1 <- merge(BB_skel_23, BB_fitness_23, all.x = TRUE, by = c("Location", "Segment", "position", "Gen", "Recipient")) %>% filter(Fitness != "missing?") #exclude missing skeletons

#Inspect df and correct any erroneously categorized NF plants


#Filter to chosen skeletons
BB_pulvsheet2 <- BB_pulvsheet1 %>% 
  filter(selected == "y") %>% 
  select(-c(7:11)) %>% 
  arrange(Location, Segment, position) %>% 
  mutate(rowN = seq(1:nrow(.))) #Create rowN var

#Create Sample IDs
#Population_MatFam_Gen_Location_SegNo+side_plantNo
#EG.
#BB_156_2_001_21R_5
BB_pulvsheet3 <- BB_pulvsheet2 %>% 
  mutate(Location_00 = sprintf("%03d", Location), SegNoPos = paste(Segment, position, sep = ""), Plant_ID_0 = sprintf("%01d", Plant_ID)) %>% 
  unite(`Sample_ID`, c(Recipient, Gen, Location_00, SegNoPos, Plant_ID_0), sep = "-", remove = FALSE) %>% 
  select(-c(Location_00, Plant_ID_0)) 

#Add Columns
BB_pulvsheet4 <- BB_pulvsheet3 %>% 
  mutate(`Tube ID` = seq_along(1:n()),
         `Total_weight(capsule+sample)(mg)` = NA, `sample weight (mg)` = NA, capsule_size = "8x5mm", capsule_weight_mg = "31.48", `Tray Name` = c(rep(c("Nemo_2023_BB_1", "Nemo_2023_BB_2", "Nemo_2023_BB_3", "Nemo_2023_BB_4"), each=96), rep("Nemo_2023_BB_5", 46)), Well_ID = NA, `Type of Material` = "plant stem tissue", Analysis = "13C, 15N", Enriched = "No") %>% 
         select(rowN, Location, SegNoPos, Recipient, Gen, Plant_ID, Sample_ID, `Tube ID`, `Total_weight(capsule+sample)(mg)`, `Tray Name`, Well_ID, capsule_size, capsule_weight_mg, `sample weight (mg)`, everything(), -c(Segment, position)) #organize columns
         

##Add in for tray number reps
 #430 rows
#BB_pulvsheet4a <- BB_pulvsheet4 %>% 
#  select(Location:Gen) %>% 
#  group_by_all() %>% 
#  filter(n()>1)

write_csv(BB_pulvsheet4, here::here("data_sheets", "2023_year2_data", "skeleton_data", "BB_pulv_blank_2023.csv"), na = "")



```

**BO**  

```{r message = FALSE, warning = FALSE}

#fitness data
BO_fitness_23 <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "BO_fitnessplants_23.csv")) %>% 
  arrange(Transect, Segment, position) %>% filter(!is.na(Fit_plant))

#skel weight data
BO_skel_23 <- read_csv(here::here("data_sheets", "2023_year2_data", "skeleton_data", "Skeleton Data 2023 - BO Skeletons.csv")) %>% rename(position = side, Gen = 4, Fitness = 7, skel_weight_mg = 8, complete_incomplete = 9, selected = 10) %>% 
  arrange(Transect, Segment, position)


BO_pulvsheet1 <- merge(BO_skel_23, BO_fitness_23, all.x = TRUE, by = c("Transect", "Segment", "position", "Gen", "Recipient")) %>% filter(Fitness != "missing?") #exclude missing skeletons

#Inspect df and correct any erroneously categorized NF plants


#Filter to chosen skeletons
BO_pulvsheet2 <- BO_pulvsheet1 %>% 
  filter(selected == "y") %>% 
  select(-c(7:12)) %>% 
  arrange(Transect, Segment, position) %>% 
  mutate(rowN = seq(1:nrow(.))) #Create rowN var


#Create Sample IDs
#Population_MatFam_Gen_Transect_SegNo+side_plantNo
#EG.
#BO_156_2_001_21R_5
BO_pulvsheet3 <- BO_pulvsheet2 %>% 
  mutate(Transect_00 = sprintf("%03d", Transect), SegNoPos = paste(Segment, position, sep = ""), Plant_ID_0 = sprintf("%01d", Plant_ID)) %>% 
  unite(`Sample_ID`, c(Recipient, Gen, Transect_00, SegNoPos, Plant_ID_0), sep = "-", remove = FALSE) %>% 
  select(-c(Transect_00, Plant_ID_0)) 

#Add Columns
BO_pulvsheet4 <- BO_pulvsheet3 %>% 
  mutate(`Tube ID` = seq_along(1:n()),
         `Total_weight(capsule+sample)(mg)` = NA, `sample weight (mg)` = NA, capsule_size = "8x5mm", capsule_weight_mg = "31.48", `Tray Name` = c(rep(c("Nemo_2023_BO_1", "Nemo_2023_BO_2"), each=96), rep("Nemo_2023_BO_3", 24)), Well_ID = NA, `Type of Material` = "plant stem tissue", Analysis = "13C, 15N", Enriched = "No") %>% 
         select(rowN, Transect, SegNoPos, Recipient, Gen, Plant_ID, Sample_ID, `Tube ID`, `Total_weight(capsule+sample)(mg)`, `Tray Name`, Well_ID, capsule_size, capsule_weight_mg, `sample weight (mg)`, everything(), -c(Segment, position)) #organize columns
         

##Add in for tray number reps
#216 rows
BO_pulvsheet4a <- BO_pulvsheet4 %>% 
  select(Transect:Gen) %>% 
  group_by_all() %>% 
  filter(n()>1)


write_csv(BO_pulvsheet4, here::here("data_sheets", "2023_year2_data", "skeleton_data", "BO_pulv_blank_2023.csv"), na = "")



```



**HR**  


```{r message = FALSE, warning = FALSE}
#fitness data
HR_fitness_23 <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "HR_fitnessplants_23.csv"))

#skel weight data
HR_skel_23 <- read_csv(here::here("data_sheets", "2023_year2_data", "skeleton_data", "Skeleton Data 2023 - HR Skeletons.csv")) %>% rename(position = side, Gen = 4, Fitness = 7, skel_weight_mg = 8, complete_incomplete = 9, selected = 10) 


HR_pulvsheet1 <- merge(HR_skel_23, HR_fitness_23, all.x = TRUE, by = c("Transect", "Segment", "position", "Gen", "Recipient")) %>% filter(Fitness != "missing", Fitness != "missing?") %>%  #exclude missing skeletons
  arrange(Transect, Segment, position)

#Inspect df and correct any erroneously categorized NF plants


#Filter to chosen skeletons
HR_pulvsheet2 <- HR_pulvsheet1 %>% 
  filter(selected == "y") %>% 
  select(-c(7:12)) %>% 
  mutate(rowN = seq(1:nrow(.))) %>%  #create row number variable
  arrange(Transect, Segment, position)


#Create Sample IDs
#Population_MatFam_Gen_Transect_SegNo+side_plantNo
#EG.
#HR_156_2_001_21R_5
HR_pulvsheet3 <- HR_pulvsheet2 %>% 
  mutate(Transect_00 = sprintf("%03d", Transect), SegNoPos = paste(Segment, position, sep = ""), Plant_ID_0 = sprintf("%01d", Plant_ID)) %>% 
  unite(`Sample_ID`, c(Recipient, Gen, Transect_00, SegNoPos, Plant_ID_0), sep = "-", remove = FALSE) %>% 
  select(-c(Transect_00, Plant_ID_0))


#Add Columns
HR_pulvsheet4 <- HR_pulvsheet3 %>% 
  mutate(`Tube ID` = seq_along(1:n()),
         `Total_weight(capsule+sample)(mg)` = NA, `sample weight (mg)` = NA, capsule_size = "8x5mm", capsule_weight_mg = "31.48", `Tray Name` = c(rep(c("Nemo_2023_HR_1", "Nemo_2023_HR_2", "Nemo_2023_HR_3", "Nemo_2023_HR_4", "Nemo_2023_HR_5", "Nemo_2023_HR_6", "Nemo_2023_HR_7"), each=96), rep("Nemo_2023_HR_7", 56)), Well_ID = NA, `Type of Material` = "plant stem tissue", Analysis = "13C, 15N", Enriched = "No") %>% 
         select(rowN, Transect, SegNoPos, Recipient, Gen, Plant_ID, Sample_ID, `Tube ID`, `Total_weight(capsule+sample)(mg)`, `Tray Name`, Well_ID, capsule_size, capsule_weight_mg, `sample weight (mg)`, everything(), -c(Segment, position)) #organize columns
         

##Add in for tray number reps
#`Tray Name` = c(rep(c("Nemo_2023_HR_1", "Nemo_2023_HR_2", "Nemo_2023_HR_3", "Nemo_2023_HR_4", "Nemo_2023_HR_5", "Nemo_2023_HR_6", "Nemo_2023_HR_7"), each=96), rep("Nemo_2023_HR_7", 57)) #727 rows

#check for dupes
HR_pulvsheet4a <- HR_pulvsheet4 %>% 
  select(Transect:Gen) %>% 
  group_by_all() %>% 
  filter(n()>1)

write_csv(HR_pulvsheet4, here::here("data_sheets", "2023_year2_data", "skeleton_data", "HR_pulv_blank_2023.csv"), na = "")



```




#### Corolla Diameter measurement sheets  

**AC**

```{r}
##Adapt above code used to create fruit/seed count data sheets to create corolla measurement data sheets
# Refer to code used to make SLA sheets; corolla photos named following standardized format


```




**HR**  


```{r}
#Load Fitness plant data




```





#### Leaf Area & Weight sheets  





