---
title: "data_org"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Description**:  

Organizational Rmd for the creation of data sheets and the fixing of typos. 

Code chunks are set to eval = FALSE once the necessary sheets or changes have been made.  

Load Packages:
```{r}
library(tidyverse)
library(janitor)
library(here)
```


**NOTE:** This code was used to organize flowering/germination data sheets early in the 2022 field season. These sheets were revised again and the versions created below were never used. Paths have been updated to indicate that much of these files are old versions.  

### Load Master Sheets  
```{r}
# 2021-2022 field season

#Main & Extra Fitness plants
nemo_allFs_AC <- read_csv(here::here("data_sheets", "compiled_sheets", "AC_mastersheet_Fitness-all_2023-02-05.csv"))

nemo_allFs_BB <- read_csv(here::here("data_sheets", "compiled_sheets", "BB_mastersheet_Fitness-all_2023-02-06.csv"))

nemo_allFs_BO <- read_csv(here::here("data_sheets", "compiled_sheets", "BO_mastersheet_Fitness-all_2023-02-06.csv"))

nemo_allFs_HR <- read_csv(here::here("data_sheets", "compiled_sheets", "HR_mastersheet_Fitness-all_2023-02-06.csv"))

```



#### Update Bodega Bay Data Sheet  

Data sheet should have one row (section) per plant, plus multiple columns to record flowering date.


```{r eval = FALSE}
#Load original germination and flowering date sheet
# - Some data present already

BB_germflow_old <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "old_versions", "Bodega Bay - Data Sheets _03-03-2022_germflow.csv")) %>% 
  rename(Block = 'Block (1-3)', Location = 'Location (1-156)', Seq = 'Seq (1-441)', Stage = 'mm/dd') %>% 
  select(-c('...16', 'Notes 12/12', 'Notes 12/19-20', 'Notes 12/30')) %>% 
  filter(!Stage == "Germ Date") %>%  #Remove germination date rows
  slice(-c(442,443)) #Remove last two rows that had notes


#Pivot/expand to create rows for all 8 plants

BB_flowdate_1 <- BB_germflow_old %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'flow_date') %>% 
  rename(date_1 = flow_date) %>% 
  mutate(date_2 = NA, date_3 = NA, date_4 = NA, date_5 = NA)

##Add in watered points

BB_gf_watered <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "old_versions", "BB_watered_Germ_Flower Dates_3-04-2022.csv")) %>% 
  rename(Seq = 'Seq (1-42)', Stage = 'Date (mm/dd)') %>% 
  select(-c('Notes 12/12', 'Notes 12/19-20', 'Notes 12/30', Seq)) %>% 
  filter(!Stage == "Germ")

BB_gf_watered$Stage <- "Flow Date"

BB_watered_piv <- BB_gf_watered %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'flow_date') %>% 
  rename(date_1 = flow_date) %>% 
  mutate(date_2 = NA, date_3 = NA, date_4 = NA, date_5 = NA, Seq = NA)


BB_pivlong_all <- rbind(BB_flowdate_1, BB_watered_piv) %>% arrange(Location)


#Modify to make rows for plants 1-4; 5-8
#Simplest method seems to be splitting this into two data frames and then pasting them together

BB_pivlong_14 <- BB_pivlong_all %>% 
  filter(Plant_ID %in% c(1:4))

BB_pivlong_58 <- BB_pivlong_all %>% 
  filter(Plant_ID %in% c(5:8)) %>% 
  select(-Block, -Seq, -Co., -Donor_ID, -Stage) %>% #Remove necessary repeated columns (retain Location & ID)
  mutate(flr_P = NA) #Column to indicate if corolla photo taken


#cbind by Location and Recipient_ID, keep all = T
BB_pivlong_split <- cbind(BB_pivlong_14, BB_pivlong_58) 

#write_csv(BB_pivlong_split, here::here("data_sheets", "blank_flowercheck_sheets", "BB_flowercheck_2022_03_08.csv"), na = "")

```


##### Update Hastings  


```{r eval = FALSE}

HR_germflow_old <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "old_versions", "Hastings Ranch- Germination Monitoring Data Sheets 2021_12_06.csv")) %>% 
  rename(Block = 'Block (1-3)', Transect = 'Transect (1-12)', Seq = 'Sequence (1-396)', Donor = "Pollen Donor ID", Recipient = "Maternal Sibship", Stage = 'mm/dd') %>% 
  select(-c('9', '10', 'Notes 12/6', 'Notes 12/17', 'Oak leaves removed when ontop of seedlings', Block, Seq)) %>% 
  filter(!Stage == "Germ Date") %>%   #Remove germination date rows
  mutate('1' = NA, '2' = NA, '3' = NA, '4' = NA, '5' = NA, '6' = NA, '7' = NA, '8' = NA)


HR_flowdate_1 <- HR_germflow_old %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'flow_date') %>% 
  rename(date_1 = flow_date) %>% 
  mutate(date_2 = NA, date_3 = NA, date_4 = NA, date_5 = NA)


##Add in watered points

HR_gf_watered <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "old_versions", "Germ_Flower Dates_watered_2021-03-06_HR.csv")) %>% 
  rename(Seq = 'Seq (1-42)', Stage = 'Date (mm/dd)') %>% 
  select(-c(Seq)) %>% 
  filter(!Stage == "Germ")

HR_gf_watered$Stage <- "Flow Date"

HR_watered_piv <- HR_gf_watered %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'flow_date') %>% 
  rename(date_1 = flow_date, Donor = Donor_ID, Recipient = Recipient_ID) %>% 
  mutate(date_2 = NA, date_3 = NA, date_4 = NA, date_5 = NA, Seq = NA) %>% 
  select(-c(Seq, Co.))


HR_pivlong_all <- rbind(HR_flowdate_1, HR_watered_piv) %>% arrange(Transect)


#Modify to make rows for plants 1-4; 5-8
HR_pivlong_14 <- HR_pivlong_all %>% 
  filter(Plant_ID %in% c(1:4))

HR_pivlong_58 <- HR_pivlong_all %>% 
  filter(Plant_ID %in% c(5:8)) %>% 
  select(-Donor, -Stage) %>% #Remove necessary repeated columns (retain Location & ID)
  mutate(flr_P = NA) #Column to indicate if corolla photo taken


HR_pivlong_split <- cbind(HR_pivlong_14, HR_pivlong_58) 
#Cannot manipulate further since columns are duplicated

#write_csv(HR_pivlong_split, here::here("data_sheets", "blank_flowercheck_sheets", "HR_flowercheck_2022_03_14.csv"), na = "")

```



##### Update Blue Oak  

```{r eval = FALSE}

BO_germflow_old <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "old_versions", "Blue Oak- Data Sheets 2021-12-21_Germ and Flow Dates.csv")) %>% 
  rename(Block = 'Block (1-3)', Transect = 'Transect (1-12)', Seq = 'Sequence (1-396)', Donor = "Pollen Donor ID", Recipient = "Maternal Sibship", Stage = 'mm/dd') %>% 
  select(-c(Block, Seq)) %>% 
  filter(!Stage == "Germ Date") %>%   #Remove germination date rows
  mutate('1' = NA, '2' = NA, '3' = NA, '4' = NA, '5' = NA, '6' = NA, '7' = NA, '8' = NA)


BO_flowdate_1 <- BO_germflow_old %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'flow_date') %>% 
  rename(date_1 = flow_date) %>% 
  mutate(date_2 = NA, date_3 = NA, date_4 = NA, date_5 = NA) %>% arrange(Transect)


##Add in watered points

BO_gf_watered <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "old_versions", "Germ_Flower Dates_watered_2021-03-06_BO.csv")) %>% 
  rename(Seq = 'Seq (1-42)', Stage = 'Date (mm/dd)') %>% 
  select(-c(Seq)) %>% 
  filter(!Stage == "Germ")

BO_gf_watered$Stage <- "Flow Date"

BO_watered_piv <- BO_gf_watered %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'flow_date') %>% 
  rename(date_1 = flow_date, Donor = Donor_ID, Recipient = Recipient_ID) %>% 
  mutate(date_2 = NA, date_3 = NA, date_4 = NA, date_5 = NA, Seq = NA) %>% 
  select(-c(Seq, Co.))


BO_pivlong_all <- rbind(BO_flowdate_1, BO_watered_piv) %>% arrange(Transect)


#Modify to make rows for plants 1-4; 5-8
BO_pivlong_14 <- BO_pivlong_all %>% 
  filter(Plant_ID %in% c(1:4))

BO_pivlong_58 <- BO_pivlong_all %>% 
  filter(Plant_ID %in% c(5:8)) %>% 
  select(-Donor, -Stage) %>% #Remove necessary repeated columns (retain Location & ID)
  mutate(flr_P = NA) #Column to indicate if corolla photo taken


BO_pivlong_split <- cbind(BO_pivlong_14, BO_pivlong_58) 
#Cannot manipulate further since columns are duplicated

#write_csv(BO_pivlong_split, here::here("data_sheets", "blank_flowercheck_sheets", "BO_flowercheck_2022_03_14.csv"), na = "")




```



##### Update Angelo Coast 

```{r eval = FALSE}

AC_germflow_old <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "old_versions", "Angelo Coast - Data Sheets 2021-10-31_Germ and Flow Dates.csv")) %>% 
  rename(Block = 'Block (1-3)', Transect = 'Transect (1-9)', Seq = 'Sequence (1-321)', Donor = "Pollen Donor ID", Recipient = "Maternal Sibship", Stage = 'mm/dd') %>% 
  select(-c(Block, Seq)) %>% 
  filter(!Stage == "Germ Date") %>%   #Remove germination date rows
  mutate('1' = NA, '2' = NA, '3' = NA, '4' = NA, '5' = NA, '6' = NA, '7' = NA, '8' = NA, '9' = NA, '10' = NA)


AC_flowdate_1 <- AC_germflow_old %>% 
  pivot_longer(cols = c('1':'10'),
               names_to = 'Plant_ID',
               values_to = 'flow_date') %>% 
  rename(date_1 = flow_date) %>% 
  mutate(date_2 = NA, date_3 = NA, date_4 = NA, date_5 = NA) %>% arrange(Transect)


AC_pivlong_all1 <- AC_flowdate_1 %>% arrange(Transect)


#Modify to make rows for plants 1-5; 6-10
AC_pivlong_15 <- AC_pivlong_all1 %>% 
  filter(Plant_ID %in% c(1:5))

AC_pivlong_610 <- AC_pivlong_all1 %>% 
  filter(Plant_ID %in% c(6:10)) %>% 
  select(-Donor, -Stage) %>% #Remove necessary repeated columns (retain Location & ID)
  mutate(flr_P = NA) #Column to indicate if corolla photo taken


##Add in watered points

AC_gf_watered <- read_csv(here::here("data_sheets", "germination_checks_2021-2022", "old_versions", "Germ_Flower Dates_watered_2021-03-06_AC.csv")) %>% 
  rename(Seq = 'Seq (1-42)', Stage = 'Date (mm/dd)') %>% 
  select(-c(Seq)) %>% 
  filter(!Stage == "Germ")

AC_gf_watered$Stage <- "Flow Date"

AC_watered_piv <- AC_gf_watered %>% 
  pivot_longer(cols = c('1':'8'),
               names_to = 'Plant_ID',
               values_to = 'flow_date') %>% 
  rename(date_1 = flow_date, Donor = Donor_ID, Recipient = Recipient_ID) %>% 
  mutate(date_2 = NA, date_3 = NA, date_4 = NA, date_5 = NA, Seq = NA) %>% 
  select(-c(Seq, Co.))


AC_pivlong_14w <- AC_watered_piv %>% 
  filter(Plant_ID %in% c(1:4))

AC_pivlong_58w <- AC_watered_piv %>% 
  filter(Plant_ID %in% c(5:8)) %>% 
  select(-Donor, -Stage) %>% #Remove necessary repeated columns (retain Location & ID)
  mutate(flr_P = NA) 



AC_pivlong_split1 <- cbind(AC_pivlong_15, AC_pivlong_610) 
#Cannot manipulate further since columns are duplicated

AC_pivlong_split2 <- cbind(AC_pivlong_14w, AC_pivlong_58w) 

AC_pivlong_split3 <- rbind(AC_pivlong_split1, AC_pivlong_split2)

#write_csv(AC_pivlong_split3, here::here("data_sheets", "blank_flowercheck_sheets", "AC_flowercheck_2022_03_15.csv"), na = "")


```


<br>  


#### Update HR skeleton weighing datasheet  
- merge in the segments for which skeletons have already been weighed  


```{r}

nemo_allFs <- read_csv(here::here("data_sheets", "compiled_sheets", "HR_mastersheet_v2_all-Fitness-only_2022-08-18.csv"))


nemo_HR_skel_temp <- read_csv(here::here("Skeleton Data - HR Skeletons.csv")) %>% 
  rename(SK_weights = `Skeleton Weights`,
         SK_notes = Notes)


nemo_HR_skel_new <- merge(nemo_allFs, nemo_HR_skel_temp, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  select(Transect, Recipient, Plant_ID, SK_weights, Units, SK_notes) %>% 
  rename(`Skeleton Weights` = SK_weights,
         Notes = SK_notes)

write_csv(nemo_HR_skel_new, here::here("data_sheets", "nemo_HR_skeleton_weights.csv"), na = "")


```



#### Create Pulverization Data Sheets  

Using Skeleton weights data sheets

**AC**  

```{r}
AC_weights <- read_csv(here::here("data_sheets", "skeleton_and_stem_data", "skeleton_weights_AC_2023_02_06.csv")) %>% select(-Units) %>% 
  filter(is.na(Notes)|Notes != "DNE") %>%  #prevent loss of NA rows in notes
  filter(`Chosen for Pulverization? (y/n)` == "y") %>% 
  select(-`Chosen for Pulverization? (y/n)`)
  
colnames(AC_weights) <- c("Transect", "Recipient", "Plant_ID", "skel_weight", "skel_notes")



# Attempt to keep only rows for which skeleton collections exist
# Exclude DNE rows
#Check and make sure ID x plant combos line up with main or extra fitness plants

mergetestAC <- merge(AC_weights, nemo_allFs_AC, by = c("Transect", "Recipient", "Plant_ID"), all.x = TRUE)
#No need right now to save this, just check that there are no all-NA rows from the master sheet (that would imply a mismatch)

#Create Sample IDs

pulv_sheet1 <- AC_weights %>% 
  mutate(Transect_00 = sprintf("%03d", Transect), Plant_ID_0 = sprintf("%02d", Plant_ID)) %>% 
  unite(`Sample ID`, c(Recipient, Transect_00, Plant_ID_0), sep = "-", remove = FALSE) %>% 
  select(Transect, Recipient, Plant_ID, `Sample ID`, skel_weight, skel_notes, -c(Transect_00, Plant_ID_0))

#Add columns for Tray ID, sample position, and sample weight (mg)
# Column names should approx. match those in the Davis Isotop sample example data sheet

pulv_sheet2 <- pulv_sheet1 %>% 
  mutate(`Tube ID` = seq_along(1:n()),
         `Weight including capsule (mg)` = NA, `Sample weight (mg)` = NA, `Tray Name` = c(rep(c("Nemo_2022_AC_1", "Nemo_2022_AC_2", "Nemo_2022_AC_3"), each=96), rep("Nemo_2022_AC_4", 15)), 
         `Well ID` = NA, 
         `Type of Material` = "plant stem tissue", Analysis = "13C, 15N", Enriched = "No") %>% 
  select(Transect, Recipient, Plant_ID, `Sample ID`, `Tube ID`, `Weight including capsule (mg)`, `Sample weight (mg)`, `Well ID`, `Tray Name`, `Type of Material`, Analysis, Enriched)

write_csv(pulv_sheet2, here::here("data_sheets", "skeleton_and_stem_data", "pulv_sample_sheet_AC.csv"), na = "")

```




#### Create SLA Data sheets (2022 data)  


**All populations**  


```{r}
#AC
AC_SLA <- nemo_allFs_AC %>% 
  distinct(Block, Transect, Recipient) %>% 
  mutate(Plant_ID = NA, leaf_area_mm2 = NA, area_notes = NA, leaf_dry_weight_mg = NA, weight_notes = NA) %>% 
  arrange(Block, Transect, Recipient)

write_csv(AC_SLA, here::here("data_sheets", "SLA_measures", "AC_SLA_data_blank.csv"), na = "")

##BB
BB_SLA <- nemo_allFs_BB %>% 
  distinct(Block, Location, Recipient) %>% 
  mutate(Plant_ID = NA, leaf_area_mm2 = NA, area_notes = NA, leaf_dry_weight_mg = NA, weight_notes = NA) %>% 
  arrange(Block, Location, Recipient)

write_csv(BB_SLA, here::here("data_sheets", "SLA_measures", "BB_SLA_data_blank.csv"), na = "")


##BO
BO_SLA <- nemo_allFs_BO %>% 
  distinct(Block, Transect, Recipient) %>% 
  mutate(Plant_ID = NA, leaf_area_mm2 = NA, area_notes = NA, leaf_dry_weight_mg = NA, weight_notes = NA) %>% 
  arrange(Block, Transect, Recipient)

write_csv(BO_SLA, here::here("data_sheets", "SLA_measures", "BO_SLA_data_blank.csv"), na = "")


##BO
HR_SLA <- nemo_allFs_HR %>% 
  distinct(Block, Transect, Recipient) %>% 
  mutate(Plant_ID = NA, leaf_area_mm2 = NA, area_notes = NA, leaf_dry_weight_mg = NA, weight_notes = NA) %>% 
  arrange(Block, Transect, Recipient)

write_csv(HR_SLA, here::here("data_sheets", "SLA_measures", "HR_SLA_data_blank.csv"), na = "")


```







