---
title: "Fitness Data Year 2"
output: html_document
date: "2023-09-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(corrplot)

```


#### Description  
This file was created to explore broad trends in our field data from the first year of the experiment (2021-2022). It was also used to explore and select different metrics of fitness (and fecundity) that would be used in later analyses.  


### Year 2 Fitness Data Sumamry  

Proportion of *Fitness Plants* that  
- produced 0 closed fruits
- produced 0 closed fruits & 0 open fruits
- produced 1+ closed fruits & 0 filled seeds

Also investigate correlations  
- between seed production (total) and estimated fecundity

**AC**
```{r message = FALSE}
#Load data (chosen fitness plants only)
AC_FitP <- read_csv(here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "AC_traits_Y2_incomplete_20230922.csv"))

#Proportion of Fitness plants with 0 closed fruits
AC_prop1 <- AC_FitP %>% 
  filter(ALL_closed_fruit == 0) %>% 
  summarize(zero_closed = n()) #128

#Proportion of Fitness plants with 0 total fruits
AC_prop2 <- AC_FitP %>% 
  filter(total_fruits == 0) %>% 
  summarize(zero_closed = 128/546,
            zero_total = n()/546) #54

#Good estimates of fecundity - including 0s
AC_prop2b <- AC_FitP %>% 
  filter(est_fecundity >= 0) %>% #NOT the same as (ALL_closed_fruit != 0 & ALL_filled_seeds >=0)!!!!!
  summarize(zero_closed = 128/546,
            real_fecundity = n()) #472

#Good estimates of fecundity - excluding 0s
AC_prop2c <- AC_FitP %>% 
  filter(est_fecundity > 0) %>% 
  summarize(zero_closed = 128/546,
            zero_total = n()) #360

AC_prop2d <- AC_FitP %>% 
  filter(is.na(est_fecundity)) %>% #should be the SAME as (ALL_closed_fruit == 0 & ALL_open_fruit >= 1)
  summarize(zero_closed = 128/546,
            fecundity_NA = n()) #74

#Proportion of Fitness plants with 0 filled seeds but 1 or more closed fruits
AC_prop3 <- AC_FitP %>% 
  filter(ALL_closed_fruit != 0 & ALL_filled_seeds == 0) %>% 
  summarize(pop = "AC",
            N_fitness_plants = 546,
            real_fecundity = 472/546,
            pos_fecundity = 360/546,
            zero_closed_fruits = 128/546,
            zero_total_fruits = 54/546,
            fecundity_NA = 74/546, #Should complement the real_fecundity column
            zero_seeds_from_nonzero_closed = n()/546)#58


```


**BB**
```{r message = FALSE}
#Load data (chosen fitness plants only)
BB_FitP <- read_csv(here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BB_traits_Y2_incomplete_20230922.csv"))

#Proportion of Fitness plants with 0 closed fruits
BB_prop1 <- BB_FitP %>% 
  filter(ALL_closed_fruit == 0) %>% 
  summarize(zero_closed = n()) #116/500

#Proportion of Fitness plants with 0 total fruits
BB_prop2 <- BB_FitP %>% 
  filter(total_fruits == 0) %>% 
  summarize(zero_closed = 116/500,
            zero_total = n()) #101

#Good estimates of fecundity - including 0s
BB_prop2b <- BB_FitP %>% 
  filter(est_fecundity >= 0) %>% #NOT the same as (ALL_closed_fruit != 0 & ALL_filled_seeds >=0)!!!!!
  summarize(real_fecdunity = n()) #485

#Good estimates of fecundity - excluding 0s
BB_prop2c <- BB_FitP %>% 
  filter(est_fecundity > 0) %>% 
  summarize(pos_fecundity = n()) #351

BB_prop2d <- BB_FitP %>% 
  filter(is.na(est_fecundity)) %>% #should be the SAME as (ALL_closed_fruit == 0 & ALL_open_fruit >= 1)
  summarize(fecundity_NA = n()) #15

#Proportion of Fitness plants with 0 filled seeds but 1 or more closed fruits
BB_prop3 <- BB_FitP %>% 
  filter(ALL_closed_fruit != 0 & ALL_filled_seeds == 0) %>% 
  summarize(pop = "BB",
            N_fitness_plants = 500,
            real_fecundity = 485/500,
            pos_fecundity = 351/500,
            zero_closed_fruits = 114/500,
            zero_total_fruits = 101/500,
            fecundity_NA = 15/500,
            zero_seeds_from_nonzero_closed = n()/500) #33


```


**BO**
```{r message = FALSE}
#Load data (chosen fitness plants only)
BO_FitP <- read_csv(here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BO_traits_Y2_incomplete_20230922.csv"))

#Proportion of Fitness plants with 0 closed fruits
BO_prop1 <- BO_FitP %>% 
  filter(ALL_closed_fruit == 0) %>% 
  summarize(zero_closed = n()) #89/257

#Proportion of Fitness plants with 0 total fruits
BO_prop2 <- BO_FitP %>% 
  filter(total_fruits == 0) %>% 
  summarize(zero_closed = 89/257,
            zero_total = n()) #44

#Good estimates of fecundity - including 0s
BO_prop2b <- BO_FitP %>% 
  filter(est_fecundity >= 0) %>% #NOT the same as (ALL_closed_fruit != 0 & ALL_filled_seeds >=0)!!!!!
  summarize(real_fecdunity = n()) #212

#Good estimates of fecundity - excluding 0s
BO_prop2c <- BO_FitP %>% 
  filter(est_fecundity > 0) %>% 
  summarize(pos_fecundity = n()) #145

BO_prop2d <- BO_FitP %>% 
  filter(is.na(est_fecundity)) %>% #should be the SAME as (ALL_closed_fruit == 0 & ALL_open_fruit >= 1)
  summarize(pos_fecundity = n()) #45

#Proportion of Fitness plants with 0 filled seeds but 1 or more closed fruits
BO_prop3 <- BO_FitP %>% 
  filter(ALL_closed_fruit != 0 & ALL_filled_seeds == 0) %>% 
  summarize(pop = "BO",
            N_fitness_plants = 257,
            real_fecundity = 212/257,
            pos_fecundity = 145/257,
            zero_closed_fruits = 89/257,
            zero_total_fruits = 44/257,
            fecundity_NA = 45/257,
            zero_seeds_from_nonzero_closed = n()/257) #23

```


**HR**
```{r message = FALSE}
#Load data (chosen fitness plants only)
HR_FitP <- read_csv(here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "HR_traits_Y2_incomplete_20230922.csv"))

#Proportion of Fitness plants with 0 closed fruits
HR_prop1 <- HR_FitP %>% 
  filter(ALL_closed_fruit == 0) %>% 
  summarize(zero_closed = n()) #197/780

#Proportion of Fitness plants with 0 total fruits
HR_prop2 <- HR_FitP %>% 
  filter(total_fruits == 0) %>% 
  summarize(zero_closed = 197/780,
            zero_total = n()) #105

#Good estimates of fecundity - including 0s
HR_prop2b <- HR_FitP %>% 
  filter(est_fecundity >= 0) %>% #NOT the same as (ALL_closed_fruit != 0 & ALL_filled_seeds >=0)!!!!!
  summarize(real_fecdunity = n()) #688

#Good estimates of fecundity - excluding 0s
HR_prop2c <- HR_FitP %>% 
  filter(est_fecundity > 0) %>% 
  summarize(pos_fecundity = n()) #505

HR_prop2d <- HR_FitP %>% 
  filter(ALL_closed_fruit == 0 & ALL_open_fruit >= 1) %>% #should be the SAME as (ALL_closed_fruit == 0 & ALL_open_fruit >= 1)
  summarize(fecundity_NA = n()) #92

#Proportion of Fitness plants with 0 filled seeds but 1 or more closed fruits
#To be combined with others
HR_prop3 <- HR_FitP %>% 
  filter(ALL_closed_fruit != 0 & ALL_filled_seeds == 0) %>% 
  summarize(pop = "HR",
            N_fitness_plants = 780,
            real_fecundity = 688/780,
            pos_fecundity = 505/780,
            zero_closed_fruits = 197/780,
            zero_total_fruits = 105/780,
            fecundity_NA = 92/780,
            zero_seeds_from_nonzero_closed = n()/780) #78

```



Combine summary from each pop
```{r}

fruits_proportions <- rbind(AC_prop3, BB_prop3, BO_prop3, HR_prop3)


```


### Correlations  


Correlations to investigate:
- closed fruits
- total fruits (open fruits)
- filled seeds [fecundity]
- seeds per fruit
- estimated fecundity

**AC**  
```{r}
#Correlation Matrix
png(filename = "corrplot_repro_fitness_AC.png", units = "in", width = 5, height = 5, res = 300)


corrplot(cor(AC_FitP %>% select(ALL_closed_fruit, total_fruits, ALL_filled_seeds, total_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red") #ignores NAs


#Excluding 0s
# corrplot(cor(filter(AC_FitP, est_fecundity != 0) %>% select(ALL_closed_fruit, total_fruits, ALL_filled_seeds, total_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red") #ignores NAs & 0s


dev.off()
```

**BB**  
```{r}
#Correlation Matrix
png(filename = "corrplot_repro_fitness_BB.png", units = "in", width = 5, height = 5, res = 300)


corrplot(cor(BB_FitP %>% select(ALL_closed_fruit, total_fruits, ALL_filled_seeds, total_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red") #ignores NAs


#Excluding 0s
 corrplot(cor(filter(BB_FitP, est_fecundity != 0) %>% select(ALL_closed_fruit, total_fruits, ALL_filled_seeds, total_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red") #ignores NAs & 0s


dev.off()
```

**BO**  
```{r}
#Correlation Matrix
png(filename = "corrplot_repro_fitness_BO.png", units = "in", width = 5, height = 5, res = 300)


corrplot(cor(BO_FitP %>% select(ALL_closed_fruit, total_fruits, ALL_filled_seeds, total_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red") #ignores NAs


#Excluding 0s
 # corrplot(cor(filter(BO_FitP, est_fecundity != 0) %>% select(ALL_closed_fruit, total_fruits, ALL_filled_seeds, total_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red") #ignores NAs & 0s


dev.off()
```



**HR**  
```{r}
#Correlation Matrix
png(filename = "corrplot_repro_fitness_HR.png", units = "in", width = 5, height = 5, res = 300)


corrplot(cor(HR_FitP %>% select(ALL_closed_fruit, total_fruits, ALL_filled_seeds, total_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red") #ignores NAs


#Excluding 0s
 # corrplot(cor(filter(HR_FitP, est_fecundity != 0) %>% select(ALL_closed_fruit, total_fruits, ALL_filled_seeds, total_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red") #ignores NAs & 0s


dev.off()
```
