---
title: "Exploratory Data Analyses"
author: "Devin Gamble"
date: "2022-07-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)

```


**Description** 
This file was created to explore broad trends in our field data from the first year of the experiment (2021-2022). It was also used to explore and select different metrics of fitness (and fecundity) that would be used in later analyses.  


**Load Data**  
```{r message = FALSE, warning = FALSE}

## AC
AC_mainF_Y1 <- read_csv(here::here("data_sheets", "compiled_sheets", "AC_mastersheet_Fitness-mains_2023-10-17.csv")) 

## BB
BB_mainF_Y1 <- read_csv(here::here("data_sheets", "compiled_sheets", "BB_mastersheet_Fitness-mains_2023-10-18.csv"))
## Already has greenhouse cohort data

## BO
BO_mainF_Y1 <- read_csv(here::here("data_sheets", "compiled_sheets", "BO_mastersheet_Fitness-mains_2023-10-18.csv")) 

## HR
HR_mainF_Y1 <- read_csv(here::here("data_sheets", "compiled_sheets", "HR_mastersheet_Fitness-mains_2023-10-18.csv"))


```



### Fitness Plants  


Proportion of *Fitness Plants* that  
- produced 0 closed fruits
- produced 0 closed fruits & 0 open fruits
- produced 1+ closed fruits & 0 filled seeds

In other words... 

**Real fecundity** estimates (Proportion of *Fitness Plants* that  
- produced 0 closed fruits
- produced 0 closed fruits & 0 open fruits
- produced 1+ closed fruits & 0 filled seeds)

which is the summation of 
- Proportion of plants with > 1 closed fruit, and with > 1 filled seeds found in closed fruit. 
- Proportion of plants with 0 fruits (closed or open)
- Proportion of plants with > 1 closed fruits but NO seeds found in these closed fruits. 

And is the complement to
- Proportion of plants with 0 closed fruits but > 1 open fruits (no fecundity estimate)


**AC**
```{r message = FALSE}

#Good estimates of fecundity - including 0s
AC_prop_a <- AC_mainF_Y1 %>% 
  filter(est_fecundity >= 0) %>% #NOT the same as (ALL_closed_fruit != 0 & ALL_filled_seeds >=0)!!!!!
  summarize(real_fecundity = n())  #307

#Good estimates of fecundity (positive) - excluding 0s
AC_prop_b <- AC_mainF_Y1 %>% 
  filter(est_fecundity > 0) %>% 
  summarize(pos_fecundity = n()) #283

#Proportion of Fitness plants with 0 total fruits
AC_prop_c <- AC_mainF_Y1 %>% 
  filter(total_fruits == 0) %>% 
  summarize(zero_total_fruits = n()) #11

# 0 fecundity: >=1 closed fruits but 0 filled seeds
AC_prop_e <- AC_mainF_Y1 %>% 
  filter(closed_fruits >= 1 & filled_seeds == 0) %>% 
  summarize(zero_seeds_from_nonzero_closed = n()) #13


#Fecundity NA | 0 closed with >=1 open fruits
AC_prop_e <- AC_mainF_Y1 %>% 
  filter(is.na(est_fecundity)) %>% #should be the SAME as (closed_fruits == 0 & open_fruits >= 1)
  summarize(fecundity_NA = n()) #12


#Proportion of Fitness plants with 0 closed fruits -- NOT used in ppt slide table
AC_prop_f <- AC_mainF_Y1 %>% 
  filter(closed_fruits == 0) %>% 
  summarize(zero_closed_fruits = n()) #23



#Proportion of Fitness plants with 0 filled seeds but 1 or more closed fruits
AC_prop_sums <- AC_mainF_Y1 %>% 
  summarize(pop = "AC",
            N_fitness_plants = 319,
            real_fecundity = 307/319,
            pos_fecundity = 283/319,
            zero_total_fruits = 11/319,
            zero_seeds_from_nonzero_closed = 13/319, 
            fecundity_NA = 12/319, #Should complement the real_fecundity column
            zero_closed_fruits = 23/319)

#Same order as the columns in ppt slide table

```



**BB**  

```{r}

#Good estimates of fecundity - including 0s
BB_prop_a <- BB_mainF_Y1 %>% 
  filter(est_fecundity >= 0) %>% #NOT the same as (ALL_closed_fruit != 0 & ALL_filled_seeds >=0)!!!!!
  summarize(real_fecundity = n())  #344

#Good estimates of fecundity (positive) - excluding 0s
BB_prop_b <- BB_mainF_Y1 %>% 
  filter(est_fecundity > 0) %>% 
  summarize(pos_fecundity = n()) #113

#Proportion of Fitness plants with 0 total fruits
BB_prop_c <- BB_mainF_Y1 %>% 
  filter(total_fruits == 0) %>% 
  summarize(zero_total_fruits = n()) #194

# 0 fecundity: >=1 closed fruits but 0 filled seeds
BB_prop_e <- BB_mainF_Y1 %>% 
  filter(closed_fruits >= 1 & filled_seeds == 0) %>% 
  summarize(zero_seeds_from_nonzero_closed = n()) #37


#Fecundity NA | 0 closed with >=1 open fruits
BB_prop_e <- BB_mainF_Y1 %>% 
  filter(is.na(est_fecundity)) %>% #should be the SAME as (closed_fruits == 0 & open_fruits >= 1)
  summarize(fecundity_NA = n()) #25


#Proportion of Fitness plants with 0 closed fruits -- NOT used in ppt slide table
BB_prop_f <- BB_mainF_Y1 %>% 
  filter(closed_fruits == 0) %>% 
  summarize(zero_closed_fruits = n()) #



#Proportion of Fitness plants with 0 filled seeds but 1 or more closed fruits
BB_prop_sums <- BB_mainF_Y1 %>% 
  summarize(pop = "BB",
            N_fitness_plants = 369,
            real_fecundity = 344/369,
            pos_fecundity = 113/369,
            zero_total_fruits = 194/369,
            zero_seeds_from_nonzero_closed = 37/369, 
            fecundity_NA = 25/369, #Should complement the real_fecundity column
            zero_closed_fruits = 219/369)

#Same order as the columns in ppt slide table


```



**BO**  

```{r}

#Good estimates of fecundity - including 0s
BO_prop_a <- BO_mainF_Y1 %>% 
  filter(est_fecundity >= 0) %>% #NOT the same as (ALL_closed_fruit != 0 & ALL_filled_seeds >=0)!!!!!
  summarize(real_fecundity = n())  #297

#Good estimates of fecundity (positive) - excluding 0s
BO_prop_b <- BO_mainF_Y1 %>% 
  filter(est_fecundity > 0) %>% 
  summarize(pos_fecundity = n()) #137

#Proportion of Fitness plants with 0 total fruits
BO_prop_c <- BO_mainF_Y1 %>% 
  filter(total_fruits == 0) %>% 
  summarize(zero_total_fruits = n()) #99

# 0 fecundity: >=1 closed fruits but 0 filled seeds
BO_prop_e <- BO_mainF_Y1 %>% 
  filter(closed_fruits >= 1 & filled_seeds == 0) %>% 
  summarize(zero_seeds_from_nonzero_closed = n()) #61


#Fecundity NA | 0 closed with >=1 open fruits
BO_prop_e <- BO_mainF_Y1 %>% 
  filter(is.na(est_fecundity)) %>% #should be the SAME as (closed_fruits == 0 & open_fruits >= 1)
  summarize(fecundity_NA = n()) #42


#Proportion of Fitness plants with 0 closed fruits -- NOT used in ppt slide table
BO_prop_f <- BO_mainF_Y1 %>% 
  filter(closed_fruits == 0) %>% 
  summarize(zero_closed_fruits = n()) #



#Proportion of Fitness plants with 0 filled seeds but 1 or more closed fruits
BO_prop_sums <- BO_mainF_Y1 %>% 
  summarize(pop = "BO",
            N_fitness_plants = 339,
            real_fecundity = 297/339,
            pos_fecundity = 137/339,
            zero_total_fruits = 99/339,
            zero_seeds_from_nonzero_closed = 61/339, 
            fecundity_NA = 42/339, #Should complement the real_fecundity column
            zero_closed_fruits = 141/339)

#Same order as the columns in ppt slide table


```




**HR**  

```{r}

#Good estimates of fecundity - including 0s
HR_prop_a <- HR_mainF_Y1 %>% 
  filter(est_fecundity >= 0) %>% #NOT the same as (ALL_closed_fruit != 0 & ALL_filled_seeds >=0)!!!!!
  summarize(real_fecundity = n())  #403

#Good estimates of fecundity (positive) - excluding 0s
HR_prop_b <- HR_mainF_Y1 %>% 
  filter(est_fecundity > 0) %>% 
  summarize(pos_fecundity = n()) #344

#Proportion of Fitness plants with 0 total fruits
HR_prop_c <- HR_mainF_Y1 %>% 
  filter(total_fruits == 0) %>% 
  summarize(zero_total_fruits = n()) #31

# 0 fecundity: >=1 closed fruits but 0 filled seeds
HR_prop_e <- HR_mainF_Y1 %>% 
  filter(closed_fruits >= 1 & filled_seeds == 0) %>% 
  summarize(zero_seeds_from_nonzero_closed = n()) #28


#Fecundity NA | 0 closed with >=1 open fruits
HR_prop_e <- HR_mainF_Y1 %>% 
  filter(is.na(est_fecundity)) %>% #should be the SAME as (closed_fruits == 0 & open_fruits >= 1)
  summarize(fecundity_NA = n()) #3


#Proportion of Fitness plants with 0 closed fruits -- NOT used in ppt slide table
HR_prop_f <- HR_mainF_Y1 %>% 
  filter(closed_fruits == 0) %>% 
  summarize(zero_closed_fruits = n()) #



#Proportion of Fitness plants with 0 filled seeds but 1 or more closed fruits
HR_prop_sums <- HR_mainF_Y1 %>% 
  summarize(pop = "HR",
            N_fitness_plants = 406,
            real_fecundity = 403/406,
            pos_fecundity = 344/406,
            zero_total_fruits = 31/406,
            zero_seeds_from_nonzero_closed = 28/406, 
            fecundity_NA = 3/406, #Should complement the real_fecundity column
            zero_closed_fruits = 34/406)

#Same order as the columns in ppt slide table


```

Combine summary from each pop
```{r}

fruits_proportions <- rbind(AC_prop_sums, BB_prop_sums, BO_prop_sums, HR_prop_sums)

```

<br>  


### Correlations  


Correlations to investigate:
- closed fruits
- total fruits (open fruits)
- filled seeds [fecundity]
- seeds per fruit
- estimated fecundity


```{r}
library(corrplot)
#Correlation Matrix

#png(filename = "corrplot_repro_fitness_AC.png", units = "in", width = 5, height = 5, res = 300)

## AC
corrplot(cor(AC_mainF_Y1 %>% select(closed_fruits, total_fruits, filled_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red", title = "AC Y1") #ignores NAs


#Excluding 0s
# corrplot(cor(filter(AC_FitP, est_fecundity != 0) %>% select(closed_fruits, total_fruits, filled_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red") #ignores NAs & 0s


#dev.off()



##BB
corrplot(cor(BB_mainF_Y1 %>% select(closed_fruits, total_fruits, filled_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red", title = "BB Y1") #ignores NAs


##BO
corrplot(cor(BO_mainF_Y1 %>% select(closed_fruits, total_fruits, filled_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red", title = "BO Y1") #ignores NAs


##HR
corrplot(cor(HR_mainF_Y1 %>% select(closed_fruits, total_fruits, filled_seeds, est_fecundity, mean_seeds_per_fruit), use="complete.obs"), method = "color", addCoef.col = "black", number.cex = .6, outline = TRUE, tl.col = "red", title = "HR Y1") #ignores NAs

```





```{r}
###NOTES 

############ Add fitness measure: Fecundity (total seed number) x mean seed mass.......

## REVIEW: mean seeds per fruit & estimated seed production are still 0s when seed_ct = NA...

#long data set
#Fitness data set


#Total number of fruits may be better than seed # - less downward bias (due to seeds lost after dehiscence)

#Typo in 'seed_count' - there are 0s that should be NAs when closed fruit == 0


#To-Do
# - single 'Fitness' data sheet with all variables required for analysis / visualizations!
# - calculate fitness metrics based on the master sheet first? includes sample sizes in df (or, merge back in the fitness sheets below..)
# - correlations among all fitness measures
#NAs vs 0s: when there were 0 closed fruit (i.e. no viable seeds collected), but possibly non-zero total fruits (open, etc.)??? Included as 0s in data, but may decide to remove

#1 or more fruits but 0 seeds --> 0 fecundity (for est seed prod)



#
# Sample Mat fams equally for seeds, but account for differences in individual Mat fam/segment fitness later on in aster
# In aster, include fruit number and/or fecundity (may be incomplete wieth 0 clsoed fruits) as hierarchical fitness estimates

```

<br>  
















------------------------------------------------------

**OLD CODE**  
The following chunks retained only to illustrate preliminary exploration of the fitness data and how best to treat it.

## Fitness Estimates  

NOTE: Use `surv_to_x` variables to delimit groups in calculating certain fitness components; NAs were originally changed to 0s for seed counts/means when 0 closed fruit were collected, though some of these instances occur for plants that did not survive to produce any fruit (i.e, total fruits = 0).  


##### Combined Fitness (Survival to flowering x Estimated seed production)

**AC**  

```{r}

#Survivorship of individual segments to flowering (maintain block)
Fitness_AC_df1 <- AC_master_df1 %>% group_by(Block, Transect, Donor, Recipient) %>% 
  summarize(
    P_surv_flower = mean(surv_to_flower)) %>% 
  filter(Block != "W") #Remove transects W1 and W2


Fitness_AC_df2 <- AC_master_df1 %>% group_by(Block, Transect, Donor, Recipient) %>%
  filter(FitP == TRUE) %>% 
  summarize(
    est_seed_prod = total_est_seed_production,
    total_fruit_prod = total_fruits_F,
    closed_fruits = closed_fruits_F,
    seed_count = seed_ct
  ) %>% ungroup() %>% 
  mutate(est_seed_prod = case_when(closed_fruits == 0 ~ as.numeric(as.character(NA)), 
                                   TRUE ~ est_seed_prod)) #option: set est_seed_prod to NA when 0 closed fruits were gathered (mean seeds per fruit not reliably estimated)
### This will introduce NAs below when before there were 0s for combined fitness (surv x est seed prod). Means in summarize() set to na.rm = TRUE to ignore these values (e.g., averages across 2 segments instead of 3...)

###
# REVISIT: Should these values be NAs or 0s?????
###




#Two Fewer rows than the summarization of P_surv_flower - possibly because two segments never had fitness plants chosen


Fitness_AC_df3 <- merge(Fitness_AC_df2, Fitness_AC_df1, by = c("Block", "Transect", "Donor", "Recipient")) %>% 
  mutate(FIT_surv_x_seedprod = P_surv_flower*est_seed_prod,
         FIT_surv_x_fruitprod = P_surv_flower*total_fruit_prod)


#Calculate Block means of fitness variables:
Fitness_AC_blockmeans <- Fitness_AC_df3 %>% group_by(Block) %>% 
  summarize(
    BLOCKmean_FIT_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    BLOCKmean_FIT_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    BLOCKmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    BLOCKmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    BLOCKmean_seed_count = mean(seed_count, na.rm = TRUE),
    BLOCKmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE)
  ) #NA values removed from these averages


#Calculate relative fitness measures for each individual segment (Block means = 1)
#Retain NAs for this df; individual segments may have produced 0 closed fruit
Fitness_AC_df3.5 <- merge(Fitness_AC_df3, Fitness_AC_blockmeans, by = "Block") %>% 
  mutate(
    Rel_FIT_surv_x_seedprod = FIT_surv_x_seedprod/BLOCKmean_FIT_surv_x_seedprod,
    Rel_FIT_surv_x_fruitprod = FIT_surv_x_fruitprod/BLOCKmean_FIT_surv_x_fruitprod,
    Rel_est_seed_prod = est_seed_prod/BLOCKmean_est_seed_prod,
    Rel_total_fruit_prod = total_fruit_prod/BLOCKmean_total_fruit_prod,
    Rel_seed_count = seed_count/BLOCKmean_seed_count,
    Rel_P_surv_flower = P_surv_flower/BLOCKmean_P_surv_flower,
  ) 


# Average across blocks - MATERNAL mean fitness estimates
Fitness_AC_df4 <- Fitness_AC_df3.5 %>% group_by(Donor, Recipient) %>% 
  summarize(
    FIT_MATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    #FIT_MATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    MATmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    MATmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    MATmean_seed_count = mean(seed_count, na.rm = TRUE),
    MATmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE), #Relative measures of fitness
    Rel_FIT_MATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod, na.rm = TRUE),
    #Rel_FIT_MATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod, na.rm = TRUE),
    Rel_MATmean_est_seed_prod = mean(Rel_est_seed_prod, na.rm = TRUE),
    Rel_MATmean_total_fruit_prod = mean(Rel_total_fruit_prod, na.rm = TRUE),
    Rel_MATmean_seed_count = mean(Rel_seed_count, na.rm = TRUE),
    #Rel_MATmean_P_surv_flower = mean(Rel_P_surv_flower, na.rm = TRUE),
    Total_seed_count_F = sum(seed_count, na.rm = TRUE) #Seed totals summed for each maternal family
    )


#Average across recipients - PATERNAL mean fitness estimates
Fitness_AC_df5 <- Fitness_AC_df3.5 %>% group_by(Donor) %>% 
  summarize(
    FIT_PATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    #FIT_PATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    PATmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    PATmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    PATmean_seed_count = mean(seed_count, na.rm = TRUE),
    PATmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE),
    Rel_FIT_PATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod, na.rm = TRUE),
    #Rel_FIT_PATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod, na.rm = TRUE),
    Rel_PATmean_est_seed_prod = mean(Rel_est_seed_prod, na.rm = TRUE),
    Rel_PATmean_total_fruit_prod = mean(Rel_total_fruit_prod, na.rm = TRUE),
    Rel_PATmean_seed_count = mean(Rel_seed_count, na.rm = TRUE),
    #Rel_PATmean_P_surv_flower = mean(Rel_P_surv_flower, na.rm = TRUE), survival probably doesn't need to be relative
    Total_seed_count_pat = sum(seed_count, na.rm = TRUE)
  )


Maternal_Paternal_Fitness <- merge(Fitness_AC_df4, Fitness_AC_df5, by = "Donor")



#Save a fitness-estimate data sheet
#Fitness components: mean survivorship, mean total fruit production, estimated total seed production (relative), Combined fitness (relative), total seed count

AC_FitSummary_1 <- Maternal_Paternal_Fitness %>% 
  select(Donor, Recipient, MATmean_P_surv_flower, MATmean_total_fruit_prod, Rel_MATmean_est_seed_prod, Rel_FIT_MATmean_surv_x_seedprod, Total_seed_count_F) %>% 
  rename(mean_surv_to_flowering = MATmean_P_surv_flower,
         mean_fruit_prod = MATmean_total_fruit_prod,
         mean_est_seed_prod_R =  Rel_MATmean_est_seed_prod,
         mean_surv_x_est_seed_prod_R = Rel_FIT_MATmean_surv_x_seedprod,
         total_seed_count = Total_seed_count_F)


#Include total seed count for every mat fam (and paternal fam)

write_csv(AC_FitSummary_1, here::here("data_sheets", "compiled_sheets", "fitness_summary_AC1.csv"))
#Data for main fitness plants only


```


<br>  


#### Visualizations    

```{r eval = FALSE}
######
### Data frame organization: 
######

# Segment Fitness: Fitness_AC_df3.5
# (Includes relative fitness measures)

# Maternal Fitness: Fitness_AC_df4

# Paternal Fitness: Fitness_AC_df5
```


**Fitness = Total Fruit Production**  

```{r}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,70)) +
  ylab("Count") +
  xlab("Segment Fruit Production (N=319)")

## Individual Segment fitness (Relative)
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,70)) +
  ylab("Count") +
  xlab("Segment Relative Fruit Production (N=319)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_total_fruit_prod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,70)) +
  ylab("Count") +
  xlab("Segment Relative Fruit Production (N=319) (By Block)")




## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = MATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,20)) +
  ylab("Count") +
  xlab("Avg. Maternal Fruit Production (N=107)")


## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_MATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,20)) +
  ylab("Count") +
  xlab("Avg. Relative Maternal Fruit Production (N=107)")



## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = PATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Fruit Production (N=40)")

## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_PATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Relative Paternal Fruit Production (N=40)")

```


**Fitness = Estimated Seed Production**  

NOTE: From here on, only fitness estimates relative to Block means are presented.  

```{r echo = FALSE}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_est_seed_prod), fill = "#A58AFF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Estimated Seed Production (N=319) (Relative)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_est_seed_prod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Estimated Seed Production (N=319) (Relative)")



## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_MATmean_est_seed_prod), fill = "#A58AFF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  ylab("Count") +
  xlab("Avg. Maternal Estimated Seed Production (N=107) (Relative)")


## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_PATmean_est_seed_prod), fill = "#A58AFF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Estimated Seed Production (N=40) (Relative)")

```



**Fitness = Survival x Estimated Seed Production**  

```{r echo = FALSE}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_seedprod), fill = "#00BE67", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Survival x Estimated Seed Production (N=319) (Relative)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_seedprod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("egment Survival x Estimated Seed Production (N=319) (Relative)")



## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_FIT_MATmean_surv_x_seedprod), fill = "#00BE67", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  ylab("Count") +
  xlab("Avg. Maternal Estimated Seed Production (N=107) (Relative)")


## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_FIT_PATmean_surv_x_seedprod), fill = "#00BE67", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Estimated Seed Production (N=40) (Relative)")

```

**Fitness = Survival x Total Fruit Production**  

```{r echo = FALSE}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_fruitprod), fill = "#FFA500", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Survival x Estimated fruit Production (N=319) (Relative)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_fruitprod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("egment Survival x Estimated fruit Production (N=319) (Relative)")
```


```{r eval = FALSE, echo = FALSE}
##These measures not estimated - commented out above

## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_FIT_MATmean_surv_x_fruitprod), fill = "#FFA500", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  ylab("Count") +
  xlab("Avg. Maternal Estimated fruit Production (N=107) (Relative)")


## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_FIT_PATmean_surv_x_fruitprod), fill = "#FFA500", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Estimated fruit Production (N=40) (Relative)")

```

<br>  


## Exploratory Data Analyses  

Phenotypic Trait and Fitness data from the 2021-2022 Nemophila Field Experiment  

Explore:
- Block effects (on fruit prod, fecundity, fitness)


**Variation in Fitness by Genotype with vs. without 0s**:  

**AC**:  

```{r}

#New df for 0s removed, only a couple fitness measures included relative to AC_df4
FITAC_no_0s <- filter(Fitness_AC_df3.5, closed_fruits > 0) %>% group_by(Donor, Recipient) %>% 
  summarize(
    MATmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    MATmean_seed_count = mean(seed_count, na.rm = TRUE),
    MATmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE), #Relative measures of fitness
    Rel_FIT_MATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod, na.rm = TRUE),
    Rel_MATmean_est_seed_prod = mean(Rel_est_seed_prod, na.rm = TRUE),
    Rel_MATmean_total_fruit_prod = mean(Rel_total_fruit_prod, na.rm = TRUE),
    Rel_MATmean_seed_count = mean(Rel_seed_count, na.rm = TRUE),
    Total_seed_count_F = sum(seed_count, na.rm = TRUE) #Seed totals summed for each maternal family
    ) 
  

ggplot() +
  geom_density(data = Fitness_AC_df4, aes(x = Rel_MATmean_total_fruit_prod), fill = "#A58AFF", color = "black", alpha = 0.3) + 
  geom_density(data = FITAC_no_0s, aes(x = Rel_MATmean_total_fruit_prod), fill = "orange", color = "black", alpha = 0.3) +
  theme_classic() +
  #scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  ylab("Count") +
  xlab("Avg. Maternal Total Fruit Production (N=107) (Relative)")

ggplot() +
  geom_histogram(data = Fitness_AC_df4, aes(x = Rel_MATmean_total_fruit_prod), fill = "#A58AFF", color = "black", alpha = 0.3) + 
  geom_histogram(data = FITAC_no_0s, aes(x = Rel_MATmean_total_fruit_prod), fill = "orange", color = "black", alpha = 0.3) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  ylab("Count") +
  xlab("Avg. Maternal Total Fruit Production (N=107) (Relative)")


#Estimated Seed production not affected by inclusion/exclusion of 0s - due to no change in seed_ct = 0 when only >1 closed fruits producers included.


```


Scatter plots:  

```{r}
AC_master_Fonly_1 <- read_csv(here::here("data_sheets", "compiled_sheets", "AC_mastersheet_v1_Fitness-only_2022-07-26.csv"))


#bivar plot for total fruit No. (open + clsoed) ~ total seed number (include sample sizes for each)

ggplot(data = AC_master_Fonly_1) +
  geom_point(aes(x = total_fruits_F, y = seed_ct)) + 
  theme_bw() +
  xlab("Total Fruit Production (open + closed) (N = 319)") +
  ylab("Seed Count")


ggplot(data = AC_master_Fonly_1) +
  geom_point(aes(x = closed_fruits_F, y = seed_ct)) + 
  theme_bw() +
  xlab("Total Closed Fruits (N = 319)") +
  ylab("Seed Count")
```



