---
title: "Exploratory Data Analyses"
author: "Devin Gamble"
date: "2022-07-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)

```


```{r}
###NOTES 



## REVIEW: mean seeds per fruit & estimated seed production are still 0s when seed_ct = NA...

#long data set
#Fitness data set


#Total number of fruits may be better than seed # - less downward bias (due to seeds lost after dehiscence)



#To-Do
# - single 'Fitness' data sheet with all variables required for analysis / visualizations!
# - calculate fitness metrics based on the master sheet first? includes sample sizes in df (or, merge back in the fitness sheets below..)
# - correlations among all fitness measures
#NAs vs 0s: when there were 0 closed fruit (i.e. no viable seeds collected), but possibly non-zero total fruits (open, etc.)??? Included as 0s in data, but may decide to remove

#1 or more fruits but 0 seeds --> 0 fecundity (for est seed prod)



#
# Sample Mat fams equally for seeds, but account for differences in individual Mat fam/segment fitness later on in aster
# In aster, include fruit number and/or fecundity (may be incomplete wieth 0 clsoed fruits) as hierarchical fitness estimates

```

**Finalized Fitness Measures**  
Specifically for estimating how many seeds two sow in year 2 of the field experiment.  

- Estimated Total Seed Production (average seeds per fruit x total fruits)  
- Combined Fitness (Estimated seed production x sibship [segment] survival to flowering)  
- Total Number of Fruits  



Load master sheets:  

`Co.` referes to the greenhouse cohort of plants used to generate G1 seeds. Cohort data is merged in below for sites where it was not tracked (AC, BO, HR). 
```{r message = FALSE}
#2021 Tray position data sheets - extract and merge in greenhouse cohort data:
AC_tray21 <- read_csv(here::here("data_sheets", "2021_tray_positions", "AC_tray_rand.csv"), skip = 1) %>% 
  select(Cohort, Donor_ID, Recipient_ID) %>% #Block and Transect not necessary to save here
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient_ID)))) %>% #Fix Recipient/Donor names
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor_ID)))) %>% 
  mutate(Donor = as.character(str_c("AC_", Donor_num))) %>%
  select(-c(Donor_num, Recipient_num, Donor_ID, Recipient_ID)) %>% 
  distinct() #remove duplicate cases

BO_tray21 <- read_csv(here::here("data_sheets", "2021_tray_positions", "BO_tray_rand.csv"), skip = 1) %>% 
  select(Cohort, Donor_ID, Recipient_ID) %>% #Block and Transect not necessary to save here
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient_ID)))) %>% #Fix Recipient/Donor names
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor_ID)))) %>% 
  mutate(Donor = as.character(str_c("BO_", Donor_num))) %>%
  select(-c(Donor_num, Recipient_num, Donor_ID, Recipient_ID)) %>% 
  distinct()

HR_tray21 <- read_csv(here::here("data_sheets", "2021_tray_positions", "HR_tray_rand.csv"), skip = 1) %>% 
  select(Cohort, Donor_ID, Recipient_ID) %>% #Block and Transect not necessary to save here
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient_ID)))) %>% #Fix Recipient/Donor names
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor_ID)))) %>% 
  mutate(Donor = as.character(str_c("HR_", Donor_num))) %>%
  select(-c(Donor_num, Recipient_num, Donor_ID, Recipient_ID)) %>% 
  distinct()





## AC
AC_master_df <- read_csv(here::here("data_sheets", "compiled_sheets", "AC_mastersheet_v1_2022-08-18.csv")) #Parsing issue with 6-22 flowering date
AC_master_df1 <- left_join(AC_master_df, AC_tray21, by = c("Donor", "Recipient")) %>% 
  rename(Co. = Cohort)


## BB
BB_master_df1 <- read_csv(here::here("data_sheets", "compiled_sheets", "BB_mastersheet_v1_2022-08-18.csv"))
#No watered points present
## Already has greenhouse cohort data


## BO
BO_master_df <- read_csv(here::here("data_sheets", "compiled_sheets", "BO_mastersheet_v1_2022-08-18.csv")) #Parsing issue with Fl_3.11 flowering date
BO_master_df1 <- left_join(BO_master_df, BO_tray21, by = c("Donor", "Recipient")) %>% 
  rename(Co. = Cohort)


## HR
HR_master_df <- read_csv(here::here("data_sheets", "compiled_sheets", "HR_mastersheet_v1_2022-08-18.csv"))
HR_master_df1 <- left_join(HR_master_df, HR_tray21, by = c("Donor", "Recipient")) %>% 
  rename(Co. = Cohort)

```


## Fitness Estimates  


NOTE: Use `surv_to_x` variables to delimit groups in calculating certain fitness components; NAs were originally changed to 0s for seed counts/means when 0 closed fruit were collected, though some of these instances occur for plants that did not survive to produce any fruit (i.e, total fruits = 0).  



##### Combined Fitness (Survival to flowering x Estimated seed production)

**AC**  

```{r}

#Survivorship of individual segments to flowering (maintain block)
Fitness_AC_df1 <- AC_master_df1 %>% group_by(Block, Transect, Donor, Recipient) %>% 
  summarize(
    P_surv_flower = mean(surv_to_flower)) %>% 
  filter(Block != "W") #Remove transects W1 and W2


Fitness_AC_df2 <- AC_master_df1 %>% group_by(Block, Transect, Donor, Recipient) %>%
  filter(FitP == TRUE) %>% 
  summarize(
    est_seed_prod = total_est_seed_production,
    total_fruit_prod = total_fruits_F,
    closed_fruits = closed_fruits_F,
    seed_count = seed_ct
  ) %>% ungroup() %>% 
  mutate(est_seed_prod = case_when(closed_fruits == 0 ~ as.numeric(as.character(NA)), 
                                   TRUE ~ est_seed_prod)) #option: set est_seed_prod to NA when 0 closed fruits were gathered (mean seeds per fruit not reliably estimated)
### This will introduce NAs below when before there were 0s for combined fitness (surv x est seed prod). Means in summarize() set to na.rm = TRUE to ignore these values (e.g., averages across 2 segments instead of 3...)

###
# REVISIT: Should these values be NAs or 0s?????
###




#Two Fewer rows than the summarization of P_surv_flower - possibly because two segments never had fitness plants chosen


Fitness_AC_df3 <- merge(Fitness_AC_df2, Fitness_AC_df1, by = c("Block", "Transect", "Donor", "Recipient")) %>% 
  mutate(FIT_surv_x_seedprod = P_surv_flower*est_seed_prod,
         FIT_surv_x_fruitprod = P_surv_flower*total_fruit_prod)


#Calculate Block means of fitness variables:
Fitness_AC_blockmeans <- Fitness_AC_df3 %>% group_by(Block) %>% 
  summarize(
    BLOCKmean_FIT_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    BLOCKmean_FIT_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    BLOCKmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    BLOCKmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    BLOCKmean_seed_count = mean(seed_count, na.rm = TRUE),
    BLOCKmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE)
  ) #NA values removed from these averages


#Calculate relative fitness measures for each individual segment (Block means = 1)
#Retain NAs for this df; individual segments may have produced 0 closed fruit
Fitness_AC_df3.5 <- merge(Fitness_AC_df3, Fitness_AC_blockmeans, by = "Block") %>% 
  mutate(
    Rel_FIT_surv_x_seedprod = FIT_surv_x_seedprod/BLOCKmean_FIT_surv_x_seedprod,
    Rel_FIT_surv_x_fruitprod = FIT_surv_x_fruitprod/BLOCKmean_FIT_surv_x_fruitprod,
    Rel_est_seed_prod = est_seed_prod/BLOCKmean_est_seed_prod,
    Rel_total_fruit_prod = total_fruit_prod/BLOCKmean_total_fruit_prod,
    Rel_seed_count = seed_count/BLOCKmean_seed_count,
    Rel_P_surv_flower = P_surv_flower/BLOCKmean_P_surv_flower,
  ) 


# Average across blocks - MATERNAL mean fitness estimates
Fitness_AC_df4 <- Fitness_AC_df3.5 %>% group_by(Donor, Recipient) %>% 
  summarize(
    FIT_MATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    #FIT_MATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    MATmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    MATmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    MATmean_seed_count = mean(seed_count, na.rm = TRUE),
    MATmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE), #Relative measures of fitness
    Rel_FIT_MATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod, na.rm = TRUE),
    #Rel_FIT_MATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod, na.rm = TRUE),
    Rel_MATmean_est_seed_prod = mean(Rel_est_seed_prod, na.rm = TRUE),
    Rel_MATmean_total_fruit_prod = mean(Rel_total_fruit_prod, na.rm = TRUE),
    Rel_MATmean_seed_count = mean(Rel_seed_count, na.rm = TRUE),
    #Rel_MATmean_P_surv_flower = mean(Rel_P_surv_flower, na.rm = TRUE),
    Total_seed_count_F = sum(seed_count, na.rm = TRUE) #Seed totals summed for each maternal family
    )


#Average across recipients - PATERNAL mean fitness estimates
Fitness_AC_df5 <- Fitness_AC_df3.5 %>% group_by(Donor) %>% 
  summarize(
    FIT_PATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    #FIT_PATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    PATmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    PATmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    PATmean_seed_count = mean(seed_count, na.rm = TRUE),
    PATmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE),
    Rel_FIT_PATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod, na.rm = TRUE),
    #Rel_FIT_PATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod, na.rm = TRUE),
    Rel_PATmean_est_seed_prod = mean(Rel_est_seed_prod, na.rm = TRUE),
    Rel_PATmean_total_fruit_prod = mean(Rel_total_fruit_prod, na.rm = TRUE),
    Rel_PATmean_seed_count = mean(Rel_seed_count, na.rm = TRUE),
    #Rel_PATmean_P_surv_flower = mean(Rel_P_surv_flower, na.rm = TRUE), survival probably doesn't need to be relative
    Total_seed_count_pat = sum(seed_count, na.rm = TRUE)
  )


Maternal_Paternal_Fitness <- merge(Fitness_AC_df4, Fitness_AC_df5, by = "Donor")



#Save a fitness-estimate data sheet
#Fitness components: mean survivorship, mean total fruit production, estimated total seed production (relative), Combined fitness (relative), total seed count

AC_FitSummary_1 <- Maternal_Paternal_Fitness %>% 
  select(Donor, Recipient, MATmean_P_surv_flower, MATmean_total_fruit_prod, Rel_MATmean_est_seed_prod, Rel_FIT_MATmean_surv_x_seedprod, Total_seed_count_F) %>% 
  rename(mean_surv_to_flowering = MATmean_P_surv_flower,
         mean_fruit_prod = MATmean_total_fruit_prod,
         mean_est_seed_prod_R =  Rel_MATmean_est_seed_prod,
         mean_surv_x_est_seed_prod_R = Rel_FIT_MATmean_surv_x_seedprod,
         total_seed_count = Total_seed_count_F)


#Include total seed count for every mat fam (and paternal fam)

write_csv(AC_FitSummary_1, here::here("data_sheets", "compiled_sheets", "fitness_summary_AC1.csv"))
#Data for main fitness plants only


```


**BB**  

```{r}
#Survivorship of individual segments to flowering (maintain block)

Fitness_BB_df1 <- BB_master_df1 %>% group_by(Block, Location, Donor, Recipient) %>% 
  summarize(
    P_surv_flower = mean(surv_to_flower)) #%>% 
  #filter(Block != "W") #Watered points already removed


Fitness_BB_df2 <- BB_master_df1 %>% group_by(Block, Location, Donor, Recipient) %>%
  filter(FitP == TRUE) %>% 
  summarize(
    est_seed_prod = total_est_seed_production,
    total_fruit_prod = total_fruits_F,
    closed_fruits = closed_fruits_F,
    seed_count = seed_ct
  ) %>% ungroup() %>% 
  mutate(est_seed_prod = case_when(closed_fruits == 0 ~ as.numeric(as.character(NA)), 
                                   TRUE ~ est_seed_prod)) #option: set est_seed_prod to NA when 0 closed fruits were gathered (mean seeds per fruit not reliably estimated)
### This will introduce NAs below when before there were 0s for combined fitness (surv x est seed prod)

###
# REVISIT: Should these values be NAs or 0s?????
###

## 71 fewer rows than survival to flowering df



Fitness_BB_df3 <- merge(Fitness_BB_df2, Fitness_BB_df1, by = c("Block", "Location", "Donor", "Recipient")) %>% 
  mutate(FIT_surv_x_seedprod = P_surv_flower*est_seed_prod,
         FIT_surv_x_fruitprod = P_surv_flower*total_fruit_prod)


#Calculate Block means of fitness variables:
Fitness_BB_blockmeans <- Fitness_BB_df3 %>% group_by(Block) %>% 
  summarize(
    BLOCKmean_FIT_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    BLOCKmean_FIT_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    BLOCKmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    BLOCKmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    BLOCKmean_seed_count = mean(seed_count, na.rm = TRUE),
    BLOCKmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE)
  ) #NA values removed from these averages


#Calculate relative fitness measures for each individual segment (Block means = 1)
#Retain NAs for this df; individual segments may have produced 0 closed fruit
Fitness_BB_df3.5 <- merge(Fitness_BB_df3, Fitness_BB_blockmeans, by = "Block") %>% 
  mutate(
    Rel_FIT_surv_x_seedprod = FIT_surv_x_seedprod/BLOCKmean_FIT_surv_x_seedprod,
    Rel_FIT_surv_x_fruitprod = FIT_surv_x_fruitprod/BLOCKmean_FIT_surv_x_fruitprod,
    Rel_est_seed_prod = est_seed_prod/BLOCKmean_est_seed_prod,
    Rel_total_fruit_prod = total_fruit_prod/BLOCKmean_total_fruit_prod,
    Rel_seed_count = seed_count/BLOCKmean_seed_count,
    Rel_P_surv_flower = P_surv_flower/BLOCKmean_P_surv_flower,
  )


# Average across blocks - MATERNAL mean fitness estimates
Fitness_BB_df4 <- Fitness_BB_df3.5 %>% group_by(Donor, Recipient) %>% 
  summarize(
    FIT_MATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    #FIT_MATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    MATmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    MATmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    MATmean_seed_count = mean(seed_count, na.rm = TRUE),
    MATmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE), #Relative measures of fitness
    Rel_FIT_MATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod, na.rm = TRUE),
    #Rel_FIT_MATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod, na.rm = TRUE),
    Rel_MATmean_est_seed_prod = mean(Rel_est_seed_prod, na.rm = TRUE),
    Rel_MATmean_total_fruit_prod = mean(Rel_total_fruit_prod, na.rm = TRUE),
    Rel_MATmean_seed_count = mean(Rel_seed_count, na.rm = TRUE),
    #Rel_MATmean_P_surv_flower = mean(Rel_P_surv_flower, na.rm = TRUE),
    Total_seed_count_F = sum(seed_count, na.rm = TRUE) #Seed totals summed for each maternal family
    )


#Average across recipients - PATERNAL mean fitness estimates
Fitness_BB_df5 <- Fitness_BB_df3.5 %>% group_by(Donor) %>% 
  summarize(
    FIT_PATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    #FIT_PATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    PATmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    PATmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    PATmean_seed_count = mean(seed_count, na.rm = TRUE),
    PATmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE),
    Rel_FIT_PATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod, na.rm = TRUE),
    #Rel_FIT_PATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod, na.rm = TRUE),
    Rel_PATmean_est_seed_prod = mean(Rel_est_seed_prod, na.rm = TRUE),
    Rel_PATmean_total_fruit_prod = mean(Rel_total_fruit_prod, na.rm = TRUE),
    Rel_PATmean_seed_count = mean(Rel_seed_count, na.rm = TRUE),
    #Rel_PATmean_P_surv_flower = mean(Rel_P_surv_flower, na.rm = TRUE), survival probably doesn't need to be relative
    Total_seed_count_pat = sum(seed_count, na.rm = TRUE)
  )


Maternal_Paternal_Fitness_BB <- merge(Fitness_BB_df4, Fitness_BB_df5, by = "Donor")



#Save a fitness-estimate data sheet
#Fitness components: mean survivorship, mean total fruit production, estimated total seed production (relative), Combined fitness (relative), total seed count

BB_FitSummary_1 <- Maternal_Paternal_Fitness_BB %>% 
  select(Donor, Recipient, MATmean_P_surv_flower, MATmean_total_fruit_prod, Rel_MATmean_est_seed_prod, Rel_FIT_MATmean_surv_x_seedprod, Total_seed_count_F) %>% 
  rename(mean_surv_to_flowering = MATmean_P_surv_flower,
         mean_fruit_prod = MATmean_total_fruit_prod,
         mean_est_seed_prod_R =  Rel_MATmean_est_seed_prod,
         mean_surv_x_est_seed_prod_R = Rel_FIT_MATmean_surv_x_seedprod,
         total_seed_count = Total_seed_count_F)


#Include total seed count for every mat fam (and paternal fam)

write_csv(BB_FitSummary_1, here::here("data_sheets", "compiled_sheets", "fitness_summary_BB1.csv"))

```


**BO**  

```{r}
#Survivorship of individual segments to flowering (maintain block)

Fitness_BO_df1 <- BO_master_df1 %>% group_by(Block, Transect, Donor, Recipient) %>% 
  summarize(
    P_surv_flower = mean(surv_to_flower)) #%>% 
  #filter(Block != "W") #Watered points already removed


Fitness_BO_df2 <- BO_master_df1 %>% group_by(Block, Transect, Donor, Recipient) %>%
  filter(FitP == TRUE) %>% 
  summarize(
    est_seed_prod = total_est_seed_production,
    total_fruit_prod = total_fruits_F,
    closed_fruits = closed_fruits_F,
    seed_count = seed_ct
  ) %>% ungroup() %>% 
  mutate(est_seed_prod = case_when(closed_fruits == 0 ~ as.numeric(as.character(NA)), 
                                   TRUE ~ est_seed_prod)) #option: set est_seed_prod to NA when 0 closed fruits were gathered (mean seeds per fruit not reliably estimated)
### This will introduce NAs below when before there were 0s for combined fitness (surv x est seed prod)

###
# REVISIT: Should these values be NAs or 0s?????
###

## 71 fewer rows than survival to flowering df



Fitness_BO_df3 <- merge(Fitness_BO_df2, Fitness_BO_df1, by = c("Block", "Transect", "Donor", "Recipient")) %>% 
  mutate(FIT_surv_x_seedprod = P_surv_flower*est_seed_prod,
         FIT_surv_x_fruitprod = P_surv_flower*total_fruit_prod)


#Calculate Block means of fitness variables:
Fitness_BO_blockmeans <- Fitness_BO_df3 %>% group_by(Block) %>% 
  summarize(
    BLOCKmean_FIT_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    BLOCKmean_FIT_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    BLOCKmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    BLOCKmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    BLOCKmean_seed_count = mean(seed_count, na.rm = TRUE),
    BLOCKmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE)
  ) #NA values removed from these averages


#Calculate relative fitness measures for each individual segment (Block means = 1)
#Retain NAs for this df; individual segments may have produced 0 closed fruit
Fitness_BO_df3.5 <- merge(Fitness_BO_df3, Fitness_BO_blockmeans, by = "Block") %>% 
  mutate(
    Rel_FIT_surv_x_seedprod = FIT_surv_x_seedprod/BLOCKmean_FIT_surv_x_seedprod,
    Rel_FIT_surv_x_fruitprod = FIT_surv_x_fruitprod/BLOCKmean_FIT_surv_x_fruitprod,
    Rel_est_seed_prod = est_seed_prod/BLOCKmean_est_seed_prod,
    Rel_total_fruit_prod = total_fruit_prod/BLOCKmean_total_fruit_prod,
    Rel_seed_count = seed_count/BLOCKmean_seed_count,
    Rel_P_surv_flower = P_surv_flower/BLOCKmean_P_surv_flower,
  )


# Average across blocks - MATERNAL mean fitness estimates
Fitness_BO_df4 <- Fitness_BO_df3.5 %>% group_by(Donor, Recipient) %>% 
  summarize(
    FIT_MATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    #FIT_MATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    MATmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    MATmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    MATmean_seed_count = mean(seed_count, na.rm = TRUE),
    MATmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE), #Relative measures of fitness
    Rel_FIT_MATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod, na.rm = TRUE),
    #Rel_FIT_MATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod, na.rm = TRUE),
    Rel_MATmean_est_seed_prod = mean(Rel_est_seed_prod, na.rm = TRUE),
    Rel_MATmean_total_fruit_prod = mean(Rel_total_fruit_prod, na.rm = TRUE),
    Rel_MATmean_seed_count = mean(Rel_seed_count, na.rm = TRUE),
    #Rel_MATmean_P_surv_flower = mean(Rel_P_surv_flower, na.rm = TRUE),
    Total_seed_count_F = sum(seed_count, na.rm = TRUE) #Seed totals summed for each maternal family
    )


#Average across recipients - PATERNAL mean fitness estimates
Fitness_BO_df5 <- Fitness_BO_df3.5 %>% group_by(Donor) %>% 
  summarize(
    FIT_PATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    #FIT_PATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    PATmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    PATmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    PATmean_seed_count = mean(seed_count, na.rm = TRUE),
    PATmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE),
    Rel_FIT_PATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod, na.rm = TRUE),
    #Rel_FIT_PATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod, na.rm = TRUE),
    Rel_PATmean_est_seed_prod = mean(Rel_est_seed_prod, na.rm = TRUE),
    Rel_PATmean_total_fruit_prod = mean(Rel_total_fruit_prod, na.rm = TRUE),
    Rel_PATmean_seed_count = mean(Rel_seed_count, na.rm = TRUE),
    #Rel_PATmean_P_surv_flower = mean(Rel_P_surv_flower, na.rm = TRUE), survival probably doesn't need to be relative
    Total_seed_count_pat = sum(seed_count, na.rm = TRUE)
  )


Maternal_Paternal_Fitness_BO <- merge(Fitness_BO_df4, Fitness_BO_df5, by = "Donor")



#Save a fitness-estimate data sheet
#Fitness components: mean survivorship, mean total fruit production, estimated total seed production (relative), Combined fitness (relative), total seed count

BO_FitSummary_1 <- Maternal_Paternal_Fitness_BO %>% 
  select(Donor, Recipient, MATmean_P_surv_flower, MATmean_total_fruit_prod, Rel_MATmean_est_seed_prod, Rel_FIT_MATmean_surv_x_seedprod, Total_seed_count_F) %>% 
  rename(mean_surv_to_flowering = MATmean_P_surv_flower,
         mean_fruit_prod = MATmean_total_fruit_prod,
         mean_est_seed_prod_R =  Rel_MATmean_est_seed_prod,
         mean_surv_x_est_seed_prod_R = Rel_FIT_MATmean_surv_x_seedprod,
         total_seed_count = Total_seed_count_F)


#Include total seed count for every mat fam (and paternal fam)

write_csv(BO_FitSummary_1, here::here("data_sheets", "compiled_sheets", "fitness_summary_BO1.csv"))


```


**HR**  
```{r}
#Survivorship of individual segments to flowering (maintain block)

Fitness_HR_df1 <- HR_master_df1 %>% group_by(Block, Transect, Donor, Recipient) %>% 
  summarize(
    P_surv_flower = mean(surv_to_flower)) %>% 
  filter(Block != "W") #Watered points already removed


Fitness_HR_df2 <- HR_master_df1 %>% group_by(Block, Transect, Donor, Recipient) %>%
  filter(FitP == TRUE) %>% 
  summarize(
    est_seed_prod = total_est_seed_production,
    total_fruit_prod = total_fruits_F,
    closed_fruits = closed_fruits_F,
    seed_count = seed_ct
  ) %>% ungroup() %>% 
  mutate(est_seed_prod = case_when(closed_fruits == 0 ~ as.numeric(as.character(NA)), 
                                   TRUE ~ est_seed_prod)) #option: set est_seed_prod to NA when 0 closed fruits were gathered (mean seeds per fruit not reliably estimated)
### This will introduce NAs below when before there were 0s for combined fitness (surv x est seed prod)


## 5 fewer rows than survival to flowering df



Fitness_HR_df3 <- merge(Fitness_HR_df2, Fitness_HR_df1, by = c("Block", "Transect", "Donor", "Recipient")) %>% 
  mutate(FIT_surv_x_seedprod = P_surv_flower*est_seed_prod,
         FIT_surv_x_fruitprod = P_surv_flower*total_fruit_prod)


#Calculate Block means of fitness variables:
Fitness_HR_blockmeans <- Fitness_HR_df3 %>% group_by(Block) %>% 
  summarize(
    BLOCKmean_FIT_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    BLOCKmean_FIT_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    BLOCKmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    BLOCKmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    BLOCKmean_seed_count = mean(seed_count, na.rm = TRUE),
    BLOCKmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE)
  ) #NA values removed from these averages


#Calculate relative fitness measures for each individual segment (Block means = 1)
#Retain NAs for this df; individual segments may have produced 0 closed fruit
Fitness_HR_df3.5 <- merge(Fitness_HR_df3, Fitness_HR_blockmeans, by = "Block") %>% 
  mutate(
    Rel_FIT_surv_x_seedprod = FIT_surv_x_seedprod/BLOCKmean_FIT_surv_x_seedprod,
    Rel_FIT_surv_x_fruitprod = FIT_surv_x_fruitprod/BLOCKmean_FIT_surv_x_fruitprod,
    Rel_est_seed_prod = est_seed_prod/BLOCKmean_est_seed_prod,
    Rel_total_fruit_prod = total_fruit_prod/BLOCKmean_total_fruit_prod,
    Rel_seed_count = seed_count/BLOCKmean_seed_count,
    Rel_P_surv_flower = P_surv_flower/BLOCKmean_P_surv_flower,
  )


# Average across blocks - MATERNAL mean fitness estimates
Fitness_HR_df4 <- Fitness_HR_df3.5 %>% group_by(Donor, Recipient) %>% 
  summarize(
    FIT_MATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    #FIT_MATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    MATmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    MATmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    MATmean_seed_count = mean(seed_count, na.rm = TRUE),
    MATmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE), #Relative measures of fitness
    Rel_FIT_MATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod, na.rm = TRUE),
    #Rel_FIT_MATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod, na.rm = TRUE),
    Rel_MATmean_est_seed_prod = mean(Rel_est_seed_prod, na.rm = TRUE),
    Rel_MATmean_total_fruit_prod = mean(Rel_total_fruit_prod, na.rm = TRUE),
    Rel_MATmean_seed_count = mean(Rel_seed_count, na.rm = TRUE),
    #Rel_MATmean_P_surv_flower = mean(Rel_P_surv_flower, na.rm = TRUE),
    Total_seed_count_F = sum(seed_count, na.rm = TRUE) #Seed totals summed for each maternal family
    )


#Average across recipients - PATERNAL mean fitness estimates
Fitness_HR_df5 <- Fitness_HR_df3.5 %>% group_by(Donor) %>% 
  summarize(
    FIT_PATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod, na.rm = TRUE),
    #FIT_PATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod, na.rm = TRUE),
    PATmean_est_seed_prod = mean(est_seed_prod, na.rm = TRUE),
    PATmean_total_fruit_prod = mean(total_fruit_prod, na.rm = TRUE),
    PATmean_seed_count = mean(seed_count, na.rm = TRUE),
    PATmean_P_surv_flower = mean(P_surv_flower, na.rm = TRUE),
    Rel_FIT_PATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod, na.rm = TRUE),
    #Rel_FIT_PATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod, na.rm = TRUE),
    Rel_PATmean_est_seed_prod = mean(Rel_est_seed_prod, na.rm = TRUE),
    Rel_PATmean_total_fruit_prod = mean(Rel_total_fruit_prod, na.rm = TRUE),
    Rel_PATmean_seed_count = mean(Rel_seed_count, na.rm = TRUE),
    #Rel_PATmean_P_surv_flower = mean(Rel_P_surv_flower, na.rm = TRUE), survival probably doesn't need to be relative
    Total_seed_count_pat = sum(seed_count, na.rm = TRUE)
  )


Maternal_Paternal_Fitness_HR <- merge(Fitness_HR_df4, Fitness_HR_df5, by = "Donor")



#Save a fitness-estimate data sheet
#Fitness components: mean survivorship, mean total fruit production, estimated total seed production (relative), Combined fitness (relative), total seed count

HR_FitSummary_1 <- Maternal_Paternal_Fitness_HR %>% 
  select(Donor, Recipient, MATmean_P_surv_flower, MATmean_total_fruit_prod, Rel_MATmean_est_seed_prod, Rel_FIT_MATmean_surv_x_seedprod, Total_seed_count_F) %>% 
  rename(mean_surv_to_flowering = MATmean_P_surv_flower,
         mean_fruit_prod = MATmean_total_fruit_prod,
         mean_est_seed_prod_R =  Rel_MATmean_est_seed_prod,
         mean_surv_x_est_seed_prod_R = Rel_FIT_MATmean_surv_x_seedprod,
         total_seed_count = Total_seed_count_F)


#Include total seed count for every mat fam (and paternal fam)

write_csv(HR_FitSummary_1, here::here("data_sheets", "compiled_sheets", "fitness_summary_HR1.csv"))


```








<br>  


#### Visualizations    

```{r eval = FALSE}
######
### Data frame organization: 
######

# Segment Fitness: Fitness_AC_df3.5
# (Includes relative fitness measures)

# Maternal Fitness: Fitness_AC_df4

# Paternal Fitness: Fitness_AC_df5
```


**Fitness = Total Fruit Production**  

```{r}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,70)) +
  ylab("Count") +
  xlab("Segment Fruit Production (N=319)")

## Individual Segment fitness (Relative)
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,70)) +
  ylab("Count") +
  xlab("Segment Relative Fruit Production (N=319)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_total_fruit_prod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,70)) +
  ylab("Count") +
  xlab("Segment Relative Fruit Production (N=319) (By Block)")




## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = MATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,20)) +
  ylab("Count") +
  xlab("Avg. Maternal Fruit Production (N=107)")


## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_MATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,20)) +
  ylab("Count") +
  xlab("Avg. Relative Maternal Fruit Production (N=107)")



## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = PATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Fruit Production (N=40)")

## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_PATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Relative Paternal Fruit Production (N=40)")

```


**Fitness = Estimated Seed Production**  

NOTE: From here on, only fitness estimates relative to Block means are presented.  

```{r echo = FALSE}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_est_seed_prod), fill = "#A58AFF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Estimated Seed Production (N=319) (Relative)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_est_seed_prod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Estimated Seed Production (N=319) (Relative)")



## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_MATmean_est_seed_prod), fill = "#A58AFF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  ylab("Count") +
  xlab("Avg. Maternal Estimated Seed Production (N=107) (Relative)")


## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_PATmean_est_seed_prod), fill = "#A58AFF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Estimated Seed Production (N=40) (Relative)")

```



**Fitness = Survival x Estimated Seed Production**  

```{r echo = FALSE}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_seedprod), fill = "#00BE67", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Survival x Estimated Seed Production (N=319) (Relative)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_seedprod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("egment Survival x Estimated Seed Production (N=319) (Relative)")



## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_FIT_MATmean_surv_x_seedprod), fill = "#00BE67", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  ylab("Count") +
  xlab("Avg. Maternal Estimated Seed Production (N=107) (Relative)")


## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_FIT_PATmean_surv_x_seedprod), fill = "#00BE67", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Estimated Seed Production (N=40) (Relative)")

```

**Fitness = Survival x Total Fruit Production**  

```{r echo = FALSE}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_fruitprod), fill = "#FFA500", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Survival x Estimated fruit Production (N=319) (Relative)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_fruitprod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("egment Survival x Estimated fruit Production (N=319) (Relative)")



## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_FIT_MATmean_surv_x_fruitprod), fill = "#FFA500", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  ylab("Count") +
  xlab("Avg. Maternal Estimated fruit Production (N=107) (Relative)")


## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_FIT_PATmean_surv_x_fruitprod), fill = "#FFA500", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Estimated fruit Production (N=40) (Relative)")

```



<br>  





## Exploratory Data Analyses  

Phenotypic Trait and Fitness data from the 2021-2022 Nemophila Field Experiment  

Explore:
- Block effects (on fruit prod, fecundity, fitness)
- 




```{r}
AC_master_Fonly_1 <- read_csv(here::here("data_sheets", "compiled_sheets", "AC_mastersheet_v1_Fitness-only_2022-07-26.csv"))


#bivar plot for total fruit No. (open + clsoed) ~ total seed number (include sample sizes for each)

ggplot(data = AC_master_Fonly_1) +
  geom_point(aes(x = total_fruits_F, y = seed_ct)) + 
  theme_bw() +
  xlab("Total Fruit Production (open + closed) (N = 319)") +
  ylab("Seed Count")


ggplot(data = AC_master_Fonly_1) +
  geom_point(aes(x = closed_fruits_F, y = seed_ct)) + 
  theme_bw() +
  xlab("Total Closed Fruits (N = 319)") +
  ylab("Seed Count")
```




#### Correlation Matrices  

```{r}



```




####  Summaries & Distributions  



#### Descriptive Stats  



#### Relationships between potential fitness measures  

Maternal Genotype Fitness Measures (to inform planting in the fall)

Segment fitness = (segment survivorship [to flowering] x total repro output [total est seed production])
Genotype fitness = average of these 3 estimates






