---
title: "Exploratory Data Analyses"
author: "Devin Gamble"
date: "2022-07-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)

```


```{r}
###NOTES 

#long data set
#Fitness data set


#Analyze fitness as fecundity alone vs surv*fecundity

#Total number of fruits may be better than seed # - less downward bias (due to seeds lost after dehiscence)



#To-Do
# - single 'Fitness' data sheet with all variables required for analysis / visualizations!
# - calculate fitness metrics based on the master sheet first? includes sample sizes in df (or, merge back in the fitness sheets below..)
# - correlations among all fitness measures
#NAs vs 0s: when there were 0 closed fruit (i.e. no viable seeds collected), but possibly non-zero total fruits (open, etc.)??? Included as 0s in data, but may decide to remove

#1 or more fruits but 0 seeds --> 0 fecundity (for est seed prod)

#Do we even need a 'fitness plant only' master sheet as calculated in the other Rmd?




#
# Sample Mat fams equally for seeds, but account for differences in individual Mat fam/segment fitness later on in aster
# In aster, include fruit number and/or fecundity (may be incomplete wieth 0 clsoed fruits) as hierarchical fitness estimates

```



Load master sheets:  
```{r message = FALSE, warning = FALSE}
## AC
AC_master_df1 <- read_csv(here::here("data_sheets", "compiled_sheets", "AC_mastersheet_v1_2022-07-26.csv")) # Fl_6.22 read in as logical - it should be a date. Parsing problem with row 1398 - unsure of issue (FFD but no Germ Date?) Or the single instance of 6-22-2022 flowering date.

AC_master_Fonly_1 <- read_csv(here::here("data_sheets", "compiled_sheets", "AC_mastersheet_v1_Fitness-only_2022-07-26.csv"))

```


## Fitness Estimates  


NOTE: Use `surv_to_x` variables to delimit groups in calculating certain fitness components; NAs were changed to 0s for seed counts/means when 0 closed fruit were collected, though some of these instances occur for plants that did not survive to produce any fruit (i.e, total fruits = 0).  

##### Fitness = Survival to flowering x Estimated seed production

**AC**  
```{r}

#Survivorship of individual segments to flowering (maintain block)

Fitness_AC_df1 <- AC_master_df1 %>% group_by(Block, Transect, Donor, Recipient) %>% 
  summarize(
    P_surv_flower = mean(surv_to_flower)) %>% 
  filter(Block != "W") #Remove transects W1 and W2


#include surv to fruit prod only?

Fitness_AC_df2 <- AC_master_df1 %>% group_by(Block, Transect, Donor, Recipient) %>%
  filter(FitP == TRUE) %>% 
  summarize(
    est_seed_prod = total_est_seed_production,
    total_fruit_prod = total_fruits_F,
    closed_fruits = closed_fruits_F,
    seed_count = seed_ct
  )

#Two Fewer rows than the summarization of P_surv_flower - possibly because two segments never had fitness plants chosen


Fitness_AC_df3 <- merge(Fitness_AC_df2, Fitness_AC_df1, by = c("Block", "Transect", "Donor", "Recipient")) %>% 
  mutate(FIT_surv_x_seedprod = P_surv_flower*est_seed_prod,
         FIT_surv_x_fruitprod = P_surv_flower*total_fruit_prod)


#Calculate Block means of fitness variables:
Fitness_AC_blockmeans <- Fitness_AC_df3 %>% group_by(Block) %>% 
  summarize(
    BLOCKmean_FIT_surv_x_seedprod = mean(FIT_surv_x_seedprod),
    BLOCKmean_FIT_surv_x_fruitprod = mean(FIT_surv_x_fruitprod),
    BLOCKmean_est_seed_prod = mean(est_seed_prod),
    BLOCKmean_total_fruit_prod = mean(total_fruit_prod),
    BLOCKmean_seed_count = mean(seed_count),
    BLOCKmean_P_surv_flower = mean(P_surv_flower)
  )


#Calculate relative fitness measures for each individual segment (Block means = 1)
Fitness_AC_df3.5 <- merge(Fitness_AC_df3, Fitness_AC_blockmeans, by = "Block") %>% 
  mutate(
    Rel_FIT_surv_x_seedprod = FIT_surv_x_seedprod/BLOCKmean_FIT_surv_x_seedprod,
    Rel_FIT_surv_x_fruitprod = FIT_surv_x_fruitprod/BLOCKmean_FIT_surv_x_fruitprod,
    Rel_est_seed_prod = est_seed_prod/BLOCKmean_est_seed_prod,
    Rel_total_fruit_prod = total_fruit_prod/BLOCKmean_total_fruit_prod,
    Rel_seed_count = seed_count/BLOCKmean_seed_count,
    Rel_P_surv_flower = P_surv_flower/BLOCKmean_P_surv_flower,
  )


# Average across blocks - MATERNAL mean fitness estimates
Fitness_AC_df4 <- Fitness_AC_df3.5 %>% group_by(Donor, Recipient) %>% 
  summarize(
    FIT_MATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod),
    FIT_MATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod),
    MATmean_est_seed_prod = mean(est_seed_prod),
    MATmean_total_fruit_prod = mean(total_fruit_prod),
    MATmean_seed_count = mean(seed_count),
    MATmean_P_surv_flower = mean(P_surv_flower), #Relative measures of fitness
    Rel_FIT_MATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod),
    Rel_FIT_MATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod),
    Rel_MATmean_est_seed_prod = mean(Rel_est_seed_prod),
    Rel_MATmean_total_fruit_prod = mean(Rel_total_fruit_prod),
    Rel_MATmean_seed_count = mean(Rel_seed_count),
    Rel_MATmean_P_surv_flower = mean(Rel_P_surv_flower)
    )


#Average across recipients - PATERNAL mean fitness estimates
Fitness_AC_df5 <- Fitness_AC_df3.5 %>% group_by(Donor) %>% 
  summarize(
    FIT_PATmean_surv_x_seedprod = mean(FIT_surv_x_seedprod),
    FIT_PATmean_surv_x_fruitprod = mean(FIT_surv_x_fruitprod),
    PATmean_est_seed_prod = mean(est_seed_prod),
    PATmean_total_fruit_prod = mean(total_fruit_prod),
    PATmean_seed_count = mean(seed_count),
    PATmean_P_surv_flower = mean(P_surv_flower),
    Rel_FIT_PATmean_surv_x_seedprod = mean(Rel_FIT_surv_x_seedprod),
    Rel_FIT_PATmean_surv_x_fruitprod = mean(Rel_FIT_surv_x_fruitprod),
    Rel_PATmean_est_seed_prod = mean(Rel_est_seed_prod),
    Rel_PATmean_total_fruit_prod = mean(Rel_total_fruit_prod),
    Rel_PATmean_seed_count = mean(Rel_seed_count),
    Rel_PATmean_P_surv_flower = mean(Rel_P_surv_flower)
  )


Maternal_Paternal_Fitness <- merge(Fitness_AC_df4, Fitness_AC_df5, by = "Donor")

#write_csv()

```

<br>  


#### Visualizations    

```{r eval = FALSE}
######
### Data frame organization: 
######

# Segment Fitness: Fitness_AC_df3.5
# (Includes relative fitness measures)

# Maternal Fitness: Fitness_AC_df4

# Paternal Fitness: Fitness_AC_df5
```


**Fitness = Total Fruit Production**  

```{r}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,70)) +
  ylab("Count") +
  xlab("Segment Fruit Production (N=319)")

## Individual Segment fitness (Relative)
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,70)) +
  ylab("Count") +
  xlab("Segment Relative Fruit Production (N=319)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_total_fruit_prod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,70)) +
  ylab("Count") +
  xlab("Segment Relative Fruit Production (N=319) (By Block)")




## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = MATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,20)) +
  ylab("Count") +
  xlab("Avg. Maternal Fruit Production (N=107)")


## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_MATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,20)) +
  ylab("Count") +
  xlab("Avg. Relative Maternal Fruit Production (N=107)")



## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = PATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Fruit Production (N=40)")

## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_PATmean_total_fruit_prod), fill = "#00A9FF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Relative Paternal Fruit Production (N=40)")

```


**Fitness = Estimated Seed Production**  

NOTE: From here on, only fitness estimates relative to Block means are presented.  

```{r echo = FALSE}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_est_seed_prod), fill = "#A58AFF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Estimated Seed Production (N=319) (Relative)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_est_seed_prod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Estimated Seed Production (N=319) (Relative)")



## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_MATmean_est_seed_prod), fill = "#A58AFF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  ylab("Count") +
  xlab("Avg. Maternal Estimated Seed Production (N=107) (Relative)")


## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_PATmean_est_seed_prod), fill = "#A58AFF", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Estimated Seed Production (N=40) (Relative)")

```



**Fitness = Survival x Estimated Seed Production**  

```{r echo = FALSE}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_seedprod), fill = "#00BE67", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Survival x Estimated Seed Production (N=319) (Relative)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_seedprod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("egment Survival x Estimated Seed Production (N=319) (Relative)")



## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_FIT_MATmean_surv_x_seedprod), fill = "#00BE67", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  ylab("Count") +
  xlab("Avg. Maternal Estimated Seed Production (N=107) (Relative)")


## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_FIT_PATmean_surv_x_seedprod), fill = "#00BE67", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Estimated Seed Production (N=40) (Relative)")

```

**Fitness = Survival x Total Fruit Production**  

```{r echo = FALSE}
## Individual Segment fitness
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_fruitprod), fill = "#FFA500", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("Segment Survival x Estimated fruit Production (N=319) (Relative)")

## Segment Fitness organized by Block
ggplot(data = Fitness_AC_df3.5) +
  geom_histogram(aes(x = Rel_FIT_surv_x_fruitprod, fill = Block), color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  ylab("Count") +
  xlab("egment Survival x Estimated fruit Production (N=319) (Relative)")



## Average Maternal Family Fitness  
ggplot(data = Fitness_AC_df4) +
  geom_histogram(aes(x = Rel_FIT_MATmean_surv_x_fruitprod), fill = "#FFA500", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  ylab("Count") +
  xlab("Avg. Maternal Estimated fruit Production (N=107) (Relative)")


## Average Paternal Family Fitness  
ggplot(data = Fitness_AC_df5) +
  geom_histogram(aes(x = Rel_FIT_PATmean_surv_x_fruitprod), fill = "#FFA500", color = "black") + 
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,10)) +
  ylab("Count") +
  xlab("Avg. Paternal Estimated fruit Production (N=40) (Relative)")

```




**SAVE CSV - MATERNAL & PATERNAL MEAN FITNESS ESTIMATES**  

<br>  



## Exploratory Data Analyses  

Phenotypic Trait and Fitness data from the 2021-2022 Nemophila Field Experiment  


```{r}
#bivar plot for total fruit No. (open + clsoed) ~ total seed number (include sample sizes for each)

ggplot(data = AC_master_Fonly_1) +
  geom_point(aes(x = total_fruits_F, y = seed_ct)) + 
  theme_bw() +
  xlab("Total Fruit Production (open + closed) (N = 319)") +
  ylab("Seed Count")


ggplot(data = AC_master_Fonly_1) +
  geom_point(aes(x = closed_fruits_F, y = seed_ct)) + 
  theme_bw() +
  xlab("Total Closed Fruits (N = 319)") +
  ylab("Seed Count")
```




#### Correlation Matrices  

```{r}



```




####  Summaries & Distributions  



#### Descriptive Stats  



#### Relationships between potential fitness measures  

Maternal Genotype Fitness Measures (to inform planting in the fall)

Segment fitness = (segment survivorship [to flowering] x total repro output [total est seed production])
Genotype fitness = average of these 3 estimates






