---
title: "Plastic Responses HR"
author: "Devin Gamble"
date: "2023-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Description:**  
This .Rmd contains analyses of data from the supplemental watering experiment in year 1 (2021-2022) of the *Nemophila menziesii* field experiment. Plasticity is estimated as the effect of watering on phenotypic traits in four populations.


**Note**: Check sample sizes for genotypes. Maybe focus on population-level responses while controlling for Recipient.  

**Note 2**: How should variables be standardized? All centered? Scaled? Both?


**Load Packages**  

```{r message = FALSE}
library(tidyverse) #For data organization and tidying
library(janitor) #more tidying functions
library(here) #great for specifying relative file paths
library(lubridate) #for adjusting 'date' values if any need cleaning up
library(lme4)
library(corrplot)
library(jtools)
#add more packages here
#library(lmerTest) #to get p-values for lmer?

```
**NOTE**: UPDATE aboveground biomass and corolla diameter/area data

```{r}
#set contrasts
options(contrasts=c(unordered="contr.sum", ordered="contr.poly"))
#DOUBLE CHECK this is correct for the below analyses
```


**Load Data**  
Hastings Reservation
These dataframes includes data on all selected fitness plants at Hastings Reservation from 2021-2022. There were 3 experimental blocks and a total of 9 + 2 transects. Plants in block 'W' received supplemental water and had, when available, 3 fitness plants chosen. Recipient ID's (Genotypes) that were present in both the regular and water blocks had, when available, 3 fitness plants chosen (Replicated = R). Important trait-related columns include germination date, the number of closed fruit, number of total fruit, seed count, mean seed mass (msm_all), mean seeds per fruit, total estimated seed production, first flowering date (FFD), days from germination to flower, flowering duration, and survival.


**Variables**  

Fitness: fecundity (total viable seed count)
Phenotypic traits:
- First flowering date  
- Flowering duration
- Aboveground biomass  
- Corolla diameter  
- Specific leaf area


**All data**  
```{r warning = FALSE}
HR_allplants <- read_csv(here::here("data_sheets", "compiled_sheets", "HR_mastersheet_full_2023-02-06.csv")) %>% select(-c(Sequence, FitP, any_FitP, F_multi, F_plant_notes, Fl_3.10:flr_P, seeds_weighed_d1:msm_d9)) %>% 
  filter(Replicated == "R") %>% #Replicated genotypes only
  mutate(FFD_c = scale(yday(FFD), scale = FALSE)[,1]) %>% #convert to Julian day of year, but don't standardize
  mutate(fl_dur_c = scale(fl_duration, scale = FALSE)[,1]) %>% # Just centers these variables
  mutate(Treatment = case_when(Block %in% c("1", "2", "3") ~ "ambient",
                               Block == "W" ~ "watered")) %>% 
  mutate(fl_duration = case_when(fl_duration == 0 ~ 1,
                                 TRUE ~ fl_duration)) %>%  #Replace durations of 0 with 1
  filter(!is.na(FFD_c))

#734 obs vs 656 for fitness plants
## Exclude genotypes without observations from both treatments?

```
**Fitness Plants Only data**  

```{r warning = FALSE}
HR_Fplants <- read_csv(here::here("data_sheets", "compiled_sheets", "HR_mastersheet_Fitness-all_2023-02-06.csv")) %>% select(-c(Sequence, FitP, any_FitP, F_multi, F_plant_notes, Fl_3.10:flr_P, seeds_weighed_d1:msm_d9))

#Standardize traits
HR_plants_z <- HR_Fplants %>% 
  filter(Replicated == "R") %>% #Replicated genotypes only
  mutate(seed_ct_r = seed_ct/mean(seed_ct)) %>% #fitness relative to pop mean
  mutate(FFD_z = scale(yday(FFD))[,1], #convert to Julian day of year
         fl_dur_z = scale(fl_duration)[,1],
         AG_biom_z = scale(AG_biomass_mg)[,1]) %>% 
  mutate(Treatment = as.factor(case_when(Block %in% c("1", "2", "3") ~ "ambient",
                               Block == "W" ~ "watered"))) %>% 
  mutate(Recipient = as.factor(Recipient)) %>% 
  mutate(fl_duration = case_when(fl_duration == 0 ~ 1,
                                 TRUE ~ fl_duration)) #Replace durations of 0 with 1

#Note: seed count includes 0s for plants that never germinated/flowered

##Corrplot of variables
corrplot(cor(na.omit(select(HR_plants_z, c(seed_ct_r, FFD_z, fl_dur_z, AG_biom_z)))), addCoef.col = "white", number.cex = 0.8, method = "square")

#Some potential for some collinearity among FFD/fl_duration, Biomass/fl_duration
```


**Visualizations**  

```{r}
##### all plants
#FFD
ggplot(data = subset(HR_allplants, !is.na(FFD_c)), aes(x=Treatment, y=FFD_c)) + 
  geom_boxplot(aes(fill = Treatment), color = "black") + labs(x= "Treatment", y = "Flowering Date") +
  theme_classic()

ggplot(data = subset(HR_allplants, !is.na(FFD)), aes(x=FFD, group = Treatment)) + 
  geom_histogram(aes(fill = Treatment), color = "black") + labs(x= "Treatment", y = "Flowering Date") +
  theme_classic()


ggplot(data = subset(HR_allplants, !is.na(FFD_c)), aes(x=Treatment, y=FFD_c, group = Recipient)) + 
  geom_line(aes(color = Recipient), linewidth = 1) + 
  labs(x= "Treatment", y = "Flowering Date") +
  theme_classic()

#Fl Duration
ggplot(data = subset(HR_allplants, !is.na(fl_dur_c)), aes(x=Treatment, y=fl_dur_c)) + 
  geom_boxplot(aes(fill = Treatment), color = "black") + labs(x= "Treatment", y = "Flowering Duration") +
  theme_classic()
ggplot(data = subset(HR_allplants, !is.na(fl_dur_c)), aes(x=Treatment, y=fl_dur_c, group = Recipient)) + 
  geom_line(aes(color = Recipient), linewidth = 1) + 
  labs(x= "Treatment", y = "Flower Duration") +
  theme_classic()


##F Duration - subset and prettier figure
HR_allplants_Fdur1 <- HR_allplants %>% group_by(Treatment, Recipient) %>% 
  summarize(fl_dur_m = mean(fl_duration)) %>% 
  select(everything()) %>% ungroup() 

ggplot(data = subset(HR_allplants_Fdur1, !is.na(fl_dur_m)), aes(x=Treatment, y=fl_dur_m, group = Recipient)) + 
  geom_line(aes(color = Recipient), linewidth = 1) + 
  labs(x= "Treatment", y = "Flowering Duration (days)") +
  theme_classic(base_size = 16) + theme(legend.position = "none")



##### fitness plants only data
ggplot(HR_plants_z, aes(x=Treatment, y=seed_ct_r)) + 
  geom_boxplot(aes(fill = Treatment), color = "black") + labs(x= "Treatment", y = "Total Seed Production") +
  theme_classic()

ggplot(HR_plants_z, aes(x=Treatment, y=seed_ct_r, group = Recipient)) + 
  geom_line(aes(color = Recipient)) + labs(x= "Treatment", y = "Total Seed Production") +
  theme_classic()


ggplot(HR_plants_z, aes(x=Treatment, y=closed_fruits_F)) + 
  geom_boxplot(aes(fill = Treatment), color = "black") + labs(x= "Treatment", y = "Total Closed Fruits") +
  theme_classic()


ggplot(data = subset(HR_plants_z, !is.na(AG_biom_z)), aes(x=Treatment, y=AG_biom_z)) + 
  geom_boxplot(aes(fill = Treatment), color = "black") + labs(x= "Treatment", y = "AG Biomass") +
  theme_classic()

```

#### Is there an overall (pop-level) effect of watering?  

**Models**  

Anovas and t-tests

```{r}
#FFD
t.test(HR_allplants$FFD_c ~ HR_allplants$Treatment)

HR_FFD_aov <- lm(FFD_c ~ Treatment, data = HR_allplants)
TukeyHSD(aov(HR_FFD_aov))


#Flowering duration *******
t.test(fl_dur_c ~ Treatment, data = HR_allplants)

HR_fdur_aov <- lm(fl_duration ~ Treatment, data = HR_allplants)
TukeyHSD(aov(HR_fdur_aov))



### Fit-plant traits

#Seed Count
t.test(seed_ct_r ~ Treatment, data = HR_plants_z)

#Fruit Count *******
t.test(closed_fruits_F ~ Treatment, data = HR_plants_z)


#Aboveground Biomass
t.test(AG_biom_z ~ Treatment, data = HR_plants_z)

HR_AGb_aov <- lm(AG_biom_z ~ Treatment, data = HR_plants_z)
TukeyHSD(aov(HR_AGb_aov))

#Corolla Diameter


#Specific Leaf Area


```

Flowering duration and Number of closed fruit are plastic at the population level
<br>  


### Estimating Reaction Norms & Plasticity  
Document variation in intercepts and slopes among genotypes re: response to water availability

Referenced Arnold et al. 2019 Notes on constructing mixed models for plasticity analysis

#### First flowering date    

```{r}
##Packing used in Arnold et al. 2019 for getting MM R-squared s:
library(MuMIn)
#r.squaredGLMM()
```

**Models**  

```{r}
# Simple Linear Model
HR_FFD_m1 <- lm(FFD_c ~ Treatment + Recipient, data = HR_allplants)

summary(HR_FFD_m1)
#pretty bad model

#
#Mixed models
HR_FFD_m1ma <- lmer(FFD_c ~ Treatment + Recipient + (1|Transect), data = HR_allplants, REML = FALSE)

summary(HR_FFD_m1ma)
r.squaredGLMM(HR_FFD_m1ma)
#r2m = marginal R2 (fixed effects only), *R2c = conditional R2 (variance due to fixed + random effects)
### Compare r2c in other models

#
#Random Intercept for Recipient
HR_FFD_m1mb <- lmer(FFD_c ~ Treatment + (1|Transect) + (1|Recipient), data = HR_allplants, REML = FALSE)
summary(HR_FFD_m1mb)
r.squaredGLMM(HR_FFD_m1mb) #better r2c than first lmer model
#Recipient mean first flower date (intercept) accounts for some variation in population level FFD

#Does adding an effect of Recipient intercepts really improve the model?
AIC(HR_FFD_m1, HR_FFD_m1ma, HR_FFD_m1mb)#Nope

#Other model comparison method: LRT
chi2 <- 1*(summary(HR_FFD_m1mb)$logLik - summary(HR_FFD_m1ma)$logLik) ## 1 indicates df diff b/w models
1-pchisq(chi2, 1)#Nope again

### Possibly because the environmental variable is binary, the potential for varying intercepts to contribute to variance in FFD is more limited.

#
#Random Intercepts & Slopes: 
#Does variation in genotype slopes explain variance in this model
HR_FFD_m1mc <- lmer(FFD_c ~ Treatment + (1|Transect) + (1+Treatment|Recipient), data = HR_allplants, REML = FALSE)
summary(HR_FFD_m1mc)
r.squaredGLMM(HR_FFD_m1mc) #better r2c (model fit), but are we overfitting? AIC suggests maybe

#Model comparisons
AIC(HR_FFD_m1ma, HR_FFD_m1mb, HR_FFD_m1mc)
#chi2 <- 3*(summary(HR_FFD_m1mc)$logLik - summary(HR_FFD_m1ma)$logLik)
#1-pchisq(chi2, 1)

#Still not better than base lmer model (no random intercepts/slopes)
# Increase in r2c could be an artifact of low sample sizes/overfitting...


#
## G x E Interactions - no random intercepts/slopes
HR_FFD_m1m_ge <- lmer(FFD_c ~ Treatment*Recipient + (1|Transect), data = HR_allplants, REML = FALSE)
summary(HR_FFD_m1m_ge)
r.squaredGLMM(HR_FFD_m1m_ge) #Nice r2c

#diagnostics
plot(HR_FFD_m1m_ge)

#Model comparisons
AIC(HR_FFD_m1ma, HR_FFD_m1mb, HR_FFD_m1mc, HR_FFD_m1m_ge)
#LRT
chi2 <- 39*(summary(HR_FFD_m1m_ge)$logLik - summary(HR_FFD_m1ma)$logLik)
1-pchisq(chi2, 1)
#LRT indicates models are s. different (they're not nested), but AIC suggests this one is not as parsimonious, even though the model fit via r2c is better
```


Best model - FFD Plasticity : Mixed-model with Random intercept for Recipient (Genotype) (although R2c is ~2 points lower than previous model)
Not enough power to investigate GxE further; attempts appear to result in model overfitting.

Conclusion: No plasticity of FFD in response to water availability at HR, but potential for genetic variation in plasticity


Estimating BLUPs
via Arnold et al. 2019
```{r eval = FALSE}
##Ranking genotype plasticity using BLUPs
#Notes: no measures of uncertainty, 
#Use in a Bayesian framework may be more appropriate - MCMCglmm package (see Hadfield et al 2010)
#Measured as the response of a given genotype to watering as the diff b/w the population mean response and genotype predicted response

Rec_BLUPs <- ranef(HR_FFD_m1mc)$Recipient #Random intercepts & slopes model
geno_index <- as.factor(c(1:21)) #Alter to maintain recipients?
geno_data <- cbind(geno_index, Rec_BLUPs)
colnames(geno_data) <- c("Recipient", "BLUP_int", "BLUP_slope")

#BLUPs by intercept
ggplot(geno_data, aes(Recipient, BLUP_int)) + 
  geom_point(aes(group = Recipient, color = Recipient)) + 
  ylab("BLUP Intercept Estimate") +
  geom_hline(yintercept = 0, lty = 2) +
  theme_classic()

#BLUP slope estimates - diff in genotype slopes realtive to pop mean slope
ggplot(geno_data, aes(Recipient, BLUP_slope)) + 
  geom_point(aes(group = Recipient, color = Recipient)) + 
  ylab("BLUP Slope") +
  geom_hline(yintercept = 0, lty = 2) +
  theme_classic()

#Add these slopes to pop average to get real genotype slopes

HR_avg_slope_FFD <- summary(HR_FFD_m1mc)$coefficients[2]
geno_data$geno_slopes <- Rec_BLUPs$Treatment1 + HR_avg_slope_FFD

#Genotype Slopes
ggplot(geno_data, aes(Recipient, geno_slopes)) + 
  geom_point(aes(group = Recipient, color = Recipient), size = 3) + 
  ylab("Plasticity (HR-slopes + BLUP slope)") +
  geom_hline(yintercept = 0, lty = 2) +
  theme_classic()
#Should reflect Recipient slopes as plotted earlier - individual values of plasticity by genotype

#BLUP Intercept x slope correlations
ggplot(geno_data, aes(BLUP_int, BLUP_slope)) + 
  geom_point(aes(group = Recipient, color = Recipient), size = 2) + 
  xlab("BLUP intercept") + ylab("BLUP slope") +
  geom_hline(yintercept = 0, lty = 2) +
  theme_classic()


##Rank genotypes in order of plasticity
geno_data$geno_ordered <- factor(geno_data$Recipient, levels = geno_data$Recipient[order(geno_data$BLUP_slope)]) #do geno_slope instead?

ggplot(geno_data, aes(geno_ordered, geno_slopes)) + 
  geom_bar(stat = "identity", aes(group = Recipient, fill = Recipient)) + 
  xlab("Recipient (least to most plastic)") +
  ylab("Plasticity (Genotype slope)") +
  geom_hline(yintercept = 0, lty = 2) +
  theme_classic()

#By looking at genotypic slopes, the most plastic Recipients will be those farthest from 0

```

<br>  


#### Flowering Duration  



```{r}
# Simple linear model
HR_fdur_m1 <- lm(fl_dur_c ~ Treatment + Recipient, data = HR_allplants)
summary(HR_fdur_m1)

#
#G x E effect
HR_fdur_m2 <- lm(fl_dur_c ~ Treatment* Recipient, data = HR_allplants)
summary(HR_fdur_m2)

#
#MM - location
HR_fdur_m1ma <- lmer(fl_dur_c ~ Treatment + Recipient + (1|Transect), data = HR_allplants, REML = FALSE)
summary(HR_fdur_m1ma) 
r.squaredGLMM(HR_fdur_m1ma)

#
#Random intercepts
HR_fdur_m1mb <- lmer(fl_dur_c ~ Treatment + (1|Recipient) + (1|Transect), data = HR_allplants, REML = FALSE)
summary(HR_fdur_m1mb) #SINGULAR
r.squaredGLMM(HR_fdur_m1mb)
#
#Random slopes & intercepts
HR_fdur_m1mc <- lmer(fl_dur_c ~ Treatment + (1+Treatment|Recipient) + (1|Transect), data = HR_allplants, REML = FALSE)
summary(HR_fdur_m1mc) #SINGULAR
r.squaredGLMM(HR_fdur_m1mc)

##Note: Issues of singularity for MMs -> singular models may be **overfit** and/or result in mis-convergence
## This was not an issue for FFD, strangely. 
#Could be avoided if using bayesian approach (e.g., blme, stan_glmer [rstanarm]), or remove complex terms (random slopes)


#Model comparison
AIC(HR_fdur_m1,HR_fdur_m2, HR_fdur_m1ma, HR_fdur_m1mb, HR_fdur_m1mc)

```
