---
title: "Plasticity_EDA_1"
author: "Devin Gamble"
date: "2022-10-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Description:**  
This .Rmd will include exploratory data analyses (EDA) on the phenotypic plasticity of traits measured in four populations of *Nemophila menziesii* during the 2021-2022 growing season of the Mazer lab's Nemophila project. 



**Load Packages**  

```{r message = FALSE}
library(tidyverse) #For data organization and tidying
library(janitor) #more tidying functions
library(here) #great for specifying relative file paths
library(lubridate) #for adjusting 'date' values if any need cleaning up
library(lme4)
library(car)
library(corrplot)
library(jtools)
library(MuMIn) # for R2s in lmer
library(lmerTest) #get P-values from lmer - must load lme4
#Provides p-values using Satterthwaite's approximation via t-test


```

```{r}
#set contrasts
options(contrasts=c(unordered="contr.sum", ordered="contr.poly"))
#DOUBLE CHECK this is correct for the below analyses
```



#### Load Fitness Plant Data  

**Bodega Bay**  

Load and tidy up data.
```{r}
## All data
BB_allplants <- read_csv(here::here("data_sheets", "compiled_sheets", "BB_mastersheet_full_2023-03-11.csv")) %>% select(-c(Sequence, Co., FitP, F_multi, F_plant_notes, Fl_2.25:flr_P, seeds_weighed_d1:msm_d9)) %>% 
  filter(Replicated == "R") %>% #Replicated genotypes only
  mutate(fl_duration = case_when(fl_duration == 0 ~ 1,
                                 TRUE ~ fl_duration)) %>%  #Replace durations of 0 with 1
  mutate(FFD_c = scale(yday(FFD), scale = FALSE)[,1], #convert to Julian day of year, but don't scale...
         fl_dur_c = scale(fl_duration, scale = FALSE)[,1], # Center all variables
         corolla_d_c = scale(corolla_diam_mm, scale = FALSE)[,1],
         AG_biom_c = scale(AG_biomass_mg, scale = FALSE)[,1],
         seed_ct_c = scale(seed_ct, scale = FALSE)[,1],
         lifet_fec_c = scale(total_est_seed_production, scale = FALSE)[,1]) %>% 
  mutate(FFD_z = scale(yday(FFD))[,1],
         fl_dur_z = scale(fl_duration)[,1], # Standardize all variables
         corolla_d_z = scale(corolla_diam_mm)[,1],
         AG_biom_z = scale(AG_biomass_mg)[,1],
         seed_ct_z = scale(seed_ct)[,1],
         lifet_fec_z = scale(total_est_seed_production)[,1]) %>% 
  mutate(seed_ct_r = seed_ct/mean(seed_ct),
         lifet_fec_r = total_est_seed_production/mean(total_est_seed_production)) %>% #fitness relative to population mean
  mutate(Treatment = as.factor(case_when(Block %in% c("1", "2", "3") ~ "ambient",
                               Block == "W" ~ "watered")),
         Treat_bin = as.numeric(case_when(Block %in% c("1", "2", "3") ~ 0, #binary var for treatment: ambient/watered = 0/1
                               Block == "W" ~ 1))) %>% 
  mutate(Recipient = as.factor(Recipient))



BB_allplants_f <- BB_allplants %>% 
  filter(!is.na(FFD_c)) #plants that flowered only

#318 obs vs 225 for fitness plants

##Fitness Plants Only
BB_Fplants <- BB_allplants %>% 
  filter(any_FitP == TRUE)

##Note on the data:
## This data frame includes data on all selected fitness plants at Bodega Bay from 2021-2022. There were 3 experimental blocks and a total of 156 (+1; 160) point-locations. Plants in block 'W' received supplemental water and had, when available, 3 fitness plants chosen. Recipient ID's (Genotypes) that were present in both the regular and water blocks had, when available, 3 fitness plants chosen (Replicated = R). Important trait-related columns include germination date, the number of closed fruit, number of total fruit, seed count, mean seed mass (msm_all), mean seeds per fruit, total estimated seed production, first flowering date (FFD), days from germination to flower, flowering duration, and survival.




######
## Exclude genotypes without observations from both treatments???


```


Correlations
```{r}
corrplot(cor(na.omit(select(BB_allplants, c(seed_ct, lifet_fec_c, FFD_c, fl_dur_c, corolla_d_c, AG_biom_c)))), addCoef.col = "white", number.cex = 0.8, method = "square")
```




#### John M's Models
Statistical approaches matched to each of John's research questions for investigating plasticity in *N. menziesii*:  


1. What is the effect of supplemental water on functional traits and fitness in Nemophila menziesii?  

I.e.: What is the magnitude & direction of plastic changes in traits between ambient (dry) and watered plots?  

- Fecundity (seed count & estimated lifetime fecundity (mean seeds/fruit x total fruit))
- First flowering date
- Flowering duration
- Corolla diameter
- Aboveground biomass

Variable Treatment: All traits mean-centered

Analysis: Linear Regression - calculate Regression coefficients for Trait ~ Treatment
Models:
- Bivariate
- Bivariate + random effects

* = Significant coefficient/trend
^ = marginally significant trend

Positive coefficients = increased value in watered treatment


###### Statistical Tests

**Fecundity**

```{r}

#Fecundity - seed count ^
PL_fec_lm1 <- lm(seed_ct_c ~ Treatment, data = BB_Fplants)
summary(PL_fec_lm1)

PL_fec_lm1_r <- lmer(seed_ct_c ~ Treatment + (1|Recipient) + (1|Block), data = BB_Fplants)
summary(PL_fec_lm1_r)
r.squaredGLMM(PL_fec_lm1_r)




#### NS
# MV - Control for FFD and biomass 
PL_fec_lm1_mv <- lm(seed_ct_c ~ Treatment + FFD_c + AG_biom_c, data = BB_Fplants)
summary(PL_fec_lm1_mv)

PL_fec_lm1_mvr <- lmer(seed_ct_c ~ Treatment  + FFD_c + AG_biom_c + (1|Recipient) + (1|Block), data = BB_Fplants)
summary(PL_fec_lm1_mvr)
r.squaredGLMM(PL_fec_lm1_mvr)
ranef(PL_fec_lm1_mvr)
#Error boundary (singular) fit: may indicate that a component has very little effect on the outcome variable - remove it
#Location as random effect removed from this model


#
#
#####
#Fecundity - lifetime ^

PL_fec_lm2 <- lm(lifet_fec_c ~ Treatment, data = BB_Fplants)
summary(PL_fec_lm2)

PL_fec_lm2_r <- lmer(lifet_fec_c ~ Treatment + (1|Recipient) + (1|Location), data = BB_Fplants)
summary(PL_fec_lm2_r)
r.squaredGLMM(PL_fec_lm2_r)
ranef(PL_fec_lm2_r)
#Block as random results in singular model


#
#MV - Control for FFD and biomass
PL_fec_lm2_mv <- lm(lifet_fec_c ~ Treatment + FFD_c + AG_biom_c, data = BB_Fplants)
summary(PL_fec_lm2_mv)

PL_fec_lm2_mvr <- lmer(lifet_fec_c ~ Treatment + FFD_c + AG_biom_c + (1|Recipient) + (1|Block), data = BB_Fplants)
summary(PL_fec_lm2_mvr)
r.squaredGLMM(PL_fec_lm2_mvr)
ranef(PL_fec_lm2_mvr)

```

**First Flowering Date**

```{r}
#FFD
PL_FFD_lm1 <- lm(FFD_c ~ Treatment, data = BB_allplants)
summary(PL_FFD_lm1)

#Genotype as main effect
#PL_FFD_lm1_m <- lm(FFD_c ~ Treatment + Recipient + Location, data = BB_allplants)
#summary(PL_FFD_lm1_m)

PL_FFD_lm1_r <- lmer(FFD_c ~ Treatment + (1|Recipient) + (1|Location), data = BB_allplants)
summary(PL_FFD_lm1_r)
r.squaredGLMM(PL_FFD_lm1_r)
ranef(PL_FFD_lm1_r) #Block removed, doesn't help explain variance - BUT LOCATION explains a LOT of variance...

AIC(PL_FFD_lm1, PL_FFD_lm1_r)
#Random effects appear to always come out best compared to other models


#Diagnostics
#plot(PL_FFD_lm1_r)
```


**Flowering Duration**

```{r}
#F duration *
PL_Fdur_lm1 <- lm(fl_dur_c ~ Treatment, data = BB_allplants)
summary(PL_Fdur_lm1)

PL_Fdur_lm1_r <- lmer(fl_dur_c ~ Treatment + (1|Recipient) + (1|Block), data = BB_allplants)
summary(PL_Fdur_lm1_r)
r.squaredGLMM(PL_Fdur_lm1_r)
ranef(PL_Fdur_lm1_r)


#Control for FFD, Biomass
PL_Fdur_lm1_mv <- lm(fl_dur_c ~ Treatment + FFD_c + AG_biom_c, data = BB_allplants)
summary(PL_Fdur_lm1_mv)

PL_Fdur_lm1_mvr <- lmer(fl_dur_c ~ Treatment + FFD_c + (1|Recipient) + (1|Block), data = BB_allplants)
summary(PL_Fdur_lm1_mvr)
r.squaredGLMM(PL_Fdur_lm1_mvr)
ranef(PL_Fdur_lm1_mvr)

```

**Corolla Diameter**

```{r}
#Corolla diam *
PL_Cdiam_lm1 <- lm(corolla_d_c ~ Treatment, data = BB_Fplants)
summary(PL_Cdiam_lm1)

PL_Cdiam_lm1_r <- lmer(corolla_d_c ~ Treatment + AG_biom_c + (1|Recipient), data = BB_Fplants)
summary(PL_Cdiam_lm1_r)
r.squaredGLMM(PL_Cdiam_lm1_r)
ranef(PL_Cdiam_lm1_r) 


#Control for biomass
PL_Cdiam_lm1_mv <- lm(corolla_d_c ~ Treatment + AG_biom_c, data = BB_Fplants)
summary(PL_Cdiam_lm1_mv)

PL_Cdiam_lm1_mvr <- lmer(corolla_d_c ~ Treatment + AG_biom_c + (1|Recipient), data = BB_Fplants)
summary(PL_Cdiam_lm1_mvr)
r.squaredGLMM(PL_Cdiam_lm1_mvr)
ranef(PL_Cdiam_lm1_mvr) #Both Location or Block result in singular model




```

**Aboveground Biomass**

```{r}
#Biomass *
PL_AGbio_lm1 <- lm(AG_biom_c ~ Treatment, data = BB_Fplants)
summary(PL_AGbio_lm1)

PL_AGbio_lm1_r <- lmer(AG_biom_c ~ Treatment + (1|Recipient), data = BB_Fplants)
summary(PL_AGbio_lm1_r)
r.squaredGLMM(PL_Cdiam_lm1_r)
ranef(PL_Cdiam_lm1_r) #Both Location or Block result in singular model

```


```{r}
##Multivariate Models

#Use Fitness plants only (BB_Fplants)! Fewer NAs and 0s to deal with - more parsimonious data set

# Doesn't always make sense to control for additional variables - trait specific models
# Control for Block & Genotype as random effects - But Block may be redundant with watering effect! Location, however, explains very little variance in traits and results in singular MMs. Include Block for now, appears more important for explaining variance in ambient locations...

```

**Notes**:

Best/simplest approach for these analyses may be to focus on Bivariate models including random effects of Recipient and Block/Location. Multivariate models may help explain trait variance, but we can discuss interactions and correlations among traits separately (e.g., watering increased plant size and corolla area, and these two traits may be linked/ are both indicators of resource acquisition...)

Correlations controlling for watering:
AG_biomass ~ corolla_diam + Treatment...


To-Do:
- Create MMs for single main effect of treatment
- Use standardized Trait variables (compare both models)
- Model diagnostics
- Explore trait correlations that may account for some trends in watered plants


```{r}
#Visualizations

# Simple Reaction Norms


```



2. Is the strength and direction of plasticity in functional traits consistent (a) among populations and (b) within populations among genotypes? 

I.e: (a) At the population level, are plastic changes in traits with water addition consistent in direction and in magnitude? (b) Do plastic changes with watering among genotypes within populations mirror one another?  

**Note:** Use standardized trait values, but note that comparisons among populations (at least in magnitude) are fraught due to local climatic differences and differences in the amount of additional soil moisture actually experienced by watered plants relative to ambient plants at each site.  

Analysis:
- (a) Extract standardized regression coefficients from the bivariate models above (controlling for random effects) and compare these among populations
- (b) Document among-genotype variation in plasticity (all traits?) within populations by comparing variance measures (HOW DO??)
--- G*E interactions?
--- Variance components from random effect of genotype?



**Some questions**:
- How many watered v.s. non-watered fitness plants (observations) are there?
- On average, how many fitness plants are there to each Recipient ID (in each block)?
- What kind of trends in quantitative/numeric trait variables can you identify?
- How do trait values differ across locations and between experiments?










<br>  

---  



### John's Models

Some commented-out lines copied from his knit Rmds


### Bodega

#### Main Effects


```{r}
# FLOWERING DURATION
#BB_final_full$log_duration <- log(BB_final_full$fl_duration+1)
#BB_duration_Model = lmer(log_duration ~ watered + (1|Block) + (1|Recipient), data = BB_final_full, REML = T)

#*Best* simple linear model -log and sqrt both work?? sqrt+1 results in nice residuals, better adj R2
PL_Fdur_lm0 <- lm(fl_duration ~ Treatment, data = BB_allplants)
summary(PL_Fdur_lm0)
plot(PL_Fdur_lm0)
hist(PL_Fdur_lm0$residuals)
##UNSURE which transformation is best!

#John's model equivalent
PL_Fdur_lm1 <- lmer(log(fl_duration+1) ~ Treatment + (1|Recipient) + (1|Block), data = BB_allplants, REML = T)
summary(PL_Fdur_lm1)
r.squaredGLMM(PL_Fdur_lm1)
#ranef(PL_Fdur_lm1)


#Better model
PL_Fdur_lm2 <- lmer(fl_duration ~ Treatment + (1|Location), data = BB_allplants)
summary(PL_Fdur_lm2)
r.squaredGLMM(PL_Fdur_lm2)
ranef(PL_Fdur_lm2)
plot(PL_Fdur_lm2)
#Log transofrmation seems more appropriate here...? Better R2c
#Location results in singular model but block does not
```


```{r}
##FIRST FLOWERING DATE
#BB_FFD_Model = lmer(FFD_DOY ~ watered + (1|Block) + (1|Recipient), data = BB_final_full, REML = T)

#best lm 
PL_FFD_lm0 <- lm(FFD_z ~ Treatment, data = BB_allplants)
summary(PL_FFD_lm0)
plot(PL_FFD_lm0)
hist(PL_FFD_lm0$residuals)

#John's model equivalent 
PL_FFD_lm1 <- lmer(yday(FFD) ~ Treatment + (1|Recipient) + (1|Block), data = BB_allplants)
summary(PL_FFD_lm1)
r.squaredGLMM(PL_FFD_lm1)
ranef(PL_FFD_lm1) #Model is singular when using Block instead of Location


#Better model - remove Recipient? Still NS, same results
PL_FFD_lm2 <- lmer(yday(FFD) ~ Treatment + (1|Recipient) + (1|Location), data = BB_allplants)
summary(PL_FFD_lm2)
r.squaredGLMM(PL_FFD_lm2)
ranef(PL_FFD_lm2)

```


```{r}
# COROLLA DIAMETER
#BB_diameter_Model = lmer(corolla_diam_mm ~ watered + (1|Block) + (1|Recipient), data = BB_final, REML = T)

#best lm
PL_Cd_lm0 <- lm(corolla_diam_mm ~ Treatment, data = BB_Fplants)
summary(PL_Cd_lm0)
plot(PL_Cd_lm0)
hist(PL_Cd_lm0$residuals)

#John's equivalent model
PL_Cdiam_lm1 <- lmer(corolla_diam_mm ~ Treatment + (1|Block) + (1|Recipient), data = BB_Fplants)
summary(PL_Cdiam_lm1)
r.squaredGLMM(PL_Cdiam_lm1)
#ranef(PL_Cdiam_lm1) #Both Location or Block result in singular model...
#Block explains variation in corolla diameter better than treatment or Recipient!


#Better model? Same results
PL_Cdiam_lm2 <- lmer(corolla_diam_mm ~ Treatment + (1|Location) + (1|Recipient), data = BB_Fplants)
summary(PL_Cdiam_lm2)
r.squaredGLMM(PL_Cdiam_lm2)
#ranef(PL_Cdiam_lm2)
```


```{r}
#ABOVEGROUND BIOMASS
#BB_final$log_biomass <- log(BB_final$AG_biomass_mg)
#BB_biomass_Model = lmer(log_biomass ~ watered + (1|Block) + (1|Recipient), data = BB_final, REML = T)

#Best lm
PL_AG_lm0 <- lm(AG_biomass_mg ~ Treatment, data = BB_Fplants)
summary(PL_AG_lm0)
plot(PL_AG_lm0)
hist(PL_AG_lm0$residuals)
#Transformation unnecessary? Better R2, marginally NS


#John Equivalent Model
PL_AGbio_lm1 <- lmer(log(AG_biomass_mg) ~ Treatment + (1|Location) + (1|Recipient), data = BB_Fplants)
summary(PL_AGbio_lm1)
r.squaredGLMM(PL_Cdiam_lm1)
ranef(PL_Cdiam_lm1) #Both Location or Block result in singular model


#Better model? #similar results as lm
PL_AGbio_lm2 <- lmer(AG_biomass_mg ~ Treatment + (1|Location), data = BB_Fplants)
summary(PL_AGbio_lm2)
r.squaredGLMM(PL_Cdiam_lm2)
ranef(PL_Cdiam_lm2)
plot(PL_AGbio_lm2)
```


```{r}
# FECUNDITY
#BB_final2$log_seedprod <- log(BB_final2$total_est_seed_production)
#BB_seed_Model = lmer(log_seedprod ~ watered + (1|Block) + (1|Recipient), data = BB_final2, REML = T)

#lm
#filtered to individuals that produced seed
PL_fec_lm0 <- lm(log(total_est_seed_production) ~ Treatment, data = filter(BB_Fplants, surv_to_seedprod == 1))
summary(PL_fec_lm0)
plot(PL_fec_lm0)
hist(PL_fec_lm0$residuals)


#John's equivalent model
#Excludes plants that didn't survive to produce seeds!!
PL_fec_lm1 <- lmer(log(total_est_seed_production) ~ Treatment + (1|Recipient) + (1|Block), data = filter(BB_Fplants, surv_to_seedprod == 1))
summary(PL_fec_lm1)
r.squaredGLMM(PL_fec_lm1)
#! N = 80 individuals! compared to 225

#Better model?
PL_fec_lm2 <- lmer(log(total_est_seed_production) ~ Treatment + (1|Recipient) + (1|Location), data = filter(BB_Fplants, surv_to_seedprod == 1))
summary(PL_fec_lm2)
r.squaredGLMM(PL_fec_lm2)

##Use GLM?





## SEED COUNT - not analyzed
#lm


```

**GxE Effects**  

Based on John's models - (minor improvements made)  

```{r}
#FFD
GxE_FFD_lm0 <- lm(yday(FFD) ~ Recipient*Treatment + Location, data = filter(BB_allplants, Recipient != "BB_161"))
summary(GxE_FFD_lm0)
plot(GxE_FFD_lm0)
hist(GxE_FFD_lm0$residuals)

Anova(GxE_FFD_lm0, type = 3)

GxE_FFD_lm_rs <- lm(yday(FFD) ~ Recipient*Treatment + Location, data = BB_allplants)


#Flowering Duration
GxE_Fdur_lm0 <- lm(fl_duration ~ Recipient*Treatment + Location, data = filter(BB_allplants, Recipient != "BB_161"))
summary(GxE_Fdur_lm0)
plot(GxE_Fdur_lm0)
hist(GxE_Fdur_lm0$residuals)
#Significant GxE effect when log-transformed but not sqrt()

Anova(GxE_Fdur_lm0, type = 3)

#Corolla Diam
GxE_Cd_lm0 <- lm(corolla_diam_mm ~ Recipient*Treatment + Location, data = BB_Fplants)
summary(GxE_Cd_lm0)
plot(GxE_Cd_lm0)
hist(GxE_Cd_lm0$residuals)

#Anova(GxE_Cd_lm0, type = 3)

#Biomass
GxE_BIO_lm1<- lm(log(AG_biomass_mg) ~ Treatment*Recipient + Location, data = BB_Fplants)
summary(GxE_BIO_lm1)
plot(GxE_BIO_lm1)
hist(GxE_BIO_lm1$residuals)

#Anova(GxE_BIO_lm1, type = 3)

#Est. Fecundity - high VIFs
GxE_FEC_lm1 <- lm(log(total_est_seed_production) ~ Treatment*Recipient, data = filter(BB_Fplants, surv_to_seedprod == 1))
summary(GxE_FEC_lm1)
plot(GxE_FEC_lm1 )
hist(GxE_FEC_lm1 $residuals)
#poor sample sizes...

#Anova(GxE_FEC_lm1, type = 3)


#not filtered version
GxE_FEC_lm1 <- lm(log(total_est_seed_production+1) ~ Treatment*Recipient + Location, data = BB_Fplants)
summary(GxE_FEC_lm1)
plot(GxE_FEC_lm1 )
hist(GxE_FEC_lm1 $residuals)

#Anova(GxE_FEC_lm1, type = 3)
```







### Blue Oak  

```{r}

BO_allplants <- read_csv(here::here("data_sheets", "compiled_sheets", "BO_mastersheet_full_2023-03-01.csv")) %>% select(-c(Sequence, FitP, F_multi, F_plant_notes, Fl_3.11:flr_P, seeds_weighed_d1:msm_d9)) %>% 
  filter(Replicated == "R") %>% #Replicated genotypes only
  mutate(fl_duration = case_when(fl_duration == 0 ~ 1,
                                 TRUE ~ fl_duration)) %>%  #Replace durations of 0 with 1
  mutate(FFD_c = scale(yday(FFD), scale = FALSE)[,1], #convert to Julian day of year, but don't scale...
         fl_dur_c = scale(fl_duration, scale = FALSE)[,1], # Center all variables
         corolla_d_c = scale(corolla_diam_mm, scale = FALSE)[,1],
         AG_biom_c = scale(AG_biomass_mg, scale = FALSE)[,1],
         seed_ct_c = scale(seed_ct, scale = FALSE)[,1],
         lifet_fec_c = scale(total_est_seed_production, scale = FALSE)[,1]) %>% 
  mutate(FFD_z = scale(yday(FFD))[,1],
         fl_dur_z = scale(fl_duration)[,1], # Standardize all variables
         corolla_d_z = scale(corolla_diam_mm)[,1],
         AG_biom_z = scale(AG_biomass_mg)[,1],
         seed_ct_z = scale(seed_ct)[,1],
         lifet_fec_z = scale(total_est_seed_production)[,1]) %>% 
  mutate(seed_ct_r = seed_ct/mean(seed_ct),
         lifet_fec_r = total_est_seed_production/mean(total_est_seed_production)) %>% #fitness relative to population mean
  mutate(Treatment = as.factor(case_when(Block %in% c("1", "2", "3") ~ "ambient",
                               Block == "W" ~ "watered")),
         Treat_bin = as.numeric(case_when(Block %in% c("1", "2", "3") ~ 0, #binary var for treatment: ambient/watered = 0/1
                               Block == "W" ~ 1))) %>% 
  mutate(Recipient = as.factor(Recipient),
         Transect = case_when(Transect == "W1" ~ "13",
                              Transect == "W2" ~ "14",
                              TRUE ~ Transect)) %>% 
  mutate(Transect = as.numeric(Transect))



BO_allplants_f <- BO_allplants %>% 
  filter(!is.na(FFD_c)) #plants that flowered only

#318 obs vs 225 for fitness plants

##Fitness Plants Only
BO_Fplants <- BO_allplants %>% 
  filter(any_FitP == TRUE)



```



#### Main Effects  

```{r}
# FLOWERING DURATION

#*Best* simple linear model -log and sqrt both work?? sqrt+1 results in nice residuals, better adj R2
PL_Fdur_lm0 <- lm(fl_duration ~ Treatment, data = BO_allplants)
summary(PL_Fdur_lm0)
plot(PL_Fdur_lm0)
hist(PL_Fdur_lm0$residuals)
##UNSURE which transformation is best!

#John's model equivalent
PL_Fdur_lm1 <- lmer(fl_duration ~ Treatment + (1|Recipient) + (1|Transect), data = BO_allplants, REML = T)
summary(PL_Fdur_lm1)
r.squaredGLMM(PL_Fdur_lm1)
plot(PL_Fdur_lm1)
#ranef(PL_Fdur_lm1)


```


```{r}
##FIRST FLOWERING DATE
#BO_FFD_Model = lmer(FFD_DOY ~ watered + (1|Block) + (1|Recipient), data = BO_final_full, REML = T)

#best lm 
PL_FFD_lm0 <- lm(yday(FFD) ~ Treatment, data = BO_allplants)
summary(PL_FFD_lm0)
plot(PL_FFD_lm0)
hist(PL_FFD_lm0$residuals)

#John's model equivalent 
PL_FFD_lm1 <- lmer(yday(FFD) ~ Treatment + (1|Recipient) + (1|Transect), data = BO_allplants)
summary(PL_FFD_lm1)
r.squaredGLMM(PL_FFD_lm1)
ranef(PL_FFD_lm1) #Model is singular when using Block instead of Transect


```


```{r}
# COROLLA DIAMETER
#BO_diameter_Model = lmer(corolla_diam_mm ~ watered + (1|Block) + (1|Recipient), data = BO_final, REML = T)

#best lm
PL_Cd_lm0 <- lm(corolla_diam_mm ~ Treatment, data = BO_Fplants)
summary(PL_Cd_lm0)
plot(PL_Cd_lm0)
hist(PL_Cd_lm0$residuals)

#John's equivalent model
PL_Cdiam_lm1 <- lmer(corolla_diam_mm ~ Treatment + (1|Transect) + (1|Recipient), data = BO_Fplants)
summary(PL_Cdiam_lm1)
r.squaredGLMM(PL_Cdiam_lm1)
#ranef(PL_Cdiam_lm1) #Both Transect or Block result in singular model...
#Block explains variation in corolla diameter better than treatment or Recipient!

```


```{r}
#ABOVEGROUND BIOMASS
#BO_final$log_biomass <- log(BO_final$AG_biomass_mg)
#BO_biomass_Model = lmer(log_biomass ~ watered + (1|Block) + (1|Recipient), data = BO_final, REML = T)

#Best lm
PL_AG_lm0 <- lm(log(AG_biomass_mg) ~ Treatment, data = BO_Fplants)
summary(PL_AG_lm0)
plot(PL_AG_lm0)
hist(PL_AG_lm0$residuals)
#Transformation unnecessary? Better R2, marginally NS


#John Equivalent Model
PL_AGbio_lm1 <- lmer(log(AG_biomass_mg) ~ Treatment + (1|Transect) + (1|Recipient), data = BO_Fplants)
summary(PL_AGbio_lm1)
r.squaredGLMM(PL_Cdiam_lm1)
ranef(PL_Cdiam_lm1) #Both Transect or Block result in singular model


#Better model? #random effects don't do it
PL_AGbio_lm2 <- lm(log(AG_biomass_mg) ~ Treatment, data = BO_Fplants)
summary(PL_AGbio_lm2)
r.squaredGLMM(PL_Cdiam_lm2)
ranef(PL_Cdiam_lm2)

```


```{r}
# FECUNDITY
#BO_final2$log_seedprod <- log(BO_final2$total_est_seed_production)
#BO_seed_Model = lmer(log_seedprod ~ watered + (1|Block) + (1|Recipient), data = BO_final2, REML = T)

#lm
#filtered to individuals that produced seed
PL_fec_lm0 <- lm(log(total_est_seed_production) ~ Treatment, data = filter(BO_Fplants, surv_to_seedprod == 1))
summary(PL_fec_lm0)
plot(PL_fec_lm0)
hist(PL_fec_lm0$residuals)

#Not filtering data (log+1):
PL_fec_lm0 <- lm(log(total_est_seed_production+1) ~ Treatment, data = BO_Fplants)
summary(PL_fec_lm0)
plot(PL_fec_lm0)
hist(PL_fec_lm0$residuals)



#John's equivalent model
#Excludes plants that didn't survive to produce seeds!!
PL_fec_lm1 <- lmer(log(total_est_seed_production) ~ Treatment + (1|Recipient) + (1|Transect), data = filter(BO_Fplants, surv_to_seedprod == 1))
summary(PL_fec_lm1)
r.squaredGLMM(PL_fec_lm1)

#Not filtering -log+1
PL_fec_lm1 <- lmer(log(total_est_seed_production+1) ~ Treatment + (1|Recipient) + (1|Transect), data = BO_Fplants)
summary(PL_fec_lm1)
r.squaredGLMM(PL_fec_lm1)


```



#### GxE Effects in BO
```{r}
#FFD
GxE_FFD_lm0 <- lm(yday(FFD) ~ Recipient*Treatment + Transect, data = BO_allplants)
summary(GxE_FFD_lm0)
plot(GxE_FFD_lm0)
hist(GxE_FFD_lm0$residuals)

#Anova(GxE_FFD_lm0, type = 3)


#Flowering Duration
GxE_Fdur_lm0 <- lm(fl_duration ~ Recipient*Treatment + Transect, data = BO_allplants)
summary(GxE_Fdur_lm0)
plot(GxE_Fdur_lm0)
hist(GxE_Fdur_lm0$residuals)
#Significant GxE effect when log-transformed but not sqrt()


#Anova(GxE_Fdur_lm0, type = 3)


#Corolla Diam
GxE_Cd_lm0 <- lm(corolla_diam_mm ~ Recipient*Treatment + Transect, data = BO_Fplants)
summary(GxE_Cd_lm0)
plot(GxE_Cd_lm0)
hist(GxE_Cd_lm0$residuals)


#Biomass
GxE_BIO_lm1<- lm(log(AG_biomass_mg) ~ Treatment*Recipient + Transect, data = BO_Fplants)
summary(GxE_BIO_lm1)
plot(GxE_BIO_lm1)
hist(GxE_BIO_lm1$residuals)



#Est. Fecundity
GxE_FEC_lm1 <- lm(log(total_est_seed_production) ~ Treatment*Recipient, data = filter(BO_Fplants, surv_to_seedprod == 1))
summary(GxE_FEC_lm1)
plot(GxE_FEC_lm1 )
hist(GxE_FEC_lm1 $residuals)
#poor sample sizes...

#not filtered version
GxE_FEC_lm1 <- lm(log(total_est_seed_production+1) ~ Treatment*Recipient + Transect, data = BO_Fplants)
summary(GxE_FEC_lm1)
plot(GxE_FEC_lm1 )
hist(GxE_FEC_lm1 $residuals)

```




### Hastings


```{r}
HR_allplants <- read_csv(here::here("data_sheets", "compiled_sheets", "HR_mastersheet_full_2023-03-01.csv")) %>% select(-c(Sequence, FitP, F_multi, F_plant_notes, Fl_3.10:flr_P, seeds_weighed_d1:msm_d9)) %>% 
  filter(Replicated == "R") %>% #Replicated genotypes only
  mutate(fl_duration = case_when(fl_duration == 0 ~ 1,
                                 TRUE ~ fl_duration)) %>%  #Replace durations of 0 with 1
  mutate(FFD_c = scale(yday(FFD), scale = FALSE)[,1], #convert to Julian day of year, but don't scale...
         fl_dur_c = scale(fl_duration, scale = FALSE)[,1], # Center all variables
         corolla_d_c = scale(corolla_diam_mm, scale = FALSE)[,1],
         AG_biom_c = scale(AG_biomass_mg, scale = FALSE)[,1],
         seed_ct_c = scale(seed_ct, scale = FALSE)[,1],
         lifet_fec_c = scale(total_est_seed_production, scale = FALSE)[,1]) %>% 
  mutate(FFD_z = scale(yday(FFD))[,1],
         fl_dur_z = scale(fl_duration)[,1], # Standardize all variables
         corolla_d_z = scale(corolla_diam_mm)[,1],
         AG_biom_z = scale(AG_biomass_mg)[,1],
         seed_ct_z = scale(seed_ct)[,1],
         lifet_fec_z = scale(total_est_seed_production)[,1]) %>% 
  mutate(seed_ct_r = seed_ct/mean(seed_ct),
         lifet_fec_r = total_est_seed_production/mean(total_est_seed_production)) %>% #fitness relative to population mean
  mutate(Treatment = as.factor(case_when(Block %in% c("1", "2", "3") ~ "ambient",
                               Block == "W" ~ "watered")),
         Treat_bin = as.numeric(case_when(Block %in% c("1", "2", "3") ~ 0, #binary var for treatment: ambient/watered = 0/1
                               Block == "W" ~ 1))) %>% 
  mutate(Recipient = as.factor(Recipient),
         Transect = case_when(Transect == "W1" ~ "10",
                              Transect == "W2" ~ "11",
                              TRUE ~ Transect)) %>% 
  mutate(Transect = as.numeric(Transect))
         



HR_allplants_f <- HR_allplants %>% 
  filter(!is.na(FFD_c)) #plants that flowered only

#318 obs vs 225 for fitness plants

##Fitness Plants Only
HR_Fplants <- HR_allplants %>% 
  filter(any_FitP == TRUE)
```


#### Main Effects - HR  


```{r}
# FLOWERING DURATION

#*Best* simple linear model -log and sqrt both work?? sqrt+1 results in nice residuals, better adj R2
PL_Fdur_lm0 <- lm(fl_duration ~ Treatment, data = HR_allplants)
summary(PL_Fdur_lm0)
plot(PL_Fdur_lm0)
hist(PL_Fdur_lm0$residuals)
##UNSURE which transformation is best!

#John's model equivalent
PL_Fdur_lm1 <- lmer(fl_duration ~ Treatment + (1|Recipient) + (1|Transect), data = HR_allplants, REML = T)
summary(PL_Fdur_lm1)
r.squaredGLMM(PL_Fdur_lm1)
plot(PL_Fdur_lm1)
#ranef(PL_Fdur_lm1)


```


```{r}
##FIRST FLOWERING DATE

#best lm 
PL_FFD_lm0 <- lm(yday(FFD) ~ Treatment + Transect, data = HR_allplants)
summary(PL_FFD_lm0)
plot(PL_FFD_lm0)
hist(PL_FFD_lm0$residuals)

#John's model equivalent 
PL_FFD_lm1 <- lmer(yday(FFD) ~ Treatment + (1|Recipient) + (1|Transect), data = HR_allplants)
summary(PL_FFD_lm1)
r.squaredGLMM(PL_FFD_lm1)
ranef(PL_FFD_lm1) #Model is singular when using Block instead of Transect


```


```{r}
# COROLLA DIAMETER
#HR_diameter_Model = lmer(corolla_diam_mm ~ watered + (1|Block) + (1|Recipient), data = HR_final, REML = T)

#best lm
PL_Cd_lm0 <- lm(corolla_diam_mm ~ Treatment, data = HR_Fplants)
summary(PL_Cd_lm0)
plot(PL_Cd_lm0)
hist(PL_Cd_lm0$residuals)

#John's equivalent model
PL_Cdiam_lm1 <- lmer(corolla_diam_mm ~ Treatment + (1|Transect) + (1|Recipient), data = HR_Fplants)
summary(PL_Cdiam_lm1)
r.squaredGLMM(PL_Cdiam_lm1)
#ranef(PL_Cdiam_lm1) #Both Transect or Block result in singular model...
#Block explains variation in corolla diameter better than treatment or Recipient!

```


```{r}
#ABOVEGROUND BIOMASS
#HR_final$log_biomass <- log(HR_final$AG_biomass_mg)
#HR_biomass_Model = lmer(log_biomass ~ watered + (1|Block) + (1|Recipient), data = HR_final, REML = T)

#Best lm
PL_AG_lm0 <- lm(log(AG_biomass_mg) ~ Treatment, data = HR_Fplants)
summary(PL_AG_lm0)
plot(PL_AG_lm0)
hist(PL_AG_lm0$residuals)
#Transformation unnecessary? Better R2, marginally NS


#John Equivalent Model
PL_AGbio_lm1 <- lmer(log(AG_biomass_mg) ~ Treatment + (1|Transect) + (1|Recipient), data = HR_Fplants)
summary(PL_AGbio_lm1)
r.squaredGLMM(PL_Cdiam_lm1)
ranef(PL_Cdiam_lm1) #Both Transect or Block result in singular model

```


```{r}
# FECUNDITY
#HR_final2$log_seedprod <- log(HR_final2$total_est_seed_production)
#HR_seed_Model = lmer(log_seedprod ~ watered + (1|Block) + (1|Recipient), data = HR_final2, REML = T)

#lm
#filtered to individuals that produced seed
PL_fec_lm0 <- lm(log(total_est_seed_production+1) ~ Treatment, data = filter(HR_Fplants, surv_to_fruitprod == 1))
summary(PL_fec_lm0)
plot(PL_fec_lm0)
hist(PL_fec_lm0$residuals)

#Not filtering data (log+1):
PL_fec_lm0 <- lm(log(total_est_seed_production+1) ~ Treatment, data = HR_Fplants)
summary(PL_fec_lm0)
plot(PL_fec_lm0)
hist(PL_fec_lm0$residuals)



#John's equivalent model
#Excludes plants that didn't survive to produce seeds!!
PL_fec_lm1 <- lmer(log(total_est_seed_production+1) ~ Treatment + (1|Recipient) + (1|Transect), data = filter(HR_Fplants, surv_to_fruitprod == 1))
summary(PL_fec_lm1)
r.squaredGLMM(PL_fec_lm1)

#Not filtering -log+1
PL_fec_lm1 <- lmer(log(total_est_seed_production+1) ~ Treatment + (1|Recipient) + (1|Transect), data = HR_Fplants)
summary(PL_fec_lm1)
r.squaredGLMM(PL_fec_lm1)


```



#### GxE Effects in BO
```{r}
#FFD
GxE_FFD_lm0 <- lm(yday(FFD) ~ Recipient*Treatment + Transect, data = HR_allplants)
summary(GxE_FFD_lm0)
plot(GxE_FFD_lm0)
hist(GxE_FFD_lm0$residuals)

Anova(GxE_FFD_lm0, type = 3)

#Flowering Duration
GxE_Fdur_lm0 <- lm(fl_duration ~ Recipient*Treatment + Transect, data = HR_allplants)
summary(GxE_Fdur_lm0)
plot(GxE_Fdur_lm0)
hist(GxE_Fdur_lm0$residuals)
#Significant GxE effect when log-transformed but not sqrt()

Anova(GxE_Fdur_lm0, type = 3)

#Corolla Diam
GxE_Cd_lm0 <- lm(corolla_diam_mm ~ Recipient*Treatment, data = HR_Fplants)
summary(GxE_Cd_lm0)
plot(GxE_Cd_lm0)
hist(GxE_Cd_lm0$residuals)

#Anova(GxE_Cd_lm0, type = 3)


##Using the smaller data frame (fitness plants only) in the following models results in aliased coefficients. Report only GxE effects for flowering date and duration?


#Biomass
GxE_BIO_lm1<- lm(log(AG_biomass_mg) ~ Treatment*Recipient, data = HR_Fplants)
summary(GxE_BIO_lm1)
plot(GxE_BIO_lm1)
hist(GxE_BIO_lm1$residuals)

#Anova(GxE_BIO_lm1, type = 3)


#Est. Fecundity
#Filtered version - high VIFs!!!
GxE_FEC_lm1 <- lm(log(total_est_seed_production+1) ~ Treatment*Recipient + Transect, data = filter(HR_Fplants, surv_to_fruitprod == 1))
summary(GxE_FEC_lm1)
plot(GxE_FEC_lm1 )
hist(GxE_FEC_lm1 $residuals)
#poor sample sizes...

#Anova(GxE_FEC_lm1, type = 3)


#not filtered version
GxE_FEC_lm1 <- lm(log(total_est_seed_production+1) ~ Treatment*Recipient + Transect, data = HR_Fplants)
summary(GxE_FEC_lm1)
plot(GxE_FEC_lm1)
hist(GxE_FEC_lm1$residuals)


#Anova(GxE_FEC_lm1, type = 3)

```



Visualizations: GxE Effects - HR
```{r}

library(RColorBrewer)

## Hastings Reservation 

HR_filter_reps <- HR_allplants %>% #Use genotype means in each treatment to make reaction norms
  filter(!is.na(FFD)) %>% 
  mutate(Treatment = fct_recode(Treatment, "Watered" = "watered", "Ambient" = "ambient")) %>% 
  group_by(Recipient, Treatment) %>% 
  summarize(
    mean_FFD = mean(yday(FFD)),
    mean_Fdur = mean(fl_duration)
  ) %>% 
  ungroup()

colpal2 <- colorRampPalette(brewer.pal(8, "Accent"))(21)

ggplot(data = HR_filter_reps, aes(x=Treatment, y=mean_FFD, group = Recipient)) + 
  geom_line(aes(color = Recipient), linewidth = 1) + 
  labs(x= "Treatment", y = "First Flowering Date") +
  scale_x_discrete(expand = c(.1, .1)) +
  scale_color_manual(values = colpal2) +
  theme_classic(base_size = 16) + 
  theme(legend.position = "none")


ggplot(data = HR_filter_reps, aes(x=Treatment, y=mean_Fdur, group = Recipient)) + 
  geom_line(aes(color = Recipient), linewidth = 1) + 
  labs(x= "Treatment", y = "Flowering Duration") +
  scale_x_discrete(expand = c(.1, .1)) +
  scale_color_manual(values = colpal2) +
  theme_classic(base_size = 16) + 
  theme(legend.position = "none")



##Bodega Bay

BB_filter_reps <- BB_allplants %>% #Use genotype means in each treatment to make reaction norms
  filter(!is.na(FFD)) %>% 
  mutate(Treatment = fct_recode(Treatment, "Watered" = "watered", "Ambient" = "ambient")) %>% 
  group_by(Recipient, Treatment) %>% 
  summarize(
    mean_FFD = mean(yday(FFD)),
    mean_Fdur = mean(fl_duration)
  ) %>% 
  ungroup()


ggplot(data = BB_filter_reps, aes(x=Treatment, y=mean_FFD, group = Recipient)) + 
  geom_line(aes(color = Recipient), linewidth = 1) + 
  labs(x= "Treatment", y = "First Flowering Date") +
  scale_x_discrete(expand = c(.1, .1)) +
  scale_color_manual(values = colpal2) +
  theme_classic(base_size = 16) + 
  theme(legend.position = "none")


ggplot(data = BB_filter_reps, aes(x=Treatment, y=mean_Fdur, group = Recipient)) + 
  geom_line(aes(color = Recipient), linewidth = 1) + 
  labs(x= "Treatment", y = "Flowering Duration") +
  scale_x_discrete(expand = c(.1, .1)) +
  scale_color_manual(values = colpal2) +
  theme_classic(base_size = 16) + 
  theme(legend.position = "none")


```




Estimating genotypic variation in the effect of watering

Random slope models

```{r}
HR_FFD_m1mc <- lmer(yday(FFD) ~ Treatment + (1|Transect) + (1+Treatment|Recipient), data = HR_allplants, REML = FALSE)
summary(HR_FFD_m1mc)
r.squaredGLMM(HR_FFD_m1mc)
ranef(HR_FFD_m1mc)

```



